### 2019-09-18 Cover_Letter.Rmd
Manuscript structure created

---
title: 'Fish do nor need Visas: Reference Letter'
author: "Juliano Palacios Abrantes"
date: '2019-09-18'
output: pdf_document
---

Should include:
- Reference to any pre-submission discussions with editors.
- The title of the paper and a statement of its main point.
- Any information needed to ensure a fair review process, including related manuscripts submitted to other journals.
- Names of colleagues who have reviewed the paper.
- A statement that none of the material has been published or is under consideration for publication elsewhere.
- Specification of where all data underlying the study are available, or will be deposited, and whether there are any restrictions on data availability such as an MTA.
- Information on any reference material or additional data files uploaded to the Auxiliary files section (see below).
- Please also upload a .docx version of your cover letter – see below.