### 2020-02-04 Manuscript_Draft.Rmd
First corrections on comments from co authors

---
title: "Fish do not need visas: the transboundary nature of the world’s exploited marine fish species (79/96 characters)"
author: "Authors: Juliano Palacios-Abrantes^1^, Gabriel Reygondeau^1,2^, Colette Wabnitz^1,3^ and William W.L. Cheung^1^"
date: "Affiliations: ^1^Institute for the Oceans and Fisheries, University of British Columbia, Vancouver, Canada, ^2^Yale University, New Haven, Connecticut, USA, ^3^Stockholm Resilience Centre, Stockholm University, Stockholm, Sweden"
# subtitle: "DRAFT DO NOT SHARE"
subtitle: "Most neighbour nations share marine species (38/40 characters)"
output:
  word_document:
    # reference_docx: ./Manuscript_Instructions/science_template.docx
  pdf_document:
    fig_caption: yes
geometry: margin=1in
editor_options:
  chunk_output_type: inline
bibliography: ./References/FishForVisa_Ref.bib
csl: ./References/science.csl
---

```{r Ch2_setup, eval = T, echo=F, warning=F,message=F, results='hide'}

#### READ ME !!! ####
# Run this chunk before knit so you make sure you have all pkgs installed in R

ipak <- function(pkg){
  new.pkg <- pkg[!(pkg %in% installed.packages()[, "Package"])]
  if (length(new.pkg)) 
    install.packages(new.pkg, dependencies = TRUE,repos = "http://cran.us.r-project.org")
  sapply(pkg, require, character.only = TRUE)
}

#### Library ####
packages <- c(
  "dplyr", # Data manipulation
  "knitr",
  "kableExtra",
  "png",
  "grid",
  "here"

)

ipak(packages)

```

# Instructions for submission

- Journal: *Science*
<!-- - Information to authors (https://www.sciencemag.org/authors/science-information-authors) -->
<!-- - There are two types of article: -->

-- **Reports** (up to ~2500 words including references, notes and captions–corresponds to ~3 printed pages in the journal) present important new research results of broad significance. Reports should include an abstract, an introductory paragraph, up to four figures or tables, and about 30 references.

<!-- Materials and Methods should be included in supplementary material -->

<!-- --- Instructions for preparing an initial manuscript https://www.sciencemag.org/authors/instructions-preparing-initial-manuscript -->

# Reviewers (5):
*Include Names, affiliations, and e-mail addresses of up to five potential reviewers and up to five excluded reviewers.*

**Note:** Please include potential reviewers you think appropiate to review the manuscript (Or confirm my suggestions)

- **Potential Reviewers (J)**  
 -- Gakushi Ishimura, Faculty of Agriculture Department of Food Production and Environmental Management, Iwate University, Morioka, Japan; gakugaku@iwate-u.ac.jp
 -- Rebecca Selden, Wellesley College, Wellesley, MA, USA; bselden@wellesley.edu
 -- Julia Xue, Shanghai Jiao Tong University; juliaxue@ouc.edu.cn
 -- Claire Armstrong, The Norwegian College of Fishery Science, Tromso, Norway claire.armstrong@uit.no

- **Excluded Reviewers (J)**

# One Sentence Summary (125/~125 characters)
*A brief teaser statement highlighting main result of the paper, understandable by a scientist not in your field, without jargon or abbreviations. This will appear online adjacent to the title and should not repeat phrases already present there*

Species shared by neighboring nations contribute over 50% of marine capture fisheries' landings and revenue with direct implications for fisheries management.

# Field code (Max. 2)

- *Fisheries*

# Abstract (125/~125 words)
*An opening sentence that sets the question that you address and is comprehensible to the general reader, background content specific to this study, results, and a concluding sentence. It should be a single paragraph.*

Regulatory boundaries and species distributions are often not aligned, especially in the case of  marine species crossing multiple Economic Exclusive Zones (EEZs). This represents a challenge for fisheries management as policies tend to focus at the national level, yet collaborations are needed to maximize long-term ecological, social and economic benefits. Here, we combined species distributions and the spatial delimitation of EEZs to identify the number of commercially exploited marine fish species that are shared between neighbouring nations.  We found that 64% of the world’s catches in the last decade, amounted to 50% of fisheries revenue, comes from 633 shared species. Some countries received over 70% of their benefits from a few shared resources. This study highlights the large and under-recognized challenge in achieving international fisheries management goals, particularly as species redistribute under climate change.

# Main Article
<!-- # Introduction (~555/500 words) -->

Distributions of marine species around the world are not constrained by human-made boundaries; rather they are shaped by biotic and abiotic factors defined by their evolutionary history [@Hutchinson:1957nt; @Nelson:2016kk]. Thus, a species can have a broader distribution (cosmopolitan), or be endemic [@Nelson:2016kk]. However, fisheries management is predicated on the definition of “stocks” that, in most cases, are delineated by human-made spatial boundaries that often do not correspond to biologically-meaningful population units [@Song:2017iua; @FredstonHermann:2018kp]. The delineation of Economic Exclusive Zones (EEZs) under the United Nations Convention on the Law of the Seas (UNCLOS) in the early 80s, for instance, established political boundaries across the distribution of many species, creating shared species between nations. Theories and empirical experience have show that fisheries targeting resources that straddle between political boundaries complicate fisheries management and potentially reduce its effectiveness to achieve stated objectives [@Munro:2002uf; @Bailey:2010ga]. For instance, climate-driven changes in species distributions have led to conflict between nations due to changes in the proportion of captures [@Munro:2002uf], quota allocation, and fishery newcomers [@Spijkers:2017ij]. Moreover, climate change is likely to exacerbate these types of conflict and presents new challenges for political relations between neighboring countries [@Pinsky:2018cb] and for fisheries management [@Miller:2013iv].

Forty years after the formal adoption of UNCLOS, accurate estimates of the number of exploited marine species that are shared by neighboring nations (hereafter called transboundary species [@Munro:2002uf]) is still unclear. **Early data limitations allowed only for an informed guessed [@Caddy:1997ue], while recent estimations, followed a litterature review and catch data to address this issue [@Teh:2015gd]. However, up to date, no method has included the actual distribution of species. We hipothesize that the methodological constraints of preious work has underestimate the contribution of these species to global catch and revenue and that the proportoin of transboundary species is larger than previouslly estimated.** Thus, we aimed to estimate the number of exploited marine fish species shared by neighbouring countries and determine their contribution to global catch and fisheries revenue. **Moreover, we estimate their importance in terms of catch and revenue by country and also carry the analysis at the ecological level.**

We overlaid the known distribution of **968** commercially valuable marine species responsable for **%%** of global catch, and the EEZ of 153 coastal countries, and developed a four-criteria methodology to determine if a species was transboundary. First we only considered shared species between neighboring EEZs, rather than the species' extended distribution (e.g. we did not include the high seas). Secondly, we considered a species to be present in a grid cell if all data sources predicted positive occurrence (see Methods). The third criteria determined whether a species was exploited within a given grid cell using spatially explicit catch data. A species was considered as present in a grid cell if both predicted species occurrences and catch data were positive. Finally, we determined a spatial threshold based on the proportion of the species' distribution within each neighboring EEZs to confirm occurance in both EEZs.

Once transboundary species were identified, we estimated their contribution to fisheries catch and revenues and categorized them according to catch trends. We define a ‘stock’ unit as a species in an EEZ, instead of using a population ecology definition (genetically or morphological distinct units) due to lack of such biological information for almost all the species included in this study [@Teh:2015gd]. While we acknowledge that a species could have multiple stocks within an EEZ, many fisheries within a country or EEZ are managed at the species instead of stock level. Also, recent research has suggested strong connectivity between stocks, even when separated by thousand of kilometers, **and providing additional ecological ground for...** [@Ramesh:2019va].

<!-- ## Results and discussion (~1000 words) -->

We identified **633** transboundary exploited species in the world, more than double of previous estimates [@Teh:2015gd], representing **65.4%** of the species analyzed (Fig. 1A). Between 2005 and 2014, national fleets targeting shared species within EEZs landed an average of **69.2** million tonnes, representing **64%** of the EEZ-derrived catches, and generating USD **79,745** million in fishing revenue. This is equivalent to **49.8%** of the global fishing revenue of the same period. These values are more than double the 34.2 million t and USD 30,700 million in fishing revenue previously estimated [@Teh:2015gd], underscoring the importance of transboundary species at the global level.

```{r Call_Map, eval = F, fig.width=6, fig.height=4, echo=FALSE, fig.pos= "H", fig.cap="Number of transboundary species and they contribution to fishing catch and revenue.** A) The number of species is represented in the EEZ while revenue from these species fished within the EEZ is represented in the land polygon. B) Regional revenue (left) and catch (right) contribution of tranbsoundary species. Points = mean +- sd."}

img <- readPNG(here("./Figures/Fig1.png"))
grid.raster(img)
```

![Fig. 1. Number of transboundary marine species and their contribution to global fisheries catches and revenue. A) The number of species and amount of revenue are represented by colour coding of EEZs and land polygons, respectively. B) Contribution of tranbsoundary species to regional revenue (left) and catch (right). Points = mean +- sd. Revenue in 2010 real USD](./Figures/Fig1.png)

**NOTE: EXPLORE THE REGIONAL SEGMENTATION YOU MADE**

Transboundary species are particularly important economically for Northern America (3,901 $\pm$ 4,497 million USD) and Eastern Asia (2,827 $\pm$ = 2,667 million USD), significantly differing from other regions in terms of revenue (one-way ANOVA; *F*(16,157) = 4.91, *p* < 0.001, $\alpha$ = 0.05) (Fig 1B). China (USD 6,279 million) in E. Asia and the US (USD 9,605 million) in N. America contribute 55% and 83% to their respective region's revenue, respectively. Russia (USD 6,361 million), Peru (USD 5,211) and Japan (USD 3,369) complete the top five countries with the most fishing revenue generated from transboundary species. These five nations are responsible for 30% of the fishing global revenue from transboundary species. Peru and Russia, responsible for the two largest fisheries in the world [@FAO:2018KK], lead transboundary catch statistics harvesting over 5,000 thousand tones each on average between 2005 and 2014 (**Supplemental figure Trans_Catch_MAP**). Peru's catches are mainly composed of Peruvian anchoveta (*Engraulis ringens*), shared with Chile, and accounting for 79% of the transboundary species production. Similarly, Russia shares the Alaskan polllock (*Theragra chalcogramma*) fishery with both Japan and the US. Despite the large catch volume of Peru and Russia, regional catches are only significantly different in E. Asia (one-way ANOVA; *F*(16,155) = 3.341, *p* < 0.001, $\alpha$ = 0.05), with China, the world's top fish producer with 15.3 million tons landed in 2016 [@FAO:2018KK] getting around 1/3 of its capture from transboundary species, followed by Japan (4.2 million tonnes) and South Korea (2.3 million tonnes). **NOTE: MAKE SURE RUSSIA IS NOT E.ASIA**

We determined the exploitation category of transboundary fisheries according to each country's catch [@Kleisner:2011wn] and also found significant differences among exploitation categories (one-way ANOVA *F*(5,446), *f* = 324.5, *p* < 0.001, $\alpha$ = 0.05; Tukey's post hoc test *p* < 0.05 for all) with most species fished to levels below 50%, sometimes lower than 10%, of their historical maximum catch (Fig.2). All regions conformed to this trend, except for Oceania, with no significant difference across continents (one-way ANOVA *F*(5,443), *f* = 1.24, *p* = 0.289, $\alpha$ = 0.05). While many environmental and social-economic factors such as temperature, markets, fishing policies affect catch volume [@Pauly:2013gs; @Branch:2008kk] previous studies demonstrated that the trends can be used to infer whether a stock is over-exploited or collapsed [@Kleisner:2011wn]. Thus suggesting that the current global exploitation rate of transboundary species could be over the maximum sustainable yield [@Kleisner:2011wn; @Pauly:2013gs].

**NOTE: THIS (L 118) PARAGRPAH NEEDS WORK**

```{r Call_Circ_Country, eval = F, fig.width=6, fig.height=4, echo=FALSE, fig.pos= "H", fig.cap="Number of transboundary species by catch category. Category A, fishing between 10%-50% of maximum catch level after a drop larger than 10% of the recorded maximum catch; Category B, fishing over 50% of maximum catch level; Category C, fishing less than 50% of maximum catch"}

img <- readPNG(here("./Figures/Fig2.png"))
grid.raster(img)
```

![Fig. 2. Number of transboundary species by catch category. Category A, fishing between 10%-50% of maximum catch level after a drop larger than 10% of the recorded maximum catch; Category B, fishing over 50% of maximum catch level; Category C, fishing less than 50% of maximum catch](./Figures/Fig2.png)

Empirical analysis suggests that in most cases, management of transboundary species will have better outcomes in terms of fish yields when nations cooperate [@Bailey:2010ga]. Examples of successful co-management include Canada and the United States over Pacific salmon [@REF] and Pacific halibut [@REF], Norway and Russia over Atlantic cod (*Gadus morhua*), haddock (*Melanogrammus aeglefinus*) and capelin (*Mallotus villosus*) in the Barents Sea [@Diekert:2010gp; @Eide:2013dha **NOTE: DOUBLE CHECK REFERENCES]**, and Namibia and South Africa over hake (*Merluccius spp*) [@Sumaila:2003vw]. Lack of collaboration over shared stocks may threaten stock sustainability [@Clark:1980js; @Nguyen:2018dk], and reduce the future profitability potential of the fishery [@Merino:2007jz]. Yet, we acknowledge that cooperation or lack thereof can be a complex process [@Jensen:2015cf] and that there are specific cases where cooperation might not be the best strategy [@REF]. Collaboration can also reduce conflict between nations over shared resources **as climate change shifts the distribution of species**. Transboundary fisheries are abundant and of economic importance for some of the top countries with past fisheries-related conflicts including Canada, the US, China and Russia [@Spijkers:2019cz]. For example, the United States and Canada had a dispute over Pacific salmons (*Oncorhynchus sp*) that lasted over 10 years and was only resolved after these two nations set a series of regulations were introduced to compensate for losses [@Munro:2002uf]. Since 2007, the European Union, Norway, Iceland, and the Faroe Islands have been at odds over the size and relative allocation of total allowable catches (TACs) for Atlantic mackerel (*Scomber scombrus*) [@Spijkers:2017ij], a species that is transboundary and contributes **698.3** (**NOTE: %%**) thousand tonnes to catches and USD **968** (**NOTE: %%**) million in revenue to the countries involved in the dispute. In addition to past and present conflicts, climate change is likely to exacerbate conflicts over transboundary species due to shifts in their distribution, requiring flexible agreements that accommodate future uncertainties [@Pinsky:2018cb].

<!-- First steps for shared management start with information sharing, then this and that. Proper management treaties not only have but can also be stronger in the fase of a canging climate. -->

Over 130 marine fish species occur in tropical and subtropical waters around the world [@Nelson:2016kk], many of which are highly migratory pelagic-oceanic species like skipjack tuna (*Katsuwonus pelamis*), other tunas (*Thunnus sp.*), sword fish (*Xiphias gladius*), and common dolphinfish (*Coryphaena hippurus*). Pelagic-oceanic species are the only group whose EEZ range (i.e., the number of EEZs where the species occur as 'transboundary') was significantly different from other species (one-way ANOVA *F*(5,109), *f* = 14.67, *p* < 0.001, $\alpha$ = 0.05; Tukey's post hoc test *p* < 0.05) with a median of 40 EEZs per species (Fig. 3).

```{r Call_Spp_Circular, eval = F, fig.width=6, fig.height=4, echo=FALSE, fig.pos= "H", fig.cap="Number of countries shared by transboundary species. Showing only species that share > 20 countries"}

img <- readPNG(here("./Figures/Fig3.png"))
grid.raster(img)
```

![Fig. 3. EEZ range of ransboundary species. Showing only species that share > 20 countries](./Figures/Fig3.png)

Due to their large migration patterns and presence in areas beyond national jurisdiction, many highly migraroty pelagic fish species are managed by Regional Fisheries Management Organizations (RFMOs). Fifteen major RFMOs in the world oversee the management of 99 species [@CullisSuzuki:2010fi]. We estimated that 81.8% of these species are shared across neighboring RFMOs, with the International Commission for the Conservation of Atlantic Tunas (ICCAT) sharing over 40 species with neighboring RFMOs (**Supplemental figure RFMO_MAP**). The effectiveness of RFMOs responsible for managing high migratory stocks has been previously questioned [@CullisSuzuki:2010fi; @Lodge:2007bt]. However, despite a global expansion of the industrial fleet in the last decade [@Coulter:2020gr], a recent analysis of 13 species (40 stocks) suggests that most tuna and billfish stocks are sustainably harvested ($B/B_{MSY} > 1$ and $F/F_{MSY} < 1$) with only three stocks currently considered as overfished ($B/B_{MSY} < 1$ and $F/F_{MSY} > 1$) [@Pons:2016ep]. Within national waters, we estimated that the catch of transboundary pelagic-oceanic species mainly originates from five large ocean states in the South Pacific, three of which (Kiribati, Solomon Islands and Papua New Guinea) are parties to the Pacific Nauru Agreement (Fig 3B). Parties to the Nauru Agreement sustainably manage 30% of global skipjack tuna [@Aqorau:2018bh], estimated to produce $2 billion worth of fish every year, representing a role model for transboundary fisheries management [@Hanich:2010ht].

**NOTE: Colette will search for a reference related a weird comment "arecenWCPF assessment highlight tt tuna in the Pacific e sustsinabl fisd, but oher species  socks in ot regionsre . I will dig up e reference and send it to you - just to make sure  the irmati overlaps"**

<!-- [@https://www.ffa.int/nauru_agreement]  -->

The median for species of all other ecosystem preferences is close to, or less than, 20, as many of these species have a narrower distribution or are less mobile (e.g. *Solea solea*, *Panulirus argus*) (Fig. 3). However, less distributed species must not be ignored. The world of fisheries is highly connected trough larvae dispersal [@Ramesh:2019va] and so, even centralized populations could be supplying populations thousands of kilometers away [@Ramesh:2019va]. Moreover, many coastal communities, specially in developing regions, and highly dependent on subsistance fishing (e.g. reef fish) for food security and livelihoods [@CisnerosMontemayor:2016gq; Wabnitz:2018gf; @Cabral:2018hc]. For example, from the top five countries fishing transboundary reef - associated species (Fig. 3B), Indonesia, Malaysia and the Philippines are among the highest fish consumption per capita in the world with 20, 60 and 40 kg per year, respectively [@Foale:2013iw]. In these countries, reeef fish contributes substantially to both food security and livelehood of coastal communities [@Cabral:2018hc].

In our analyses we did not seek to include climate change as a driver of species’ distribution. Distribution shifts are the most commonly recorded response of marine species to the effects of climate change - chiefly warming - in the ocean [@Poloczanska:2016kk]. Catch composition might have already changed in the last five decades [@Cheung:2013gk], and will continue throuugh the $21^{st}$ century [@Cheung:2016jd]. Previous study projected that climate change will bring at least one new transboundary species per EEZ in the future [@Pinsky:2018cb], while also change the proportion of the distribution (e.g. Area Index) that each county’s share to a degree that could influence fisheries management (*Palacios-Abrantes et.al. under review*). Exploring the detailed impacts of climate change in share species is key to develop local adaptation methods to minimize the effect of such distributional change on the conservation of species as well as the economy and food security of coastal nations.

Our study identified species currently shared by neighbouring coastal nations and highlights the importance of these species' contribution to global capture and revenue from wild fisheries. Our results show that captures and revenues from shared species are higher than previously estimated and particularly important to countries in Northern America and Eastern Asia. This result highlights the importance of these fisheries and their potential contribution to regional food security and wellbeing. Moreover, the vast majority of transboundary species are harvest at low rates with no geographic variation. Previous work has found that collaboration will result in better outcomes for shared marine resources. Thus, countries should pay specific attention in the management of the identified transboundary stocks potentially creating collaborative management plans. Identifying existing transboundary species is the first step towards joint management frameworks that are precautionary, strive for sustainability, and can be flexible to accommodate the uncertain future driven by climate change.

<!-- # **Acknowledgments* * -->

<!-- ## Funding: include complete funding information;  -->
<!-- ## Authors contributions: a complete list of contributions to the paper (we encourage you to follow the CRediT model), -->
<!-- ## Competing interests: competing interests of any of the authors must be listed (all authors must also fill out the Conflict of Interest form). Where authors have no competing interests, this should also be declared. -->

## Data and materials availability

All data is available from the authors upon request.

# Materials and Methods

To ascertain whether or not a species can be considered transboundary, we developed a method based on a series of criteria and the overlay of species' distributions and the spatial boundaries of the world’s EEZs. All analyses were conducted at the species level and we consider transboundary species as those that occur within the Exclusive Economic Zone (EEZ) of neighboring countries [@Munro:2002uf]. All data was scaled to $0.5^o$ x $0.5^o$ lat-long grid cells.

## Databases of species' geographic distribution

To determine the number of transboundary marine species exploited by fisheries within each of the world's EEZs we used the list of commercial marine fish and invertebrate species from the *Sea Around Us* (SAU) database (http://www.seaaroundus.org) and determined their distributions. To this end, we used four data sources of species-distributions: (*i*) occurence data, (*ii*) an Ensemble Environmental Niche Model (ENMs), (*iii*) a life-history-based distribution model, and (*iv*) fisheries catch data (**Table x**). Each source represents a different method of estimating the distribution of a given species and thus, provides a more robust result. Only commercial fished species with data from all four sources were included in the analysis **NOTE: WC "Need to discuss how this affect the samples of transboundary species."**. The final dataset comprised 968 marine species (including anadromous species like salmononids), **XX%** of the total number of reported marine species in the SAU database and **XX%** of the reported marine taxa of included by the Food and Agricultural Organization of the United Nations (FAO).

<!-- **NOTE: for transboundary lobsters https://books.google.com.ar/books?id=XEElAwNrKZsC&pg=PA224&lpg=PA224&dq=belize+mexico+lobster+shared+stock&source=bl&ots=lsmUH743d_&sig=ACfU3U1P1t7_L8q_YRh3oQGszUO8Lg1jFw&hl=en&sa=X&ved=2ahUKEwiv36qbwbjnAhX3HrkGHYIbCTgQ6AEwFHoECA0QAQ#v=onepage&q=belize%20mexico%20lobster%20shared%20stock&f=false** -->

### *Occurence data*

Occurrence data was collected from five publicly available repositories: Fishbase (http://fishbase.org), the Global Biodiversity Information Facility (GBIF; https://www.gbif.org/), the Ocean Biogeographic Information System (OBIS; https://obis.org/),  the Intergovernmental Oceanographic Comission (IOC; http://ioc-unesco.org), and the International Union for Conservation of Nature (IUCN; https://www.iucn.org/technical-documents/spatial-data) [@Reygondeau:2019uh].

### *Distribution models*

In addition to occurrence data, we used two different methods to estimate species distributions, hereafter referred to as Ecological Niche Model-Nereus (ENM-Nereus) and Species Distribution Model-SAU (SDM-SAU). Although they use the same occurrence and environmental data, the models are structurally different complementing each other and providing robustness to the results.

The ENM-Nereus consists in a multimodel approach based on a Bioclim and a Boosted Regression Tree model [@Thuiller:2009gp], a Maxent model [@Phillips:2006ff], and a Non-Parametric Probabilistic Ecological Niche Model [@Beaugrand:2011fd]. Environmental variables utilized in the models include sea surface temperature, surface pH, surface oxygen concentration, and vertically integrated (0–100 m) net primary production (NPP) [@Asch:2018ca]. The ENM-Nereus employed all of the observational data-sources previously mentioned. Global environmental conditions were average for the period 1970-2000 and averaged for three Earth System Models developed by the Geophysical Fluid Dynamics Laboratory (GFDL- https://www.gfdl.noaa.gov/earth-system-model/), the Institute Pierre Simon Laplace (IPSL- www.icmc.ipsl.fr/), and the Max Planck Institute for Meteorology (MPI- www.mpimet.mpg.de/en/science/models/). See [@Asch:2018ca; @Reygondeau:2019uh] for model details.

The SDM-SAU model follows a five-steps process based on species-specific life history data, rather than environmental variables [@Close:2006ux; @Palomares:wg]. For each commercial species, the model first uses the FAO major fishing areas and countries' EEZs to determine a broad distribution. It then uses life history information to delimit its range within the FAO fishing area (e.g., thermal preference, depth limit). The range is delimited even further by expert-review polygons and compared with that of AquaMaps [@Kaschner:2016tl], OBIS and GBIF occurrence data. The model then determines a species habitat preference based on the assumption that the relative abundance of a species is determined by the number of habitats in a grid cell and the distance of the species to each habitat, as well as the importance of the habitat to the species size. Finally, the species equatorial submergence (e.g. the latitudinal region where a species is not seen in between poles) is estimated for each species. See [@Close:2006ux; @Palomares:wg] for model details.

### *Catch data*

The previous models combine observational data with a series of biotic and abiotic information to determine the probability that a species will be found in a given space at a given time. However, this does not mean that the species in question will actually be there. While the models do use approaches to double-check species occurrences (e.g., ENM-Nereus uses four different species distribution algorithms and SPD-SAU undertakes validation by means of other models), we used a fourth data source to corroborate the models’ outputs. The Sea Around Us estimates total reconstructed catches - catches based on all publicly available information sources and including discards, as well as unreported and illegal catches that are not included in available FAO data - for each country. Catches are also spatially allocated [@Zeller2016]. Thus, we used the Sea Around Us catch reconstruction database from 2005 to 2014 as the fourth dataset to estimate transboundary species and to estimate their catch contribution within EEZs. **We selected this time frame to investigate the recent (our time frame includes the last ten years of data) contribution of transboundary fisheries and to reduce the uncertanty embeded in the recosntruction process (See Key uncertanties).**

```{r Data_Table, eval = T, echo = F}

Dasets_names <- c(
  "Occurrence",
  "ENM-Nereus",
  "SDM-SAU",
  "Catch-SAU"
)
Brieff_Description <- c(
  "Occurrence data from multiple sources",
  "Environmental niche model based on environmental variables and different algorithms",
  "Species distribution model based on species traits",
  "Spatial catch allocation based on country-by-country reconstructions"
)

Sources <-  c(
  "Reygondeau 2019",
  "Asch et al., 2017 & Reygondeau 2019",
  "Palomares et al., 2016",
  "Zeller et al., 2016"
)

data.frame(Dasets_names,
           Brieff_Description,
           Sources) %>% 
  kable("latex", booktabs = T,
        caption = "Summary table of data sources used in the current analysis") %>%
  kable_styling(full_width = F) %>%
  column_spec(2,width = "8cm") %>% #wraps text in column2
  column_spec(3,width = "3cm") #%>% #wraps text in column2

```

## Determining transboundary species trait

We developed a four-criteria methodology to determine whether or not a species can be considered transboundary. Only species that meet all criteria were considered as transboundary and the analysis was undertaken only within the EEZs of coastal states. In some cases, criteria can also be interpreted as indices to measure the uncertainty in the analysis as they provide a continuous index (e.g. from 0 to 1) rather than a discrete, yes or no, index. All analyses were run using the statistical software *R version 3.5.2 (2018-12-20)* with the packages *data.table*[@Packagedatatable:2019uh], *janitor*[@PackagejanitorSim:2018ur], *wesanderson*[@Packagewesanderson:2018vn], *rfishbase*[@PackagerfishbaseR:2019th], *R.matlab*[@PackageRmatlabRe:2018un], *sf*[@PackagesfSimpleF:2018vp], *sp*[@PackagespClasses:2019tn], *tidiverse*[@PackagetidyverseE:2017vq], *tidytext*[@PackagetidytextTe:2019vk], and *zoo*[@PackagezooSInfr:2019tm]. All code is available at https://github.com/jepa/FishForVisa.

### Criteria 1; Neighboring EEZs

As previously mentioned, we define transboundary species as those marine species that occur within the EEZs of two or more neighboring countries. Hence, according to this criteria a species was only considered as transboundary if it was shared between two neighboring countries, regardless of the species extended distribution. We defined EEZ boundaries using the Sea Around Us shapefile (updated 1 July 2015, available from http://www.seaaroundus.org), noting that it subdivides EEZs by regions (e.g., Mexico Pacific and Mexico Atlantic), and determined the intersections between polygons using the R package sf. When estimating transboundary species, we filtered out those shared by EEZs sub-regions, and when aggregating results by country, species that occurred in more than one sub-region were only accounted for once. Species that were present in EEZs that were non-continental territories neighboring other countries were kept separately (e.g., Argentina and Falkland Islands), but removed in cases where the non-continental territory belonged to the same nation (e.g., Brazil and Fernando de Noronha). Associated states like Puerto Rico and Hong Kong were added to their associated Sovereign (e.g. Puerto Rico with the United States and Hong Kong with China).

### Criteria 2; Data agreement

We used the occurrence database, the ENM-Nereus model, and SDM-SAU model to determine the presence of each species within each of the $0.5^{\circ}$ x $0.5^{\circ}$ grid cells. We computed a *Species Index* by dividing the number of datasets that confirmed the presence of a species in a grid cell over the total number of datasets (n = 3). The index is unit-less and represents a scale from 0.3 to 1, with 0.3 indicating that only one dataset reported the presence of the species in a grid cell, 0.6 two out of three sources, and 1 pointing to agreement across all 3 datasets. All analyses only considered cases where *Species Index* = 1 to obtain a more conservative estimate of transboundary species.

### Criteria 3; Modeling verification

We assumed that a species was only present in a given grid-cell if it was reported in the Sea Around Us catch database. Therefore, all species that were not reported as caught in any single year between the reference years (2005 to 2014) in a given grid-cell were dropped, regardless of the *Species Index*. This rule assumes that if a commercial species is projected within any fishing country, such a species would have been fished, and thus reported at some point over a period of time (here 2005 - 2014), thus validating the models. See [@Pauly:2019bf] and [@Zeller2016] for catch reconstruction and spatial allocation details.

### Criteria 4; Spatial distribution

Finally, to have a more robust result and not categorize a species as transboundary based on its presence in a single $0.5^{\circ}$ x $0.5^{\circ}$ grid cell within an EEZ, we computed an *Area Index*. The *Area Index* consist in the proportion of a given species' overarching shared distribution between neighbouring EEZs accounted for by each individual EEZ. We classified a species as transboundary if both neighboring EEZs enclosed over 25% of the species joint shared distribution. This threshold can be lowered for a more relaxed result or increased for a more conservative estimate (**S2_Hist_per_Treshold**).

## Fisheries trends

We estimated the economic contribution in 2010 real USD of transboundary species for each country using global ex-vessel price data [@Tai2017]. The database we draw from includes ex-vessel price from multiple sources and a structured interpolation method (e.g. similar countries, species) to fill in data gaps [@Sumaila:2015uc]. The contribution of transboundary species was derived by aggregating the catch of species occurring within the EEZs of countries sharing that species. For this study, we did not include catch nor revenue from areas beyond national jurisdiction. Finally, we used the catch data to determine the exploitation category of each species within a EEZ. Although this method has previously been used to estimate stock status [@Grainger:1996tk], the categories presented here are intended as catch trends, and not status for each species. We only assessed species within each EEZ for which at least ten years of data were available between the first and last reported catch and with at least five consecutive years of data. Three final categories were drawn for each species depending on the catch volume within each EEZ (e.g. present, maximum and minimum EEZ's historical catch) and the year (e.g. year of maximum historical catch of the species within that EEZ) (**Table XX**) [@Kleisner:2011wn]. Finally, we reported the predominant category over the last ten years of catch data (2005-2014). We performed a series of one-way analysis of variance (one-way ANOVA) to determine statistically significant differences between the means of different groups (e.g., regions, species, exploitation category) of transboundary species and their contribution to catch and fishing revenue.

```{r Status_Table, eval = T, echo = F}

### Original version

# Stock_Status <- c("Rebuilding","Developing","Max Exploited","Over Exploited","Collapsed","No Status")

# Rules <- c(
#   "(Year of Catch > Year Post Max. Min. & Post Max Min Catch < (Max Catch*0.10)) & (Catch > (Max Catch*0.10) & Catch < (Max Catch*0.50))", # Reb.
#   "Year of Catch < Year of Max. Catch & Catch <= (Max Catch*0.50) or Year of Max Catch = Last Year of data)", #Dev.
#   "Catch > (Max Catch*0.50)", #Max Exp.
#   "Year of Catch > Year of Max. Catch & (Catch > (Max. Catch*0.10) & Catch < (Max. Catch*0.50)", # Over.
#   "Year or Catch > Year of Max. Catch & Catch < Max. Catch*0.10", # Colla.
#   "Non of the above rules applied"
# )

# data.frame(Stock_Status,
#            Rules) %>% 
#   kable("latex", booktabs = T,
#         caption = "Rules to determine the status of each transboundary species") %>%
#   kable_styling(full_width = F) %>%
#   column_spec(1,width = "3cm") %>% #wraps text in column2
#   column_spec(2,width = "13cm") #%>% #wraps text in column2

### Grouped version

Categories <- c("A - Growing Fishery",
                  "B - Stable Fishery",
                  "C - Fishery under 10% of Peak",
                  "No Status")

Rules <- c(
  "(Year of Catch > Year Post Max. Min. & Post Max Min Catch < (Max Catch*0.10)) & (Catch > (Max Catch*0.10) & Catch < (Max Catch*0.50)) or Year of Catch < Year of Max. Catch & Catch <= (Max Catch*0.50) or Year of Max Catch = Last Year of data)", #Rebuilding or #Developing
  "Catch > (Max Catch*0.50)", #Max Expploitation
  "Year of Catch > Year of Max. Catch & (Catch > (Max. Catch*0.10) & Catch < (Max. Catch*0.50) or Catch < Max. Catch*0.10", #Over explpoited or collapsed
  "Non of the above rules applied"
)

data.frame(Categories,
           Rules) %>% 
  kable("latex", booktabs = T,
        caption = "Rules to determine the category of each transboundary species") %>%
  kable_styling(full_width = F) %>%
  column_spec(1,width = "3cm") %>% #wraps text in column2
  column_spec(2,width = "13cm") #%>% #wraps text in column2

```

## Key uncertainties
<!-- # Related to Criteria 1 (Neighbours) -->
We have identified four key uncertainties in the method that may affect the estimation of transboundary stocks. Firstly, we ran the analysis at the species level due to limited spatial-specific data on species sub populations (stocks), therefore we are not able to identify transboundary stocks within EEZs. However, we defined transboundary only between  neighbouring countries (e.g. US and Canada) and did not include countries that were separated by another country (e.g. Canada and Mexico) and/or the high seas (e.g. Canada and Spain).
<!-- Related to Criteria 2 (Models) -->
Secondly, the predicted species distribution is affected by the uncertainties of the environmental data and models used for such predictions. Structural differences within Earth System Models results in variations of oceanic conditions, which undoubtedly affect the ENM-Nereus. We averaged the results from the three models in an effort to capture the structural variation across models. Climate natural variability is a major driver of marine species distribution, potentially removing a species from an EEZ for long periods (e.g. Anchovies and Sardines "regimes" in the Eastern Pacific are strongly influenced by water temperature decal oscillations [@Chavez:2003jr]). Thus, a species’ distribution can potentially be reduced or shifted in such way that it only covers one EEZ until oceanic conditions change again and the associated species' distribution expands. To account for such climate variability, we derived the ENM-Nereus results as an average of oceanic conditions between 1970 and 2000. This is not an issue for the SDM-SAU as it does not directly require environmental variables [@Palomares:wg].
<!-- Related to Criteria 3 (SAU catch data) -->
Thirdly, we assumed that, if the Sea Around Us reconstructed data recorded a species as caught in any given grid cell, then the species was present within that grid cell. While catch data is not exempt of uncertainty, in most cases, the differences between the Sea Around Us and the FAO self-reported data is smaller towards recent years [@Pauly:2019bf]. Thus, we limited the catch data reference period in our analyses to between 2005 and 2014, the last ten years of data available. Likewise, the spatial allocation of the catch is subjected to imprecisions, mainly produced by differences in the spatial scale of the original data and the method employed by the Sea Around Us. However, this discrepancy is mainly applicable to data provided by RFMOs in the high seas, which were not included in this study (see http://www.seaaroundus.org/data/#/spatial-catch and @Zeller2016 for mode details in the spatial allocation of the SAU).
<!-- Related to Criteria 4; Spatial distribution -->
Finally, the results of this study only considered those species that had a *Species Index* of 1 (e.g. all datasets agree on presence) and at least 25% of *Area Index* (e.g. the species shared distribution was at least 25% in each EEZ). This presents a relative conservative estimate of the number of transboundary species in the world. However, using a different *Species Index* will give results based on less datasets while a smaller *Area Index* will estimate more transboundary species (**S3_Area.png**).

# Supplements

```{r Call_Sup_1, eval = F, fig.width=6, fig.height=4, echo=FALSE, fig.pos= "H", fig.cap="Number of transboundary species per EEZ and they contribution to countries' EEZ catch. The number of species is represented in the EEZ while the catch (thousand tonnes) is represented in the land polygon."}

img <- readPNG(here("./Figures/S1_Catch_EEZ.png"))
grid.raster(img)
```

![**Number of transboundary species per EEZ and their contributions to countries’ EEZ catch. The number of species is displayed in the EEZ while catch (in thousand tonnes) is displayed in the land polygon.**](./Figures/S1_Catch_EEZ.png)

```{r Call_Sup_2, eval = F, fig.width=6, fig.height=4, echo=FALSE, fig.pos= "H", fig.cap="Number of shared species by RFMOs. Analysis included with 85 species managed by Fourteen RFMOs"}

img <- readPNG(here("./Figures/S2_RFMO.png"))
grid.raster(img)
```

![**Number of shared species across RFMOs. Analysis included with 99 species managed by Fourteen RFMOs **](./Figures/S2_RFMO.png)

The number of gridcells in which a species is present (*Area Index*) is expected to directly influence whether or not a species is considered transboundary. Therefore, we tested the sensitivity of our results to the *Area Index* by estimating the number of transboundary species along a gradient of values (*S3 Index Area*). As expected, using a more relaxed value (e.g. the species is present within 10% or more of the EEZ) will result in a greater number of transboundary species (median of 40 EEZs shared per species), while a more restrictive value (e.g., 50%) results in a median estimate of 4. The current analysis utilised a 25% *Area Index* threshold (*S3 Index Area* - darker histogram).

```{r Call_Sup_3, eval = F, fig.width=6, fig.height=4, echo=FALSE, fig.pos= "H", fig.cap="Histogram of transboundary species using different values of Area Index. Darker histogram represents used value (25%) in the analysis"}

img <- readPNG(here("./Figures/S3_Area.png"))
grid.raster(img)
```

![**Histogram of number of transboundary species using different Area Index threshold values. The darker histogram represents the threshold (25%) used in the present analysis**](./Figures/S3_Area.png)

# References (37/about 30)

<!-- # Recicled text (Not to be included) -->

<!-- From 1973 to 1982, members of the United Nations held a series of meetings to discuss regulations regarding the high seas, a region of international common property, at that time, consisting of waters from 12 miles from shore [@McRaeandMunro1989]. The establishment of the UN Convention on the Law of the Sea (UNCLOS) allowed coastal states to claim jurisdiction over the exploration and exploitation of marine resources over 200 nautical miles off their coasts [@UN:1982]. While intended to improve fisheries management by granting property rights over shared resources, this arbitrary delimitation of management areas was not informed by bio-geography, essentially ignoring species distributions and creating what we know today as shared fish stocks [@Gulland, @Munro2002]. Fish stocks, are comonly refered in fisheries management as a particular population that is more or less isolated from other stocks of the same species and hence self-sustaining [@FAO:1999tk]. The Food and Agricultural Organization (FAO) recognizes three types of shared stocks: (*i*) transboundary stocks, shared by 2 or more neighboring coastal nations; (*ii*) straddling stocks, occurring in two or more adjacent national jurisdictions and the high seas; and (*iii*) highly migratory stocks, found in the EEZs of coastal nations that are not necessarily adjoining, and the high seas. The inception of shared stocks called for the establishment of new fisheries management methods. Managers and scientists adopted a game theory approach – or means to analyze strategic interactions among decision-makers – which demonstrates that collaboration is most likely to result in the best overall outcome for nations sharing a common resource [@Bailey:2010ga,@Eide:2013dh]. Lack of collaboration in shared stocks may threaten stock sustainability [@Clark:1980js; @Nguyen:2018dk], reduce the profitability potential of the fishery [@Merino:2007jz], and lead to conflict between coastal nations [@Spijkers:2017ij], although cooperation and non-cooperation can be more complex [@Jensen:2015cf]. Moreover, under climate change, species distributions will shift [@Cheung2016], resulting in new shared stocks [@Pinsky:2018cb] and highlighting the key role cooperative fisheries management can play in maintaining stocks and profitability [@Nguyen:2018dk] as well as hopefully reducing fisheries conflict [@Spijkers:2019cz]. -->

<!-- There is a gap in assessing the nature and number of transboundary species in the world based on their spatial distribution. Caddy [@Caddy:1997ue] in 1997 estimated that there could be up to 1,500 transboundary fish species in the world. However, such estimation lacked a proper assessment due to limited information on the distribution of species at the time [@Caddy:1997ue]. More recently, Teh et al. [@Teh:2015gd] estimated that 206 species can be considered shared, accounting for a total catch of 34.2 million t and a global landed value of USD 30.7 trillion. While this represents a good starting point, the analysis was based on a literature review of shared species and did not include a mechanistic way of determining the transboundary nature of the species [@Teh:2015gd]. Thus, despite recent research highlighting the interconnection of marine species [@Ramesh:2019va], the total number of transboundary species is yet to be identified at a global scale. Yet, such an understanding is critical to effectively managing fisheries, as needs differ substantially between discreet and shared species [@Mills] and climate change will yield new challenges associated with new shared fisheries [@Pinsky:2018cb]. -->

<!-- *Getting accurate estimates of the number of shared species and where they are located can help develop fisheries management and conservation plans that account for the challenges from transboundary resources management.*  -->

<!-- The Food and Agricultural Organization (FAO) recognizes three types of shared stocks: (i) transboundary stocks, shared by 2 or more neighboring coastal nations; (ii) straddling stocks, occurring in two or more adjacent national jurisdictions and the high seas; and (iii) highly migratory stocks, found in the EEZs of coastal nations that are not necessarily adjoining, and the high seas. We focus on estimating the number of species that meet the first category. -->

<!-- We used four-criteria that combined multiple species distribution data-sources and the world's EEZs to determine if a species can be considered a “shared species”. The first criteria consisted in only considering shared species between neighbouring EEZs, rather than the extended distribution. For the second criteria we considered a species to be present in a grid cell if all data sources predicted positive occurrence. Current species distributions were predicted from occurrence records obtained from publicly available observational databases of the world’s marine species and two structurally distinct species distribution models [@Asch:2018ca; Reygondeau:2019uh; @Zeller2016] (see Methods). The third criteria determined whether the species were exploited at the grid cell using spatially explicit catch data [@Zeller2016]. A species was kept if both catch data and predicted species occurrences were positive at the same spatial cells. Finally, the last criteria determined a spatial treshold to confirm that a species occurs in both neighboring EEZs. Once we have identified the shared stocks, we estimated the contribution of the species to fishing revenues and landings using the SAU database [@Zeller2016] and categorized them according to catch trends [@Kleisner:2013gh].  -->

<!-- Only Bluefin (*Thunnus thynnus*) in the Pacific, Bigeye (*Thunnus obesus*) in the west Pacific, and Yellowfin (*Thunnus albacares*) in the Atlantic still experiencing overfishing  -->