### 2020-07-31 Scripts/Initial_Analysis.Rmd
Fixed all mf anoying country names

---
title: "Initial Analysis"
author: "Juliano Palacios-Abrantes"
date: '2019-02-04'
output: html_document
editor_options: 
  chunk_output_type: console
---

# Read Me
This script runs the analysis for the manuscript *Marine species do not need visas: the transboundary nature of the world’s exploited marine species* from the data loading to the results analysis. Is divided in two parts; *Methods* where the raw data is analyzed and *Results* where the results from the analysis are processed.

```{r setup, eval = T, echo=F, warning=F,message=F, results='hide'}

#### READ ME !!! ####
library(MyFunctions)

#### Library ####
packages <- c(
  "readxl", # Read dataframe
  "data.table", # Read dataframe (Fast!)
  "wesanderson",
  "tidyverse", # for all data wrangling and ggplot
  "janitor", # for data cleaning
  "tidytext", # to order the facet wrap https://juliasilge.com/blog/reorder-within/
  "cowplot", # for figures 1 and 3
  # "ggimage", #for reading images to the circular plot
  # "ggrepel", # for nice plot labels
  # "ggsflabel", # for nice sf_plots labels
  # "spdep", # for poly2nb old
  "sf", #Spatial analysis 
  "sp", #Spatial analysis 
  # "purrr",#Spatial analysis
  "rgdal", #Spatial analysis
  "tools", #Spatial analysis 
  "parallel", # for parallelization
  # "taxize", # For getting species names
  "rfishbase", # for species ecosystem affinity
  "zoo", #for runing mean
  "pgirmess" # for dune test after kurtis wallas
)

my_lib(packages)

#### Set paths for computer ####

Jursidiction <- "EEZ"

# Set paths depending on machine Beast (jepa88), "carmelia" or Hall1000
# if(Sys.info()[7] == "jepa88"){
#   Data_Path <- "Z:/JULIANO_NEYMAR/FishForVisa/"
# }
# if(Sys.info()[7] == "carmelia"){
#   Data_Path <- paste(here::here(),"/Temporal_Data/",sep = "")
# }
# if(Sys.info()[7] == "hall1000"){
#   Data_Path <- "/Volumes/DATA/JULIANO_NEYMAR/FishForVisa/"
# }

# Setting up paths 
# If path exists overrite it, otherwise creats a new path for that year

# Path for saving Results
Results_Path = paste(Data_Path,"Results/",sep = "")
if(file.exists(Results_Path) == TRUE){
  Results_Path = paste(Data_Path,"Results/",sep = "")
}else{
  dir.create(paste(Data_Path,"Results/",sep = ""))
  Results_Path = paste(Data_Path,"Results/",sep = "")
}

# Path for saving Figures
Figures_Path = paste(Data_Path,"Figures/",sep = "")

if(file.exists(Figures_Path) == TRUE){
  Figures_Path = paste(Data_Path,"Figures/",sep = "")
}else{
  dir.create(paste(Data_Path,"Figures/",sep = ""))
  Figures_Path = paste(Data_Path,"Figures/",sep = "")
}

ggtheme_map <- function(base_size = 9, Region = "NA") {

  theme(text             = element_text(#family = "Helvetica",
    color = "gray30", size = base_size),
    plot.title       = element_text(size = rel(1.25), hjust = 0, face = "bold"),
    panel.background = element_blank(),
    panel.border     = element_blank(),
    panel.grid.minor = element_blank(),
    panel.grid.major = element_line(color = "transparent"),
    strip.background =element_rect(fill = "transparent"),
    strip.text.x = element_text(size = 18, colour = "black",face= "bold.italic", angle = 0),
    axis.line        = element_blank(), # element_line(colour = "grey30", size = .5))
    axis.ticks       = element_blank(),
    axis.text        = element_blank(),
    axis.title       = element_blank(),
    legend.key       = element_rect(colour = NA, fill = NA, size = 4),
    legend.position = "bottom",
    legend.key.width =unit(5,"line")
  )
}

### Message for my brain

print("_____________________ WARNING DO NOT FORGET TO PULL/PUSH BEFORE YOU START_____________________")

```

# Methods

These are the methods to follow:

1. Determine which countries have Neighboring EEZs  
1.1. For this we used the Sea Around Us EEZ map that discriminate by oceans and a routine (*Estimate Transboundary Species*) to identify Neighbouring polygons,
1.2. Then we match the SAU spatial information (e.g. INDEX within EEZ and coordinates) with the Neighbouring data

2. Overlap distribution maps with Neighbors data  
2.1. For each species we merged the Neighbors data with their known distribution (SAU distribution and catch,ESMs, and Observation)  
3. Determine transboundary if:  
3.1. Species is present in both Neighbouring EEZs by INDEX

We create two "confidence level" based on data and area covered.

## Determine species list

This part is to determine the list of species to be analyzed in the project. Initially this list was limited by the species data from the DBEM, however, as we are not using the DBEM, the limit is now the SAU.

```{r Speceis_List_For_Data, eval = F, echo = T}

# Get the DBEM list of species to extract from other modesl
#GFDL
DBEM_species_G <- data_frame(TaxonKey = list.files("/Volumes/DATA/DATA/DBEM/GFDL26F1"))
#IPSL
DBEM_species_I <- data_frame(TaxonKey = list.files("/Volumes/DATA/DATA/DBEM/IPSL26F1"))
#MPI
DBEM_species_M <- data_frame(TaxonKey = list.files("/Volumes/DATA/DATA/DBEM/MPI26F1"))

# Merge the, and save to send to gabs

DBEM_Species_List <- DBEM_species_I %>% 
  bind_rows(DBEM_species_G,
            DBEM_species_M) %>% 
  group_by(TaxonKey) %>% 
  summarise() %>% 
  filter(str_detect(TaxonKey, "^6"))

# Include species name, as well...

exploited_species <- read.csv("/Volumes/DATA/PROJECTION exploited species ENM/exploited_species_list.csv")

DBEM_Species_List <- DBEM_Species_List %>% 
  left_join(exploited_species,
            by ="TaxonKey")

# write.csv(DBEM_Species_List,
#           "DBEM_Species_List.csv",
#           row.names = F)

# Send second set of species

# Get species already modelled
# Removes everything before the number
Step_One <- gsub(".*_","",list.files(paste(Data_Path,"/Distribution_Data/SAU_Distribution/",sep="")))

# Species List
Species_List <- gsub("\\..*","",Step_One)

exploited_species <- read.csv("/Volumes/DATA/PROJECTION exploited species ENM/exploited_species_list.csv") #%>% 
# left_join(Spp_Clasification) %>% 
# filter(!is.na(DemPel)) # remove freshwater spp

# Get all species in SAU database (2080)
Species_list <- exploited_species %>% 
  rename(scientific_name = TaxonName) %>% 
  pull(scientific_name)

# get their environ. association according to FishBase (species())

Species_environ_list <- rbind(species(Species_list, 
                                      fields = species_fields$habitat)
) %>% 
  mutate(TaxonName = Species_list) %>% 
  left_join(exploited_species,  by ="TaxonName") %>% 
  filter(!TaxonKey %in% Species_List) %>% 
  filter(Saltwater == -1) %>%  # remove those that are not happy with saltwater
  select(TaxonKey,TaxonName,CommonName)

# write.csv(Species_environ_list,
#           "Transboundary_Spp_Set2.csv",
#           row.names = F)

# Sent to Gabs Seprember 16th

### Check Gabs Data ####

# Does all data have the same number of files?

# SAU <- length(list.files("/Volumes/DATA/JULIANO_NEYMAR/FishForVisa/Distribution_Data/SAU_Distribution/"))
# ENM <- length(list.files("/Volumes/DATA/JULIANO_NEYMAR/FishForVisa/Distribution_Data/ENM/"))
# Occ <- length(list.files("/Volumes/DATA/JULIANO_NEYMAR/FishForVisa/Distribution_Data/Occurence/"))

####### Yes they do! ####

```

## Determine Neighbors

This is the first step of the analysis and will determine what EEZs are neighbors. The routine creates a final data-set that we then use to estimate transboundary species.

### Function `EstNeighbours`

This function determines what EEZs are Neighbors and returns a data.frame of it

```{r funEstNeighbours, eval = F, echo=T}
# Determine which countries are Neighbours

EstNeighbours <- function(Shapefile, Partial = "NA"){

  # Get sparse matrix
  int = suppressMessages(st_intersects(Shapefile))
  # Warnning message
  # although coordinates are longitude/latitude, st_intersects assumes that they are planar

  # Intitalize new object
  mat = NULL

  # Fill
  for(i in 1:NROW(int)){
    #Create matrix with ID and intersecting IDs
    tmp = cbind(i, int[[i]])

    # Join to Initalized object
    mat  = rbind(mat, tmp)
  }

  # Build final data.frame substituting ID for name (in this case nc$name)
  Result_table = data.frame(Name = Shapefile$Name[mat[,1]], 
                            Neighbour_Territory = Shapefile$Name[mat[,2]]
  )

  # Remove duplicates... eg of course an object intersects with itself...
  Result_table = Result_table[Result_table$Name != Result_table$Neighbour_Territory,]

  if(Partial == T){
    # If only partial analysis just returns Neighbour list
    return(Result_table)

  }else{

    # Merge them with SAU data so we have the INDEX and lat long info
    suppressWarnings(
      # warning for characters in name, no problem
      DBEM_Neighbours_Data <- Result_table %>% 
        left_join(EEZIDs_List, 
                  by = "Name") %>% # Asign EEZ id to each country
        left_join(EEZ_CellID, 
                  by = "EEZID") %>% # Asign EEZ id to each country
        left_join(Coor,
                  by = "INDEX") %>%  # Include lat long
        rename(Country_Territory = Name)

    )

    return(DBEM_Neighbours_Data)

  } # end ifelse
} # end function

# subset <- World_EEZs %>% filter(Name %in% c("Chile","Peru"))
# EstNeighbours(subset)
# EstNeighbours(subset,T)

```

### Sub Routine; Neighbors

```{r Neighbours_Control_Pannel, eval = F, echo = F}

###___________________Read Me_____________________#
#### Using the SAU shapefile that contains EEZ per ocean (E.G. Mexico Pacific and Mexico Atlantic)
###_______________________________________________#

# SAU shapefile

# The path
path_world <- paste(Data_Path,"Spatial_Data/SAU_Shapefile/",sep="")

# The File
fnam_world <- "SAUEEZ_July2015.shp"

# Load it!
World_EEZs <- st_read(dsn = path_world,
                      layer =file_path_sans_ext(fnam_world))

#### Grid cell information ####

# List of DBEM INDEX within the EEZ 
#Data provided by Vicki Lam

EEZIDs_List <- read_excel(paste(Data_Path,"Spatial_Data/V_Lam/Updated_EEZList_17June2016.xlsx",sep=""))
EEZ_CellID <- read_excel(paste(Data_Path,"Spatial_Data/V_Lam/EEZ_CellID.xlsx",sep=""))

colnames(EEZ_CellID) <- c("EEZID","INDEX")

# DBEM Coordinate system
Coor <- fread(paste(Data_Path,"Spatial_Data/Lon_Lat_DBEM.txt",sep=""),header = FALSE)
colnames(Coor) <- c("INDEX","Longitude","Latitude")

```

```{r Neighbours_Subroutine, eval = F, echo = T}

# Run Function (Ran and saved May 27 2019 at 6pm)
system.time(
  EEZ_Neighbour_List <- EstNeighbours(World_EEZs, Partial = T)
)
# 8203.05    0.25 8204.51 (around 2 hours 30 min)

write.csv(EEZ_Neighbour_List,
          paste(Data_Path,"Spatial_Data/EEZ_Neighbour_List.csv", sep =""),
          row.names = F)

####_________________________________________ TESTING PERIOD_________________________________________###
# I'm substetting central america to do the whole analysis (smaller EEZs = faster computation) once I have figure out things I'll run for all

### Data
# For now using only the subdata

## Spatial Test Data
# SAU_Regions <- as.data.frame(unique(World_EEZs$Name))

# Test_Countries <-c("Guatemala (Caribbean)","Guatemala (Pacific)","Belize","Honduras (Pacific)","El Salvador","Costa Rica (Pacific)","Costa Rica (Caribbean)", "Panama (Caribbean)", "Panama (Pacific)")

# Test <- World_EEZs[World_EEZs@data$Name %in% Test_Countries, ]

```

## Determine Species Status

This sub routine determines the species exploitation status based on the SAU methodology. **Note:** For the current project we do not consider this method to determine the species exploitation status *per se*, but rather a catch trend that *could* indicate that a species is under/over/max exploited.

In the original method the authors analyze stocks, while we analyze at the species level within each EEZ. Also (n), the number of species, is defined as a time series of a given species, genus or family (higher and pooled groups have been excluded) for which the first and last reported landings are at least 10 years apart, for which there are at least 5 years of consecutive catches and for which the catch in a given area (LME) is at least 1000 tonnes. We removed the last one (LME) and computed all of the species that only followed the first two;

1) The first and last reported landings are at least ten years apart;

2) There are at least five years of consecutive catches

Once those thresholds are met, the status of the species is determined as follows:

- Rebuilding (Recovering),	Year of landing > year of post-max. min. landing AND post-max. min. landing < 10% of max. landing AND landing is 10-50% of max. landing
- Developing,	Year of landing < year of max. landing AND landing is < or = 50% of max. landing OR year of max. landing = final year of landing
- Max Exploited,	Landing > 50% of max. landing
- Over exploited,	Year of landing > year of max. landing AND landing is between 10-50% of max. landing
- Collapsed,	Year of landing > year of max. landing AND landing is < 10% of max. landing

*Notes* = 3 years average mean

Finally we choose the reporting year based on the most frequent status in the last 10 years of data. In cases where there was a mach between status (e.g. Cayman Islands for species 6000006 had 5 *Max Exploited* and 5 *Overexploted*) then we report the status at the last year of data.

### Function `GetSppDist`

This function reads in species distribution from the observation data, ESM models and SAU data.

```{r GetSppDist, eval = T, echo = T}

#### The function
# NOTES
# For now we're using only the last 10 years average, basicaaly if the species has been fished in any of these years, its considered present

GetSppDist=function(Spp,Model,Coord){

  Distpath <- paste(Data_Path,"Distribution_Data/",sep="")
  INDEX = seq(1,259200,1)
  # INDEX <- Coordinates$INDEX

  # SAU Distributions
  if(Model == "SAU_D"){
    File_Name <- paste("SAU_Distribution/DIST_GOOD_SP_")
  }

  # SAU Catch
  if(Model == "SAU_C"){
    File_Name <- paste("SAU_data_per_species/CATCH_SP_")
  }

  # Occurence
  if(Model == "Occ"){
    File_Name <- paste("Occurence/OCCURENCE_JULIANO_")
  }

  # ENM Model
  if(Model == "ENM"){
    File_Name <- paste("ENM/ENM_JULIANO_")
  }

  # DBEM Data
  # if(Model == "DBEM"){
  #   Distpath <- "/Volumes/DATA/JULIANO_NEYMAR/PristineSeasData
  #   Final_Path <- "ENM"
  #   File_Name <- paste("ENM_JULIANO",Spp,".mat",sep="")
  # }

  if(Model == "All"){

    Models_List <- c(paste(Distpath,"SAU_Distribution/DIST_GOOD_SP_",Spp,".mat",sep =""),
                     paste(Distpath,"Occurence/OCCURENCE_JULIANO_",Spp,".mat",sep=""),
                     paste(Distpath,"ENM/ENM_JULIANO_",Spp,".mat",sep="")
    )

    # Jumps species not modeled. NOTE: We only need one as ENM, Occ and SAU Dis have all the same 939 spp
    if(file.exists(Models_List[1])){

      Load <- lapply(Models_List, FUN=R.matlab::readMat, na.strings=0)

      sppdist <- as.data.frame(bind_cols(Load)) %>% 
        mutate(
          INDEX = INDEX,
          TaxonKey = Spp
        )
      colnames(sppdist) <- c("SAU_D","Occ","ENM","INDEX","TaxonKey")

      #### Step for SAU catch data that has to be averaged
      File_Name <- paste("SAU_data_per_species/CATCH_SP_")
      SppPath <- paste(Distpath,File_Name,Spp,".mat",sep="")

      # For now we're using only the last 10 years average, basicaaly if the species has been fished in any of these years, its considered present
      SAU_C_data <- as.data.frame(R.matlab::readMat(SppPath)) %>% 
        select(CATCH.1:CATCH.65) %>% 
        mutate(INDEX = INDEX) %>% 
        gather("Year","Catch",CATCH.56:CATCH.65) %>% # Last 10 years of data
        group_by(INDEX) %>% 
        summarise(SAU_C = mean(Catch,na.rm=T))

      # Join both tables
      sppdist <- sppdist %>% 
        left_join(SAU_C_data,
                  by = "INDEX") %>% 
        select(TaxonKey,INDEX,everything()) %>% 
        left_join(CoorG,
                  by = "INDEX")

      # Fix coordinate system incompatibility between Gab and DBEM
      sppdist <- suppressWarnings(sppdist[order(sppdist$Latitude, rev(sppdist$Longitude),decreasing=TRUE), ] %>% 
                                    mutate(INDEX = seq(1,259200,1)) %>% 
                                    gather("Model","Value",3:6) %>%
                                    mutate(Value = ifelse(is.na(Value), 0, Value)) # Converting NA's to ceros
      )

      getSppDist = sppdist

    }

  }else{

    # Merge paths
    SppPath <- paste(Distpath,File_Name,Spp,".mat",sep="")

    if(file.exists(SppPath) == TRUE){

      #Install (if needed) R.matlab package
      if(!require(R.matlab)){
        install.packages("R.matlab")
      }

      # Read Files

      sppdist <- as.data.frame(R.matlab::readMat(SppPath)) %>% 
        mutate(INDEX = INDEX,
               Species = Spp) %>% 
        left_join(Coord)

      # Fix coordinate system incompatibility between Gab and DBEM
      sppdist <- sppdist[order(sppdist$Latitude, rev(sppdist$Longitude),decreasing=TRUE), ] %>% 
        mutate(INDEX = seq(1,259200,1))

      # Return 
      getSppDist=sppdist

    }else{
      print(paste("No info for this species",Spp, "in",Model))
    }

  }
}

#______ Test ______#

# Variables outside function
# head(GetSppDist(600004,"SAU_D")) # Works
# head(GetSppDist(600004,"SAU_C")) # Works
# head(GetSppDist(600004,"Occurence")) # Works
# head(GetSppDist(600004,"ENM")) # Works
# head(GetSppDist(600004,"All")) # Works

```

### Function `EstSppStatus`

This function estimates the exploitation status of a species based on catch data.

```{r funEstSppStatus, eval = T, echo =F}

# Spp <- 601417
# Coord = CoorG

EstSppStatus <- function(Spp,Coord, Save_Path = "NA", Per_Taxa = NA){

  # Load species' data
  Data <- GetSppDist(Spp, Model = "SAU_C", Coord = Coord) %>% 
    set_names(c(seq(1950,2014,1),"INDEX","Species","Longitude","Latitude")) %>% 
    gather("Year","Value",`1950`:`2014`) %>%
    arrange(INDEX) %>% 
    left_join(Index_Code, # Match species catch with INDEX
              by = "INDEX") %>% 
    filter(!is.na(EEZID),
           Value > 0) %>% 
    group_by(TaxonKey=Species,Year,Territory) %>% # Estimates total catch per country
    summarise(Total_Catch = sum(Value,na.rm=T)) %>% 
    group_by(TaxonKey,Territory,Year) %>% # Running mean as Kleisner & Pauly 2012
    mutate(RMean = rollmean(x = Total_Catch, 3,
                            align = "left",
                            fill = Total_Catch)
    ) %>% 
    select(-Total_Catch) %>% 
    ungroup() %>% 
    mutate(Year = as.numeric(Year))

  # Tresholds to evaluate stocks if these are not met, then soecies status = "No Status"
  Tresholds_Data <- Data %>% 
    arrange(Territory) %>% 
    group_by(TaxonKey,Territory, grp = cumsum(c(1, diff(Year) != 65))) %>% 
    group_by(TaxonKey,Territory) %>% 
    summarise(Years_appart = ifelse(max(Year)-min(Year) >= 0,"Yes","No"), # Step 1. first and last reported landings are at least 10 years apart
              Consecutive = ifelse(max(grp) >= 5,"Yes","No")#, # Step 2. There are at least 5 consecutive catch years
              # Catch_over = ifelse(max(RMean >= 1000),"Yes","No") # Step 3. 3) The catch in a particular area (LME) is at least 1,000 tonnes. (CURRENTLY OFF)
    ) %>% 
    gather("Step","Treshold",Years_appart:Consecutive) %>% 
    group_by(TaxonKey,Territory) %>% 
    summarise(Met_Criteria = paste(unique(Treshold),collapse =";")
    )

  # Remove species that do not met the criterias

  No_Status_Data <- Tresholds_Data %>% 
    filter(Met_Criteria != "Yes") %>% 
    select(TaxonKey,Territory) %>% 
    mutate(Status = "No Trend")

  # Evaluate species that met the criteria

  # Get Variables to perform status metric
  Status_Data <- Data %>% 
    # filter(Territory == "Israel (Red Sea)") %>%  # For testing
    anti_join(No_Status_Data, # Removes no status data
              by = c("TaxonKey","Territory")
    ) %>% 
    # Get Max catch and max catch year
    group_by(TaxonKey,Territory) %>% 
    top_n(1,RMean) %>% 
    group_by(TaxonKey,Territory,RMean) %>%  # Solves for years with the same Max catch data (e.g. Cayman Isl all years)
    summarise(Year = min(Year)) %>%
    rename(Max_Peak = RMean,
           Y_Max_Peak = Year) %>% 
    # Get Min catch after peak catch
    left_join(Data,
              by = c("TaxonKey","Territory")
    ) %>% 
    filter(Year >= Y_Max_Peak) %>% # Filter years after peak year
    group_by(TaxonKey,Territory) %>% 
    top_n(-1,RMean) %>% # find min value
    group_by(TaxonKey,Y_Max_Peak,Territory,Max_Peak,RMean) %>%  # Solves for years with the same min after peak catch data (e.g. Cayman Isl 2013-2014)
    summarise(Year = min(Year)) %>% 
    rename(Post_Max_Min = RMean,
           Y_Post_Max_Min = Year) %>% 
    left_join(Data,
              by = c("TaxonKey","Territory")
    ) %>% 
    # Final Species Status Evaluation 
    mutate(Status = ifelse(Year > Y_Post_Max_Min & Post_Max_Min < (Max_Peak*0.10) & ((RMean > (Max_Peak*0.10) & RMean < (Max_Peak*0.50))),"Rebuilding",
                           ifelse((Year < Y_Max_Peak & RMean <= (Max_Peak*0.50) | Y_Max_Peak == max(Year)),"Developing",
                                  ifelse(RMean > (Max_Peak*0.50),"Max Exploited",
                                         ifelse(Year > Y_Max_Peak & ((RMean > (Max_Peak*0.10) & RMean < (Max_Peak*0.50))),"Over Exploited",
                                                ifelse(Year > Y_Max_Peak & RMean < (Max_Peak*0.10),"Collapsed",
                                                       "No Trend")
                                         )
                                  )
                           )
    )
    )

  #_______________#
  # Function Return 
  #_______________#

  ### Save the data for each species in all years if Per_taxa = T

  if(Per_Taxa == T){

    # Select onlyZ 
    Status_Data <- Status_Data %>%
      ungroup() %>%
      select(TaxonKey,Territory,Year, Status)

    if(is.na(Save_Path) == TRUE){

      Save_Path <- Results_Path # standar result's path in my Drobo
    }

    # File name for each species
    Name <- paste(Save_Path,Spp,"_Status.csv",sep="")

    # Save file for each species
    write_csv(Status_Data,
              Name)
  }else{

    ### Determine the final status using the top category over the last 10 years

    Determine_Status_Data <- Status_Data %>% 
      filter(Year > 2004) %>% # last 10 years of data (2005-2014)
      group_by(TaxonKey,Territory,Status) %>%
      summarise(n_Status= n()) %>%
      group_by(TaxonKey,Territory) %>%
      top_n(1,n_Status)

    # Fix cases where you have equal categories.
    suppressMessages( # supresses "No duplicate combinations found of: Territ"
      Duplicates <- Determine_Status_Data %>% 
        ungroup() %>% 
        janitor::get_dupes(Territory) %>% 
        select(TaxonKey,Territory) %>% 
        pull(Territory) %>% 
        unique()
    )

    # In case we have a tide, we use the last year's category
    Duplicates_Status <- Status_Data %>% 
      filter(Territory %in% Duplicates,
             Year == 2014) %>%
      ungroup() %>%
      select(TaxonKey,Territory,Status)

    # Build Final Status database
    Final_Status_Data <- Determine_Status_Data %>% 
      filter(!Territory %in% Duplicates) %>% 
      bind_rows(Duplicates_Status)

    # If both datasets have value RETURN BOTH
    if(nrow(No_Status_Data) > 0 & nrow(Final_Status_Data) > 0){

      Fun_Result <- bind_rows(No_Status_Data,
                              Final_Status_Data)
    }

    # If only No status data, then just return No Status Data
    if(nrow(No_Status_Data) > 0 & nrow(Final_Status_Data) == 0){
      return(No_Status_Data)
    }else{
      return(Final_Status_Data)
    }
  } # close if per taxa
}

# Testing
# Test_1 <- EstSppStatus(Species_List[1],CoorG)

```

### Data Species Status

### Sub routine for Species Status

```{r Spp_Status_Control_Pannel, eval = T, echo = F}

# SAU relations between INDEX and Country's EEZs
EEZIDs_List <- read_excel(paste(Data_Path,"Spatial_Data/V_Lam/Updated_EEZList_17June2016.xlsx",sep=""))
EEZ_CellID <- read_excel(paste(Data_Path,"Spatial_Data/V_Lam/EEZ_CellID.xlsx",sep=""))
colnames(EEZ_CellID) <- c("EEZID","INDEX")

Index_Code <- EEZIDs_List %>% 
  left_join(EEZ_CellID) %>% 
  rename(Territory = Name)

# Gabriel's coordinate system
CoorG <- read.csv(paste(Data_Path,"Spatial_Data/coordinates_gab.csv",sep="")) %>%
  mutate(INDEX = seq(1,259200,1))

# Removes everything before the number
Step_One <- gsub(".*_","",list.files(paste(Data_Path,"/Distribution_Data/SAU_Distribution/",sep="")))

# Species List
Species_List <- gsub("\\..*","",Step_One)

```

```{r Spp_Status_Sub_Routine, eval = T, echo = F}
# Needs to have trans species 
Species_Status <- bind_rows(
  mclapply(
    Species_List[1:2],
    FUN = EstSppStatus,
    Coord = CoorG,  # Gabriels coordinate system for read species dist
    Per_Taxa = T,
    Save_Path = "Z:/Vicky/Stock_Status/"
  )
)

# Save data
readr::write_csv(Species_Status,
                 paste(Results_Path,"Species_Status.csv",sep="")
)
# head(Species_Status)
```

## Get Catch Data

### Function `EstCatch`

This function just creates a database of total catch per species within each EEZ

```{r EstCatch, eval = F, echo = T}

EstCatch <- function(Spp, Coord, Index_Code){

  SppDist <- GetSppDist(Spp,Model="All",Coord) %>% 
    filter(
      Model == "SAU_C",
      Value > 0) %>% 
    left_join(Index_Code,
              by = "INDEX") %>% 
    group_by(eez_name = Territory,
             taxon_key = TaxonKey) %>% 
    summarise(total_catch = sum(Value))

  return(SppDist)

}

# EstCatch(Spp,Coord,Index_Code)

```

### Sub routine for catch data

```{r control_panel, eval = F, echo = T}

# Removes everything before the number
Step_One <- gsub(".*_","",list.files(paste(Data_Path,"/Distribution_Data/SAU_data_per_species/",sep="")))

# Species List
Species_List <- gsub("\\..*","",Step_One)

CoorG <- read.csv(paste(Data_Path,"Spatial_Data/coordinates_gab.csv",sep="")) %>%
  mutate(INDEX = seq(1,259200,1))

# SAU relations between INDEX and Country's EEZs
EEZIDs_List <- read_excel(paste(Data_Path,"Spatial_Data/V_Lam/Updated_EEZList_17June2016.xlsx",sep=""))
EEZ_CellID <- read_excel(paste(Data_Path,"Spatial_Data/V_Lam/EEZ_CellID.xlsx",sep=""))
colnames(EEZ_CellID) <- c("EEZID","INDEX")

Index_Code <- EEZIDs_List %>% 
  left_join(EEZ_CellID) %>% 
  rename(Territory = Name)

```

```{r subroutine_get_catch_data, eval = F, echo = T}
# Notes, needs GetSppDist
# Note, It averages the LAST 10 years oe ach gridcell and then adds them together
#  I think NA is the ABNJ

system.time(
  SAU_Catch_Data <- bind_rows(
    mclapply(
      Species_List,
      FUN = EstCatch,
      Coord = CoorG,  # Gabriels coordinate system for read species dist
      Index_Code = Index_Code
    )
  )
)

# Save dataframe for future
Save_Path <- paste(Data_Path,"Distribution_Data/sau_catch_country_taxon.csv",sep="")
write_csv(SAU_Catch_Data,
          Save_Path)

```

## Estimate Transboundary Species

This is the main step of the project. In this segment we estimate the number of transboundary species existing today in each of the worlds EEZs.

The data-sets SAU_D, OCC, and ESM are used as species presence while SAU_C is used as a confirmation of that. The `Model Index` is the first threshold and estimates the uncertainty of models. Goes from 0 - 1 as follows:

- 1 or 0; All models agree that in the presence of the species
- 0.75; Three out of four models agree on species presence
- 0.50; Two out of four models agree on species presence
- 0.25; Only one model flags a species presence
- 0; All models agree there is no species there

The `Area_Index` is the second threshold and estimates the proportion of each species distribution in each EEZ. It only take into consideration the distribution of the species within the two EEZs, rather than the complete distribution. Hence, is mirror (e.g. if EEZ#1 is 0.40, EEZ#2 is 0.6)

** NOTE THAT THE FUNCTION IS SETTED TO OUTPUT RESULTS FOR ALL MODELS **

### Function `EstTransIndex`

This is the main function of the project, the one that actually estimates if a species is transboundary or not

```{r funEstTransIndex, eval= T ,echo = T,warning = F,message = F}

# Function to determine whether or not a species is transboundary

# Varibales needed
# Spp: Species to be analized, data with presence/absence per gridcell
# Index_EEZ: Reference list of all INDEX that fall within EEZs
# EEZ_Size: Number of grid-cells that each EEZ has
# Neighbours: Reference list of neighbouring countries

# Needs getSppDist

#### For testing 
# Spp <- 600004  # Anchovie Equator, Chile and Peru
# Spp <- 600107 # Tuna Many RFMOs
# Spp <- 600004 # Argument 1 must have names (RFMO)
# Spp <- 600464 # Discrete species 
# Spp <- 600051 # One of Pierrs massive (wrong) discrete spp
# x <- c(600464, # Discreat spp in one nation
# 600059, # nowhere to be found spp
# 601352, # Discrete spp in two nations
# 600004) # transboundary spp
# Model <- "All"
# Coord <- CoorG
# 
# # For EEZ
# Neighbours <- Neighbours_Data
# Index_Code <- Index_Code

# __________________________________ #

# The function
EstTransIndex <- function(Spp,Model = "All",Neighbours,Coord,Index_Code,Save="Y",Jurisdiction_Type){

  # Get model data from spps
  SppDist <- GetSppDist(Spp,Model,Coord)

  # Result 1. Number of Countries that share the species

  #____________ ESTIMATING MODEL INDEX (TRESHOLD 1)_________ #
  Trans_Spp <- SppDist %>%
    filter(INDEX %in% Neighbours$INDEX,# Filter data only located within EEZs
           Model != "SAU_C") %>% # Only using observational and modelled data
    mutate(Value = ifelse(Value > 0, 1,0)) %>%  
    filter(Value > 0) %>%
    group_by(TaxonKey,
             INDEX
    ) %>%
    summarise(Model_Index = sum(Value,na.rm=T)/3
    ) %>%
    filter(Model_Index > 0.4) %>% # at least 2 sources agree
    # ____________ ESTIMATING FUNDAMENTAL NICHE (TRESHOLD 2)_________ #
    left_join(SppDist,
              by = c("TaxonKey","INDEX")) %>%
    filter(Model == "SAU_C", # Only keeping cells where SAU catch exists
           Value > 0) %>%
    mutate(Model_Index = Model_Index*100) %>%
    select(-Model, -Value, -Latitude,-Longitude) %>%
    left_join(Index_Code,
              by = "INDEX")

  # MODEL INDEX (TRESHOLDS 1 & 2) #
  Model_Index_D <- Trans_Spp %>% 
    group_by(Territory,
             TaxonKey,
             Model_Index
    ) %>% 
    summarise(n_cells_spp = n()) %>% 
    select(-n_cells_spp)

  #____________ ESTIMATING DISTRIBUTION INDEX (TRESHOLD 3)_________ #
  # The number of species' cells present within each country's EEZ

  # if(n_Territory > 1){

  #Step 1.  Get EEZ id and Neighbour
  Neighbours_List <- Neighbours %>% 
    group_by(Territory,Neighbour) %>% 
    summarise(n=n()) %>% 
    ungroup() %>% 
    select(-n)

  # Step 2. Determines the amount of grids present in each country
  Spp_Grid <- Trans_Spp %>% 
    group_by(TaxonKey,
             Model_Index, # Un-comment after producing models x datasets
             Territory) %>% 
    summarise(n_spp_eez = length(unique(INDEX))) %>% 
    left_join(Neighbours_List,
              by = "Territory") %>% 
    filter(Territory %in% Trans_Spp$Territory, #Filter out unwanted Neighbours (those who don't have grids within but get included because they are Neighbours)
           Neighbour %in% Trans_Spp$Territory)

  # IF SPECIES DISCRETE AND APEARS IN ONLY 1 REGION OR > 1 REGION BUT NOT NEIGHBOURS #

  # Step 2.1. Determines discrete spp
  if(nrow(Spp_Grid) == 0 & Save == "Y"){

    #  Get the Territory where species happen
    Discrete_Territory <- unique(Model_Index_D$Territory)
    n_Territory <- length(Discrete_Territory)

    # Reads dataset of Non transboundary species
    Non_Transboundary_Data <- fread(paste(Results_Path,"Trans_Results/Non_Transboundary_Data.csv",sep=""))

    # Set new data
    Current_Spp <- tibble(TaxonKey = as.integer(Spp),
                          Territory = Discrete_Territory)

    # Combine new spp with others
    New_Non <- Non_Transboundary_Data %>% 
      bind_rows(Current_Spp)

    # Set a name for file and saves data
    New_Non_Name <- paste(Results_Path,"Trans_Results/Non_Transboundary_Data.csv",sep="")
    write.csv(New_Non, New_Non_Name, row.names = F)

    EstTransIndex=Current_Spp

  }

  # Step 3. Sum total grids per Neighbours
  if(nrow(Spp_Grid) > 0 & Save == "Y"){

    # Split dataframes to merge latter
    Territory_T <- Spp_Grid %>% 
      ungroup() %>% 
      select(
        Model_Index, # Un-comment after producing models x datasets
        TaxonKey,
        Name=Territory,
        n_spp_eez
      )

    Neighbour_T <- Spp_Grid %>% 
      ungroup() %>% 
      select(
        Model_Index,
        TaxonKey,n_spp_eez,
        Name=Neighbour,
        Territory
      )

    # Merge dataframes to get totals per Neighbourds
    Area_Index_D <- full_join(Territory_T,
                              Neighbour_T, 
                              by = c("Model_Index","Name","TaxonKey")
    ) %>%
      rowwise() %>%
      mutate(Spp_Total = sum(n_spp_eez.x,n_spp_eez.y,na.rm=T)) %>% # Total gridcelles per Neighbours
      distinct() %>% # Removes false duplicates from `full_join()`
      rename(Territory = Name,
             Neighbour =Territory,
             n_spp_Country = n_spp_eez.x,
             n_spp_Neighbour = n_spp_eez.y) %>% 
      mutate(Area_Index = n_spp_Country/Spp_Total) %>%  #Estimates the proportion of grids per country
      select(1:3,6,8)

    #____________ FINAL INDEX TABLE_________ #
    Indexes <- Model_Index_D %>% 
      full_join(Area_Index_D,
                by = c("TaxonKey","Territory","Model_Index")
      ) %>% 
      filter(!is.na(Area_Index),
             !is.na(Model_Index) # Uncomment after producing Models x datasets
      ) %>%  # Remove cases where the country does not pass the area_index and viceversa
      select(TaxonKey,Territory,Neighbour,everything())

    ### Save spp dataframe

    if(nrow(Indexes) > 0 & Save == "Y"){

      File_Name <- paste(Spp,"_Transboundary.csv",sep = "")
      Save_Path <- paste(Results_Path,"Trans_Results/",File_Name,sep="")

      write_csv(Indexes,
                Save_Path)

    }

    if(nrow(Indexes) > 0 & Save == "N"){
      EstTransIndex=Indexes # A dataframe with   
    }

    ### Function return
    EstTransIndex=Indexes # A dataframe with

  } # Closes Treshold (Step 3)
} # closes function

# __________________________________
# Test function
# __________________________________

# Spp <-  # Tuna Many RFMOs
# Model <- "All"
# Coord = CoorG
# #
# 
# # For EEZ
# Neighbours <- Neighbours_EEZ
# Index_Code <- EEZ_Index_Code
# 
# 
# EstTransIndex(Spp = 600107,
#               Mode = "All",
#               Neighbours = ,Coord,Index_Code,Save="Y")

```

### Function `Mclapply_Hack`

This function will hack the `mclappy()` function to work in PC and use multiple cores for the analysis. It was created by someone else and the original version is here: 
https://www.r-bloggers.com/implementing-mclapply-on-windows-a-primer-on-embarrassingly-parallel-computation-on-multicore-systems-with-r/

```{r Mclapply_Hack_Fun, eval= T , echo = T, warning = F, message = F}

# The hack

Mclapply_Hack <- function(...){
  ## Create a cluster
  size.of.list <- length(list(...)[[1]])

  cl <- makeCluster(min(size.of.list, n_cores))

  ## Find out the names of the loaded packages 
  loaded.package.names <- c(
    ## Base packages
    sessionInfo()$basePkgs,
    ## Additional packages
    names(sessionInfo()$otherPkgs))
  tryCatch( {

    ## Copy over all of the objects within scope to
    ## all clusters. 
    this.env <- environment()
    while( identical( this.env, globalenv() ) == FALSE ) {
      clusterExport(cl,
                    ls(all.names=TRUE, env=this.env),
                    envir=this.env)
      this.env <- parent.env(environment())
    }
    clusterExport(cl,
                  ls(all.names=TRUE, env=globalenv()),
                  envir=globalenv())

    ## Load the libraries on all the clusters
    ## N.B. length(cl) returns the number of clusters
    parLapply( cl, 1:length(cl), function(xx){
      lapply(loaded.package.names, function(yy) {
        require(yy , character.only=TRUE)})
    })

    ## Run the lapply in parallel 
    return( parLapply( cl, ...) )
  }, finally = {        
    ## Stop the cluster
    stopCluster(cl)
  })

  ## Warn the user if they are using Windows
  if( Sys.info()[['sysname']] == 'Windows' ){
    message(paste(
      "\n", 
      "   *** Microsoft Windows detected ***\n",
      "   \n",
      "   For technical reasons, the MS Windows version of mclapply()\n",
      "   is implemented as a serial function instead of a parallel\n",
      "   function.",
      "   \n\n",
      "   As a quick hack, we replace this serial version of mclapply()\n",
      "   with a wrapper to parLapply() for this R session. Please see\n\n",
      "     http://www.stat.cmu.edu/~nmv/2014/07/14/implementing-mclapply-on-windows \n\n",
      "   for details.\n\n"))
  }

  ## If the OS is Windows, set mclapply to the
  ## the hackish version. Otherwise, leave the
  ## definition alone. 
  mclapply <- switch( Sys.info()[['sysname']],
                      Windows = {Mclapply_Hack}, 
                      Linux   = {mclapply},
                      Darwin  = {mclapply})

}
## end mclapply.hack.R

```

### **Principal Routine**

```{r Trans_Spp_Control_Pannel, eval = F, echo = F}

# Set Jurisdiction to EEZ or RFMO
Jurisdiction <- "EEZ"

# Determine forlder for saving results
Results_Path <- paste(Data_Path,"Results/",Jurisdiction,"/",sep = "")

# Removes everything before the number
Step_One <- gsub(".*_","",list.files(paste(Data_Path,"/Distribution_Data/SAU_Distribution/",sep="")))

# Species List
Species_List <- gsub("\\..*","",Step_One)

# Set Spceies list
Exploited_Species <- fread(paste(Data_Path,"Distribution_Data/Species/exploited_species_list.csv",sep="")) %>%
  filter(TaxonKey %in% Species_List)

# Step to run function for species not computed
Computed_Spp <- as.integer(gsub("\\_.*","",list.files(Results_Path)))

Exploited_Species <- Exploited_Species %>%
  filter(!TaxonKey %in% Computed_Spp)

# Gabriel's coordinate system
CoorG <- read.csv(paste(Data_Path,"Spatial_Data/coordinates_gab.csv",sep="")) %>%
  mutate(INDEX = seq(1,259200,1))

# For Function
Exploited_Species_List <- Exploited_Species$TaxonKey

# Load Jursidiction-dependant data

# Load neighbours data from previouse routine
Neighbours_Data <- read_excel(paste(Data_Path,"Spatial_Data/EEZ_Neighbour_List.xlsx", sep =""))

# Modifications have to be made if ran un Hall
# R<e9>union (Réunion) and C<f4>te d'Ivoire (Côte d'Ivoire)

# SAU relations between INDEX and Country's EEZs
EEZIDs_List <- read_excel(paste(Data_Path,"Spatial_Data/V_Lam/Updated_EEZList_17June2016.xlsx",sep=""))
EEZ_CellID <- read_excel(paste(Data_Path,"Spatial_Data/V_Lam/EEZ_CellID.xlsx",sep=""))
colnames(EEZ_CellID) <- c("EEZID","INDEX")

Index_Code <- EEZIDs_List %>% 
  left_join(EEZ_CellID) %>% 
  rename(Territory = Name)

# Creates dummy dataset for Non transboundary species
write.csv(tibble(), paste(Results_Path,"Non_Transboundary_Data.csv",sep=""), row.names = F)

# Number of cores to us 
n_cores <- detectCores()

# Just a warning for me not to forget
print(paste("You are running the model for",nrow(Exploited_Species),"Species at the",Jurisdiction,"level and will be saved at",Results_Path))
```

```{r Trans_Spp_Routine, eval = F, echo = F}

# Run model in parallel
system.time(
  Dis_Trans <- bind_rows(
    mclapply(
      Exploited_Species_List,
      FUN = EstTransIndex,
      Model = "All", # all data sources
      Neighbours = Neighbours_Data,
      Coord = CoorG,  # Gabriels coordinate system for read species dist
      Index_Code = Index_Code, 
      Save = "Y",  # For saving
      Jurisdiction_Type = Jurisdiction # For saving in proper path
    )
  )
)

head(Dis_Trans)
# Data for Hall with 8 cores on RFMOs
# user    system   elapsed 
# 11375.652  3726.515  8342.296

# Data on time for the Beast with 30 cores
# user   system  elapsed 
# 8117.72  1991.91 10296.00

# Data on time for the Beast with 32 cores
#  user  system elapsed 
# 7831.10 1327.27 9261.48

```

# Results

## Functions needed

### Function `GetTransResults`

This function simply reads in the files and import them all as on large d.f

```{r Fun_GetTransResults, eval = F}

GetTransResults=function(Spp){

  # Set the path for each file
  Distpath <- paste(Results_Path,"Trans_Results/",Spp,sep="")

  # Loads all files in a df
  Load_Data <- bind_rows(lapply(Distpath, FUN=fread))

  return(Load_Data)

}

```

### Function `CleanResults`

There are quite a few inconsistencies between the names of territories within the SAU shape file, the distribution data and the UN database. The following function cleans up the names, some of them manually some of the automatic. It also removes the "transboundary nature" of those species shared by two territories that belong to the same nation (e.g. Brazil mainland and Brazil (Fernando de Noronha) or US Gulf of Mexico and US East coast). Finally, there is an annoying thing coding issue between MAC and Windows related to characters so I have to also fix for that, manually.

```{r funCleanResults, eval = F, echo = F}

# Requires Clean_sau_eez_fes
# Data <- Results_Trans
CleanResults <- function(Data){

  # Clean out countries that share with their own borders (e.g. Brazil mainland and Brazil (Fernando de Noronha))
  Auto_Remove_Territories <- Data %>% 
    filter(!is.na(Territory),
           !is.na(Neighbour)) %>% 
    rename(eez_name=Territory) %>% 
    left_join(Clean_sau_eez_fes,
              by = "eez_name") %>% 
    select(TaxonKey,
           Territory= eez_name,
           eez_name=Neighbour, 
           Model_Index,
           Area_Index,
           CT_fishing_entity= fishing_entity) %>% 
    left_join(Clean_sau_eez_fes,
              by = "eez_name") %>% 
    select(1,2, 
           Neighbour = eez_name,
           Model_Index:CT_fishing_entity,
           NT_fishing_entity= fishing_entity,
           everything()) %>% 
    mutate(
      CT_fishing_entity = ifelse(is.na(CT_fishing_entity),paste(Territory),paste(CT_fishing_entity)),
      NT_fishing_entity = ifelse(is.na(NT_fishing_entity),paste(Neighbour),paste(NT_fishing_entity)),
      Same_Nation = ifelse(CT_fishing_entity==NT_fishing_entity,"Yes","No")
    )

  #### Manual removal of those left...

  # List of territories to eliminate (No other neighobors but mainland)
  Islands_to_Remove <- c(
    "Brazil (Fernando de Noronha)",
    "Brazil (St Paul and St. Peter Archipelago)",
    "Hawaii Main Islands (USA)",
    "Hawaii Northwest Islands (USA)",
    "Juan Fernandez Islands (Chile)",
    "Kermadec Isl. (New Zealand)"
  )

  # Complete list of territories to remove from dataset (Including islands)
  Territories_to_Remove <- Data %>% 
    filter(
      Territory %in% Islands_to_Remove |
        Neighbour %in% Islands_to_Remove |
        Territory == "Balearic Island (Spain)" & Neighbour == "Spain (Mediterranean and Gulf of Cadiz)" |
        Territory == "British Virgin Isl. (UK)" & Neighbour == "Anguilla" |
        Neighbour == "British Virgin Isl. (UK)" & Territory == "Anguilla" |
        Territory == "Corsica (France)" & Neighbour == "France (Mediterranean)" |
        Neighbour == "Corsica (France)" & Territory == "France (Mediterranean)" |
        Territory == "Crete (Greece)" & Neighbour == "Grece" |
        Neighbour == "Crete (Greece)" & Territory == "Grece" |
        Territory == "Cyprus (North)" & Neighbour == "Cyprus (South)" |
        Neighbour == "Cyprus (North)" & Territory == "Cyprus (South)" |
        Territory == "Channel Isl. (UK)" & Neighbour == "United Kingdom" |
        Neighbour == "Channel Isl. (UK)" & Territory == "United Kingdom" |
        Territory == "Glorieuse Islands (France)" & Neighbour == "Mayotte (France)" |
        Neighbour == "Glorieuse Islands (France)" & Territory == "Mayotte (France)" |
        Territory == "Guam (USA)" & Neighbour == "Northern Marianas (USA)" |
        Neighbour == "Guam (USA)" & Territory == "Northern Marianas (USA)" |
        Territory == "Italy" & Neighbour == "Sardinia (Italy)" |
        Neighbour == "Italy" & Territory == "Sardinia (Italy)" |
        Territory == "Italy" & Neighbour == "Sicily (Italy)" |
        Neighbour == "Italy" & Territory == "Sicily (Italy)" |
        Territory == "Japan (Daito Islands)" & Neighbour == "Japan (main islands)"|
        Neighbour == "Japan (Daito Islands)" & Territory == "Japan (main islands)"|
        Territory == "Japan (Daito Islands)" & Neighbour == "Japan (Ogasawara Islands)"|
        Neighbour == "Japan (Daito Islands)" & Territory == "Japan (Ogasawara Islands)"|
        Territory == "Norway" & Neighbour == "Svalbard Isl. (Norway)" |
        Neighbour == "Norway" & Territory == "Svalbard Isl. (Norway)" |
        Territory == "Sicily (Italy)" & Neighbour == "Sardinia (Italy)" |
        Neighbour == "Sicily (Italy)" & Territory == "Sardinia (Italy)" |
        Territory == "South Africa (Atlantic Coast)" & Neighbour == "South Africa (Indian Ocean Coast)" |
        Neighbour == "South Africa (Atlantic Coast)" & Territory == "South Africa (Indian Ocean Coast)" |
        Territory == "USA (Alaska, Arctic)" & Neighbour == "USA (Alaska, Subarctic)" |
        Neighbour == "USA (Alaska, Arctic)" & Territory == "USA (Alaska, Subarctic)"
    ) %>%
    group_by(Territory,Neighbour) %>% 
    summarise(n())

  # Cleaned results 
  Clean_Results <- Auto_Remove_Territories %>% 
    anti_join(Territories_to_Remove,
              by = c("Territory","Neighbour")
    ) %>% 
    filter(Same_Nation == "No") %>% 
    select(-8,-9)

  # Fix coding issues between MAC and Windows

  Clean_Results$Territory <- Clean_Results$Territory %>% 
    gsub("Cura<e7>ao (Netherlands)","Curaçao (Netherlands)",.) %>% 
    gsub("C<f4>te d'Ivoire","Côte d'Ivoire",.) %>% 
    gsub("R<e9>union (France)","Réunion",.)

  Clean_Results$Neighbour <- Clean_Results$Neighbour %>% 
    gsub("Cura<e7>ao (Netherlands)","Curaçao (Netherlands)",.) %>% 
    gsub("C<f4>te d'Ivoire","Côte d'Ivoire",.) %>% 
    gsub("R<e9>union (France)","Réunion",.)

  Clean_Results$CT_fishing_entity <- Clean_Results$CT_fishing_entity %>% 
    gsub("Cura<e7>ao (Netherlands)","Curaçao (Netherlands)",.) %>% 
    gsub("C<f4>te d'Ivoire","Côte d'Ivoire",.) %>% 
    gsub("R<e9>union (France)","Réunion",.)

  return(Clean_Results)

}

```

### Function `Figure CreateFigureOne`

This function will create figure one for both catch or value

```{r fun_Figure_1, eval =F, echo =F}

CreateFigureOne <- function(Variable_Selected){

  if(Variable_Selected == "catch"){
    Map_Label <- "Catch (Thousand Tons)"
    Plot_Name <- "Catch_plot.png"
  }else{
    Map_Label <- "Revenue (Million USD)"
    Plot_Name <- "Rev_plot.png"
  }

  # Filter vatch/valye data for the variable selecter (Map)
  Transboundary_Data_Plot <- Value_Trans_Country %>% 
    filter(Variable %in% Variable_Selected)

  # Total catch/Revenue shapefile
  World_sf_si_land <- World_land_sf %>%
    # st_transform(crs = 4326) %>% # 4326
    st_transform(crs = "+proj=eck4") %>%
    st_simplify(preserveTopology = TRUE, dTolerance = 0.1) %>% 
    left_join(Value_Trans_Country, by ="fishing_entity") %>% 
    filter(Variable %in% Variable_Selected)

  # Get the NA's to be grey
  Grey_land <- World_land_sf %>%  
    filter(!fishing_entity %in% World_sf_si_land$fishing_entity)

  Bins <- Value_Trans_Country %>% 
    filter(Variable %in% Variable_Selected) %>% 
    arrange(Bins) %>% 
    pull(Bins) %>% 
    unique()

  Labels_Trans <- c("(1-29]",paste(Bins_Trans[2:5]))
  Labels <- c("(1-10]",paste(Bins[2:5]))

  Map <- ggplot() +
    geom_sf(data = World_sf_si_land,
            aes(fill = reorder(Bins_Plot,Value)), 
            colour = "grey80",
            size = 0.1) +
    geom_sf(data = World_sf_si,
            aes(fill =  reorder(Bins_Plot,n_trans_spp)),
            colour = "grey20",
            size = 0.1) + # n_trans_spp or Bins
    geom_sf(data = Grey_land,aes(fill = Bins),fill = "grey80", colour = "grey50",size = 0.1) + # n_trans_spp or Bins
    scale_fill_manual("",values = c("#5BBCD6",#"#5BBCD6", # blue
                                    "#00A08A",#"#00A08A", #green
                                    "#F2AD00",#"#F2AD00", # Yellow
                                    "#F98400",#"#F98400", #Orange
                                    "#FF0000"#,"#FF0000" #Red
    ),
    na.translate = F
    ) +
    ggtheme_map() +
    # Transboundary labels
    annotate("text", 
             # x =(c(-85.5,-30,25,81,136)+5),
             # y = -93.5,
             x =(c(-8500000,-3000000,2500000,7500000,12800000)),
             y = -9050000,
             label = Labels_Trans, 
             color="black", 
             size=3, 
             angle=0, 
             # fontface="bold",
             hjust=0) +
    # Revenue labels
    annotate("text",
             # x =(c(-85.5,-31,20,75,131)+5),
             # y = -110,
             x =(c(-8500000,-3200000,2000000,7000000,12500000)),
             y = -11000000,
             label = Labels,
             color="black",
             size=3,
             angle=0,
             # fontface="bold",
             hjust=0) +
    # Legend Titles
    annotate("text", 
             # x =-180,
             # y = -93.5,
             x =-17000000,
             y = -9050000,
             label = "Transboundary Spp. (n)", 
             color="black", 
             size=3, 
             angle=0, 
             fontface="bold",
             hjust=0) +
    annotate("text", 
             # x =-180,
             # y = -110,
             x =-17000000,
             y = -11000000,
             label = Map_Label,
             color="black", 
             size=3, 
             angle=0, 
             fontface="bold",
             hjust=0) +
    theme(
      legend.box.margin = margin(t = -58, l = 120),
      legend.text = element_blank(),
    ) +
    guides(fill = guide_legend(nrow = 1))

  ####___________________________###
  ### Now create the Two circluar bar plots
  ####___________________________###

  # Revenue Plot

  # Set plot limits

  max_Rev_lim <- Subregion_Trans_Value %>%
    filter(Variable == "value") %>%
    group_by() %>%
    summarise(max(Value_sum)) %>%
    pull()

  Subregion_Rev_Data <- Subregion_Trans_Value %>% 
    filter(Variable == "value")

  Rev_Plot <- ggplot(
    # Regional plots
    data = Subregion_Rev_Data,
    aes(
      x=reorder(id,Value_sum),
      y=Value_sum, 
      fill=Value_sum
    )
  ) +    
    ylim(-1500, max_Rev_lim) +
    geom_bar(stat="identity") +
    # Regional averages
    geom_point(data = Subregion_Rev_Data,
               aes(
                 x=reorder(id,Value_sum),
                 y=Value_m
               ),
               colour = "black",
               alpha = 0.5
    ) +
    # Regional sds
    geom_errorbar(data = Subregion_Rev_Data,
                  aes(
                    ymin=ifelse(Value_m-Value_sd <= 0,0,Value_m-Value_sd), # sets lower to 0
                    ymax=Value_m+Value_sd
                  ), 
                  width=.2,
                  position=position_dodge(.9),
                  alpha = 0.5) +
    # Looking good
    theme_minimal() +
    theme(
      legend.position = c(0.73,0.60),
      axis.text = element_blank(),
      axis.title = element_blank(),
      panel.grid = element_blank(),
      plot.margin = margin(0,-10,-6,-10,"cm")
    )  +
    # Base line
    # geom_segment(aes(x = 1, y = 1, xend = 17, yend = 17),  
    #              colour = "black", size=0.1, inherit.aes = FALSE) +
    coord_polar(start = 0) + 
    geom_text(data=label_data_rev, 
              aes(x=id,
                  y=Value_sum,
                  label=sub_region,
                  hjust=hjust), 
              color="black",
              fontface="bold",
              size=2, 
              angle= label_data_rev$angleb, 
              inherit.aes = FALSE) +
    scale_fill_gradientn("",
                         colours = pal,
                         limits = c(0,14000),
                         breaks = seq(0,14000,3500)
                         ) +
    guides(fill = guide_colorbar(label.position = "left")
    );Rev_Plot

  ###____________###
  ### Catch plot
  ###____________###

  max_Catch_lim <- Subregion_Trans_Value %>% 
    filter(Variable == "catch") %>% 
    group_by() %>% 
    summarise(max(Value_sum)) %>% 
    pull()

  Subregion_Catch_Data <- Subregion_Trans_Value %>% 
    filter(Variable == "catch")

  Catch_Plot <- ggplot(data = Subregion_Catch_Data,
                     aes(x=reorder(id,Value_sum),
                         y=Value_sum, 
                         fill=Value_sum
                     )
) +    
  ylim(-1.5, max_Catch_lim) +
  geom_bar(stat="identity") +
  # Regional averages
  geom_point(data = Subregion_Catch_Data,
             aes(
               x=reorder(id,Value_sum),
               y=Value_m
             ),
             colour = "black",
             alpha = 0.5
  ) +
  # Regional sd's
  geom_errorbar(data = Subregion_Catch_Data,
    aes(
      ymin=ifelse(Value_m-Value_sd <= 0,0,Value_m-Value_sd),
      ymax=Value_m+Value_sd
    ), 
    width=.2,
    position=position_dodge(.9),
    alpha = 0.5) +
  theme_minimal() +
  theme(
    legend.position = c(0.1,0.60),
    axis.text = element_blank(),
    axis.title = element_blank(),
    panel.grid = element_blank(),
    plot.margin = margin(0,-10,-6,-5,"cm")
  )  +
  # Base line
  # geom_segment(aes(x = 0, y = 1, xend = 17, yend = 17),
  #              colour = "black", size=0.1, inherit.aes = FALSE) +
  coord_polar(start = 0) + 
  geom_text(data=label_data_catch, 
            aes(x=id,
                y=ifelse(Value_sum >= 12,12,Value_sum+0.1), # Fix names overlap
                label=sub_region,
                hjust=hjust), 
            color="black",
            fontface="bold",
            size=2, 
            angle= label_data_catch$angleb, 
            inherit.aes = FALSE) +
  scale_fill_gradientn("",
                       limits = c(0,12),
                       breaks = seq(0,12,3),
                       colours = pal
  );Catch_Plot

  ## Joining everything in one ploit

  plot_all <- ggdraw() +
  # map
  draw_plot(Map, x = 0, y = 0.45, width = 1, height = 0.58) +
  # Revenue circular
  draw_plot(Rev_Plot, x = 0.1, y = 0.08, width = 0.5, height = 0.4) +
  # Catch circular
  draw_plot(Catch_Plot, x = 0.45, y = 0.08, width = 0.5, height = 0.4) +
  # Labeks
  draw_plot_label(label = c("A", "B"), 
                  size = 20,
                  x = c(0, 0), 
                  y = c(1, 0.45)
  ) +
  draw_plot_label(label = c("Revenue (USD) | Catch (Tons)","(Million)"), 
                  size = 8,
                  x = c(0.330,0.485), 
                  y = c(0.31,0.29)
  )

  gc()
  save_plot(
    # paste(Figures_Path,"100_25_Bins_Value_Region.png",sep=""),
    Plot_Name,
    plot_all,
    base_height = 8,
    base_width = 8
  )

} # End of function
```

### Clean names table

Create a table that matches all names

```{r names_table, eval =F, echo =F}

Clean_sau_eez_fes <- sau_eez_fes %>% 
  gather("Type","fishing_region",2,4) %>% 
  left_join(sau_eez_fes,
            by = ("fishing_entity_id")
  ) %>% 
  group_by(eez_name,fishing_entity,fishing_entity_id) %>% 
  summarise(n()) %>% 
  ungroup() %>% 
  ### Manually Fix names for estimating catch
  mutate(sf_eez_name = ifelse(eez_name =="Greece (without Crete)","Greece", # Vicky
                              ifelse(eez_name =="Greenland (Denmark)","Greenland",
                                     ifelse(eez_name =="United Kingdom (UK)","United Kingdom",
                                            ifelse(fishing_entity =="Italy","Italy",
                                                   ifelse(eez_name =="Chile (mainland)","Chile",
                                                          ifelse(fishing_entity =="Brazil","Brazil",
                                                                 ifelse(eez_name == "Ecuador (mainland)", "Ecuador",
                                                                        ifelse(eez_name =="Portugal (mainland)","Portugal",
                                                                               ifelse(fishing_entity == "Korea (North)","Korea (North)",
                                                                                      ifelse(fishing_entity == "Guadeloupe  (France)","Guadeloupe (France)",
                                                                                             ifelse(eez_name == "South Africa (Atlantic and Cape)","South Africa (Atlantic Coast)",
                                                                                                    ifelse(eez_name =="Viet Nam","Vietnam",
                                                                                                           ifelse(eez_name =="US Virgin Isl.","US Virgin Islands",
                                                                                                                  ifelse(eez_name == "Spain (mainland, Med and Gulf of Cadiz)", "Spain",
                                                                                                                         paste(eez_name)
                                                                                                                  )
                                                                                                           )
                                                                                                    )
                                                                                             ))))))))))
  ),	
  fishing_entity = ifelse(fishing_entity == "Viet Nam","Vietnam",fishing_entity
  )
  )

# UN Regions dataset
UN_Regions <- read.csv(paste(my_path("D"),"Spatial/UN_Regions.csv",sep=""), header = TRUE) %>% 
  clean_names() %>% 
  mutate(un_fishing_entity = country_or_area) %>% 
  group_by(un_fishing_entity,
           fishing_entity = country_or_area,
           georegion = region_name,
           sub_region = sub_region_name) %>% 
  summarise(n=n()) %>% #merge china's 3 rows
  select(-n) %>% 
  filter(fishing_entity != "") %>%  # there are three rows with "developing" status as in, the code
  ungroup() %>% 
  # Change names to match SAU data and for shorter versions
  mutate(
    fishing_entity = gsub("\\s*\\([^\\)]+\\)","",fishing_entity),
    fishing_entity = gsub("Curaçao","Curacao",fishing_entity),
    fishing_entity = gsub("United Kingdom of Great Britain and Northern Ireland","United Kingdom",fishing_entity),
    fishing_entity = gsub( "United States of America", "USA",fishing_entity),
    fishing_entity = gsub("Democratic People's Republic of Korea", "Korea (North)",fishing_entity),
    fishing_entity = gsub("Republic of Korea", "Korea (South)",fishing_entity),
    fishing_entity = gsub("Iran (Islamic Republic of)", "Iran",fishing_entity),
    fishing_entity = gsub("Democratic Republic of the Congo", "Congo (ex-Zaire)",fishing_entity),
    # fishing_entity = gsub("Congo", "Congo, R. of",fishing_entity),
    fishing_entity = gsub("United Republic of Tanzania", "Tanzania",fishing_entity),
    fishing_entity = gsub("Cabo Verde", "Cape Verde",fishing_entity),
    fishing_entity = gsub("Timor-Leste", "Timor Leste",fishing_entity),
    fishing_entity = gsub(" and ", " & ",fishing_entity),
    # fishing_entity = gsub("Puerto Rico (USA)", "Puerto Rico",fishing_entity),
    fishing_entity = gsub("Taiwan", "Taiwan (China)",fishing_entity),
    fishing_entity = gsub("Côte d’Ivoire", "Côte d'Ivoire",fishing_entity),
    # fishing_entity = gsub("Russian Federation", "Russia",fishing_entity)
    fishing_entity = gsub("Viet Nam", "Vietnam",fishing_entity),
    fishing_entity = gsub("Solomon Islands", "Solomon Isl.",fishing_entity),
    fishing_entity = gsub("Marshall Islands", "Solomon Isl.",fishing_entity),
  ) %>% 
  mutate(
    fishing_entity = ifelse(un_fishing_entity == "Congo","Congo, R. of",
                            ifelse(un_fishing_entity == "Democratic Republic of the Congo	","Congo (ex-Zaire)",paste(fishing_entity))
    )
  )

# The path
path_world <- paste(my_path("G"),"Spatial/SAU/SAU_Shapefile",sep="")

# The File
fnam_world <- "SAUEEZ_July2015.shp"

# Load it!
World_sf <- st_read(dsn = path_world,
                    layer =file_path_sans_ext(fnam_world)) %>% 
  rename(eez_name = Name)

World_sf_df <- as.tibble(World_sf) %>% 
  select(sf_eez_name=eez_name,sf_objectid=OBJECTID,sf_eezid=EEZID)

# Vicky
Clean_sau_eez_fes_new <- Clean_sau_eez_fes %>% 
  left_join(UN_Regions) %>% 
  select(-`n()`) %>% 
  full_join(World_sf_df) %>% 
  select(1:3,4,8,9,everything())

#  Now check tim sau data
SAU_Tim_Data %>% 
  anti_join(Clean_sau_eez_fes_new) %>% 
  group_by(fishing_entity_id) %>% 
  summarise(n())

write_csv(Clean_sau_eez_fes_new,
          "matching_names.csv")

```

## Load Results data

In here we load and cleaned the multiple external data sets required to produce the manuscript figures

```{r Load_Results_Data, eval = F}

Results_Path <- my_path("R")

# Set Jurisdiction to EEZ or RFMO
Jurisdiction <- "EEZ"
# Jurisdiction <- "RFMO"

# Updates result's path for the jurisdiction
Results_Path <- paste(my_path("R"),Jurisdiction,"/",sep="")

# Species list to get species names
Exploited_Species <- fread(paste(my_path("D"),"Distribution/Species/exploited_species_list.csv",sep="")) %>% 
  clean_names()

Species_Sampled <- fread(paste(my_path("D"),"Distribution/Sampled_Species.csv", sep =""))

Spp_Clasification <- fread(paste(my_path("D"),"Distribution/Species/SppTaxonName.csv",sep="")) %>% 
  clean_names()

# Conuntry Names relation

matching_country_names <- my_path("D","Spatial","matching_names.csv", read = T) %>%
  mutate(eez_name = ifelse(is.na(eez_name),sf_eez_name,eez_name))

# Spatial Data needed to remove duplicates
# sau_eez_fes <- fread(paste(my_path("G"),"Spatial/SAU/sau_eez_fes.csv",sep="")) %>% 
  # clean_names()

# New dara produced by Vicky 
sau_eez_fes <- read_excel("~/Desktop/EEZ_FishingEntityID_v2_9June2020.xlsx")

Clean_sau_eez_fes <- sau_eez_fes

# UN Regions dataset
UN_Regions <- read.csv(paste(my_path("D"),"Spatial/UN_Regions.csv",sep=""), header = TRUE) %>% 
  clean_names() %>% 
  # mutate(un_fishing_entity = country_or_area) %>% 
  group_by(#un_fishing_entity,
    fishing_entity = country_or_area,
    georegion = region_name,
    sub_region = sub_region_name) %>% 
  summarise(n=n()) %>% #merge china's 3 rows
  select(-n) %>% 
  filter(fishing_entity != "") %>%  # there are three rows with "developing" status as in, the code
  ungroup() %>% 
  # Change names to match SAU data and for shorter versions
  mutate(
    fishing_entity = gsub("","",fishing_entity),
    fishing_entity = gsub("British Indian Ocean Territory","Brit. Indian Ocean Terr. (UK)",fishing_entity),
    fishing_entity = gsub("Bermuda","Bermuda (UK)",fishing_entity),
    fishing_entity = gsub("British Virgin Islands","British Virgin Isl. (UK)",fishing_entity),
    fishing_entity = gsub("Cayman Islands","Cayman Isl. (UK)",fishing_entity),
    fishing_entity = gsub("Christmas Island","Christmas Isl. (Australia)",fishing_entity),
    fishing_entity = gsub("Mayotte","Mayotte (France)",fishing_entity),
    fishing_entity = gsub("Guadeloupe","Guadeloupe (France)",fishing_entity),
    fishing_entity = gsub("Guam","Guam (USA)",fishing_entity),
    fishing_entity = gsub("Martinique","Martinique (France)",fishing_entity),
    fishing_entity = gsub("Montserrat","Montserrat (UK)",fishing_entity),
    fishing_entity = gsub("New Caledonia","New Caledonia (France)",fishing_entity),
    fishing_entity = gsub("Niue","Niue (New Zealand)",fishing_entity),
    fishing_entity = gsub("Norfolk Island","Norfolk Isl. (Australia)",fishing_entity),
    fishing_entity = gsub("Northern Mariana Islands","North Marianas (USA)",fishing_entity),
    fishing_entity = gsub("Pitcairn","Pitcairn (UK)",fishing_entity),
    fishing_entity = gsub("Puerto Rico","Puerto Rico (USA)",fishing_entity),
    fishing_entity = gsub("Saint Helena","Saint Helena (UK)",fishing_entity),
    fishing_entity = gsub("Anguilla","Anguilla (UK)",fishing_entity),
    fishing_entity = gsub("Svalbard and Jan Mayen Islands","Svalbard Isl. (Norway)",fishing_entity),
    fishing_entity = gsub("Tokelau","Tokelau (New Zealand)",fishing_entity),
    fishing_entity = gsub("Turks & Caicos Islands","Turks & Caicos Isl. (UK)",fishing_entity),
    fishing_entity = gsub("United States Virgin Islands","US Virgin Isl.",fishing_entity),
    fishing_entity = gsub("Wallis and Futuna Islands","Wallis & Futuna Isl. (France)",fishing_entity),
    fishing_entity = gsub("Saint Barthélemy","St Barthelemy (France)",fishing_entity),
    fishing_entity = gsub("Iran (Islamic Republic of)","Iran",fishing_entity),
    fishing_entity = gsub("Bonaire","Bonaire (Netherlands)",fishing_entity),
    fishing_entity = gsub("Aruba","Aruba (Netherlands)",fishing_entity),
    fishing_entity = gsub("Curaçao","Curacao",fishing_entity),
    fishing_entity = gsub("United Kingdom of Great Britain and Northern Ireland","United Kingdom",fishing_entity),
    fishing_entity = gsub("United States of America", "USA",fishing_entity),
    fishing_entity = gsub("Democratic People's Republic of Korea", "Korea (North)",fishing_entity),
    fishing_entity = gsub("Republic of Korea", "Korea (South)",fishing_entity),
    fishing_entity = gsub("United Republic of Tanzania", "Tanzania",fishing_entity),
    fishing_entity = gsub("Cabo Verde", "Cape Verde",fishing_entity),
    fishing_entity = gsub("Timor-Leste", "Timor Leste",fishing_entity),
    fishing_entity = gsub(" and ", " & ",fishing_entity),
    # fishing_entity = gsub("Taiwan", "Taiwan (China)",fishing_entity),
    fishing_entity = gsub("Côte d’Ivoire", "Côte d'Ivoire",fishing_entity),
    fishing_entity = gsub("Vietnam", "Viet Nam",fishing_entity),
    fishing_entity = gsub("Solomon Islands", "Solomon Isl.",fishing_entity),
    # fishing_entity = gsub("Marshall Islands", "Solomon Isl.",fishing_entity),
    fishing_entity = gsub("Congo","Congo, R. of",fishing_entity),
    # fishing_entity = gsub("Réunion","Réunion (France)",fishing_entity),
  ) %>%
  # Some of them do not work with gsub
  mutate(
    fishing_entity = ifelse(fishing_entity == "Venezuela (Bolivarian Republic of)","Venezuela",
                                 ifelse(fishing_entity == "Democratic Republic of the Congo, R. of","Congo (ex-Zaire)",
                                        ifelse(fishing_entity == "Iran (Islamic Republic of)","Iran",
                                               ifelse(fishing_entity == "Saint Martin (French Part)","St Martin",
                                                      ifelse(fishing_entity == "Cocos (Keeling) Islands","Cocos (Keeling) Isl. (Australia)",
                                                             ifelse(fishing_entity == "Falkland Islands (Malvinas)","Falkland Isl. (UK)",
                                                                    ifelse(fishing_entity =="Micronesia (Federated States of)","Micronesia",
                                                                           ifelse(fishing_entity == "Saint Pierre & Miquelon","Saint Pierre & Miquelon (France)",
                                                                                  ifelse(fishing_entity == "Sint Maarten (Dutch part)","Sint Maarten",
                                                                                         paste(fishing_entity)
                                                                                  )
                                                                           )
                                                                    )
                                                             )
                                                      )
                                               )
                                        )
                                 )
    )
  ) %>%
  filter(fishing_entity != "Solomon Isl." | sub_region != "Micronesia")

# Load Land shapefile
World_land_sf <- rnaturalearth::ne_countries(scale = 'medium', returnclass = c("sf")) %>% 
  st_transform(crs = "+proj=eck4") %>%
  # st_transform(crs = 4326) %>% # 4326
  rename(fishing_entity = name_long) %>% 
  mutate(
    fishing_entity = gsub(" and ", " & ",fishing_entity),
    fishing_entity = gsub("Anguilla", "Anguilla (UK)",fishing_entity),
    fishing_entity = gsub("American Samoa (USA of America)", "American Samoa",fishing_entity),
    fishing_entity = gsub("Antigua & Barb.", "Antigua & Barbu",fishing_entity),
    fishing_entity = gsub("Aruba","Aruba (Netherlands)",fishing_entity),
    fishing_entity = gsub("Bermuda","Bermuda (UK)",fishing_entity),
    fishing_entity = gsub("Bonaire","Bonaire (Netherlands)",fishing_entity),
    fishing_entity = gsub("British Indian Ocean Territory","Brit. Indian Ocean Terr. (UK)",fishing_entity),
    fishing_entity = gsub("British Virgin Islands","British Virgin Isl. (UK)",fishing_entity),
    fishing_entity = gsub("Cayman Islands","Cayman Isl. (UK)",fishing_entity),
    fishing_entity = gsub("Christmas Island","Christmas Isl. (Australia)",fishing_entity),
    fishing_entity = gsub("Curaçao","Curacao",fishing_entity),
    fishing_entity = gsub("Democratic Republic of the Congo", "Congo (ex-Zaire)",fishing_entity),
    fishing_entity = gsub("Dem. Rep. Korea", "Korea (North)",fishing_entity),
    fishing_entity = gsub("Falkland Islands","Falkland Isl. (UK)",fishing_entity),
    fishing_entity = gsub("Faeroe Islands","Faeroe Isl. (Denmark)",fishing_entity),
    fishing_entity = gsub("Federated States of Micronesia", "Micronesia",fishing_entity),
    fishing_entity = gsub("Guam","Guam (USA)",fishing_entity),
    fishing_entity = gsub("Northern Mariana Islands","North Marianas (USA)",fishing_entity),
    fishing_entity = gsub("Marshall Islands", "Marshall Isl.",fishing_entity),
    fishing_entity = gsub("Montserrat", "Montserrat (UK)",fishing_entity),
    fishing_entity = gsub("New Caledonia", "New Caledonia (France)",fishing_entity),
    fishing_entity = gsub("Niue", "Niue (New Zealand)",fishing_entity),
    fishing_entity = gsub("Norfolk Island", "Norfolk Isl. (Australia)",fishing_entity),
    fishing_entity = gsub("Pitcairn Islands","Pitcairn (UK)",fishing_entity),
    fishing_entity = gsub("Puerto Rico", "Puerto Rico (USA)",fishing_entity),
    fishing_entity = gsub("Republic of Congo", "Congo, R. of",fishing_entity),
    fishing_entity = gsub("Republic of Korea","Korea (South)",fishing_entity),
    fishing_entity = gsub("Saint Helena","Saint Helena (UK)",fishing_entity),
    fishing_entity = gsub("Saint-Martin","St Martin",fishing_entity),
    fishing_entity = gsub("Saint-Barthélemy","St Barthelemy (France)",fishing_entity),
    fishing_entity = gsub("Saint Pierre & Miquelon","Saint Pierre & Miquelon (France)",fishing_entity),
    fishing_entity = gsub("Solomon Islands", "Solomon Isl.",fishing_entity),
    fishing_entity = gsub("Syria", "Syrian Arab Republic",fishing_entity),
    fishing_entity = gsub("Timor-Leste", "Timor Leste",fishing_entity),
    fishing_entity = gsub("Turks & Caicos Islands", "Turks & Caicos Isl. (UK)",fishing_entity),
    fishing_entity = gsub("The Gambia","Gambia",fishing_entity),
    fishing_entity = gsub("United States", "USA",fishing_entity),
    fishing_entity = gsub("Vietnam", "Viet Nam",fishing_entity),
    fishing_entity = gsub("Wallis & Futuna Islands","Wallis & Futuna Isl. (France)",fishing_entity),
    fishing_entity = ifelse(fishing_entity == "USA Virgin Islands","US Virgin Isl.",fishing_entity)
  )

# View(unique(World_land_sf$fishing_entity))

# head(World_land_sf)

# Load SAU shapefile name and paths

# The path
path_world <- paste(my_path("G"),"Spatial/SAU/SAU_Shapefile",sep="")

# The File
fnam_world <- "SAUEEZ_July2015.shp"

# Load it!
World_sf <- st_read(dsn = path_world,
                    layer =file_path_sans_ext(fnam_world)) %>% 
  rename(eez_name = Name,
         eez_id = EEZID) %>% 
  st_transform(crs = "+proj=eck4") %>% 
  mutate(
    eez_id = ifelse(eez_name == "Russia (Laptev to Chukchi Sea)",913,eez_id)
  )

# Regional UN-sub region area
Regional_Area <- World_sf %>% 
  as.data.frame() %>% 
  left_join(Clean_sau_eez_fes) %>% 
  left_join(UN_Regions) %>% 
  filter(fishing_entity %in% Clean_sau_eez_fes$fishing_entity) %>% 
  group_by(sub_region) %>% 
  summarise(Size_Region = sum(Area_km2)) %>% 
  mutate(sub_region = ifelse(is.na(sub_region), "Other",paste(sub_region))) %>% 
  arrange(sub_region)

# ---------------------------- #
# Supplement table 2
# write_csv(Regional_Area,
          # "Table S1.csv")
# ---------------------------- #

# ---------------------------- #
# Transboundary Results ####
# ---------------------------- #

# Read Sau Value Data (Thank you Tim!) reconstructede data
# Gross Domestic Product Implicit Price Deflator according to the US Saint Louis Federal Reserve Bank (average of all 2019 quarters) to transform the original 2010 real USD revenue value to 2019 real USD (https://fred.stlouisfed.org/searchresults?st=GDP+Implicit+price+deflator).*

SAU_Tim_Data <- fread(paste(my_path("D"),"Distribution/sau_catch_value_country_taxon_JEPA.csv", sep ="")) %>%
  filter(!fishing_entity_id %in% c(213,223,207,208)) %>%  # Remove high seas and 207 and 208 that we do not know 
  mutate(fishing_entity_id = ifelse(fishing_entity_id %in% c(219,220), 183,fishing_entity_id)) %>% 
  mutate(value = 1.16 * value)

# Read SAU non-reconstructed catch data from 2006 for comparrisonwith Sumaila
# SAU_Vicky_Data <- fread(paste(my_path("D"),"Distribution/ReportCatch_FEYrEEZTaxon_2010_14.csv", sep =""))

# Set global values for catch and revenue (within EEZ with re-constructed catch)
SAU_Global_Data <- SAU_Tim_Data %>%
  group_by(fishing_entity_id,taxon_key) %>% 
  summarise_at(c("catch","value"),
               mean,na.rm=T) %>% 
  gather("Variable","Value",catch:value)

# Read Species Status Data (Sub-routine Species Status)
Status_Data <- fread(paste(my_path("R"),"EEZ/Species_Status.csv", sep ="")) %>% 
  clean_names() %>% 
  rename(eez_name=territory) #%>% 
  # mutate(eez_name = ifelse(eez_name == "Spain (Mediterranean and Gulf of Cadiz)","Spain",eez_name)
         # )

# List of species
# Spp <- list.files(paste(Results_Path,"Trans_Results/",sep=""))[-829]

# Read Results
# Clean_Results_Trans <- GetTransResults(Spp) %>% 
#   CleanResults(.) %>% 
#   clean_names() %>% 
#   rename(eez_name = territory,
#          fishing_entity = ct_fishing_entity,
#          eez_neighbour = neighbour) %>% 
#   mutate(fishing_entity = ifelse(eez_name == "Spain (Mediterranean and Gulf of Cadiz)","Spain",paste(fishing_entity)))

# Read Results For Carmelia

Clean_Results_Trans <- fread(paste(my_path("R"),"/EEZ/Clean_Results_Trans.csv", sep ="")) %>% 
  mutate(eez_name = ifelse(eez_name == "Cote dIvoire","Côte d'Ivoire",
                           ifelse(eez_name == "Curacao (Netherlands)","Curaçao (Netherlands)",eez_name)
  ),
  fishing_entity = ifelse(fishing_entity == "Cote dIvoire","Côte d'Ivoire",
                          ifelse(fishing_entity == "Curacao (Netherlands)","Curaçao (Netherlands)",
                                 ifelse(eez_name == "Spain (mainland, Med and Gulf of Cadiz)","Spain",fishing_entity)))
  )

```

## Determine Thresholds

Determine transboundary stocks and countries based on selected thresholds. We created two thresholds (levels of uncertainty), in here we can chose which level we wanna work with.

```{r Results_Tresholds, eval = F, echo = F}

#### Select Tresholds ###

# Area Index
Min_AI <- 0.25 # Min area index
Max_AI <- 1-Min_AI # Max Area Index
# Model Index
MI <- 100 # Model Index

#### Determine transboundary and species and nations

Tresholded_T_Data <- Clean_Results_Trans %>% 
  filter(
    area_index >= Min_AI,
    area_index <= Max_AI,
    model_index >= MI
  )

```

## Overall results

```{r Transboundary_Data_per_Territory, eval = F, echo = F}

# -------------------------- #
# Number of EEZs and countries
# -------------------------- #

# number of eezs
length(unique(World_sf$eez_name))

# Number of coastal territories
length(unique(sau_eez_fes$fishing_entity))

# -------------------------- #
# Number of Transboundary and Discrete Species
# -------------------------- #

n_trans <- Tresholded_T_Data %>% 
  pull(taxon_key) %>% 
  unique() %>% 
  length()

# Percentage of transboundary species from sampled
n_trans/nrow(Species_Sampled)*100

# -------------------------- #
# Percentage of capture analyzed
# -------------------------- #

# Filter Tim's data for only those species we estimate
Total_SAU_Sampled <- SAU_Global_Data %>%
  filter(taxon_key %in% Species_Sampled$TaxonKey) %>% 
  group_by(Variable) %>% 
  summarize(Total_Sampled = sum(Value,na.rm=T),
            Sampled_nspp = length(unique(taxon_key))
            )

# Get the total values for the world excluding not identified species
Total_SAU_id <- SAU_Global_Data %>% 
  left_join(Exploited_Species) %>% # list of SAU exploited species identified
  filter(!is.na(taxon_name)) %>% # reove those NAs 
  group_by(Variable) %>% 
  summarize(Total_id = sum(Value,na.rm=T),
            id_nspp = length(unique(taxon_key))
            )

# Get the total values for the world including non id species
Total_SAU_Global <- SAU_Global_Data %>% 
  group_by(Variable) %>% 
  summarize(Total_Global = sum(Value,na.rm=T),
            Global_nspp = length(unique(taxon_key))
            )

# Percentage catch of species analyzed from the identified species
Total_Catch_Spp <- Total_SAU_Sampled$Total_Sampled[1]/Total_SAU_id$Total_id[1] *100 #96.52%

# Percentage species analyzed from the identified species
Total_spp_analyzed <-Total_SAU_Sampled$Sampled_nspp[1]/Total_SAU_id$id_nspp[1]*100 #63.22%

# When we do not remove the groups that are not species level the number sampled drops to ~60%

### So what species are NOT there? 
#  Absten species
# Missing_Species <- SAU_Global_Data %>%
#   filter(!taxon_key %in% Species_Sampled$TaxonKey) %>% 
#   left_join(Exploited_Species) %>% 
#   filter(!is.na(taxon_name)) %>% 
#   group_by(taxon_key,taxon_name,common_name) %>% 
#   summarise(sum(Value))

# -------------------------- #
# Total and percentage Catch and Revenue from Transboundary species
# -------------------------- #

# Filter SAU data for only those sampled species
SAU_Sampled_Data <- Clean_sau_eez_fes %>%
  group_by(fishing_entity,fishing_entity_id) %>% 
  summarise(n=n()) %>% 
  left_join(SAU_Tim_Data, by = "fishing_entity_id") %>% 
  left_join(Exploited_Species) %>% 
  filter(!is.na(taxon_name)) %>% 
  clean_names() %>% 
  group_by(fishing_entity_id,taxon_key) %>% 
  summarise_at(c("catch","value"),
               mean,na.rm=T) %>% 
  gather("Variable","Value",catch:value)

# Numbers for paper
Summary_Trans <- Clean_sau_eez_fes %>%
  group_by(fishing_entity,fishing_entity_id) %>% 
  summarise(n=n()) %>% 
  left_join(SAU_Sampled_Data) %>% 
  filter(taxon_key %in% Species_Sampled$TaxonKey) %>%
  semi_join(Tresholded_T_Data,
            by = c("taxon_key","fishing_entity")
  ) %>% # only counting the catch from transboundary speices (e.g. anchoveta from chile and peru but not from equador)
  group_by(Variable) %>%
  summarise(Total_Trans = sum(Value, na.rm=T)) %>% 
  left_join(Total_SAU_Sampled) %>% 
  left_join(Total_SAU_Global) %>% 
  mutate(Per_sampled = Total_Trans/Total_Sampled*100,
         Per_SAU = Total_Trans/Total_Global*100)

```

```{r comparriosn_wtih_teh}

# Average data for the last 10 years

SAU_Global_Vicky <- SAU_Vicky_Data %>% 
  filter(#eezid != 0,
         year == 2006) %>%
  group_by(fishing_entity_id,taxon_key,eezid) %>% 
  summarise_at(c("sumcatch"),
               mean,na.rm=T)

## Get the total of global data
Total_SAU_Vicky <- sum(SAU_Global_Vicky$sumcatch,na.rm = T)

Summary_Trans_vicky <- Tresholded_T_Data %>% 
  select(taxon_key,eez_name) %>% 
  group_by(taxon_key,eez_name) %>% 
  summarise(n=n()) %>% 
  ungroup() %>% 
  select(-n) %>% 
  left_join(Clean_sau_eez_fes,
            by = "eez_name") %>% 
  group_by(taxon_key,fishing_entity_id) %>% 
  summarise(n()) %>% 
  left_join(SAU_Global_Vicky,
            by = c("taxon_key","fishing_entity_id")) %>% 
  # filter(eezid != 0) %>% # I think they used high seas data as well?
  group_by() %>%
  summarise(Total_Trans = sum(sumcatch, na.rm=T)) %>% 
  mutate(Per_sampled = Total_Trans/Total_SAU_Vicky*100); Summary_Trans_vicky

```

```{r Discrete_Data_per_Territory, eval = F, echo = F}

Species_Left_Out <- Species_Sampled %>% 
  filter(!TaxonKey %in% Clean_Results_Trans$taxon_key) %>% 
  select(taxon_key = TaxonKey)

###_________________###
# Discrete species
# Species that are transboundary in some cases but discrete in other cases
###_________________###

Discrete <- Clean_Results_Trans %>% 
  filter(
    area_index < Min_AI |
    area_index > Max_AI |
    model_index < MI
  ) %>% 
  group_by(taxon_key,fishing_entity,eez_name) %>% 
  summarise(n()) %>% 
  anti_join(Tresholded_T_Data,
            by = c("taxon_key","fishing_entity"))

# Both categories (trans and Discrete)
n_discrete <- Tresholded_T_Data %>% 
  semi_join(Discrete,
            by = "taxon_key") %>% # those that are not only discrete
  group_by(taxon_key) %>% 
  summarise(n()) %>% 
  pull(unique(taxon_key)) %>% 
  length()

Summary_Discrete <- Clean_sau_eez_fes %>%
  group_by(fishing_entity,fishing_entity_id) %>% 
  summarise(n=n()) %>% 
  left_join(SAU_Sampled_Data) %>% 
  semi_join(Discrete,
            by = c("taxon_key","fishing_entity")
            ) %>% # only counting the catch from transboundary speices (e.g. anchoveta from chile and peru but not from equador)
  group_by(Variable) %>%
  summarise(Total_Dis = sum(Value, na.rm=T)) %>% 
  left_join(Total_SAU_Sampled) %>% 
  left_join(Total_SAU_Global) %>% 
  mutate(Per_sampled = Total_Dis/Total_Sampled*100,
         Per_SAU = Total_Dis/Total_Global*100)

###_________________###
# Discrete Exclusive
# Species that are only discrete and are not present in the transboundary category
###_________________###

  # Discrete species (not in Tresholded)
Discrete_Exclusive <- Discrete %>% 
  filter(!taxon_key %in% Tresholded_T_Data$taxon_key) %>% 
  bind_rows(Species_Left_Out) %>%
  group_by(taxon_key) %>% 
  summarise(n()) %>% 
  left_join(Exploited_Species)

## double check

Discrete_Exclusive %>% 
  semi_join(Tresholded_T_Data, by = "taxon_key") # 0 ok thes

# Number of discrete species
n_discrete_exclusive <- length(unique(Discrete_Exclusive$taxon_key))

# Total species
n_discrete_exclusive + n_trans # 938

Summary_Discrete_Exclusive <- Discrete_Exclusive %>% 
  left_join(SAU_Sampled_Data) %>% 
  group_by(Variable) %>%
  summarise(Total_Discrete = sum(Value, na.rm=T)) %>% 
  left_join(Total_SAU_Sampled) %>% 
  left_join(Total_SAU_Global) %>% 
  mutate(Per_sampled = Total_Discrete/Total_Sampled*100,
         Per_SAU = Total_Discrete/Total_Global*100)

```

```{r test, eval = F, echo = F}

Tra <- Summary_Trans %>% 
  mutate(Type = "Trans")

Dis <- Summary_Discrete %>% 
  mutate(Type = "Dis")

Solo <- Summary_Discrete_Exclusive %>% 
  mutate(Type = "Solo")

All <- bind_rows(Tra,Dis,Solo) %>% 
  select(1:7) %>% 
  group_by(taxon_key,fishing_entity) %>% 
  summarise(n())

missing <- anti_join(Total_SAU_Sampled,All,
          by = c("fishing_entity_id","taxon_key"))

missing %>% 
  group_by(Variable) %>% 
  summarise(sum(Value))

# Now
58242635
# Goal
58986239

58986239-58242635

```

## Results per Territory

Determine the number of species per territory

```{r Data_per_Territory, eval = F, echo = F}

# Data per Territory (EEZ)
Transboundary_Spp <- Tresholded_T_Data %>% 
  group_by(eez_name) %>% 
  summarise(
    n_trans_spp = length(unique(taxon_key)),
    n_neighbours = length(unique(eez_neighbour)),
    trans_Rrate = n_trans_spp/n_neighbours
  ) %>% 
  left_join(EEZ_Area, 
            by = "eez_name") %>% 
  mutate(
    Bins = cut(n_trans_spp,breaks = 5),
    Bins_Plot = cut(n_trans_spp,breaks = 5,
                    labels = c('super low', 'low','median', 'high','super high')),
    # Include the weighted method for supplementals
    Weight_Spp = ifelse(n_trans_spp/Size_EEZ < 1,1, n_trans_spp/Size_EEZ),
    # Weight_Bins = cut(Weight_Spp,breaks = 5),
    Weight_Bins_Plot = cut(Weight_Spp,breaks = 5,
                           labels = c('super low', 'low','median', 'high','super high')
    )
  )

# Get Catch and Revenue data for mapping

# Set manual breaks for bins
Map_Breaks <- c(-1,10,100,1000,2000,12000) 
# Map_Breaks <- c(-1,20,200,2000,1000,20000)

###

Total_Value_EEZ <- sau_eez_fes %>% 
  group_by(fishing_entity_id, fishing_entity) %>% 
  summarise(n=n()) %>% # Get all of the fishing entity and id
  select(-n) %>% 
  left_join(SAU_Global_Data) %>% 
  group_by(fishing_entity,Variable) %>% 
  summarise(Total_Value = sum(Value,na.rm=T))

Value_Trans_Country <- sau_eez_fes %>% 
  group_by(fishing_entity_id, fishing_entity) %>% 
  summarise(n=n()) %>% # Get all of the fishing entity and id
  select(-n) %>% 
  left_join(Tresholded_T_Data,
            by = "fishing_entity") %>% 
  group_by(fishing_entity_id,
           fishing_entity,
           taxon_key) %>% 
  summarise(n()) %>% # get each taxa per nation
  left_join(SAU_Sampled_Data,
            by = c("fishing_entity_id","taxon_key")) %>% 
  group_by(fishing_entity,Variable) %>% # Estimate the amount of dollars of each nation 
  summarise(
    Trans_Entity_n = length(unique(taxon_key)),
    Trans_Total = sum(Value,na.rm=T), #million
  ) %>%
  ungroup() %>%
  filter(!is.na(Variable)) %>%
  mutate(
    Trans_Total = ifelse(Variable == "value",Trans_Total/1000000, #million
                         ifelse(Variable == "catch",Trans_Total/1000, #thousand
                                Trans_Total)
                         )
  ) %>% 
  spread(Variable,Trans_Total) %>% 
  # left_join(Entity_Area,
  #           by = "fishing_entity") %>%
  # # Include the weighted method for supplementals
  # mutate_at(
  #   vars(catch,value),
  #   funs(weight= ./(Size_Entity/1000000))
  # ) %>%
  # gather("Variable","Value",catch:value,catch_weight:value_weight) %>%
  gather("Variable","Value",catch:value) %>%
  group_by(fishing_entity,Variable,Value) %>% 
  mutate(
    Bins = cut(Value, breaks = Map_Breaks, dig.lab = 5),
    Bins_Plot = cut(Value, breaks = Map_Breaks,
                    labels = c('super low', 'low','median','high','super high'),
                    dig.lab = 5 # Removes scientific notaion
    )
  ) %>% 
  left_join(Total_Value_EEZ,
            by = c("fishing_entity", "Variable") 
            ) %>% # uncomment for estimating percentage
  # Convert catch to million tonnes
  ungroup() %>% 
  mutate(
    Value = ifelse(Variable == "catch",Value/1000, #million
                         Value),
    Trans_Per = (Value/Total_Value)*100000000 #Return to percentage
  )

```

```{r Data_per_Continent, eval = F, echo = F}

# Continental level for transboundary species
Continent_Trans_Value <- Value_Trans_Country %>% 
  left_join(UN_Regions,
            by = "fishing_entity") %>% 
  mutate(georegion = ifelse(is.na(georegion), "Other",paste(georegion))) %>% 
  group_by(georegion,Variable) %>% 
  summarise_if(is.numeric,sum,na.rm=T) #%>% 
  # select(-Size_Entity,-Trans_Entity_n)

## Sub region leven

Subregion_Trans_raw_Value <- Value_Trans_Country %>% 
  left_join(UN_Regions,
            by = "fishing_entity") %>% 
  mutate(sub_region = ifelse(is.na(sub_region), "Other",paste(sub_region))) %>% 
  group_by(sub_region,Variable) %>% 
  summarise_if(is.numeric,c(sum = sum,
                            m = mean,
                            sd = sd),
               na.rm=T) %>%
  # select(-Size_Entity,-Trans_Entity_n) %>% 
  arrange(Variable) %>%
  ungroup() %>% 
  mutate(id = rep(seq(1,17),2)) %>% 
  mutate(Value_sum=ifelse(Variable == "catch" & Value_sum >= 12,12,Value_sum))

# Fix names for plot
Subregion_Trans_Value <- Subregion_Trans_raw_Value

Subregion_Trans_Value$sub_region <- Subregion_Trans_raw_Value$sub_region %>% 
  gsub("Northern","N.",.) %>% 
  gsub("Southern","S.",.) %>% 
  gsub("Western","W.",.) %>% 
  gsub("Eastern","E.",.)

# Transboundary Species shapefile
World_sf_si <- World_sf %>%
  st_simplify(preserveTopology = TRUE, dTolerance = 0.1) %>%
  left_join(Transboundary_Spp, by ="eez_name") %>% 
  st_transform(crs = "+proj=eck4")
  # st_transform(crs = 4326)

### Prepare data for plot

number_of_bar <- 17 # Number of regions

# Label info Rev
label_data_rev <- Subregion_Trans_Value %>% 
  filter(Variable == "value") %>% 
  arrange(Value_sum) %>% # soo it looks like a nice snail :)
  mutate(
    id = seq(1,number_of_bar),
    angle = 90 -360 * (id-0.5)/number_of_bar,
    hjust = ifelse(angle < -90,1,0),
    angleb = ifelse(angle < -90, (angle +180), angle)
  ) %>% 
  mutate(
    sub_region = ifelse(str_detect(sub_region,"Latin"),"Ltn. Ame. &\nthe Car.",
                        ifelse(sub_region =="Northern America","Northern\nAmerica",
                               paste(sub_region))
    )#,
    # sub_region_n = paste(sub_region," (n=",Trans_Entity_n_sum,")",sep = "")
  )

# Label info Catch
label_data_catch <- Subregion_Trans_Value %>% 
  filter(Variable == "catch") %>% 
  arrange(Value_sum) %>% # soo it looks like a nice snail :)
  mutate(
    id = seq(1,number_of_bar),
    angle = 90 -360 * (id-0.5)/number_of_bar,
    hjust = ifelse(angle < -90,1,0),
    angleb = ifelse(angle < -90, (angle +180), angle)
  ) %>% 
  mutate(
    sub_region = ifelse(str_detect(sub_region,"Latin"),"Ltn. Ame. &\nthe Car.",
                        ifelse(sub_region =="Northern America","Northern\nAmerica",
                               paste(sub_region))
    )#,
    # sub_region_n = paste(sub_region," (n=",Trans_Entity_n_sum,")",sep = "")
  )

# Create labels for legend
Bins_Trans <- Transboundary_Spp %>% 
  arrange(Bins) %>% 
  pull(Bins) %>% 
  unique()

```

### EEZ Map of Transboundary Species

Produce the map of transboundary species within each EEZ and the value in catch/revenue within each country

```{r Figure_1}

# Set color pal
pal <- wes_palette("Zissou1",100,type = "continuous")

# Variable to plot (value or catch)
Variable_plot <- "value"
CreateFigureOne(Variable_plot)

```

```{r Weight_Revenue, eval =F, echo =F}

Weighted_values <-c("value_weight","catch_weight")

# data
Weighted_Data <- Subregion_Trans_Value %>% 
  filter(Variable %in% Weighted_values)

# Labels

label_data_rev <- Subregion_Trans_Value %>% 
  filter(Variable == Weighted_values[1]) %>% 
  arrange(Value_sum) %>% # soo it looks like a nice snail :)
  mutate(
    id = seq(1,number_of_bar),
    angle = 90 -360 * (id-0.5)/number_of_bar,
    hjust = ifelse(angle < -90,1,0),
    angleb = ifelse(angle < -90, (angle +180), angle)
  ) %>% 
  mutate(
    sub_region = ifelse(str_detect(sub_region,"Latin"),"Ltn. Ame. &\nthe Car.",
                        ifelse(sub_region =="Northern America","Northern\nAmerica",
                               paste(sub_region)
                               )
    )
  )

# Label info Catch
label_data_catch <- Subregion_Trans_Value %>% 
  filter(Variable == Weighted_values[2]) %>% 
  arrange(Value_sum) %>% # soo it looks like a nice snail :)
  mutate(
    id = seq(1,number_of_bar),
    angle = 90 -360 * (id-0.5)/number_of_bar,
    hjust = ifelse(angle < -90,1,0),
    angleb = ifelse(angle < -90, (angle +180), angle)
  ) %>% 
  mutate(
    sub_region = ifelse(str_detect(sub_region,"Latin"),"Ltn. Ame. &\nthe Car.",
                        ifelse(sub_region =="Northern America","Northern\nAmerica",
                               paste(sub_region))
    )
  )

# Revenue Plot

  # Set plot limits

  max_Rev_lim <- Subregion_Trans_Value %>%
    filter(Variable == Weighted_values[1]) %>%
    group_by() %>%
    summarise(max(Value_sum)) %>%
    pull()

  Subregion_Rev_Data <- Subregion_Trans_Value %>% 
    filter(Variable == Weighted_values[1])

  Rev_Plot <- ggplot(
    # Regional plots
    data = Subregion_Rev_Data,
    aes(
      x=reorder(id,Value_sum),
      y=Value_sum, 
      fill=Value_sum
    )
  ) +    
    ylim(-50, 250) +
    geom_bar(stat="identity") +
    # Regional averages
    geom_point(data = Subregion_Rev_Data,
               aes(
                 x=reorder(id,Value_sum),
                 y=Value_m
               ),
               colour = "black",
               alpha = 0.5,
               size = 0.5
    ) +
    # Regional sds
    geom_errorbar(data = Subregion_Rev_Data,
                  aes(
                    ymin=ifelse(Value_m-Value_sd <= 0,0,Value_m-Value_sd), # sets lower to 0
                    ymax=Value_m+Value_sd
                  ), 
                  width=.2,
                  position=position_dodge(.9),
                  alpha = 0.5) +
    # Looking good
    theme_minimal() +
    theme(
      legend.position = "",
      axis.text = element_blank(),
      axis.title = element_blank(),
      panel.grid = element_blank(),
      plot.margin = margin(1,0,0,-3,"cm")
    )  +
    coord_polar(start = 0) + 
    geom_text(data=label_data_rev, 
              aes(x=id,
                  y=Value_sum+1,
                  label=sub_region,
                  hjust=hjust), 
              color="black",
              fontface="bold",
              size=2, 
              angle= label_data_rev$angleb, 
              inherit.aes = FALSE) +
    scale_fill_gradientn("",
                         colours = pal,
                         limits = c(0,250),
                         breaks = seq(0,250,50)
                         );Rev_Plot

  ###____________###
  ### Catch plot
  ###____________###

  max_Catch_lim <- Subregion_Trans_Value %>% 
    filter(Variable == Weighted_values[2]) %>% 
    group_by() %>% 
    summarise(max(Value_sum)) %>% 
    pull()

  Subregion_Catch_Data <- Subregion_Trans_Value %>% 
    filter(Variable == Weighted_values[2])

  Catch_Plot <- ggplot(data = Subregion_Catch_Data,
                     aes(x=reorder(id,Value_sum),
                         y=Value_sum, 
                         fill=Value_sum
                     )
) +    
  ylim(-50, 251) +
  geom_bar(stat="identity") +
  # Regional averages
  geom_point(data = Subregion_Catch_Data,
             aes(
               x=reorder(id,Value_sum),
               y=Value_m
             ),
             colour = "black",
             alpha = 0.5,
             size = 0.5
  ) +
  # Regional sd's
  geom_errorbar(data = Subregion_Catch_Data,
    aes(
      ymin=ifelse(Value_m-Value_sd <= 0,0,Value_m-Value_sd),
      ymax=Value_m+Value_sd
    ), 
    width=.2,
    position=position_dodge(.9),
    alpha = 0.5) +
  theme_minimal() +
  theme(
    legend.position = "bottom",
    legend.key.width =unit(2,"line"),
    legend.key.height =unit(0.5,"line"),
    axis.text = element_blank(),
    axis.title = element_blank(),
    panel.grid = element_blank(),
    plot.margin = margin(1,0,0,0,"cm")
  )  +
  coord_polar(start = 0) + 
  geom_text(data=label_data_catch, 
            aes(x=id,
                y=ifelse(sub_region=="N. Europe",250,Value_sum+1), # Fix names overlap
                label=sub_region,
                hjust=hjust), 
            color="black",
            fontface="bold",
            size=2, 
            angle= label_data_catch$angleb, 
            inherit.aes = FALSE) +
  scale_fill_gradientn("",
                       limits = c(0,251),
                       breaks = seq(0,250,50),
                       colours = pal
  ); Catch_Plot

  ## Joining everything in one ploit

  # Get legend for cowplot
legend_plot <- get_legend(Catch_Plot)

# Re-plot without legend
Catch_Plot <- Catch_Plot +
  theme(legend.position = "")

  p<- ggdraw() +
  # Revenue circular
  draw_plot(Rev_Plot, x = 0, y = 0, width = 0.8, height = 1) +
  # Catch circular
  draw_plot(Catch_Plot, x = 0.4, y = 0, width =0.8, height = 1) +
  # Legend
  draw_plot(legend_plot, x = 0.2, y = 0.1, width = 0.59, height = 0.1) +
  # Labeks
  draw_plot_label(label = c("A", "B"),
                  size = 15,
                  x = c(0.1, 0.6),
                  y = c(1, 1)
  ) +
  draw_plot_label(
    label = c("Revenue (USD) | Catch (Tons)","Million per Km2"),
                  size = 6,
                  x = c(0.3,0.40),
                  y = c(0.25,0.22)
  )

  save_plot(
    paste(Figures_Path,"weight_plot.png",sep=""),
    p,
    base_height = 4,
    base_width = 6
  )

```

### China, US, Russia, and Peru

```{r Map_Particulars, eval =F, echo =F}

# Revenue details

Top_Rev_Countries <- c("China", "Peru", "Russian Federation", "Chile", "USA","Japan")

# Total revenue from Top countries
Total_Top_Rev <- Value_Trans_Country %>% 
  filter(Variable == "value") %>% 
  filter(fishing_entity %in% Top_Rev_Countries) %>% 
  left_join(UN_Regions,
            by = "fishing_entity") %>% 
  left_join(Subregion_Trans_raw_Value) %>% 
  mutate(Per_Top = Value/Value_sum*100) %>% 
  select(1:3,Per_Top)

Top_Rev <- sau_eez_fes %>% 
  filter(fishing_entity %in% Top_Rev_Countries) %>% 
  group_by(fishing_entity_id, fishing_entity) %>% 
  summarise(n=n()) %>% # Get all of the fishing entity and id
  select(-n) %>% 
  full_join(Tresholded_T_Data,
            by = "fishing_entity") %>% 
  group_by(fishing_entity_id,
           fishing_entity,
           taxon_key) %>% 
  summarise(n()) %>% # get each taxa per nation
  left_join(SAU_Sampled_Data,
            by = c("fishing_entity_id","taxon_key")) %>% 
  filter(Variable == "value") %>% 
  left_join(Exploited_Species) %>% 
  group_by(fishing_entity) %>%
  top_n(16,Value) %>%
  select(fishing_entity,Variable,Value,taxon_name) %>% 
  mutate(Top_Value = Value/1000000) %>% # Million of USD
  select(-Value) %>% 
  left_join(Total_Top_Rev,
            by = c("fishing_entity","Variable")
            ) %>% 
  mutate(Ratio = Top_Value/Value*100)

Top_Rev %>% 
  group_by(fishing_entity) %>% 
  summarise(sum(Ratio))

### Total revenue from the top 5

Top5 <- c("Peru", "Russian Federation", "Japan", "USA","China")

Total_Top5 <- Value_Trans_Country %>% 
  filter(Variable == "value") %>% 
  filter(fishing_entity %in% Top_Rev_Countries) %>% 
  ungroup() %>% 
  select(fishing_entity,Value) %>% 
  group_by() %>% 
  summarise(sum(Value)) %>% 
  pull()/(Total_SAU_Sampled$Total_Sampled[2]/1000000)*100

```

### Statistical differences within regions
We used a one-way ANOVA to determine significant differences within regions for transboundary fisheries catches and revenue with a Tukey's HSD post-hoc test.

```{r statistics_fig1, eval = F, echo = T}

### Using ANOVA to estimate significan differences within regions for revenue and catch of transboundary species

### Reveneu ANOVA

Subregion_Rev_stats <- Subregion_Data_stats %>% 
  filter(Variable == "value")

Rev_aov <- aov(Value~sub_region, data = Subregion_Rev_stats)
Rev_aov

summary(Rev_aov)
#              Df    Sum Sq Mean Sq F value   Pr(>F)    
# sub_region   16  75738982 4733686   4.948 3.46e-08 ***
# Residuals   158 151144071  956608

#  Post-hoc analysis
TukeyHSD(Rev_aov) %>% 
  tidy() %>% 
  filter(adj.p.value < 0.05)

### Catch ANOVA

Subregion_Catch_stats <- Subregion_Data_stats %>% 
  filter(Variable == "catch")

Catch_aov <- aov(Value~sub_region, data = Subregion_Catch_stats)
Catch_aov

summary(Catch_aov)
#  Df Sum Sq Mean Sq F value  Pr(>F)   
# sub_region   16  27.51  1.7195   2.265 0.00539 **
# Residuals   158 119.97  0.7593

#  Post-hoc analysis
TukeyHSD(Catch_aov) %>% 
   tidy() %>% 
  filter(adj.p.value < 0.05) # Eastern Asia

# Weight ANOVA
Subregion_Weigh_stats <- Subregion_Data_stats %>% 
  filter(Variable == "value_weight")

Weight_aov <- aov(Value~sub_region, data = Subregion_Weigh_stats)
Weight_aov
summary(Weight_aov)

#  Df Sum Sq Mean Sq F value  Pr(>F)   
# sub_region   16   6722   420.1   2.411 0.00289 **
# Residuals   158  27532   174.3                   
# ---
# Signif. codes:  0 ‘***’ 0.001 ‘**’ 0.01 ‘*’ 0.05 ‘.’ 0.1 ‘ ’ 1

#### There are significant differences. But between what?

#  Post-hoc analysis
TukeyHSD(Weight_aov) %>% 
  tidy() %>% 
  filter(adj.p.value < 0.05) # Northern Europe
```

### OHI-inspired Circular plot for Territories

```{r Status_Plot_Data, eval = F, echo = F}

# -------------------------- #
# Transboundary data
# -------------------------- #

# Stock status Data

Status_Data_Plot <- Tresholded_T_Data %>% 
  left_join(Status_Data, 
            by =c("eez_name","taxon_key")
  ) %>%
  mutate(status = ifelse(is.na(status),"No Status",status)) %>% 
  group_by(fishing_entity,status) %>% 
  summarise(value = length(unique(taxon_key))) %>% 
  left_join(UN_Regions, by ="fishing_entity") %>%
  mutate(Georegion = ifelse(is.na(georegion),"Other",paste(georegion))) %>% 
  mutate(value = ifelse(value > 100,100,value)) %>% 
  filter(!is.na(fishing_entity),
         !str_detect(fishing_entity, ".\\(.") # remove associated territories
  ) %>% 
  select(individual = fishing_entity,observation = status,group = Georegion,value) %>% 
  arrange(group,value) %>% 
  mutate(value = ifelse(observation %in% c("Collapsed","Over Exploited"),value*-1,value),
         observation = ifelse(observation %in% c("Developing","Rebuilding"),"Category A",
                              ifelse(observation == "Max Exploited","Category B",
                                     ifelse(observation %in% c("Over Exploited","Collapsed","Category C"),"Category C", "No Category")
                              )
         )
  ) %>% 
  group_by(individual,observation,group) %>% 
  summarise(value = sum(value))

Status_Plot_Data <- tibble(
  fishing_entity = rep(unique(Status_Data_Plot$individual),4),
  observation = rep(unique(Status_Data_Plot$observation),153)
) %>% 
  left_join(UN_Regions) %>% 
  rename(individual = fishing_entity) %>% 
  full_join(Status_Data_Plot,
            by = c("individual","observation")
  ) %>% 
  mutate(value = ifelse(is.na(value),0,
                        ifelse(value < -100,-100, # if less than 100, 100
                               ifelse(value > 100,100, # if more than 100, 100
                                      value)
                        )#, # Ser NAs to 0's
                        # group = ifelse(is.na(group),"Other",group)
  )
  ) %>% 
  select(individual,observation,group=georegion,value) %>% 
  filter(group != "other")

# Summary for manuscipt

T_Summary <- Status_Data_Plot %>% 
  mutate(value = ifelse(value <0, value*-1,value)) %>% 
  group_by(observation) %>% 
  summarise(tot=sum(value),
            mean(value),
            sd(value)) %>% 
  arrange(desc(tot))

# -------------------------- #
# Discrete data
# -------------------------- #

Discrete_Status_Data <- Discrete %>% 
  left_join(Status_Data, 
            by =c("eez_name","taxon_key")
  ) %>%
  mutate(status = ifelse(is.na(status),"No Status",status)) %>% 
  group_by(fishing_entity,status) %>% 
  summarise(value = length(unique(taxon_key))) %>% 
  left_join(UN_Regions, by ="fishing_entity") %>%
  mutate(Georegion = ifelse(is.na(georegion),"Other",paste(georegion))) %>% 
  # mutate(value = ifelse(value > 100,100,value)) %>% 
  filter(!is.na(fishing_entity),
         !str_detect(fishing_entity, ".\\(.") # remove associated territories
  ) %>% 
  select(individual = fishing_entity,observation = status,group = Georegion,value) %>% 
  arrange(group,value) %>% 
  mutate(value = ifelse(observation %in% c("Collapsed","Over Exploited"),value*-1,value),
         observation = ifelse(observation %in% c("Developing","Rebuilding"),"Category A",
                              ifelse(observation == "Max Exploited","Category B",
                                     ifelse(observation %in% c("Over Exploited","Collapsed","Category C"),"Category C", "No Category")
                              )
         )
  ) %>% 
  group_by(individual,observation,group) %>% 
  summarise(value = sum(value))

# Summary for manuscipt

D_Summary <- Discrete_Status_Data %>% 
  mutate(value = ifelse(value <0, value*-1,value)) %>% 
  group_by(observation) %>% 
  summarise(tot=sum(value),
            mean(value),
            sd(value)) %>% 
  arrange(desc(tot))

#### Supplemental Table 3

# T_Summary %>%
#   left_join(D_Summary, by = "observation") %>%
#   mutate(
#     "Mean Shared" = paste(round(`mean(value).x`,2)," \u00B1 (",round(`sd(value).x`,2),")",sep=""),
#     "Mean Discrete" = paste(round(`mean(value).y`,2)," \u00B1 (",round(`sd(value).y`,2),")",sep="")
#     ) %>%
#   select(1,8:9) %>%
#   filter(observation != "No Category") %>%
#   rename(Category = observation) %>%
#   arrange(Category) %>%
#   write_csv("Category_summary.csv")

```

```{r Territories_Circluar_Plot, eval = F, echo = T}

# -------------------------- #
# Data manipulation for plot
# -------------------------- #

# Set a number of 'empty bar' to add at the end of each Georegion
Char <- as.factor(Status_Plot_Data$group) # Set groups to factors

##### COPY PASTE

# Transform data in a tidy format (long format)
# data <- Status_Data_Plot

# Set a number of 'empty bar' to add at the end of each group
empty_bar <- 2
nObsType <- nlevels(as.factor(Status_Plot_Data$observation))
to_add <- data.frame(matrix(NA, empty_bar*nlevels(Char)*nObsType, ncol(Status_Plot_Data)) )
colnames(to_add) <- colnames(Status_Plot_Data)
to_add$group <- rep(levels(Char), each=empty_bar*nObsType)
Status_Plot_Data <- rbind(Status_Plot_Data, to_add) # Problem converts to list
# Status_data <- bind_rows(data,to_add)
Status_Plot_Data <- Status_Plot_Data %>% arrange(group, individual)
Status_Plot_Data$id <- rep( seq(1, nrow(Status_Plot_Data)/nObsType) , each=nObsType)

# to_add$value <- NA
# to_add$observation <- NA
# to_add$individual <- to_add$group

# Status_data <- Status_Data_Plot %>%
#   bind_rows(to_add) %>%
#   ungroup() %>%
#   mutate(individual = forcats::fct_explicit_na(individual),
#          group = forcats::fct_explicit_na(group)) %>%
#   group_by(group,individual) %>%
#   mutate(id = as.integer(individual),
#          order = as.integer(group))

# Get the name and the y position of each label
label_data <- Status_Plot_Data %>% 
  # filter(value > 0) %>% 
  group_by(id, individual) %>% 
  summarize(tot=sum(value,na.rm=T)) %>% 
  mutate(
    individual = ifelse(individual == "Saint Vincent & the Grenadines","Snt. Vincent & Grenadines", individual)
    )

number_of_bar <- nrow(label_data)
angle <- 90 - 360 * (label_data$id-0.5) /number_of_bar     # I substract 0.5 because the letter must have the angle of the center of the bars. Not extreme right(1) or extreme left (0)
label_data$hjust <- ifelse( angle < -90, 1, 0)
label_data$angle <- ifelse(angle < -90, angle+180, angle)

label_data <- label_data %>% 
  mutate(individual = ifelse(individual == "(Missing)",paste(" "),paste(individual)))

# prepare a data frame for base lines
base_data <-  Status_Plot_Data %>% 
  group_by(group) %>% 
  summarize(start=min(id,na.rm=T), 
            end=max(id,na.rm=T) - empty_bar,
            median = median(value, na.rm = T)) %>% 
  rowwise() %>% 
  mutate(title=mean(c(start, end)))

# prepare a data frame for grid (scales)
grid_data <- base_data
grid_data$end <- grid_data$end[ c( nrow(grid_data), 1:nrow(grid_data)-1)] + 1
grid_data$start <- grid_data$start - 1
grid_data <- grid_data[-1,]

Status_Plot_Data %>%
  ggplot() + 
  geom_bar(aes(x=id,
               y=value, 
               fill=observation),
           stat="identity", 
           alpha=0.5) +
  # scale_fill_manual("Catch Category",
  #                   values = c(rev(wes_palette(n = 5, name = "Rushmore1")),"grey")
  #                   ) +
  scale_fill_manual("Catch Category",
                    values = c(
                      "#046C9A", # Developing / Rebulding / Cat A / blue
                      "#0B775E", #Max exp. / Cat B / green
                      "#F2300F", # collapsed / Overexp. / Cat. C / Red
                      "#8D8680"
                    ),
                    na.value = "white"
  ) +
  # Add a val=100/75/50/25 lines. I do it at the beginning to make sur barplots are OVER it.
  geom_segment(data=grid_data, aes(x = end, y = 0, xend = start, yend = 0), colour = "black", alpha=1, size=0.3 , inherit.aes = FALSE ) +
  geom_segment(data=grid_data, aes(x = end, y = 50, xend = start, yend = 50), colour = "black", alpha=1, size=0.3, inherit.aes = FALSE ) +
  geom_segment(data=grid_data, aes(x = end, y = 100, xend = start, yend = 100), colour = "black", alpha=1, size=0.3, inherit.aes = FALSE ) +
  geom_segment(data=grid_data, aes(x = end, 
                                   y = -50, 
                                   xend = start, 
                                   yend = -50), colour = "black", alpha=1, size=0.3 , inherit.aes = FALSE ) +
  # Add text showing the value of each 100/75/50/25 lines
  ggplot2::annotate("text", 
                    # x = rep(max(data$id),3),
                    x =c(0.2,0,0,0.2,0.2),
                    y = c(-100,-50, 0, 50, 100),
                    label = c(">100","50","0", "50", ">100") , 
                    color="black", 
                    size=5, 
                    angle=0, 
                    # fontface="bold",
                    hjust=1) +
  # Add labels on top of each bar
  geom_text(data=label_data,
            aes(
              x = id,
              y = 105,
              label=individual,
              hjust=hjust),
            color="black",
            # fontface="bold",
            alpha=1,
            size=4,
            angle= label_data$angle,
            inherit.aes = FALSE ) +
  # Continents labels
  geom_text(data=base_data,
            aes(x = title,
                y = 85, 
                label=group), 
            hjust=c(1,1,1,0,0,0),
            colour = "black",
            # alpha=0, 
            size=5,
            fontface="bold",
            inherit.aes = FALSE) +
  ylim(-100,105) +
  theme_minimal() +
  theme(legend.position = c(0.93,0.14),
        legend.key.size = unit(1,"cm"),
        legend.title = element_text(size = 16),
        legend.text = element_text(size = 14),
        axis.text = element_blank(),
        axis.title = element_blank(),
        panel.grid = element_blank(),
        plot.margin = unit(rep(-1,4), "cm")) +
  coord_polar()

# Save plot

ggsave(
  plot = last_plot(),
  width = 17,
  height = 15,
  units = "in",
  # filename = "Plot.png"
  filename = "Territory_circular_status.png"
)

```

### Statistical differences between species status

Using MANOVA to investigate differences in catch category between discrete and transboundary species

```{r statistics_fig2, eval = F, echo = T}

#-----------------# 
# Get Data
#-----------------#

# Transboundary Data
Transboundary_Data_Stats <- Status_Plot_Data %>% 
filter(observation != "No Category") %>%
  mutate(value = abs(value))

# Discrete Data
Discrete_Data_Stats <- Discrete_Status_Data %>% 
  filter(observation != "No Category") %>%
  mutate(value = abs(value))

#### By Category (Manova)

# Use manova to test for a significant effect of Catch category on discrete and transboundary species simultaneously. Takes into account possible correlations between dependent variables.
# Decreases Tipe I errors

Manova_Differences <- Transboundary_Data_Stats %>% 
  rename(Trans_value = value) %>% 
  left_join(Discrete_Data_Stats, by = c("individual","observation","group")
  ) %>% 
  rename(Discrete_value= value) %>% 
  mutate(Discrete_value = ifelse(is.na(Discrete_value),0,Discrete_value))

#-----------------# 
## Test assumptions
#-----------------#

Manova_Differences %>% 
  gather("cat","val",Trans_value,Discrete_value) %>% 
  ggplot() +
  geom_histogram(
    aes(
      x = val
    )
  ) +
  facet_wrap(~observation+cat)

### Correlation

plot(Manova_Differences$Trans_value~Manova_Differences$Discrete_value)
cor(Manova_Differences$Trans_value,Manova_Differences$Discrete_value)

#-----------------# 
## MANOVA test
#-----------------#

mano <- manova(cbind(Discrete_value,Trans_value) ~ observation, data = Manova_Differences)

summary(mano)
# Df  Pillai approx F num Df den Df    Pr(>F)    
# observation   2 0.21621   27.091      4    894 < 2.2e-16 ***
# Residuals   447

summary.aov(mano)
#  Response Discrete_value :
#              Df Sum Sq Mean Sq F value    Pr(>F)    
# observation   2   7088  3544.1  26.305 1.584e-11 ***
# Residuals   447  60226   134.7                      
# ---
# Signif. codes:  0 ‘***’ 0.001 ‘**’ 0.01 ‘*’ 0.05 ‘.’ 0.1 ‘ ’ 1
# 
#  Response Trans_value :
#              Df Sum Sq Mean Sq F value    Pr(>F)    
# observation   2  22778 11389.2  55.005 < 2.2e-16 ***
# Residuals   447  92555   207.1

#-----------------# 
## Run non-parametric to make sure
# Same result
#-----------------#

# kruskal.test(Trans_value~Discrete_value, data = Manova_Differences)
# Kruskal-Wallis chi-squared = 100.24, df = 45, p-value = 4.358e-06

#-----------------# 
## ANOVA test for each category
#-----------------#

# Transboundary Species

#### Test for normallity 
Transboundary_Data_Stats %>%
  ggplot() +
  geom_histogram(
    aes(
      x = value
    )
  ) +
  facet_wrap(~observation)

bartlett.test(value~observation, Transboundary_Data_Stats)

# Differences between categories of transboundary species

Stat_Cat_aov <- aov(value~observation, data = Transboundary_Data_Stats)
Stat_Cat_aov
summary(Stat_Cat_aov)
#              Df Sum Sq Mean Sq F value Pr(>F)    
# observation   2  22778   11389   55.01 <2e-16 ***
# Residuals   447  92555     207

TukeyHSD(Stat_Cat_aov)
#                            diff       lwr       upr     p adj
# Category B-Category A  5.946667  2.039431  9.853902 0.0011125
# Category C-Category A 17.160000 13.252765 21.067235 0.0000000
# Category C-Category B 11.213333  7.306098 15.120569 0.0000000

#-----------------# 
# Double check the Non parametic test
# same result
#-----------------#

# Stat_Cat_kt <- kruskal.test(observation~value, data = Transboundary_Data_Stats)
# Stat_Cat_kt
# 
# # Post hock test
# kruskalmc(Transboundary_Data_Stats$value,Transboundary_Data_Stats$observation)

# -------------------------- #
# Discrete species
# -------------------------- #

#### Test for normallity 
Discrete_Data_Stats %>%
  ggplot() +
  geom_histogram(
    aes(
      x = value
    )
  ) +
  facet_wrap(~observation)

bartlett.test(value~observation, Discrete_Data_Stats)

# ANOVA of Discrete categories
Stat_discrete_cat_aov <- aov(value~observation, data = Discrete_Data_Stats)
Stat_discrete_cat_aov
summary(Stat_discrete_cat_aov)

#             Df Sum Sq Mean Sq F value   Pr(>F)    
# observation   2   6442    3221   21.86 1.03e-09 ***
# Residuals   380  55992     147

TukeyHSD(Stat_discrete_cat_aov)
# 
#  diff         lwr       upr     p adj
# Category B-Category A 3.627759 -0.02854633  7.284065 0.0523425
# Category C-Category A 9.917391  6.31117452 13.523608 0.0000000
# Category C-Category B 6.289632  2.79874173  9.780522 0.0000835

#-----------------# 
# Double check the Non parametic test
# All different... But it does not show any distance, just the categorial difference
#-----------------#

Stat_discrete_cat_kt <- kruskal.test(observation~value, data = Discrete_Data_Stats)
Stat_discrete_cat_kt

kruskalmc(Discrete_Data_Stats$value,Discrete_Data_Stats$observation)

```

## Results per Species

### Mackerel (*Scomber scombrus*) war

```{r mackrell, eval = F, echo =F}

Exploited_Species %>% 
  filter(taxon_name == "Scomber scombrus") # 600118

NW_Europe <- UN_Regions %>% 
  filter(sub_region  %in% c("Western Europe", "Northern Europe")) %>% 
  pull(fishing_entity) %>% 
  unique()

Conflict_nations <- c(NW_Europe,"Portugal","Spain")

Mackerel <- Clean_sau_eez_fes %>% 
  group_by(fishing_entity, fishing_entity_id) %>% 
  summarise(n()) %>% 
  left_join(SAU_Sampled_Data, by = "fishing_entity_id") %>% 
  filter(taxon_key == 600118,
         fishing_entity %in% Conflict_nations
  ) %>% 
  group_by(Variable) %>% 
  summarise(Mac = sum(Value))

### Mackerel percentage of region's total catch

Conflict_Values <- Clean_sau_eez_fes %>% 
  group_by(fishing_entity, fishing_entity_id) %>% 
  summarise(n()) %>% 
  left_join(SAU_Sampled_Data, by = "fishing_entity_id") %>% 
  filter(
         fishing_entity %in% Conflict_nations
  ) %>% 
  group_by(Variable) %>% 
  summarise(Conflict= sum(Value))

Mackerel %>% 
  left_join(Conflict_Values) %>% 
  mutate(Mac/Conflict*100)

```

Create figures that are based in species (e.g. number of countries each species has)

```{r Circular_Spp_Data, eval = F, echo =F}

# Number of countries shared by each Species
Species_Trans <- Tresholded_T_Data %>% 
  group_by(taxon_key) %>% 
  summarise(
    n_countries = length(unique(fishing_entity))
  ) %>% 
  left_join(Exploited_Species) %>%
  left_join(Spp_Clasification)

# hist(Species_Trans$n_countries)

# Use rfishbase to get species environment
Species_list <- Species_Trans %>% 
  mutate(scientific_name = paste(genus,species)) %>% 
  pull(scientific_name)

Species_environ_list <- rbind(species(Species_list, #species from rfishbase package
                                      fields = species_fields$habitat[4])
) %>% 
  mutate(scientific_name = Species_list)

Raw_Species_Data <- Tresholded_T_Data %>% 
  group_by(taxon_key) %>% 
  summarise(n_trans_spp = length(unique(fishing_entity))) %>% 
  left_join(Exploited_Species) %>%
  rename(scientific_name = taxon_name) %>% 
  left_join(Species_environ_list) %>%
  mutate(DemersPelag = ifelse(is.na(DemersPelag),"Other",
                              ifelse(DemersPelag %in% c("bathypelagic","benthopelagic"), "bathy-bentho-pelagic",paste(DemersPelag)
                              )
  ),
  n_trans_spp = ifelse(n_trans_spp > 100, 100, n_trans_spp),
  DemersPelag = str_to_title(DemersPelag)
  ) %>% 
  select(individual=scientific_name,
         value = n_trans_spp,
         group = DemersPelag) %>% 
  arrange(group,value)  %>% 
  mutate(
    individual = ifelse(individual == "Chloroscombrus chrysurus", "C. chrysurus",
                        ifelse(individual == "Scomberomorus commerson", "S. commerson",
                               individual)
    )
  ) %>% 
  filter(!group %in% c("Bathydemersal","Pelagic"))

#### Ecosystem based plots

Top_5_Ecosystem <- Raw_Species_Data %>% 
  rename(taxon_name = individual) %>% 
  group_by(group) %>% 
  top_n(5,value)

# If you want to show revenue
# Country_Spp_Data <- sau_eez_fes %>% 
#   group_by(fishing_entity_id,fishing_entity) %>% 
#   summarise(n=n()) %>% 
#   select(-n) %>% 
#   left_join(SAU_Revenue_Data) %>% 
#   left_join(Tresholded_T_Data) %>% 
#   left_join(Exploited_Species) %>% 
#   left_join(Top_5_Ecosystem) %>% # includes the groups
#   filter(taxon_name %in% Top_5_Ecosystem$taxon_name) %>% 
#   group_by(fishing_entity,group) %>% 
#   summarise(top_catch = sum(mean_value)/1000000) %>%  # Million
#   group_by(group) %>% 
#   top_n(5)

# If you want to show catch  
Country_Spp_Data <- Tresholded_T_Data %>%
  left_join(Exploited_Species) %>%
  left_join(Top_5_Ecosystem) %>% # includes the groups
  left_join(Clean_sau_eez_fes) %>% 
  left_join(SAU_Sampled_Data,
            by = c("fishing_entity_id","taxon_key")
  ) %>%
  filter(taxon_name %in% Top_5_Ecosystem$taxon_name,
         Variable == "catch") %>%
  group_by(fishing_entity,group) %>%
  summarise(top_catch = sum(Value)) %>%
  group_by(group) %>%
  top_n(5) %>% 
  mutate(
    fishing_entity = ifelse(fishing_entity == "Korea (South)", "S. Korea",
                            ifelse(fishing_entity == "Russian Federation","Russia",
                                   ifelse(fishing_entity == "Papua New Guinea", "N.Guinea",
                                          ifelse(fishing_entity == "Solomon Isl.", "Solomon",fishing_entity)))
    )
  )

```

### OHI-inspired Circular plot Species

```{r species_circluar_plot, eval = F, echo = T}

Median_trans <- median(Species_Data$value)
# hist(Species_Data$value)
# unique(Species_Data$group)

Species_Data <- Raw_Species_Data %>%
  filter(value > 20)

# Set a number of 'empty bar' to add at the end of each group
Char <- as.factor(Species_Data$group)
empty_bar <- 2
to_add <- data.frame( matrix(NA, empty_bar*nlevels(Char), ncol(Species_Data)) )
colnames(to_add) <- colnames(Species_Data)
to_add$group <- rep(levels(Char), each=empty_bar)
Species_Data <- rbind(Species_Data, to_add)
Species_Data <- Species_Data %>% arrange(group,value)
Species_Data$id <- seq(1, nrow(Species_Data))

# Get the name and the y position of each label

# Create labels for all but top 5
label_data <- Species_Data 
number_of_bar <- nrow(label_data)
angle <- 90 - 360 * (label_data$id-0.5) /number_of_bar     # I substract 0.5 because the letter must have the angle of the center of the bars. Not extreme right(1) or extreme left (0)
label_data$hjust <- ifelse( angle < -90, 1, 0)
label_data$angle <- ifelse(angle < -90, angle+180, angle)

# Get the name and the y position of each label
# Top n of each group label
Top_spp <- Species_Data %>% 
  group_by(group) %>% 
  top_n(5,value)

# Create labels for all but top 5
label_data <- Species_Data
number_of_bar <- nrow(label_data)
angle <- 90 - 360 * (label_data$id-0.5) /number_of_bar     # I substract 0.5 because the letter must have the angle of the center of the bars. Not extreme right(1) or extreme left (0)
label_data$hjust <- ifelse( angle < -90, 1, 0)
label_data$angle <- ifelse(angle < -90, angle+180, angle)

# Remove top 5
label_data_rest <- label_data %>% 
  filter(!individual %in% Top_spp$individual)

# Create labels for top 5
Top_label_data <- label_data %>% 
  filter(individual %in% Top_spp$individual)

# prepare a data frame for base lines
base_data <- Species_Data %>% 
  group_by(group) %>% 
  summarize(start=min(id), 
            end=max(id) - empty_bar, 
            median = median(value, na.rm = T)
  ) %>% 
  rowwise() %>% 
  mutate(title=mean(c(start, end)))

# prepare a data frame for grid (scales)
grid_data <- base_data
grid_data$end <- grid_data$end[ c( nrow(grid_data), 1:nrow(grid_data)-1)] + 1
grid_data$start <- grid_data$start - 1
# grid_data <- grid_data[-1,]
grid_data[1,2] <- 5 # Manually include bathydemerssal
grid_data[1,3] <- 1

# Median data
median_data <- grid_data %>% 
  mutate_at(vars(start,end),
            .funs = as.numeric)

Seq_axis_spp <- seq(20,100,20)
Plot_Limit <- max(Species_Data$value,na.rm=T)
Ticks_Color <- "black"

### ------------------- ###
# Make the plot
### ------------------- ###
Spp_circular <- ggplot(Species_Data, 
                       aes(x=as.factor(id), 
                           y=value,
                           fill=group)
) +       # Note that id is a factor. If x is numeric, there is some space between the first bar
  geom_bar(aes(x=as.factor(id),
               y=value,
               fill=group),
           stat="identity",
           alpha=0.5) +
  # Add a val=100/75/50/25 lines. I do it at the beginning to make sur barplots are OVER it.
  geom_segment(data=grid_data, 
               aes(x = end, y = 100, xend = start, yend = 100), 
               colour = Ticks_Color, alpha=1, size=0.3, inherit.aes = FALSE) +
  geom_segment(data=grid_data,
               aes(x = end, y = 80, xend = start, yend = 80),  
               colour = Ticks_Color, alpha=1, size=0.3, inherit.aes = FALSE) +
  geom_segment(data=grid_data, 
               aes(x = end, y = 60, xend = start, yend = 60),  
               colour = Ticks_Color, alpha=1, size=0.3, inherit.aes = FALSE) +
  geom_segment(data=grid_data, 
               aes(x = end, y = 40, xend = start, yend = 40),  
               colour = Ticks_Color, alpha=1, size=0.3, inherit.aes = FALSE) +
  geom_segment(data=grid_data, 
               aes(x = end, y = 20, xend = start, yend = 20), 
               colour = Ticks_Color, alpha=1, size=0.3, inherit.aes = FALSE) +
  # Add median lines for each gropup
  geom_segment(data=median_data, 
               aes(x = end, y = median, xend = start+5, yend = median, colour = group),
               alpha=1, size=0.5 , inherit.aes = FALSE) +
  # Add text showing the value of each line
  annotate("text", 
           x = rep(max(Species_Data$id),5),
           y = Seq_axis_spp,
           label = c(Seq_axis_spp[-5],">100"),
           color=Ticks_Color, 
           size=3, 
           angle=0, 
           fontface="bold",
           hjust=1) +
  geom_bar(aes(x=as.factor(id),
               y=value,
               fill=group),
           stat="identity",
           alpha=0.5) +
  # Species names
  geom_text(data=label_data_rest,
            aes(x=id, y=101,
                label=individual,
                hjust=hjust), 
            color="black",
            fontface="italic",
            alpha=1,
            size=3,
            angle= label_data_rest$angle,
            inherit.aes = FALSE) +
  # Bold top 3 
  geom_text(data=Top_label_data,
            aes(x=id, y=101,
                label=individual,
                hjust=hjust#,
                # color = group
            ),
            color="black",
            fontface="bold.italic",
            alpha=1,
            size=3,
            angle= Top_label_data$angle,
            inherit.aes = FALSE,
            show.legend = FALSE) +
  # Add base line information
  geom_segment(data=base_data, 
               aes(x = start, 
                   y = -5, 
                   xend = end, 
                   yend = -5
               ), 
               colour = "black",
               alpha=0.8, 
               size=0.6 , 
               inherit.aes = FALSE) +
  ylim(-50,110) + # sets the circle the first value the central the last value how wide
  theme_minimal() +
  theme(
    legend.position = c(1,0.11),
    legend.text = element_text(size = 13),
    legend.title = element_text(size = 14, face = "bold"),
    axis.text = element_blank(),
    axis.title = element_blank(),
    panel.grid = element_blank(),
    plot.margin = unit(rep(-3.5,4.5),"cm")
  ) +
  coord_polar() + 
  scale_fill_manual("Ecosystem Preference",
                    values = c(wes_palette(n = 5, name = "Darjeeling1"),
                               wes_palette(n = 3, name = "Darjeeling2")
                    ),
                    na.value = "white"
  ) +
  scale_colour_manual("Ecosystem Preference",
                      values = c(wes_palette(n = 5, name = "Darjeeling1"),
                                 wes_palette(n = 3, name = "Darjeeling2")
                      ),
                      na.value = "white"
  )

#### Ecosystem  preference based plots

Ecosystem_plots <- Country_Spp_Data %>%
  ungroup() %>% 
  mutate(name = reorder_within(fishing_entity,top_catch,group)) %>% # to order the facet wrap
  ggplot() +
  geom_bar(
    aes(
      x = name,
      y = top_catch/1000,
      fill = group
    ), 
    stat = "identity"
  ) +
  labs(x = "",
       y = "Catch (Thousand tonnes)") +
  scale_fill_manual("",
                    values = c(wes_palette(n = 5, name = "Darjeeling1"),
                               wes_palette(n = 3, name = "Darjeeling2")
                    ),
                    na.value = "white"
  ) +
  facet_wrap(~group,
             nrow=2,
             scales = "free") +
  # ggtheme_plot() +
  my_ggtheme_p() +
  scale_x_reordered() +
  theme(legend.position = "",
        axis.text.x = element_text(size = 11,
                                   angle = 0,
                                   face = "plain"),
        strip.background = element_blank(),
        strip.text.x = element_blank()
  )

### Combined plot

spp_plot <- ggdraw() +
  # Circular plot
  draw_plot(Spp_circular, x = 0, y = 0.305, width = 1, height = 0.60) +
  # Environmental maps
  draw_plot(Ecosystem_plots, x = 0.02, y = -0.029, width = 0.95, height = 0.26) +
  # Labeks
  draw_plot_label(label = c("A", "B"), 
                  size = 30,
                  x = c(0, 0), 
                  y = c(1, 0.27)
  )

save_plot(
  paste("spp_circular_catch_country.png",sep=""),
  # "spp_plot.png",
  spp_plot,
  base_height = 14,
  base_width = 14
)

```

### Statistics on Species Ecosystem Preference

```{r Spp_Stats, eval = T, echo = F}

head(Raw_Species_Data)

# Differences between groups
Stat_spp_aov <- aov(value~group, data = Raw_Species_Data)
Stat_spp_aov
summary(Stat_spp_aov)
#              Df Sum Sq Mean Sq F value Pr(>F)    
# group         5  56721   11344   53.82 <2e-16 ***
# Residuals   597 125844     211

TukeyHSD(Stat_spp_aov) %>% 
  tidy() %>% 
  filter(adj.p.value < 0.05) # Pelagic-Oceanic

```

# Sensitivity Analysis

```{r Sensitivity_Area_Treshold, eval = T, echo = F}

# Model Index
MI <- 100 # Model Index

####_____End Select Tresholds_______ ###

Area_Index_Level <- c(seq(.1,0.5,0.1),0.25)
result <- c(40,35,28,18,4,31.5)

plot(result~Area_Index_Level)

p <- list()
i = NULL

for(i in 1:length(Area_Index_Level)){

  data <- Clean_Results_Trans %>% 
    filter(
      area_index >= Area_Index_Level[i],
      area_index <= 1-Area_Index_Level[i],
      model_index == MI
    ) %>% 
    group_by(eez_name) %>% 
    summarise(
      n_trans_spp = length(unique(taxon_key)),
      n_Neighbours = length(unique(eez_neighbour)),
      Trans_Rate = n_trans_spp/n_Neighbours
    )

  Median_data <- median(data$n_trans_spp)
  print(Median_data)

  p[[i]] <- ggplot(data) +
    geom_histogram(aes(n_trans_spp),
                   binwidth = 10, 
                   colour = "white",
                   fill = ifelse(i!= 6,"grey80","grey20")
    ) +
    labs(
      x = "Transboundary species (n)",
      y = "Frequency of EEZs",
      title= paste((Area_Index_Level[i]*100),"% area threshold"," (Median ",Median_data,")",sep="")
    ) +
    # ggtitle(paste("Using a",(Area_Index_Level[i]*100),"% area treshold. (Median",Median_data,")")) +
    ggtheme_plot()
}

ggsave(
  plot = do.call(gridExtra::grid.arrange,p),
  width = 8,
  height = 8,
  units = "in",
  filename = "Area_Thresholds.png",
  path = "/Volumes/DATA/JULIANO_NEYMAR/FishForVisa/Sensitivity_Analysis"
)

```

#Problem solving

### First world Anchovies

In a first run the model suggests anchovies "Engraulis ringens" (600004) is transboundary between US-Mex-Can. I'm exploring this since all four models agree...

Found the problem, Gab's coordinate system, hence the data, are in different system than DBEM and SAU data...

```{r Trans_Spp_Function, eval= F ,echo = T,warning = F,message = F}

### DATA EXTRA NEEDED

# DBEM Coordinate system
Coor <- fread(paste(Data_Path,"Spatial_Data/Lon_Lat_DBEM.txt",sep=""),header = FALSE)
colnames(Coor) <- c("INDEX","Longitude","Latitude")
EEZIDs_List <- read_excel(paste(Data_Path,"Spatial_Data/V_Lam/Updated_EEZList_17June2016.xlsx",sep=""))
EEZ_CellID <- read_excel(paste(Data_Path,"Spatial_Data/V_Lam/EEZ_CellID.xlsx",sep=""))
colnames(EEZ_CellID) <- c("EEZID","INDEX")

CoorG <- read.csv(paste(Data_Path,"Spatial_Data/coordinates_gab.csv",sep="")) %>% 
  mutate(INDEX = seq(1,259200,1))

# colnames(CoorG) <- c("INDEX","Longitude","Latitude")
coords <- CoorG[order(CoorG$Latitude, rev(CoorG$Longitude),decreasing=TRUE), ]

# Explore model step by step

# Get Catch Data
Anchoveta_Dist <- GetSppDist(600004,"All",CoorG)

Anchoveta_Dist_Coor <- Anchoveta_Dist %>% 
  left_join(CoorG)

# Swap coordinate system to match DBEM and SAU
Anchoveta_Dist_Coor_Fix <- Anchoveta_Dist_Coor[order(Anchoveta_Dist_Coor$Latitude, rev(Anchoveta_Dist_Coor$Longitude),decreasing=TRUE), ]

x <- Anchoveta_Dist %>% 
  left_join(EEZ_CellID) %>% 
  left_join(EEZIDs_List) %>% 
  filter(
    # Model == "Occ",
    Value >0)

# head(x)
# Exploring using Vicky's data

ggplot(x) +
  geom_tile(aes(
    x = Longitude,
    y = Latitude,
    fill = Name,
    color = Name
  )
  ) +
  facet_wrap(~Model)

Spp_Trans <- bind_rows(
  mclapply(
    600004, FUN =EstTransIndex, Model = "All", Neighbours = Neighbour_List, Coord = CoorG
  )
)

# Estimate total spp distribution within neighbours

Trans_Spp <- SppDist %>%
  filter(INDEX %in% Neighbours$INDEX) %>% # Filter data only located within EEZs
  mutate(Value = ifelse(Value > 0, 1,0)) %>%  # For now using SAU_C as presence absence
  group_by(TaxonKey,
           INDEX
  ) %>%
  summarise(Model_Index = mean(Value,na.rm=T)) %>%
  filter(Model_Index > 0) %>%
  mutate(Model_Index = Model_Index*100) %>%
  left_join(Neighbours,
            by = "INDEX")

Test <- Trans_Spp %>% 
  filter(Name == "USA (West Coast)" & Model_Index ==100)

# Explore only SAU catch data

# Get Catch Data
Anchoveta_Dist <- GetSppDist(Spp,"SAU_C")

# Get Mex, Can and USA W coast Grid
Anchoveta_Dist_EEZ <- Anchoveta_Dist %>%

  Anchoveta_NA <-  Anchoveta_Dist%>% 
  filter(INDEX %in% North_America$INDEX) %>% 
  gather("Year","Tonnes",CATCH.1:CATCH.65) %>% 
  group_by(Year) %>% 
  summarise(Sum_T = sum(Tonnes, na.rm =T))

```

### Over-counting transboundary stocks

In a first run, Australia results in many transboundary stocks, I believe this is because it is counting the number of countries that that the species share (e.g. one species 6 countries = 6 transboundary spp) rather than 1 species 6 countries = 1 trans spp.

**Problem fixed.** Now we can estimate the number of transboundary species as well as the number of countries/regions that each country shares per species.

```{Australia_case, eval =F, echo = F}

Australia <- Results_Data %>% 
  filter(Name == "Australia")

Original_Australia <-Australia %>% 
  filter(Model_Index >= 50#,
         # Trans_Index >= 0.10
  ) %>% 
  group_by(Name,Model_Index) %>% 
  summarise(n_Trans_Spp = n()) %>% 
  filter(Model_Index == 100)

## Testing the waters

Explore_Australia <- Australia %>% 
  filter(Model_Index >= 50#,
         # Trans_Index >= 0.10
  ) %>% 
  slice(1:20) %>% 
  group_by(Name,Model_Index) %>% 
  summarise(
    n_trans = length(unique(TaxonKey)),
    n_countries = length(unique(Neighbour_Territory)),
  )

```

### Fixing counts within EEZs

In the first run. Countries where EEZs are divided but touching (e.g. South Africa Indian and Atlantic coasts) are double counted.

This might be fixed with Tyler's data of EEZs to ISOs... He did it manually so it does not work!

```{South_Africa_case, eval =F, echo = F}

# Tyler E.'s' data
EEZ_ISO <- read.csv("/Volumes/DATA/JULIANO_NEYMAR/FishForVisa/Spatial_Data/EEZID_ISO_TylEd.csv", comment.char="#")

#SAU / Vicky's data
EEZIDs_List <- read_excel(paste(Data_Path,"Spatial_Data/V_Lam/Updated_EEZList_17June2016.xlsx",sep=""))

EEZ_CellID <- read_excel(paste(Data_Path,"Spatial_Data/V_Lam/EEZ_CellID.xlsx",sep=""))
colnames(EEZ_CellID) <- c("EEZID","INDEX")

# Worlds adata

# The path
path_world <- paste(Data_Path,"Spatial_Data/World_EEZ_v10_2018",sep="")
# #The File
fnam_world <- "eez_v10.shp"
# #Load it!
# 
World_EEZs <- st_read(dsn = path_world,
                      layer =file_path_sans_ext(fnam_world))

head(World_EEZs)

World_EEZs %>% 
  filter(
    # Territory1 == "South Africa",
    ISO_Ter1 == "ZAF") %>% 
  head()

# Get ISO and territory names from world map
ISO_Ter <- World_EEZs %>%
  select(ISO_Ter1,Territory1) %>% 
  as.data.frame() %>% 
  select(ISO_Ter1,Territory1)

head(ISO_Ter)

# I think EEZID equals ISO when not devided...

EZZID_Yes_ISO <- ISO_Ter %>% 
  semi_join(ISO_Ter,
            by ="ISO_Ter1") %>% 
  rename(Name = Territory1) %>% 
  # head() %>% 
  left_join(EEZIDs_List,
            by = "Name")

EEZ_ISO_Clean <- ISO_Ter %>% 
  left_join(EEZ_ISO,
            by = "ISO_Ter1")

SA <- Results_Data %>% 
  filter(stringr::str_detect(Name,"South Africa"))

## Creating a subset dataset so i can work on my laptop and not remoteley

SubsetData <- Clean_Results %>% 
  filter(Country_Territory=="USA (East Coast)")

write.csv(SubsetData,
          "SubsetData.csv",
          row.names = F) # Using USA East coast and Gulf as example

###_____________________________ ###

# Loading subsetted data

Clean_Results <- read.csv("~/GitHub/FishForVisa/Temporal_Data/SubsetData.csv")

sau_eez_fes <- read.csv("~/GitHub/FishForVisa/Temporal_Data/sau_eez_fes.csv")

Neighbourd_Territory_T <- Clean_Results %>% 
  group_by(eez_name=Neighbour_Territory) %>% 
  summarise() %>%
  left_join(sau_eez_fes,
            by = "eez_name") %>% 
  select(Neighbour_Territory= eez_name,
         NT_fishing_entity=fishing_entity
  )

Country_Territory_T <- Clean_Results %>% 
  group_by(eez_name=Country_Territory) %>% 
  summarise() %>%
  left_join(sau_eez_fes,
            by = "eez_name")  %>% 
  select(Country_Territory= eez_name,
         CT_fishing_entity=fishing_entity)

Non_duplicate_Data <- Clean_Results %>%   
  left_join(Neighbourd_Territory_T,
            by ="Neighbour_Territory") %>% 
  left_join(Country_Territory_T,
            by ="Country_Territory") %>% 
  mutate(
    Same_Nation = ifelse(CT_fishing_entity==NT_fishing_entity,"Yes","No")
  ) %>% 
  filter(Same_Nation == "No")

# FIXED!!!!!

```

- **Done debugging**

### Missing stocks

There are places like Peru-Chile or Alaska-Canada that should have trans spp... In Alaska the issue is that Alaska doesn't actually exist as a separate shapefile...

```{r Missing_stocks, eval =F}

Exploited_Species <- fread(paste(Data_Path,"Distribution_Data/exploited_species_list.csv",sep=""))

Clean_sau_eez_fes <- sau_eez_fes %>% 
  gather("Type","FishingRegion",2,4) %>% 
  left_join(sau_eez_fes,
            by = ("fishing_entity_id")
  ) %>% 
  group_by(FishingRegion,fishing_entity) %>% 
  summarise(n())

# The Alaska paradox
# It's a problem of the NAs, fixed.

USA_West <- Results_Data %>% 
  filter(Country_Territory == "USA (West Coast)") %>% 
  left_join(Exploited_Species,
            by = "TaxonKey")

# Northern sea Russia
Russia_map <- World_EEZs %>% 
  filter(stringr::str_detect(Name,"Russia")) %>% 
  st_simplify(preserveTopology = TRUE, dTolerance = 100)

head(Russia_map)

ggplot(Russia_map) +
  geom_sf(aes(fill = Name)) 
# Shapefiles not matching
# Russia (Laptev to Chukchi Sea)
# Russia (Kara Sea)

Russian_Regions <- c("Russia (Laptev to Chukchi Sea)","Russia (Kara Sea)")

Russia_Results <- Results_Data %>% 
  filter(stringr::str_detect(Country_Territory,"Russia"))

unique(Russia_Results$Country_Territory)
# Missing Laptev to Chukchi sea

### What about Vicky's data
#SAU / Vicky's data
EEZIDs_List <- read_excel(paste(Data_Path,"Spatial_Data/V_Lam/Updated_EEZList_17June2016.xlsx",sep=""))

Russia_EEZlist <- EEZIDs_List %>% 
  filter(stringr::str_detect(Name,"Russia"))

unique(Russia_EEZlist$Name)

### So maybe there are no Neighbours?
Neighbour_List <- fread(paste(Data_Path,"Spatial_Data/Neighbour_List.csv", sep =""))

Russia_Neighbour <- Neighbour_List %>% 
  filter(stringr::str_detect(Name,"Russia"))

unique(Russia_Neighbour$Country_Territory)

# OK So Kara Sea has no borders outside Russia, hence white. Now Laptev does have borders but No species... which is weird... Paific Halibut?

# I manually ran the function to see what's going on

# ggplot() +
#   geom_sf(data = Russia_map, aes(colour = Name)) +
#   geom_tile(data = subset(SppDist, Value !=0),
#             aes(
#     x = Longitude,
#     y = Latitude,
#     fill = log10(Value)
#   ) 
#   ) +
#   facet_wrap(~Model)

# The main issue is that there is no catch data there...

```

### Pierrs massive discrete stocks

```{r Pierrs_Discretion, eval = F ,echo = T,warning = F,message = F}

Pierre <-  Clean_Results_Trans %>% 
  filter(Territory == "Saint Pierre & Miquelon (France)") %>% 
  left_join(Exploited_Species,
            by = "TaxonKey")

# Very wrong!
# using 600051 to see what's the problem with the Area index

# Found the problem, Territory and Neighbour were switched un the last step of the EstTransSpp function

```

### Côte de Ivory missing data

For some reason Côte de ivory is missing data. It was fixed at the beginning when I was averaging 1951 to 1961 instead of 2004 to 2014, but now is somehow back to zero.

```{Cote_Deivory, eval =F, echo =F}

# It might be a problem in the name côte de'ivory....
# Nope, both INDEX and neighbours datasets are names alike

# Run getSppDistribution function with this:
Model="All"
Coord=CoorG

# Spp <- 600423 #This species should work

# Its a problem with compabillity between Windows and MAC. Fixed by changing the neighbours csv. to excell

# Now Check Vietnam...Curacao and Réunion...
# In INDEXES "Vietnam" in all others "Viet Nam"
# Fixed directly on the excell

# The russioan probelmo

# Load neighbours data from previouse routine
Neighbours_Data <- read_excel(paste(Data_Path,"Spatial_Data/EEZ_Neighbour_List.xlsx", sep =""))

EEZCellIUD913 <- EEZ_CellID %>% 
  filter(EEZID == 913)

R912 <- Neighbours_Data %>% 
  filter(ID == 912)

R913 <- Neighbours_Data %>% 
  filter(ID == 913)

anti_join(R912,R913)

Index_Code913 <- Index_Code %>% 
  filter(EEZID == 913)

Neighbours_Data %>% 
  filter(INDEX %in% Index_Code913$INDEX)

# SAU relations between INDEX and Country's EEZs
EEZIDs_List <- read_excel(paste(Data_Path,"Spatial_Data/V_Lam/Updated_EEZList_17June2016.xlsx",sep=""))
# EEZIDs_List <- read_excel(paste(Data_Path,"Spatial_Data/V_Lam/Updated_EEZList_JEPA_92019.xlsx",sep=""))
EEZ_CellID <- read_excel(paste(Data_Path,"Spatial_Data/V_Lam/EEZ_CellID.xlsx",sep=""))
# EEZ_CellID <- read_excel(paste(Data_Path,"Spatial_Data/V_Lam/Updated_EEZList_JEPA_92019.xlsx",sep=""))
colnames(EEZ_CellID) <- c("EEZID","INDEX")

Index_Code <- EEZIDs_List %>% 
  left_join(EEZ_CellID) %>% 
  rename(Territory = Name)

# The path
path_world <- paste(Data_Path,"Spatial_Data/SAU_Shapefile/",sep="")
# #The File

fnam_world <- "SAUEEZ_July2015.shp"
World_sf <- st_read(dsn = path_world,
                    layer =file_path_sans_ext(fnam_world)) %>% 
  rename(Territory = Name)

x <- Neighbours_Data %>%
  select(-Neighbour) %>% 
  group_by(Territory) %>% 
  distinct() %>% 
  anti_join(Index_Code) %>% 
  group_by(Territory) %>% 
  summarise(n());x

# World_sf %>% 
#   filter(Territory %in% x$Territory)

# Fuck it! Re-run Neighbours code...

Russians <- c("Russia (Laptev to Chukchi Sea)", #913
              "US Virgin Islands") #912

SF_Names <- 850

Russian_Index <- Index_Code %>% 
  filter(Territory %in% Russians)

unique(Russian_Index$EEZID)
unique(Russian_Index$Territory)

World_sf %>% 
  # filter(Territory == "Russia (Laptev to Chukchi Sea)") # EEZid 912
  # filter(Territory == "Russia (Kara Sea)") # EEZid also 912
  filter(Territory %in% Russians) %>% 
  ggplot() +
  geom_sf(aes(fill = Territory))

unique(Russian_sf$Territory)

# split shapefile by DF
Russian_id <- World_sf %>% 
  filter(Territory %in% Russians) %>% 
  left_join(Russian_Index, by= "Territory") %>%  
  mutate(EEZID = ifelse(Territory == "Russia (Kara Sea)",912,913)) %>% 
  filter(EEZID == 913) %>% 
  st_simplify(preserveTopology = TRUE, dTolerance = 1000)

ggplot(Russian_id) +
  geom_sf(aes())

# ok!

# split shapefile by DF
Russia_id_912_913 <- World_sf %>% 
  filter(Territory %in% Russians) %>% 
  left_join(Russian_Index, by= "Territory") %>%  
  mutate(EEZID = ifelse(Territory == "Russia (Kara Sea)",912,913)) %>% 
  as.data.frame() %>% 
  ungroup() %>% 
  select(Territory,INDEX,EEZID)

write.csv(Russia_id_912_913,
          "Russia_id_912_913.csv",
          row.names = F)

head(x)

```

### Number of Collapsed USA species

```{r Collapsed_USA, eval = F ,echo = T,warning = F,message = F}

#  At the end, the ram legacy and the SAU are tide in results...

USA_Collapsed <- Species_Status %>% 
  filter(str_detect(Territory, pattern="USA")#,
         # Status == "Collapsed"
  ) %>% 
  ungroup() %>% 
  mutate(TaxonKey = as.numeric(TaxonKey)) %>%
  left_join(Exploited_Species)

# Comparring to RAM legacy in the W pacific coast

# Sabelfish match
# Pacifc sardine no data ram
# Chillipepper match
# Lincod no match althou circular path in RAM
# Dover no match by far...
# Chub mackerel... match
# Petral metch
# Black Rockfish No mach...ish
# Widow rockfish no match
# Starry flounder no match
# Rex sole no data RAM
# Pacific ocean percCollapsed Vs Overfished
#

```

### Fixing Russia Moneeeeey

```{r russia}

# Shapefile names
Russia_Sf <- World_land_sf %>% 
  filter(str_detect(fishing_entity,"Russia")) %>% 
  pull(fishing_entity)

"Russia"

Russia_Trans <- Revenue_Country %>% 
  filter(str_detect(fishing_entity,"Russia")) %>% 
  pull(fishing_entity)

"Russian Federation"
### Laptev and Chukchi are the issue, No Alaska'?

### Wake
Wake_ssf <- World_sf %>% 
  filter(str_detect(eez_name,"Wake")) %>% 
  pull(eez_name)

"Wake Isl. (USA)"

Wake_Trans <- Transboundary_Spp %>% 
  filter(str_detect(eez_name,"Wake")) %>% 
  pull(eez_name)

"NA"

Wake_Tresh <- Tresholded_T_Data %>% 
  filter(str_detect(eez_name,"Wake")) %>% 
  pull(eez_name)

"NA"

Wake_Raw <- Clean_Results_Trans %>% 
  filter(str_detect(eez_name,"Wake"))

#### Son todas under 25%.... Por eso

#### Macquarie Island
Macquarie_Raw <- Clean_Results_Trans %>% 
  filter(str_detect(eez_name,"Macquarie"))

# Under 25% Bye

### Minomi

Neighbours_Data %>% 
  filter(str_detect(Territory,"Japan")) %>% 
  pull(Territory) %>% 
  unique()

# Not in our shaopefile

```

### UN Data exploration

```{r}

UN_Regions %>% 
  filter(fishing_entity %in% Clean_sau_eez_fes$fishing_entity) %>% 
  group_by(sub_region) %>% 
  summarise(n_countries = length(unique(fishing_entity))) %>% 
  arrange(n_countries)

World_sf %>% 
  as.data.frame() %>% 
  left_join(Clean_sau_eez_fes) %>% 
  left_join(UN_Regions) %>% 
  filter(fishing_entity %in% Clean_sau_eez_fes$fishing_entity) %>% 
  group_by(sub_region) %>% 
  summarise(Size_EEZ = sum(Area_km2)) %>% 
  arrange(desc(Size_EEZ))

```

# Extra Analysis
## Extra Analysis; Functions

```{r Trans_Spp_Function_All, eval = F ,echo = T,warning = F,message = F}

# Function to determine whether or not a species is transboundary for each model

# Varibales needed
# Spp: Species to be analized, data with presence/absence per gridcell
# Index_EEZ: Reference list of all INDEX that fall within EEZs
# EEZ_Size: Number of grid-cells that each EEZ has
# Neighbours: Reference list of neighbouring countries

# Needs getSppDist

#### For testing 
# Spp <- 600059
# Model <- "All"
# Neighbours <- Neighbour_List
# Coord = CoorG
# __________________________________

# The function
EstTransIndex <- function(Spp,Model,Neighbours,Coord,Save="Y"){

  # Get model data from spps
  SppDist <- GetSppDist(Spp,Model,Coord)

  # Result 1. Number of Countries that share the species

  #____________ ESTIMATING MODEL INDEX (TRESHOLD 1)_________ #
  Trans_Spp <- SppDist %>%
    filter(INDEX %in% Neighbours$INDEX,# Filter data only located within EEZs
           Model != "SAU_C") %>% # Only using observational and modelled data
    mutate(Value = ifelse(Value > 0, 1,0)) %>%
    left_join(EEZ_CellID,
              by = "INDEX") %>% 
    left_join(EEZIDs_List,
              by = "EEZID")

  # MODEL INDEX (TRESHOLDS 1 & 2) #
  Model_Index_D <- Trans_Spp %>% 
    group_by(Name,
             TaxonKey,
             Model
    ) %>% 
    summarise(n_cells_spp = n()) %>% 
    select(-n_cells_spp)

  #____________ ESTIMATING DISTRIBUTION INDEX (TRESHOLD 3)_________ #
  # The number of species' cells present within each country's EEZ

  #Step 1.  Get EEZ id and Neighbour
  Neighbours_EEZ <- Neighbours %>% 
    group_by(EEZID,Country_Territory,Neighbour_Territory) %>% 
    summarise(n=n()) %>% 
    ungroup() %>% 
    select(-n)

  # Step 2. Determines the amount of grids present in each country
  Spp_Grid <- Trans_Spp %>% 
    group_by(TaxonKey,
             Model, 
             EEZID) %>% 
    summarise(n_spp_eez = length(unique(INDEX))) %>% 
    left_join(Neighbours_EEZ,
              by = "EEZID") %>% 
    filter(Country_Territory %in% Trans_Spp$Name, #Filter out unwanted Neighbours (those who don't have grids within but get included because they are Neighbours)
           Neighbour_Territory %in% Trans_Spp$Name)

  # Step 3. Sum total grids per Neighbours

  # Split dataframes to merge latter
  Country_Territory_T <- Spp_Grid %>% 
    ungroup() %>% 
    select(
      Model_Index, # Un-comment after producing models x datasets
      TaxonKey,Name=Country_Territory,n_spp_eez
    )

  Neighbour_Territory_T <- Spp_Grid %>% 
    ungroup() %>% 
    select(
      Model_Index,
      TaxonKey,n_spp_eez,Name=Neighbour_Territory, Country_Territory
    )

  # Merge dataframes to get totals per Neighbourds
  Area_Index_D <- full_join(Country_Territory_T,
                            Neighbour_Territory_T, by = c("Name","TaxonKey","Model")) %>%
    rowwise() %>%
    mutate(Spp_Total = sum(n_spp_eez.x,n_spp_eez.y,na.rm=T)) %>% # Total gridcelles per Neighbours
    distinct() %>% # Removes false duplicates from `full_join()`
    rename(Neighbour_Territory = Name,
           n_spp_Country = n_spp_eez.x,
           n_spp_Neighbour = n_spp_eez.y) %>% 
    mutate(Area_Index = n_spp_Country/Spp_Total) %>%  #Estimates the proportion of grids per country
    select(1:4,6,8)

  #____________ FINAL INDEX TABLE_________ #
  Indexes <- Model_Index_D %>% 
    rename(Country_Territory = Name) %>% 
    full_join(Area_Index_D,
              by = c("TaxonKey","Country_Territory","Model")
    ) %>% 
    filter(!is.na(Area_Index))%>%  # Remove cases where the country does not pass the area_index and viceversa
    select(Model,TaxonKey,Country_Territory,Neighbour_Territory,everything())

  ### Save spp dataframe

  if(nrow(Indexes) > 0 & Save == "Y"){

    File_Name <- paste(Spp,"_Transboundary.csv",sep = "")
    Save_Path <- paste(Results_Path,"All_Models/",File_Name,sep="")

    write.csv(Indexes,
              Save_Path,
              row.names = F)

  }

  if(nrow(Indexes) == 0 & Save == "Y"){

    # Reads dataset of Non transboundary species
    Non_Transboundary_Data <- fread(paste(Results_Path,"All_Models/","Non_Transboundary_Data.csv",sep=""))

    # Get current Spp that is not transboundary 
    Current_Spp <- tibble(Non_trans = Spp)

    # Combine new spp with others
    New_Non <- Non_Transboundary_Data %>% 
      bind_rows(Current_Spp)

    # Set a name for file and saves data
    New_Non_Name <- paste(Results_Path,"All_Models/","Non_Transboundary_Data.csv",sep="")
    write.csv(New_Non, New_Non_Name, row.names = F)

  }

  ### Function return

  EstTransIndex=Indexes # A dataframe with

}

### Testing Function ####
# Spp <- 600006
# Model <- "All"
# Neighbours <- Neighbour_List
# 
# Tryout <- EstTransIndex(Spp,Model,Neighbours,CoorG,Save="N")
# 
# Over_20 <-Tryout %>% 
#     filter(Area_Index > 0.20,
#            Area_Index < 0.80)

# Spp <- NULL
# Model <- NULL

```

## Extra Analysis; Results

### Maps for each model/data-set
```{Tran_country_Map_All, eval =F, echo =F}

World_EEZs_simple <- World_EEZs %>%
  st_transform(crs = 4326) %>% # 4326
  st_simplify(preserveTopology = TRUE, dTolerance = 0.1) %>%
  left_join(TenTwenyfive, by ="Name") #%>%
# filter(!is.na(n_trans)) #%>%
# st_transform(crs = 3832)

# head(World_EEZs_simple)
# World_EEZs_simple %>% 
#   filter(Name == "Chile")

pal <- wes_palette("Zissou1", 100, type = "continuous")

Mod <- unique(TenTwenyfive$Model)

# Loop through all variables
for(v in 1:length(Mod)){

  Model_List <- Mod[v]

  plot <- ggplot(data = subset(World_EEZs_simple,Model %in% Model_List)) +
    geom_sf(aes(fill = n_trans)) +
    scale_fill_gradientn(colours = pal,
                         na.value = "white") +
    ggtheme_map() +
    ggtitle(paste(Model_List))

  # Name plot
  Plot_Name <- paste(Model_List,"plot.png",sep="_")

  # Save plot
  ggsave(
    plot = plot,
    width = 14,
    height = 10,
    units = "in",
    filename = Plot_Name,
    path = "/Volumes/DATA/JULIANO_NEYMAR/FishForVisa/Figures/All_Models/"
  )

}

```

```{r Continent_PNG, eval =F }

names(UN_Regions)
# Load Land shapefile
path_world_land <- paste(Data_Path,"Spatial_Data/ne_50m_admin_0_countries",sep="")
# #The File
fnam_world_land <- "ne_50m_admin_0_countries.shp"

World_land_sf <- st_read(dsn = path_world_land,
                         layer =file_path_sans_ext(fnam_world_land)) %>% 
  st_transform(crs = 4326)

head(World_land_sf)

Regions <- unique(World_land_sf$REGION_UN)
# r= 1

for(r in 1:length(Regions)){

  Oceania <- World_land_sf %>%
    st_transform(crs = 3832)

  Oceania %>% 
    filter(REGION_UN %in% Regions[5]) %>% 
    ggplot() +
    geom_sf(aes(),fill="black",colour = "black") +
    ggtheme_map() #+
  # coord_sf(xlim = c(-180, 90), ylim = c(-90, 90), expand = FALSE)

  Plot_Name <-paste(Regions[5],".png",sep="")

  ggsave(Plot_Name,
         plot = last_plot(),
         width = 4,
         height = 4,
         units = "in",
         path = "/Volumes/DATA/JULIANO_NEYMAR/FishForVisa/Images_Data/"
  )

}

```

```{r round_map, eval = F, echo = F}
# #### Round map tryout

# ### EEZ shapefile
# eez  = readRDS(paste(Data_Path,"Spatial_Data/eez_simplified.rds",sep="")) #taken from cotterel https://github.com/cottrellr/shocks
# 
# eez = spTransform(eez, CRS("+proj=longlat"))
# eez = spTransform(eez, CRS("+proj=wintri"))
# eez = fortify(eez) 
# 
# World_sf_si <- World_sf %>%
#   st_transform(crs = 4326
#

# mapworld = fortify(spTransform(rworldmap::getMap(), CRS("+proj=wintri")))
# 
# mapworld<- mapworld %>% 
#   rename(Territory = id) %>% 
#   left_join(Data,
#             by = "Territory") %>% 
#   left_join(Data_EEZ,
#             by = "Territory") %>% 
#   rename(id = Territory)
# 
# # head(mapworld)
# 
# #create map boundary
# ll.to.wt <- function (points)
#   as.data.frame(spTransform(SpatialPoints(points, CRS("+proj=longlat")),
#                             CRS("+proj=wintri")))
#  
# lseq = seq(-85, 85, by=.25)
# boundary <- ll.to.wt(data.frame(
#   long = c(rep(-180, length(lseq)), rep(180, length(lseq)), -180),
#   lat  = c(lseq,                    rev(lseq),          lseq[1])))
#  
# # New facet label names for dose variable
# ggplot()+
#   geom_polygon(data=boundary, aes(x=long, y=lat), fill="azure") +
#   geom_map(data=mapworld, map=mapworld, aes(map_id=id, x=long, y=lat), fill ="gray93", colour=NA) +
#   geom_polygon(data = mapworld, 
#                aes(
#                  x = long,
#                  y = lat,
#                  group = group,
#                  fill = BinsLan
#                ),
#                colour="grey") +
#   geom_polygon(data = mapworld, 
#                aes(
#                  x = long,
#                  y = lat,
#                  group = group,
#                  fill = Bins)
#                ) +
#   geom_path(data=eez, aes(x=long,y=lat, group=group),colour="grey") +
# geom_point(data=ui, aes(x=long, y=lat, size=M, color=M), alpha=0.9) +
# geom_text_repel(data=ui,aes(x=long, y=lat, label=un.name), size=2.5, color = 'black',
#                 box.padding = unit(0.5, "lines"),
#                 point.padding = unit(0.5, "lines"),
#                 segment.color = 'grey50') +
# scale_radius(range = c(1, 2)) #+
# scale_colour_distiller(palette ="Spectral", direction = 1, limits=c(0,1), breaks=c(c(0,1))) + # or direction=1
# theme_map()+
#   theme(legend.key = element_blank(),
#         strip.background = element_blank(),
#         legend.position = "none",
#         text = element_text(size=12,face="bold"))+
#   facet_wrap(~variable, scales = "free", ncol=1)+scale_shape(guide=FALSE)+scale_size(guide=FALSE)
# y=tag_facet(q, tag_pool = c("d","e","f"))

```

# Vulnerability Assessment

```{r Vulnerabillity, eval = T, echo = F}

trans_vul <- Tresholded_T_Data %>% 
  group_by(taxon_key) %>% 
  summarise(
    n_countries = length(unique(eez_name))
  ) %>% 
  left_join(cc_vul) %>% 
  left_join(fish_vul) %>% 
  ggplot() +
  geom_point(
    aes(
      x = fishvul,
      y = n_countries
    )
  )

# climate change vulnerabillity?
modcc <- lm(n_countries~ccvul, data = trans_vul)
modcc
summary(modcc)

# fishing vulnerabillity?
modfis <- lm(n_countries~fishvul, trans_vul)
modfis
summary(modfis)

# Together?

modccfis <- lm(n_countries~fishvul + ccvul, trans_vul)
modccfis
summary(modccfis)

## Using a MANOVA to test the significant effect of fishvul and ccvul to catchstatus

# Prepare data

Status_Per_Species <- Tresholded_T_Data %>% 
  left_join(Status_Data, 
            by =c("eez_name","taxon_key")
  ) %>%
  mutate(status = ifelse(is.na(status),"No Status",status)) %>% 
  group_by(taxon_key,status) %>% 
  summarise(value = length(unique(fishing_entity))) %>% 
  mutate(Categories = ifelse(status %in% c("Developing","Rebuilding"),"Increasing",
                             ifelse(status == "Max Exploited","Constant",
                                    ifelse(status %in% c("Over Exploited","Collapsed","Category C"),"Decreasing", "No Category")
                             )
  )
  ) %>% 
  left_join(cc_vul) %>% 
  left_join(fish_vul) %>% 
  filter(status != "No Category",
         Categories != "No Category")

# Test for assumptions
# - Data Normally distributed (YES)

qqnorm(Status_Per_Species$ccvul) # yes
qqnorm(Status_Per_Species$fishvul) # yes

ggplot(Status_Per_Species) +
  geom_histogram(
    aes(
      # x = fishvul # Yes
      # x = ccvul #yes
    )
  ) +
  facet_wrap(~status)

# -  n for each group > total of dependent variables (YES)
# - All groups have common variance-covariance – usually fine if sample sizes for each group are the same or very similar. Otherwise, Box’s M can be used to test. (YES, SAMPLE SIZES ARE THE SAME)
# - Homoscedasticity
# - There is a moderate linear correlation between the dependent variables (if too weak, or too strong, a series of one-way ANOVAs may be more appropriate) (NO!)

man_vul <- manova(cbind(ccvul,fishvul)~status,data = Status_Per_Species)
summary(man_vul)

summary.aov(man_vul)

## Since we don't have linear correlation between dependent variables, we test anovas

### fishvul ANOVA

Fishvul_aov <- aov(fishvul~Categories, data = Status_Per_Species)
summary(Fishvul_aov)

#   Df Sum Sq Mean Sq F value Pr(>F)
# Categories     2   2698  1349.2   2.286  0.102 # NO SIG
# Residuals   1392 821450   590.1               
# 294 observations deleted due to missingness

### ccvul ANOVA (N)

Ccvul_aov <- aov(ccvul~Categories, data = Status_Per_Species)
summary(Ccvul_aov)
# Df Sum Sq Mean Sq F value Pr(>F)
# Categories     2   1798   899.0   2.181  0.113 # NO SIG
# Residuals   1686 695057   412.3

```

# Old code
```{r OLD_funEstNeighbours, eval = F, echo=T}
# Determine which countries are Neighbours

# Method adopted from Adrew Heiss
# https://gist.github.com/andrewheiss/926b9d60a26e29f6bf32

EstNeighbours <- function(Shapefile){

  # Get the list using the poly2nb function

  Neighbour_List <- poly2nb(Shapefile)

  # Convert te nb data to a Matrix of crossing references
  Neighbour_Matrix <- nb2mat(Neighbour_List,
                             style="B", 
                             zero.policy=TRUE)

  # Rename the matrix axis
  colnames(Neighbour_Matrix) <- rownames(Neighbour_Matrix)

  # Get the names of the countries for identifying them
  Country_Names <- tibble(id = row.names(Shapefile@data),
                          Country_Territory = as.character(Shapefile@data$Name),
                          Neighbour_Territory = Country_Territory,
                          EEZID = Shapefile@data$EEZID # specific of SAU sf
  )

  # Clean up and transform the Neighbour matrix
  Neighbours <- as.data.frame(Neighbour_Matrix) %>%
    mutate(country = row.names(.)) %>%  # Convert row names to actual column
    gather(Neighbour, present, -country) %>%  # Convert to long
    filter(present == 1) %>%  # Only look at cells with a match
    # Add country names
    left_join(select(Country_Names,
                     -Neighbour_Territory), by=c("country" = "id")) %>%
    left_join(select(Country_Names, 
                     -Country_Territory), by=c("Neighbour" = "id")) %>% 
    select(EEZID = EEZID.x,Country_Territory,Neighbour_Territory) # select final column

  # Merge them with SAU data

  DBEM_Neighbours_Data <- EEZIDs_List %>% 
    left_join(EEZ_CellID, 
              by = "EEZID") %>% # Asign EEZ id to each country
    left_join(Neighbours, 
              by = "EEZID") %>% # Include all Neighbouring countries 
    filter(Neighbour_Territory != "NA") %>% 
    left_join(Coor,
              by = "INDEX") # Inclu

  return(DBEM_Neighbours_Data)

}

### Test function

# Subset data for testing function
# SAU_Regions <- as.data.frame(unique(World_EEZs$Name))

# Test_Countries <-c("Mexico (Atlantic)","Mexico (Pacific)","Guatemala (Caribbean)","Guatemala (Pacific)","Belize")
# Test <- World_EEZs[World_EEZs@data$Name %in% Test_Countries, ]

# Running function
# getNeighbours(Test)

#### Works beautifully!

```

```{r Trans_Spp_Function, eval= T ,echo = T,warning = F,message = F}

# Function to determine whether or not a species is transboundary

# Varibales needed
# Spp: Species to be analized, data with presence/absence per gridcell
# Index_EEZ: Reference list of all INDEX that fall within EEZs
# EEZ_Size: Number of grid-cells that each EEZ has
# Neighbours: Reference list of neighbouring countries

# Needs getSppDist

#### For testing 
# Spp <- 600059
#Model <- "All"
#Neighbours <- Neighbour_List
#Coord = CoorG
# __________________________________

# The function
EstTransIndex <- function(Spp,Model,Neighbours,Coord,Save="Y"){

  # Get model data from spps
  SppDist <- GetSppDist(Spp,Model,Coord)

  # Result 1. Number of Countries that share the species

  #____________ ESTIMATING MODEL INDEX (TRESHOLD 1)_________ #
  Trans_Spp <- SppDist %>%
    filter(INDEX %in% Neighbours$INDEX,# Filter data only located within EEZs
           Model != "SAU_C") %>% # Only using observational and modelled data
    mutate(Value = ifelse(Value > 0, 1,0)) %>%  
    filter(Value > 0) %>%
    group_by(TaxonKey,
             INDEX
    ) %>%
    summarise(Model_Index = mean(Value,na.rm=T)
    ) %>%
    filter(Model_Index > 0) %>%
    # ____________ ESTIMATING FUNDAMENTAL NICHE (TRESHOLD 2)_________ #
    left_join(SppDist,
              by = c("TaxonKey","INDEX")) %>%
    filter(Model == "SAU_C", # Only keeping cells where SAU catch exists
           Value > 0) %>%
    mutate(Model_Index = Model_Index*100) %>%
    select(-Model, -Value, -Latitude,-Longitude) %>%
    left_join(EEZ_CellID,
              by = "INDEX") %>% 
    left_join(EEZIDs_List,
              by = "EEZID")

  # MODEL INDEX (TRESHOLDS 1 & 2) #
  Model_Index_D <- Trans_Spp %>% 
    group_by(Name,
             TaxonKey,
             Model_Index
    ) %>% 
    summarise(n_cells_spp = n()) %>% 
    select(-n_cells_spp)

  #____________ ESTIMATING DISTRIBUTION INDEX (TRESHOLD 3)_________ #
  # The number of species' cells present within each country's EEZ

  #Step 1.  Get EEZ id and Neighbour
  Neighbours_EEZ <- Neighbours %>% 
    group_by(EEZID,Country_Territory,Neighbour_Territory) %>% 
    summarise(n=n()) %>% 
    ungroup() %>% 
    select(-n)

  # Step 2. Determines the amount of grids present in each country
  Spp_Grid <- Trans_Spp %>% 
    group_by(TaxonKey,
             Model_Index, # Un-comment after producing models x datasets
             EEZID) %>% 
    summarise(n_spp_eez = length(unique(INDEX))) %>% 
    left_join(Neighbours_EEZ,
              by = "EEZID") %>% 
    filter(Country_Territory %in% Trans_Spp$Name, #Filter out unwanted Neighbours (those who don't have grids within but get included because they are Neighbours)
           Neighbour_Territory %in% Trans_Spp$Name)

  # Step 3. Sum total grids per Neighbours

  # Split dataframes to merge latter
  Country_Territory_T <- Spp_Grid %>% 
    ungroup() %>% 
    select(
      Model_Index, # Un-comment after producing models x datasets
      TaxonKey,
      Name=Country_Territory,
      n_spp_eez
    )

  Neighbour_Territory_T <- Spp_Grid %>% 
    ungroup() %>% 
    select(
      Model_Index,
      TaxonKey,n_spp_eez,
      Name=Neighbour_Territory, 
      Country_Territory
    )

  # Merge dataframes to get totals per Neighbourds
  Area_Index_D <- full_join(Country_Territory_T,
                            Neighbour_Territory_T, 
                            by = c("Model_Index","Name","TaxonKey")
  ) %>%
    rowwise() %>%
    mutate(Spp_Total = sum(n_spp_eez.x,n_spp_eez.y,na.rm=T)) %>% # Total gridcelles per Neighbours
    distinct() %>% # Removes false duplicates from `full_join()`
    rename(Neighbour_Territory = Name,
           n_spp_Country = n_spp_eez.x,
           n_spp_Neighbour = n_spp_eez.y) %>% 
    mutate(Area_Index = n_spp_Country/Spp_Total) %>%  #Estimates the proportion of grids per country
    select(1:3,6,8)

  #____________ FINAL INDEX TABLE_________ #
  Indexes <- Model_Index_D %>% 
    rename(Country_Territory = Name) %>% 
    full_join(Area_Index_D,
              by = c("TaxonKey","Country_Territory","Model_Index")
    ) %>% 
    filter(!is.na(Area_Index),
           !is.na(Model_Index) # Uncomment after producing Models x datasets
    )%>%  # Remove cases where the country does not pass the area_index and viceversa
    select(TaxonKey,Country_Territory,Neighbour_Territory,everything())

  ### Save spp dataframe

  if(nrow(Indexes) > 0 & Save == "Y"){

    File_Name <- paste(Spp,"_Transboundary.csv",sep = "")
    Save_Path <- paste(Results_Path,File_Name,sep="")

    write.csv(Indexes,
              Save_Path,
              row.names = F)

  }

  if(nrow(Indexes) == 0 & Save == "Y"){

    # Reads dataset of Non transboundary species
    Non_Transboundary_Data <- fread(paste(Results_Path,"Non_Transboundary_Data.csv",sep=""))

    # Get current Spp that is not transboundary 
    Current_Spp <- tibble(TaxonKey = Spp)

    # Combine new spp with others
    New_Non <- Non_Transboundary_Data %>% 
      bind_rows(Current_Spp)

    # Set a name for file and saves data
    New_Non_Name <- paste(Results_Path,"Non_Transboundary_Data.csv",sep="")
    write.csv(New_Non, New_Non_Name, row.names = F)

  }

  ### Function return

  EstTransIndex=Indexes # A dataframe with

}
```

```{habitat_barplot, eval =F, echo =F}

Species_data <- Species_Trans %>% 
  mutate(scientific_name = paste(Genus,Species)) %>%
  left_join(Species_environ_list) %>% 
  mutate(DemersPelag = replace_na(DemersPelag, "Other")) %>% 
  mutate(Distance = ifelse(DemersPelag == "reef-associated", 1,
                           ifelse(DemersPelag =="pelagic-neritic",2,
                                  ifelse(DemersPelag == "demersal",3,
                                         ifelse(DemersPelag == "pelagic-oceanic",4,
                                                ifelse(DemersPelag =="benthopelagic",5,
                                                       ifelse(DemersPelag =="bathypelagic",6,
                                                              ifelse(DemersPelag =="bathydemersal",7,
                                                                     8)))))))
  ) %>% 
  group_by(DemersPelag,Distance) %>% 
  summarise(Countries_Genus = sum(n_countries,na.rm=T),
            Mean_Countries_Genus = mean(n_countries,na.rm=T),
            sd_Countries_Genus = sd(n_countries,na.rm=T),
            N_spp =length(unique(TaxonName))
  )

ggplot(Species_data,
       aes(
         x = reorder(DemersPelag,desc(Distance)),
         y = Mean_Countries_Genus,
         fill = Countries_Genus#,
         # colour = Countries_Genus
       )
) +
  geom_bar(stat = "identity") +
  # geom_errorbar(data = Species_data,
  #               aes(
  #                 x = reorder(DemersPelag,desc(Distance)),
  #                 ymin=Mean_Countries_Genus-sd_Countries_Genus, # mean +- 2*sd
  #                 ymax=Mean_Countries_Genus+sd_Countries_Genus
  #               ),
  #               colour = "black",
  #               width=.2,
  #               linetype = 2,
  #               size = 0.6,
  #               alpha = 0.5
# ) +
geom_text(data = Species_data, 
          aes(
            x = reorder(DemersPelag,desc(Distance)),
            y = 2,
            label = paste(as.character(N_spp), "species"),
          ),
          colour = "black"
) +
  coord_flip() +
  scale_fill_gradientn("Total number\nof countries",
                       colours = wes_palette("Zissou1"),
                       limits = c(0,1500),
                       breaks = seq(0,1500,250)
  ) +
  scale_colour_gradientn("Total number\nof countries",
                         colours = wes_palette("Zissou1"),
                         limits = c(0,1500),
                         breaks = seq(0,1500,250)
  ) +
  scale_y_continuous(breaks=seq(0, 40, 5), 
                     limits=c(0, 40)
  )+
  labs(
    x = "Habitat type",
    y = "Mean number of countries per habitat"
  ) +
  ggtheme_plot()

```

```{r Territories_Circluar_Plot, eval = T, echo = T}

# -------------------------- #
# Data manipulation for plot
# -------------------------- #

# Set a number of 'empty bar' to add at the end of each group
Char <- as.factor(Countries_Data_Plot$group) # Set groups to factors
empty_bar <- length(unique(Countries_Data_Plot$group))
to_add <- data.frame( matrix(NA, empty_bar*nlevels(Char), ncol(Countries_Data_Plot)) )
colnames(to_add) <- colnames(Countries_Data_Plot)
to_add$group <- rep(levels(Char), each=empty_bar)
Countries_Data_Plot <- bind_rows(Countries_Data_Plot, to_add)
Countries_Data_Plot <- Countries_Data_Plot %>% arrange(group,value)
Countries_Data_Plot$id <- seq(1, nrow(Countries_Data_Plot))

# Get the name and the y position of each label

# Top n of each group label
Top_Countries <- Countries_Data_Plot %>% 
  group_by(group) %>% 
  top_n(5,value)

# Create labels for all but top 5
label_data <- Countries_Data_Plot 
number_of_bar <- nrow(label_data)
angle <- 90 - 360 * (label_data$id-0.5) /number_of_bar     # I substract 0.5 because the letter must have the angle of the center of the bars. Not extreme right(1) or extreme left (0)
label_data$hjust <- ifelse( angle < -90, 1, 0)
label_data$angle <- ifelse(angle < -90, angle+180, angle)

# Remove top 5
label_data_rest <- label_data %>% 
  filter(!individual %in% Top_Countries$individual)

# Create labels for top 5
Top_label_data <- label_data %>% 
  filter(individual %in% Top_Countries$individual)

# prepare a data frame for base lines
base_data <- Countries_Data_Plot %>% 
  group_by(group) %>% 
  summarize(start=min(id), 
            end=max(id) - empty_bar, 
            median = median(value, na.rm = T)
  ) %>% 
  rowwise() %>% 
  mutate(title=mean(c(start, end)))

# prepare a data frame for grid (scales)
grid_data <- base_data
grid_data$end <- grid_data$end[ c( nrow(grid_data), 1:nrow(grid_data)-1)] + 1
grid_data$start <- grid_data$start - 1
grid_data[1,2] <- 5 # Manually include Africa for the median
grid_data[1,3] <- 1

# Median data
median_data <- grid_data %>% 
  mutate_at(vars(start,end),
            .funs = as.numeric)

# -------------------------- #
# Get id to Discrete data
# -------------------------- #

Discrete_Data_Plot <- Discrete_Data_Plot %>% 
  left_join(Countries_Data_Plot,
            by =c("individual","group")
  ) %>%
  mutate(value = ifelse(value.x <= -100,-100,value.x)) %>%
  select(individual,group,value,id) %>%
  arrange(group,value)

### make sure they have all same countries

Discrete_Data_Plot %>% 
  anti_join(Countries_Data_Plot,
            by = c("individual","group","id"))

# -------------------------- #
# Make the plot
# -------------------------- #

# Global Variables
Seq_axis <- seq(-100,100,20)
Plot_Limit <- max(Countries_Data_Plot$value,na.rm=T)
Plot_Limit_Low <- -110
Ticks_Color <- "grey50"

# Base plot
ggplot(Countries_Data_Plot,
       aes(x=as.factor(id), 
           y=value,
           fill=group)
) +       # Note that id is a factor. If x is numeric, there is some space between the first bar
  geom_bar(aes(x=as.factor(id),
               y=value,
               fill=group),
           stat="identity",
           alpha=0.5) +
  geom_bar(data = Discrete_Data_Plot,
           aes(x=as.factor(id),
               y=value),
           stat="identity",
           alpha=0.5) +
  # Add a val=100/75/50/25 lines. I do it at the beginning to make sur barplots are OVER it.
  geom_segment(data=grid_data, aes(x = end, y = 100, xend = start, yend = 100), colour = Ticks_Color, alpha=1, size=0.3 , inherit.aes = FALSE ) +
  geom_segment(data=grid_data, aes(x = end, y = 80, xend = start, yend = 80), colour = Ticks_Color, alpha=1, size=0.3 , inherit.aes = FALSE ) +
  geom_segment(data=grid_data, aes(x = end, y = 60, xend = start, yend = 60), colour = Ticks_Color, alpha=1, size=0.3 , inherit.aes = FALSE ) +
  geom_segment(data=grid_data, aes(x = end, y = 40, xend = start, yend = 40), colour = Ticks_Color, alpha=1, size=0.3 , inherit.aes = FALSE ) +
  geom_segment(data=grid_data, aes(x = end, y = 20, xend = start, yend = 20), colour = Ticks_Color, alpha=1, size=0.3 , inherit.aes = FALSE ) +
  geom_segment(data=grid_data, aes(x = end, y = 0, xend = start, yend = 0), colour = "black", alpha=0, size=0.3 , inherit.aes = FALSE ) + # Cero
  geom_segment(data=grid_data, aes(x = end, y = -100, xend = start, yend = -100), colour = Ticks_Color, alpha=1, size=0.3 , inherit.aes = FALSE ) +
  geom_segment(data=grid_data, aes(x = end, y = -80, xend = start, yend = -80), colour = Ticks_Color, alpha=1, size=0.3 , inherit.aes = FALSE ) +
  geom_segment(data=grid_data, aes(x = end, y = -60, xend = start, yend = -60), colour = Ticks_Color, alpha=1, size=0.3 , inherit.aes = FALSE ) +
  geom_segment(data=grid_data, aes(x = end, y = -40, xend = start, yend = -40), colour = Ticks_Color, alpha=1, size=0.3 , inherit.aes = FALSE ) +
  geom_segment(data=grid_data, aes(x = end, y = -20, xend = start, yend = -20), colour = Ticks_Color, alpha=1, size=0.3 , inherit.aes = FALSE ) +
  # Add median lines for each gropup
  geom_segment(data=median_data,
               aes(x = end,
                   y = median,
                   xend = start+5, 
                   yend = median,
                   colour = group),
               alpha=1,
               size=0.5 , 
               inherit.aes = FALSE ) +
  # geom_hline(yintercept = Median_trans, colour = "black", alpha=1, size=0.3,linetype="dashed") + # All groups median
  # Add text showing the value of each line
  annotate("text", label = "Transboundary", x = 214, y = 45, size = 5, colour = "grey10", angle = 90) +
  annotate("text", label = "Discrete", x = 210, y = -45, size = 5, colour = "grey10", angle = 90) +
  annotate("text", 
           x = rep(max(Countries_Data_Plot$id),length(Seq_axis)),
           y = Seq_axis,
           label = c("<-100",as.character(Seq_axis)[2:(length(Seq_axis)-1)],">100"),
           color = Ticks_Color, 
           size= 4, 
           angle = 0, 
           # fontface="bold",
           hjust=1) +
  geom_bar(aes(x=as.factor(id),
               y=value,
               fill=group),
           stat="identity",
           alpha=0.5) +
  #Country labels
  geom_text(data=label_data_rest,
            aes(x=id, y=Plot_Limit,
                label=individual,
                hjust=hjust), 
            color="black",
            # fontface="bold",
            alpha=1,
            size=2.5,
            angle= label_data_rest$angle,
            inherit.aes = FALSE) +
  # Bold top 3 
  geom_text(data=Top_label_data,
            aes(x=id, y=Plot_Limit,
                label=individual,
                hjust=hjust), 
            color="black",
            fontface="bold",
            alpha=1,
            size=2.5,
            angle= Top_label_data$angle,
            inherit.aes = FALSE) +
  # Add base line information
  geom_segment(data=base_data, 
               aes(x = start, 
                   y = 0, 
                   xend = end, 
                   yend = 0
               ), 
               colour = "black",
               alpha=0.8, 
               size=0.6 , 
               inherit.aes = FALSE) +
  ylim(Plot_Limit_Low,Plot_Limit) + # sets the circle the first value the central the last value how wide
  theme_minimal() +
  theme(
    # legend.position = c(0.5,0.5),
    legend.position = "",
    axis.text = element_blank(),
    axis.title = element_blank(),
    panel.grid = element_blank(),
    plot.margin = unit(rep(-1,4), "cm") 
  ) +
  coord_polar() + 
  scale_fill_manual("Continent",
                    values = c(wes_palette(n = 5, name = "Darjeeling1"),
                               wes_palette(n = 4, name = "Darjeeling2")
                    ),
                    na.value = "white"
  ) +
  scale_colour_manual("Continent",
                      values = c(wes_palette(n = 5, name = "Darjeeling1"),
                                 wes_palette(n = 4, name = "Darjeeling2")
                      ),
                      na.value = "white")

# Save plot

# ggsave(
#     plot = last_plot(),
#     width = 12,
#     height = 12,
#     units = "in",
#     filename = "Territory_circular_both.png",
#     path = "/Volumes/DATA/JULIANO_NEYMAR/FishForVisa/Figures/"
#   )

```

```{r Top_country_Circle_Diagram, eval =F, echo = F}

# Top 10 species

# Number of countries shared by each Species
Top_10_Countries <- Clean_Results_Trans %>% 
  filter(
    Area_Index < Min_AI |
      Area_Index > Max_AI,
    Model_Index == MI,
    Same_Nation == "No"
  ) %>% 
  group_by(Country_Territory) %>% 
  summarise(
    n_countries = length(unique(TaxonKey))
  ) %>% 
  top_n(10,n_countries)

# Countries that share the top 10
Top_Ten_Countries <- Clean_Results_Trans %>% 
  filter(
    Area_Index < Min_AI |
      Area_Index > Max_AI,
    Model_Index == MI,
    Same_Nation == "No",
    Country_Territory %in% Top_10_Countries$Country_Territory
  ) %>% 
  group_by(from=CT_fishing_entity,
           to=NT_fishing_entity
  ) %>% 
  summarise(value=n()
  ) %>% 
  ungroup() %>% 
  mutate(from = paste("F", from),
         to = paste("T", to)
  )

####________________###
# Circulize package ####
# https://jokergoo.github.io/circlize_book/book/
####________________###

# To set links collor 
Bins <- as.numeric(cut(Top_Ten_Countries$value,breaks = 5))

# Basic plot
circos.clear()
chordDiagram(Top_Ten_Countries, 
             annotationTrack = c("grid", "axis"), # keep grid and axis but not labels
             annotationTrackHeight = c(0.001, 0.05), # set links close to the axis
             directional = 1, # direction of arrows
             direction.type = c("diffHeight", "arrows"), # use arrows
             grid.col = "white",
             col = ifelse(Bins == 1,"#046C9A", # deep blue
                          ifelse(Bins == 2,"#78B7C5", #light blue
                                 ifelse(Bins == 3,"#EBCC2A", #yellow
                                        ifelse(Bins == 4, "#E1AF00", #cream
                                               "#F21A00")))), #red
             link.arr.col = wes_palette("Zissou1"),
             link.arr.type = "big.arrow",# make links big arrows
             diffHeight = uh(1, "mm"), # brings the arrow to the bottom of the source
             # annotationTrack = "grid", # no labels
             scale = FALSE, # fraction of total instead of absolute value
             preAllocateTracks = list(track.height = max(strwidth(unlist(dimnames(Top_Ten_Countries))))) # I have no idea but helps with the labels
) 
#  Rest of eddits
circos.track(track.index = 1, panel.fun = function(x, y) {
  xlim = get.cell.meta.data("xlim")
  xplot = get.cell.meta.data("xplot")
  ylim = get.cell.meta.data("ylim")
  sector.name = get.cell.meta.data("sector.index")

  circos.text(mean(xlim),
              ylim[1], 
              sector.name, 
              facing = "clockwise",
              niceFacing = TRUE,
              adj = c(-0.2, 0.5),
              col = "black")

}, bg.border = NA)

```

```{r Top_spp_Circle_Diagram, eval =F, echo = F}

# Top 5 shared species per marine environment (number of countries sharing the species)
Top_Species <- Species_Trans %>% 
  mutate(scientific_name = paste(Genus,Species)) %>%
  left_join(Species_environ_list) %>% 
  mutate(DemersPelag = replace_na(DemersPelag, "Other")) %>% 
  # group_by(TaxonKey,TaxonName,CommonName,DemersPelag) %>% 
  # summarise(Sum_Countries =sum(n_countries,na.rm=T)) %>% 
  group_by(DemersPelag) %>% 
  top_n(5,n_countries)

# Countries that share the top 10
Top_Spp_Countries <- Clean_Results_Trans %>% 
  filter(
    Area_Index < Min_AI |
      Area_Index > Max_AI,
    Model_Index == MI,
    Same_Nation == "No",
    TaxonKey %in% Top_Species$TaxonKey
  ) %>% 
  left_join(Species_Trans) %>% 
  rename(scientific_name =TaxonName) %>%
  left_join(Species_environ_list) %>% 
  mutate(from = paste(Genus,"\n",Species,sep=""),
         DemersPelag = replace_na(DemersPelag, "Other")) %>%
  group_by(
    # from,
    from = scientific_name,
    to=CT_fishing_entity,
    DemersPelag
  ) %>% 
  summarise(value=n()
  ) #%>%
# filter(DemersPelag == "demersal") %>%
# select(-DemersPelag)

####________________###
# Circulize package ####
# https://jokergoo.github.io/circlize_book/book/
####________________###

# To set links color 
Bins <- as.numeric(cut(Top_Spp_Countries$value,breaks = 5)) # By size of sharing

# To get country's flags... https://flagpedia.net/download
flags <- c(paste("Temporal_Data/flags-mini/",list.files("Temporal_Data/flags-mini/"),sep=""))

# List of Ecosystems
Classification <- Top_Species %>% group_by(DemersPelag) %>% summarise(n()) %>%  pull(DemersPelag)

####________________ PLOT PER ECOSYSTEM _____________________ ###

for(s in 1:2){ # sclae loop
  for(i in 1:length(Classification)) {
    # for(i in 1:2) { # for testing

    # Set scales to plott both absolut and proportional values
    if(s == 1){
      Scales <- "TRUE"
    }else{
      Scales <- "FALSE"
    }

    # Saving the plot
    Plot_name <- paste(Figures_Path,Classification[i],Scales,".tiff",sep="")
    tiff(Plot_name, width = 5, height = 5, units = 'in', res = 500)

    # Subsetting the data for each plot
    DemersPelag_i <- Classification[i]

    Data <- Top_Ten_Spp_Countries %>% 
      filter(DemersPelag %in% DemersPelag_i) %>% 
      select(-3)

    # Use to set links' colors to each species
    Speceis_cols <- Data %>% group_by(from) %>% summarise(n()) %>% pull(from)

    # The base plot
    circos.clear()
    chordDiagram(Data, 
                 annotationTrack = c("grid", "axis"), # keep grid and axis but not labels
                 annotationTrackHeight = c(0.001, 0.1), # set links close to the axis
                 directional = 1, # direction of arrows
                 direction.type = c("diffHeight", "arrows"), # use arrows
                 grid.col = "grey",
                 # if col == Bins then colors will change depending of value
                 # col = ifelse(Bins == 1,"#046C9A", # deep blue
                 #              ifelse(Bins == 2,"#78B7C5", #light blue
                 #                     ifelse(Bins == 3,"#EBCC2A", #yellow
                 #                            ifelse(Bins == 4, "#E1AF00", #cream
                 #                                   "#F21A00")))), #red
                 # if col == Species_cols then colors will change depending of species
                 col = ifelse(Data$from == Speceis_cols[1],"#FF0000", # deep blue
                              ifelse(Data$from == Speceis_cols[2],"#00A08A", #light blue
                                     ifelse(Data$from == Speceis_cols[3],"#F2AD00", #yellow
                                            ifelse(Data$from == Speceis_cols[4], "#F98400", #cream
                                                   "#5BBCD6")))), #red
                 # link.arr.col = wes_palette("Zissou1"), 
                 link.arr.type = "big.arrow",# make links big arrows
                 diffHeight = uh(-0.5, "mm"), # brings the arrow to the bottom of the source
                 scale = Scales, # fraction of each species to total share of each country instead of absolute value
                 preAllocateTracks = list(track.height = max(strwidth(unlist(dimnames(Data))))) # I have no idea but helps with the labels
    )

    #  Function to inlcude other layers
    circos.track(track.index = 1, panel.fun = function(x, y) {
      xlim = get.cell.meta.data("xlim")
      xplot = get.cell.meta.data("xplot")
      ylim = get.cell.meta.data("ylim")
      sector.name = get.cell.meta.data("sector.index")

      # Add labels
      circos.text(mean(xlim),
                  ylim[1],
                  sector.name,
                  facing = "clockwise",
                  niceFacing = TRUE,
                  adj = c(-0.2, 0.5),
                  col = "black",
                  cex = 0.6) # size of labels

      # Add country's flags instead of names (still need to link names to falgs)

      # library(EBImage)
      # pos = circlize:::polar2Cartesian(circlize(CELL_META$xcenter, CELL_META$ycenter))
      # image = EBImage::readImage(flags[CELL_META$sector.numeric.index])
      # 
      # rasterImage(image,
      #             xleft = pos[1, 1] - 0.006, 
      #             ybottom = pos[1, 2] - 0.005,
      #             xright = pos[1, 1] + 0.05,
      #             ytop = pos[1, 2]+ 0.04
      #             )

      # END FLAGS TRYOUT

      # Set plot Title
      Title <- paste("Top",Classification[i], "Species",sep =" ")
      title(Title)

    }, bg.border = NA) #end of circos.track

    dev.off()
  } # close for loop
} #close scale loop

```

### Species status (SAU methods)

Note that (n), the number of ‘stocks’ is defined as a time series of a given species, genus or family (higher and pooled groups have been excluded) for which the first and last reported landings are at least 10 years apart, for which there are at least 5 years of consecutive catches and for which the catch in a given area is at least 1000 tonnes.

1) The first and last reported landings are at least ten years apart;

2) There are at least five years of consecutive catches; and

3) The catch in a particular area (LME) is at least 1,000 tonnes.

Once those thresholds are met, the status of the species is determined as follows:

- Rebuilding (Recovering)	Year of landing > year of post-max. min. landing AND post-max. min. landing < 10% of max. landing AND landing is 10-50% of max. landing
- Developing	Year of landing < year of max. landing AND landing is < or = 50% of max. landing OR year of max. landing = final year of landing
- Exploited	Landing > 50% of max. landing
- Over exploited	Year of landing > year of max. landing AND landing is between 10-50% of max. landing
- Collapsed	Year of landing > year of max. landing AND landing is < 10% of max. landing

*Notes* = 3 years average mean

```{r Species_Status, eval = T, echo =F}

CoorG <- read.csv(paste(Data_Path,"Spatial_Data/coordinates_gab.csv",sep="")) %>%
  mutate(INDEX = seq(1,259200,1))

# SAU relations between INDEX and Country's EEZs
EEZIDs_List <- read_excel(paste(Data_Path,"Spatial_Data/V_Lam/Updated_EEZList_17June2016.xlsx",sep=""))
EEZ_CellID <- read_excel(paste(Data_Path,"Spatial_Data/V_Lam/EEZ_CellID.xlsx",sep=""))
colnames(EEZ_CellID) <- c("EEZID","INDEX")

Index_Code <- EEZIDs_List %>% 
  left_join(EEZ_CellID) %>% 
  rename(Territory = Name)

Spp <- GetSppDist(600004, Model = "SAU_C", Coord = CoorG)

Data <- Spp %>% 
  gather("Year","Value",CATCH.1:CATCH.65) %>% 
  arrange(INDEX) %>% 
  mutate(Year = rep(seq(1951,2014,1),263250)) %>% 
  left_join(Index_Code) %>% 
  filter(!is.na(EEZID),
         Value > 0) %>% 
  group_by(Species,Year,EEZID,Territory) %>% # Total catch per country
  summarise(Total_Catch = sum(Value,na.rm=T)) %>% 
  group_by(Species,Territory,EEZID,Year) %>% 
  mutate(RMean = rollmean(x = Total_Catch, 5, align = "right", fill = Total_Catch)) %>% 
  group_by(Species,Territory,EEZID)

# head(Data)

# Tresholds to evaluate stocks if these are not met, then soecies status = "No Status"
Tresholds <- Data %>% 
  arrange(Territory) %>% 
  group_by(Species,Territory, grp = cumsum(c(1, diff(Year) != 65))) %>% 
  group_by(Species,Territory) %>% 
  summarise(Years_appart = ifelse(max(Year)-min(Year) >= 0,"Yes","No"), # Step 1. first and last reported landings are at least 10 years apart
            Consecutive = ifelse(max(grp) >= 5,"Yes","No"), # Step 2. There are at least 5 consecutive catch years
            Catch_over = ifelse(max(RMean >= 1000),"Yes","No") # Step 3. 3) The catch in a particular area (LME) is at least 1,000 tonnes.
  ) %>% 
  gather("Step","Treshold",Years_appart:Catch_over) %>% 
  filter(Treshold == "Yes")

# Evaluate stocks

# Filter data that does not satisfy tresholds

Landings_Peaks <- Data %>% 
  group_by(Species,Territory) %>% 
  summarise(Max_Peak = max(RMean,na.rm=T))

Eval_Data <- Data %>% 
  semi_join(Tresholds,
            by = c("Species","Territory")) %>% 
  left_join(Landings_Peaks,
            by = c("Species","Territory"))

Peak_Year <- Eval_Data %>% 
  filter(RMean == Max_Peak) %>% 
  select(Species,Territory,EEZID,Y_Max_Peak = Year)

Post_Max_Min <- Eval_Data %>% 
  left_join(Peak_Year) %>% 
  filter(Year > Y_Max_Peak) %>% 
  group_by(Species,Territory,EEZID) %>% 
  summarise(Post_Max_Min = min(RMean,na.rm=T))

Y_Post_Max_Min <- Eval_Data %>% 
  left_join(Post_Max_Min) %>% 
  filter(RMean == Post_Max_Min) %>% 
  select(Species,Territory,EEZID,Y_Post_Max_Min = Year)

Data_Status <- Eval_Data %>% 
  left_join(Peak_Year,
            by = c("Species","Territory","EEZID")) %>%
  left_join(Post_Max_Min,
            by = c("Species","Territory","EEZID")) %>% 
  left_join(Y_Post_Max_Min,
            by = c("Species","Territory","EEZID")) %>% 
  rename(Y_RMean = Year) %>% 
  mutate(Status = ifelse(Y_RMean > Y_Post_Max_Min & Post_Max_Min < (Max_Peak*0.10) & (RMean > (Max_Peak*0.10) & RMean < (Max_Peak*0.50)),"Rebuilding",
                         ifelse((Y_RMean < Y_Max_Peak & RMean <= (Max_Peak*0.50) | Y_Max_Peak == max(Y_RMean)),"Developing",
                                ifelse(RMean > Max_Peak*0.50,"Max Exploited",
                                       ifelse(Y_RMean > Y_Max_Peak & (RMean > (Max_Peak*0.10) & RMean < (Max_Peak*0.50)),"Over Exploited",
                                              ifelse(Y_RMean > Y_Max_Peak & RMean < Max_Peak*0.10,"Collapsed",
                                                     "No Status")
                                       )
                                )
                         )
  )
  ) %>% 
  filter(Y_RMean > 2004) %>% 
  group_by(Species,EEZID,Territory,Status) %>% 
  summarise(n_Status= n()) %>% 
  group_by(Species,EEZID,Territory) %>% 
  top_n(1,n_Status)

# Manual estimation

data.frame(
  Y_RMean = 1980,
  RMean = 642187,
  Max_Peak = 1298712,
  Y_Max_Peak = 1990,
  Post_Max_Min = 681967,
  Y_Post_Max_Min = 2010,
  Y_RMean = 2014
) %>% 
  mutate(Status = ifelse(Y_RMean > Y_Post_Max_Min & Post_Max_Min < (Max_Peak*0.10) & (RMean > (Max_Peak*0.10) & RMean < (Max_Peak*0.50)),"Rebuilding",
                         ifelse((Y_RMean < Y_Max_Peak & RMean <= (Max_Peak*0.50) | Y_Max_Peak == max(Y_RMean)),"Developing",
                                ifelse(RMean > Max_Peak*0.50,"Max Exploited",
                                       ifelse(Y_RMean > Y_Max_Peak & (RMean > (Max_Peak*0.10) & RMean < (Max_Peak*0.50))," Over Exploited",
                                              ifelse(Y_RMean > Y_Max_Peak & RMean < Max_Peak*0.10,"Collapsed",
                                                     "NA")
                                       )
                                )
                         )
  )
  )

```

```{r Sub-regional_differences, eval = F}

#-----------------#
# Sub-regional differences
#-----------------#

Subregion_Data_stats <- Value_Trans_Country %>%
  left_join(UN_Regions,
            by = "fishing_entity") %>%
  mutate(sub_region = ifelse(is.na(sub_region), "Other",paste(sub_region))) %>%
  select(fishing_entity,sub_region,Variable,Value) %>%
  filter()

# Revenue ANOVA
Subregion_Rev_stats <- Subregion_Data_stats %>%
  filter(Variable == "value")

Rev_aov <- aov(Value~sub_region, data = Subregion_Rev_stats)
Rev_aov
summary(Rev_aov)

# Df    Sum Sq Mean Sq F value  Pr(>F)    
# sub_region   16  75628526 4726783    4.91 4.2e-08 ***
# Residuals   157 151140183  962676

#### There are significant differences. But between what?

#  Post-hoc analysis
TukeyHSD(Rev_aov)

# Northern America sig than everyone except for Eastern Asia
# Eastern Asia sig with everyone except with Eastern Europe and North Am.

```

# RFMO Analysis

```{r RFMO_Data, eval = F}

RFMO_Data <- read.delim("/Volumes/DATA/JULIANO_NEYMAR/FishForVisa/Spatial_Data/RFMO/RFMO_procedure_outcome.csv")

x <- RFMO_Data %>% 
  mutate(
    x = gsub("<.*?>", "", RFMO_Data$primary_species),
    y = gsub("[\\(\\)]", "", regmatches(x, gregexpr("\\(.*?\\)", x)))
  )

# Save to remove "" and pass to colomuns in excell
# write.csv(x,
#           "rfmos.csv")

rfmos <- read.csv("~/Github/FishForVisa/rfmos.csv")
View(rfmos)

list <- rfmos$Taxa

clean_rfmos <- rfmos %>% 
  gather("taxa","taxon_name",y:X.21) %>% 
  filter(taxon_name != "") %>% 
  select(RFMO_name = name,
         contracting_parties,
         area,
         taxon_name)

# write_csv(clean_rfmos,
#           "clean_rfmos.csv")

### Join with exploited species

clean_rfmos <- clean_rfmos %>% 
  left_join(Exploited_Species)

```

## RFMO specifics

```{r RFMOs, eval = T, echo = F}

### Results exploration

ICCAT_List <- rfmo_list %>% 
  filter(rfmo_name == "ICCAT") %>% 
  pull(taxon_key) %>% 
  unique()

Tresholded_T_Data %>% 
  filter(taxon_key %in% ICCAT_List) %>% 
  group_by(taxon_key) %>% 
  summarise(
    n_countries = length(unique(fishing_entity))
  )

ICCAT_Trans <- Species_Data %>% 
  rename(taxon_name = individual) %>% 
  left_join(Exploited_Species) %>% 
  filter(taxon_key %in% ICCAT_List)

```

## RFMO Map of Transboundary Species

```{r RFMO_Map, eval =F, echo =F}

Tresholded_rfmo_Data <- Results_rfmo %>% 
  filter(
    area_index >= Min_AI,
    area_index <= Max_AI,
    model_index >= MI
  ) %>% 
  group_by(rfmo_name) %>% 
  summarise(
    n_trans_spp = length(unique(taxon_key)),
    n_neighbours = length(unique(neighbour)),
    trans_Rrate = n_trans_spp/n_neighbours
  )

# Number of RFMOs
rfmo_list %>% 
  pull(rfmo_name) %>% 
  unique() %>% 
  length() # 15

# Number of total taxa
rfmo_list %>% 
  pull(taxon_key) %>% 
  unique() %>% 
  length() # 99

# Number of taxa per RFMO
rfmo_list %>% 
  group_by(rfmo_name) %>% 
  summarise(in_trans_spp = length(unique(taxon_key)))

# Number of shared taxa
Results_rfmo %>% 
  pull(taxon_key) %>% 
  unique() %>% 
  length() # 81

### Bar plot

Tresholded_rfmo_Data %>% 
  ggplot() +
  geom_bar(
    aes(
      x = reorder(rfmo_name,-n_trans_spp),
      y = n_trans_spp
    ),
    stat = "identity"
  ) +
  coord_flip()

head(RFMOs_sf)

# Transboundary Species shapefile
World_sf_si <- RFMOs_sf %>%
  filter(!Name %in% c("IPHC","PSC","NASCO")) %>% 
  rename(rfmo_name = Name) %>% 
  st_transform(crs = 4326) %>% # 4326
  # st_simplify(preserveTopology = TRUE, dTolerance = 0.1) %>%
  left_join(Tresholded_rfmo_Data, by ="rfmo_name")

# Total catch/Revenue shapefile
World_sf_si_land <- World_land_sf %>%
  st_transform(crs = 4326) #%>% # 4326
# st_simplify(preserveTopology = TRUE, dTolerance = 0.1)

# Get the NA's to be grey
# Grey_land <- World_sf_si_land %>% 
#   filter(is.na(Bins))

# Make map
Plot <- ggplot() +
  geom_sf(data = World_sf_si,
          aes(
            fill =  n_trans_spp,
            colour = n_trans_spp
          ), 
          size = 0.1
  ) + # n_trans_spp or Bins
  geom_sf(data = World_sf_si_land) +
  # geom_sf_label(data = World_sf_si,
  #               aes(label = eez_name)
  #               ) +
  scale_fill_gradientn("Number of Shared Species",
                       colours = wes_palette("Zissou1", 100, type = "continuous"),
                       limits = c(0,45),
                       breaks = seq(0,45,5)
  ) +
  scale_colour_gradientn("Number of Shared Species",
                         colours = wes_palette("Zissou1", 100, type = "continuous"),
                         limits = c(0,45),
                         breaks = seq(0,45,5)
  ) +
  ggtheme_map()

Plot_Name <- paste("S2_RFMO.png", sep = "")

ggsave(Plot_Name,
       plot = Plot,
       width = 16,
       height = 8,
       units = "in",
       path = Figures_Path
)
```