### 2022-06-30 Scripts/Initial_Analysis.Rmd
Incorporated straddling analysis to transboundary analysis

---
title: "Initial Analysis"
author: "Juliano Palacios-Abrantes"
date: '2019-02-04'
output: html_document
editor_options: 
  chunk_output_type: console
---

# Read Me
This script runs the analysis for the manuscript *Marine species do not need visas: the transboundary nature of the world’s exploited marine species* from the data loading to the results analysis. Is divided in two parts; *Methods* where the raw data is analyzed and *Results* where the results from the analysis are processed.

```{r setup, eval = T, echo=F, warning=F,message=F, results='hide'}

#### READ ME !!! ####
library(MyFunctions)

#### Library ####
packages <- c(
  "readxl", # Read dataframe
  "data.table", # Read dataframe (Fast!)
  "wesanderson",
  "tidyverse", # for all data wrangling and ggplot
  "janitor", # for data cleaning
  "tidytext", # to order the facet wrap https://juliasilge.com/blog/reorder-within/
  "cowplot", # for figures 1 and 3
  # "ggimage", #for reading images to the circular plot
  # "ggrepel", # for nice plot labels
  # "ggsflabel", # for nice sf_plots labels
  # "spdep", # for poly2nb old
  "sf", #Spatial analysis 
  "sp", #Spatial analysis 
  # "purrr",#Spatial analysis
  "rgdal", #Spatial analysis
  "tools", #Spatial analysis 
  "parallel", # for parallelization
  # "taxize", # For getting species names
  "rfishbase", # for species ecosystem affinity
  "zoo", #for runing mean
  "pgirmess" # for dune test after kurtis wallas
)

my_lib(packages)

#### Set paths for computer ####

Jursidiction <- "EEZ"

# Set paths depending on machine Beast (jepa88), "carmelia" or Hall1000
# if(Sys.info()[7] == "jepa88"){
#   Data_Path <- "Z:/JULIANO_NEYMAR/FishForVisa/"
# }
# if(Sys.info()[7] == "carmelia"){
#   Data_Path <- paste(here::here(),"/Temporal_Data/",sep = "")
# }
# if(Sys.info()[7] == "hall1000"){
#   Data_Path <- "/Volumes/DATA/JULIANO_NEYMAR/FishForVisa/"
# }

# Setting up paths 
# If path exists overrite it, otherwise creats a new path for that year

# Path for saving Results
Results_Path = paste(Data_Path,"Results/",sep = "")
if(file.exists(Results_Path) == TRUE){
  Results_Path = paste(Data_Path,"Results/",sep = "")
}else{
  dir.create(paste(Data_Path,"Results/",sep = ""))
  Results_Path = paste(Data_Path,"Results/",sep = "")
}

# Path for saving Figures
Figures_Path = paste(Data_Path,"Figures/",sep = "")

if(file.exists(Figures_Path) == TRUE){
  Figures_Path = paste(Data_Path,"Figures/",sep = "")
}else{
  dir.create(paste(Data_Path,"Figures/",sep = ""))
  Figures_Path = paste(Data_Path,"Figures/",sep = "")
}

ggtheme_map <- function(base_size = 9, Region = "NA") {

  theme(text             = element_text(#family = "Helvetica",
    color = "gray30", size = base_size),
    plot.title       = element_text(size = rel(1.25), hjust = 0, face = "bold"),
    panel.background = element_blank(),
    panel.border     = element_blank(),
    panel.grid.minor = element_blank(),
    panel.grid.major = element_line(color = "transparent"),
    strip.background =element_rect(fill = "transparent"),
    strip.text.x = element_text(size = 18, colour = "black",face= "bold.italic", angle = 0),
    axis.line        = element_blank(), # element_line(colour = "grey30", size = .5))
    axis.ticks       = element_blank(),
    axis.text        = element_blank(),
    axis.title       = element_blank(),
    legend.key       = element_rect(colour = NA, fill = NA, size = 4),
    legend.position = "bottom",
    legend.key.width =unit(5,"line")
  )
}

### Message for my brain

print("_____________________ WARNING DO NOT FORGET TO PULL/PUSH BEFORE YOU START_____________________")

```

# Methods

These are the methods to follow:

1. Determine which countries have Neighboring EEZs  
1.1. For this we used the Sea Around Us EEZ map that discriminate by oceans and a routine (*Estimate Transboundary Species*) to identify Neighbouring polygons,
1.2. Then we match the SAU spatial information (e.g. INDEX within EEZ and coordinates) with the Neighbouring data

2. Overlap distribution maps with Neighbors data  
2.1. For each species we merged the Neighbors data with their known distribution (SAU distribution and catch,ESMs, and Observation)  
3. Determine transboundary if:  
3.1. Species is present in both Neighbouring EEZs by INDEX

We create two "confidence level" based on data and area covered.

## Determine species list

This part is to determine the list of species to be analyzed in the project. Initially this list was limited by the species data from the DBEM, however, as we are not using the DBEM, the limit is now the SAU.

```{r Speceis_List_For_Data, eval = F, echo = T}

# Get the DBEM list of species to extract from other modesl
#GFDL
DBEM_species_G <- data_frame(TaxonKey = list.files("/Volumes/DATA/DATA/DBEM/GFDL26F1"))
#IPSL
DBEM_species_I <- data_frame(TaxonKey = list.files("/Volumes/DATA/DATA/DBEM/IPSL26F1"))
#MPI
DBEM_species_M <- data_frame(TaxonKey = list.files("/Volumes/DATA/DATA/DBEM/MPI26F1"))

# Merge the, and save to send to gabs

DBEM_Species_List <- DBEM_species_I %>% 
  bind_rows(DBEM_species_G,
            DBEM_species_M) %>% 
  group_by(TaxonKey) %>% 
  summarise() %>% 
  filter(str_detect(TaxonKey, "^6"))

# Include species name, as well...

exploited_species <- read.csv("/Volumes/DATA/PROJECTION exploited species ENM/exploited_species_list.csv")

DBEM_Species_List <- DBEM_Species_List %>% 
  left_join(exploited_species,
            by ="TaxonKey")

# write.csv(DBEM_Species_List,
#           "DBEM_Species_List.csv",
#           row.names = F)

# Send second set of species

# Get species already modelled
# Removes everything before the number
Step_One <- gsub(".*_","",list.files(paste(Data_Path,"/Distribution_Data/SAU_Distribution/",sep="")))

# Species List
Species_List <- gsub("\\..*","",Step_One)

exploited_species <- read.csv("/Volumes/DATA/PROJECTION exploited species ENM/exploited_species_list.csv") #%>% 
# left_join(Spp_Clasification) %>% 
# filter(!is.na(DemPel)) # remove freshwater spp

# Get all species in SAU database (2080)
Species_list <- exploited_species %>% 
  rename(scientific_name = TaxonName) %>% 
  pull(scientific_name)

# get their environ. association according to FishBase (species())

Species_environ_list <- rbind(species(Species_list, 
                                      fields = species_fields$habitat)
) %>% 
  mutate(TaxonName = Species_list) %>% 
  left_join(exploited_species,  by ="TaxonName") %>% 
  filter(!TaxonKey %in% Species_List) %>% 
  filter(Saltwater == -1) %>%  # remove those that are not happy with saltwater
  select(TaxonKey,TaxonName,CommonName)

# write.csv(Species_environ_list,
#           "Transboundary_Spp_Set2.csv",
#           row.names = F)

# Sent to Gabs Seprember 16th

### Check Gabs Data ####

# Does all data have the same number of files?

# SAU <- length(list.files("/Volumes/DATA/JULIANO_NEYMAR/FishForVisa/Distribution_Data/SAU_Distribution/"))
# ENM <- length(list.files("/Volumes/DATA/JULIANO_NEYMAR/FishForVisa/Distribution_Data/ENM/"))
# Occ <- length(list.files("/Volumes/DATA/JULIANO_NEYMAR/FishForVisa/Distribution_Data/Occurence/"))

####### Yes they do! ####

```

## Determine Neighbors

This is the first step of the analysis and will determine what EEZs are neighbors. The routine creates a final data-set that we then use to estimate transboundary species.

### Function `EstNeighbours`

This function determines what EEZs are Neighbors and returns a data.frame of it

```{r funEstNeighbours, eval = F, echo=T}
# Determine which countries are Neighbours

EstNeighbours <- function(Shapefile, Partial = "NA"){

  # Get sparse matrix
  int = suppressMessages(st_intersects(Shapefile))
  # Warnning message
  # although coordinates are longitude/latitude, st_intersects assumes that they are planar

  # Intitalize new object
  mat = NULL

  # Fill
  for(i in 1:NROW(int)){
    #Create matrix with ID and intersecting IDs
    tmp = cbind(i, int[[i]])

    # Join to Initalized object
    mat  = rbind(mat, tmp)
  }

  # Build final data.frame substituting ID for name (in this case nc$name)
  Result_table = data.frame(Name = Shapefile$Name[mat[,1]], 
                            Neighbour_Territory = Shapefile$Name[mat[,2]]
  )

  # Remove duplicates... eg of course an object intersects with itself...
  Result_table = Result_table[Result_table$Name != Result_table$Neighbour_Territory,]

  if(Partial == T){
    # If only partial analysis just returns Neighbour list
    return(Result_table)

  }else{

    # Merge them with SAU data so we have the INDEX and lat long info
    suppressWarnings(
      # warning for characters in name, no problem
      DBEM_Neighbours_Data <- Result_table %>% 
        left_join(EEZIDs_List, 
                  by = "Name") %>% # Asign EEZ id to each country
        left_join(EEZ_CellID, 
                  by = "EEZID") %>% # Asign EEZ id to each country
        left_join(Coor,
                  by = "INDEX") %>%  # Include lat long
        rename(Country_Territory = Name)

    )

    return(DBEM_Neighbours_Data)

  } # end ifelse
} # end function

# subset <- World_EEZs %>% filter(Name %in% c("Chile","Peru"))
# EstNeighbours(subset)
# EstNeighbours(subset,T)

```

### Sub Routine; Neighbors

```{r Neighbours_Control_Pannel, eval = F, echo = F}

###___________________Read Me_____________________#
#### Using the SAU shapefile that contains EEZ per ocean (E.G. Mexico Pacific and Mexico Atlantic)
###_______________________________________________#

# SAU shapefile

# The path
path_world <- paste(Data_Path,"Spatial_Data/SAU_Shapefile/",sep="")

# The File
fnam_world <- "SAUEEZ_July2015.shp"

# Load it!
World_EEZs <- st_read(dsn = path_world,
                      layer =file_path_sans_ext(fnam_world))

#### Grid cell information ####

# List of DBEM INDEX within the EEZ 
#Data provided by Vicki Lam

EEZIDs_List <- read_excel(paste(Data_Path,"Spatial_Data/V_Lam/Updated_EEZList_17June2016.xlsx",sep=""))
EEZ_CellID <- read_excel(paste(Data_Path,"Spatial_Data/V_Lam/EEZ_CellID.xlsx",sep=""))

colnames(EEZ_CellID) <- c("EEZID","INDEX")

# DBEM Coordinate system
Coor <- fread(paste(Data_Path,"Spatial_Data/Lon_Lat_DBEM.txt",sep=""),header = FALSE)
colnames(Coor) <- c("INDEX","Longitude","Latitude")

```

```{r Neighbours_Subroutine, eval = F, echo = T}

# Run Function (Ran and saved May 27 2019 at 6pm)
system.time(
  EEZ_Neighbour_List <- EstNeighbours(World_EEZs, Partial = T)
)
# 8203.05    0.25 8204.51 (around 2 hours 30 min)

write.csv(EEZ_Neighbour_List,
          paste(Data_Path,"Spatial_Data/EEZ_Neighbour_List.csv", sep =""),
          row.names = F)

####_________________________________________ TESTING PERIOD_________________________________________###
# I'm substetting central america to do the whole analysis (smaller EEZs = faster computation) once I have figure out things I'll run for all

### Data
# For now using only the subdata

## Spatial Test Data
# SAU_Regions <- as.data.frame(unique(World_EEZs$Name))

# Test_Countries <-c("Guatemala (Caribbean)","Guatemala (Pacific)","Belize","Honduras (Pacific)","El Salvador","Costa Rica (Pacific)","Costa Rica (Caribbean)", "Panama (Caribbean)", "Panama (Pacific)")

# Test <- World_EEZs[World_EEZs@data$Name %in% Test_Countries, ]

```

## Determine Catch Trends

This sub routine determines the species exploitation trend based on the SAU methodology. **Note:** For the current project we do not consider this method to determine the species exploitation trend *per se*, but rather a catch trend that *could* indicate that a species is under/over/max exploited.

In the original method the authors analyze stocks, while we analyze at the species level within each EEZ. Also (n), the number of species, is defined as a time series of a given species, genus or family (higher and pooled groups have been excluded) for which the first and last reported landings are at least 10 years apart, for which there are at least 5 years of consecutive catches and for which the catch in a given area (LME) is at least 1000 tonnes. We removed the last one (LME) and computed all of the species that only followed the first two;

1) The first and last reported landings are at least ten years apart;

2) There are at least five years of consecutive catches

Once those thresholds are met, the trend of the species is determined as follows:

- Rebuilding (Recovering),	Year of landing > year of post-max. min. landing AND post-max. min. landing < 10% of max. landing AND landing is 10-50% of max. landing
- Developing,	Year of landing < year of max. landing AND landing is < or = 50% of max. landing OR year of max. landing = final year of landing
- Max Exploited,	Landing > 50% of max. landing
- Over exploited,	Year of landing > year of max. landing AND landing is between 10-50% of max. landing
- Collapsed,	Year of landing > year of max. landing AND landing is < 10% of max. landing

*Notes* = 3 years average mean

Finally we choose the reporting year based on the most frequent trend in the last 10 years of data. In cases where there was a mach between trend (e.g. Cayman Islands for species 6000006 had 5 *Max Exploited* and 5 *Overexploted*) then we report the trend at the last year of data.

Categories are:

- Increasing Catch 
- COnstant Catch
- Decreasing Catch

### Function `GetSppDist`

This function reads in species distribution from the observation data, ESM models and SAU data.

```{r GetSppDist, eval = T, echo = T}

#### The function
# NOTES
# For now we're using only the last 10 years average, basicaaly if the species has been fished in any of these years, its considered present

GetSppDist=function(Spp,Model,Coord){

  Distpath <- paste(Data_Path,"Distribution_Data/",sep="")
  INDEX = seq(1,259200,1)
  # INDEX <- Coordinates$INDEX

  # SAU Distributions
  if(Model == "SAU_D"){
    File_Name <- paste("SAU_Distribution/DIST_GOOD_SP_")
  }

  # SAU Catch
  if(Model == "SAU_C"){
    File_Name <- paste("SAU_data_per_species/CATCH_SP_")
  }

  # Occurence
  if(Model == "Occ"){
    File_Name <- paste("Occurence/OCCURENCE_JULIANO_")
  }

  # ENM Model
  if(Model == "ENM"){
    File_Name <- paste("ENM/ENM_JULIANO_")
  }

  # DBEM Data
  # if(Model == "DBEM"){
  #   Distpath <- "/Volumes/DATA/JULIANO_NEYMAR/PristineSeasData
  #   Final_Path <- "ENM"
  #   File_Name <- paste("ENM_JULIANO",Spp,".mat",sep="")
  # }

  if(Model == "All"){

    Models_List <- c(paste(Distpath,"SAU_Distribution/DIST_GOOD_SP_",Spp,".mat",sep =""),
                     paste(Distpath,"Occurence/OCCURENCE_JULIANO_",Spp,".mat",sep=""),
                     paste(Distpath,"ENM/ENM_JULIANO_",Spp,".mat",sep="")
    )

    # Jumps species not modeled. NOTE: We only need one as ENM, Occ and SAU Dis have all the same 939 spp
    if(file.exists(Models_List[1])){

      Load <- lapply(Models_List, FUN=R.matlab::readMat, na.strings=0)

      sppdist <- as.data.frame(bind_cols(Load)) %>% 
        mutate(
          INDEX = INDEX,
          TaxonKey = Spp
        )
      colnames(sppdist) <- c("SAU_D","Occ","ENM","INDEX","TaxonKey")

      #### Step for SAU catch data that has to be averaged
      File_Name <- paste("SAU_data_per_species/CATCH_SP_")
      SppPath <- paste(Distpath,File_Name,Spp,".mat",sep="")

      # For now we're using only the last 10 years average, basicaaly if the species has been fished in any of these years, its considered present
      SAU_C_data <- as.data.frame(R.matlab::readMat(SppPath)) %>% 
        select(CATCH.1:CATCH.65) %>% 
        mutate(INDEX = INDEX) %>% 
        gather("Year","Catch",CATCH.56:CATCH.65) %>% # Last 10 years of data
        group_by(INDEX) %>% 
        summarise(SAU_C = mean(Catch,na.rm=T))

      # Join both tables
      sppdist <- sppdist %>% 
        left_join(SAU_C_data,
                  by = "INDEX") %>% 
        select(TaxonKey,INDEX,everything()) %>% 
        left_join(CoorG,
                  by = "INDEX")

      # Fix coordinate system incompatibility between Gab and DBEM
      sppdist <- suppressWarnings(sppdist[order(sppdist$Latitude, rev(sppdist$Longitude),decreasing=TRUE), ] %>% 
                                    mutate(INDEX = seq(1,259200,1)) %>% 
                                    gather("Model","Value",3:6) %>%
                                    mutate(Value = ifelse(is.na(Value), 0, Value)) # Converting NA's to ceros
      )

      getSppDist = sppdist

    }

  }else{

    # Merge paths
    SppPath <- paste(Distpath,File_Name,Spp,".mat",sep="")

    if(file.exists(SppPath) == TRUE){

      #Install (if needed) R.matlab package
      if(!require(R.matlab)){
        install.packages("R.matlab")
      }

      # Read Files

      sppdist <- as.data.frame(R.matlab::readMat(SppPath)) %>% 
        mutate(INDEX = INDEX,
               Species = Spp) %>% 
        left_join(Coord)

      # Fix coordinate system incompatibility between Gab and DBEM
      sppdist <- sppdist[order(sppdist$Latitude, rev(sppdist$Longitude),decreasing=TRUE), ] %>% 
        mutate(INDEX = seq(1,259200,1))

      # Return 
      getSppDist=sppdist

    }else{
      print(paste("No info for this species",Spp, "in",Model))
    }

  }
}

#______ Test ______#

# Variables outside function
# head(GetSppDist(600004,"SAU_D")) # Works
# head(GetSppDist(600004,"SAU_C")) # Works
# head(GetSppDist(600004,"Occurence")) # Works
# head(GetSppDist(600004,"ENM")) # Works
# head(GetSppDist(600004,"All")) # Works

```

### Function `EstCatchTrend`

This function estimates the exploitation trend of a species based on catch data.

```{r funEstCatchTrend, eval = T, echo =F}

# Spp <- 601417
# Coord = CoorG

EstCatchTrend <- function(Spp,Coord, Save_Path = "NA", Per_Taxa = NA){

  # Load species' data
  Data <- GetSppDist(Spp, Model = "SAU_C", Coord = Coord) %>% 
    set_names(c(seq(1950,2014,1),"INDEX","Species","Longitude","Latitude")) %>% 
    gather("Year","Value",`1950`:`2014`) %>%
    arrange(INDEX) %>% 
    left_join(Index_Code, # Match species catch with INDEX
              by = "INDEX") %>% 
    filter(!is.na(EEZID),
           Value > 0) %>% 
    group_by(TaxonKey=Species,Year,Territory) %>% # Estimates total catch per country
    summarise(Total_Catch = sum(Value,na.rm=T)) %>% 
    group_by(TaxonKey,Territory,Year) %>% # Running mean as Kleisner & Pauly 2012
    mutate(RMean = rollmean(x = Total_Catch, 3,
                            align = "left",
                            fill = Total_Catch)
    ) %>% 
    select(-Total_Catch) %>% 
    ungroup() %>% 
    mutate(Year = as.numeric(Year))

  # Tresholds to evaluate stocks if these are not met, then soecies Trend = "No Trend"
  Tresholds_Data <- Data %>% 
    arrange(Territory) %>% 
    group_by(TaxonKey,Territory, grp = cumsum(c(1, diff(Year) != 65))) %>% 
    group_by(TaxonKey,Territory) %>% 
    summarise(Years_appart = ifelse(max(Year)-min(Year) >= 0,"Yes","No"), # Step 1. first and last reported landings are at least 10 years apart
              Consecutive = ifelse(max(grp) >= 5,"Yes","No")#, # Step 2. There are at least 5 consecutive catch years
              # Catch_over = ifelse(max(RMean >= 1000),"Yes","No") # Step 3. 3) The catch in a particular area (LME) is at least 1,000 tonnes. (CURRENTLY OFF)
    ) %>% 
    gather("Step","Treshold",Years_appart:Consecutive) %>% 
    group_by(TaxonKey,Territory) %>% 
    summarise(Met_Criteria = paste(unique(Treshold),collapse =";")
    )

  # Remove species that do not met the criterias

  No_Trend_Data <- Tresholds_Data %>% 
    filter(Met_Criteria != "Yes") %>% 
    select(TaxonKey,Territory) %>% 
    mutate(Trend = "No Trend")

  # Evaluate species that met the criteria

  # Get Variables to perform Trend metric
  Trend_Data <- Data %>% 
    # filter(Territory == "Israel (Red Sea)") %>%  # For testing
    anti_join(No_Trend_Data, # Removes no Trend data
              by = c("TaxonKey","Territory")
    ) %>% 
    # Get Max catch and max catch year
    group_by(TaxonKey,Territory) %>% 
    top_n(1,RMean) %>% 
    group_by(TaxonKey,Territory,RMean) %>%  # Solves for years with the same Max catch data (e.g. Cayman Isl all years)
    summarise(Year = min(Year)) %>%
    rename(Max_Peak = RMean,
           Y_Max_Peak = Year) %>% 
    # Get Min catch after peak catch
    left_join(Data,
              by = c("TaxonKey","Territory")
    ) %>% 
    filter(Year >= Y_Max_Peak) %>% # Filter years after peak year
    group_by(TaxonKey,Territory) %>% 
    top_n(-1,RMean) %>% # find min value
    group_by(TaxonKey,Y_Max_Peak,Territory,Max_Peak,RMean) %>%  # Solves for years with the same min after peak catch data (e.g. Cayman Isl 2013-2014)
    summarise(Year = min(Year)) %>% 
    rename(Post_Max_Min = RMean,
           Y_Post_Max_Min = Year) %>% 
    left_join(Data,
              by = c("TaxonKey","Territory")
    ) %>% 
    # Final Catch Trend Evaluation 
    mutate(Trend = ifelse(Year > Y_Post_Max_Min & Post_Max_Min < (Max_Peak*0.10) & ((RMean > (Max_Peak*0.10) & RMean < (Max_Peak*0.50))),"Rebuilding",
                           ifelse((Year < Y_Max_Peak & RMean <= (Max_Peak*0.50) | Y_Max_Peak == max(Year)),"Developing",
                                  ifelse(RMean > (Max_Peak*0.50),"Max Exploited",
                                         ifelse(Year > Y_Max_Peak & ((RMean > (Max_Peak*0.10) & RMean < (Max_Peak*0.50))),"Over Exploited",
                                                ifelse(Year > Y_Max_Peak & RMean < (Max_Peak*0.10),"Collapsed",
                                                       "No Trend")
                                         )
                                  )
                           )
    )
    )

  #_______________#
  # Function Return 
  #_______________#

  ### Save the data for each species in all years if Per_taxa = T

  if(Per_Taxa == T){

    # Select onlyZ 
    Trend_Data <- Trend_Data %>%
      ungroup() %>%
      select(TaxonKey,Territory,Year, Trend)

    if(is.na(Save_Path) == TRUE){

      Save_Path <- Results_Path # standar result's path in my Drobo
    }

    # File name for each species
    Name <- paste(Save_Path,Spp,"_Trend.csv",sep="")

    # Save file for each species
    write_csv(Trend_Data,
              Name)
  }else{

    ### Determine the final Trend using the top category over the last 10 years

    Determine_Trend_Data <- Trend_Data %>% 
      filter(Year > 2004) %>% # last 10 years of data (2005-2014)
      group_by(TaxonKey,Territory,Trend) %>%
      summarise(n_Trend= n()) %>%
      group_by(TaxonKey,Territory) %>%
      top_n(1,n_Trend)

    # Fix cases where you have equal categories.
    suppressMessages( # supresses "No duplicate combinations found of: Territ"
      Duplicates <- Determine_Trend_Data %>% 
        ungroup() %>% 
        janitor::get_dupes(Territory) %>% 
        select(TaxonKey,Territory) %>% 
        pull(Territory) %>% 
        unique()
    )

    # In case we have a tide, we use the last year's category
    Duplicates_Trend <- Trend_Data %>% 
      filter(Territory %in% Duplicates,
             Year == 2014) %>%
      ungroup() %>%
      select(TaxonKey,Territory,Trend)

    # Build Final Trend database
    Final_Trend_Data <- Determine_Trend_Data %>% 
      filter(!Territory %in% Duplicates) %>% 
      bind_rows(Duplicates_Trend)

    # If both datasets have value RETURN BOTH
    if(nrow(No_Trend_Data) > 0 & nrow(Final_Trend_Data) > 0){

      Fun_Result <- bind_rows(No_Trend_Data,
                              Final_Trend_Data)
    }

    # If only No trend data, then just return No Trend Data
    if(nrow(No_Trend_Data) > 0 & nrow(Final_Trend_Data) == 0){
      return(No_Trend_Data)
    }else{
      return(Final_Trend_Data)
    }
  } # close if per taxa
}

# Testing
# Test_1 <- EstSppTrend(Species_List[1],CoorG)

```

### Data Catch Trend

### Sub routine for Catch Trend

```{r Spp_Trend_Control_Pannel, eval = T, echo = F}

# SAU relations between INDEX and Country's EEZs
EEZIDs_List <- read_excel(paste(Data_Path,"Spatial_Data/V_Lam/Updated_EEZList_17June2016.xlsx",sep=""))
EEZ_CellID <- read_excel(paste(Data_Path,"Spatial_Data/V_Lam/EEZ_CellID.xlsx",sep=""))
colnames(EEZ_CellID) <- c("EEZID","INDEX")

Index_Code <- EEZIDs_List %>% 
  left_join(EEZ_CellID) %>% 
  rename(Territory = Name)

# Gabriel's coordinate system
CoorG <- read.csv(paste(Data_Path,"Spatial_Data/coordinates_gab.csv",sep="")) %>%
  mutate(INDEX = seq(1,259200,1))

# Removes everything before the number
Step_One <- gsub(".*_","",list.files(paste(Data_Path,"/Distribution_Data/SAU_Distribution/",sep="")))

# Species List
Species_List <- gsub("\\..*","",Step_One)

```

```{r Spp_trend_Sub_Routine, eval = T, echo = F}
# Needs to have trans species 
Species_Trend <- bind_rows(
  mclapply(
    Species_List[1:2],
    FUN = EstCatchTrend,
    Coord = CoorG,  # Gabriels coordinate system for read species dist
    Per_Taxa = T,
    Save_Path = "Z:/Vicky/Stock_Trend/"
  )
)

# Save data
readr::write_csv(Catch_Trend,
                 paste(Results_Path,"Species_Trend.csv",sep="")
)
# head(Species_Trend)
```

## Get Catch Data

### Function `EstCatch`

This function just creates a database of total catch per species within each EEZ

```{r EstCatch, eval = F, echo = T}

EstCatch <- function(Spp, Coord, Index_Code){

  SppDist <- GetSppDist(Spp,Model="All",Coord) %>% 
    filter(
      Model == "SAU_C",
      Value > 0) %>% 
    left_join(Index_Code,
              by = "INDEX") %>% 
    group_by(eez_name = Territory,
             taxon_key = TaxonKey) %>% 
    summarise(total_catch = sum(Value))

  return(SppDist)

}

# EstCatch(Spp,Coord,Index_Code)

```

### Sub routine for catch data

```{r control_panel, eval = F, echo = T}

# Removes everything before the number
Step_One <- gsub(".*_","",list.files(paste(Data_Path,"/Distribution_Data/SAU_data_per_species/",sep="")))

# Species List
Species_List <- gsub("\\..*","",Step_One)

CoorG <- read.csv(paste(Data_Path,"Spatial_Data/coordinates_gab.csv",sep="")) %>%
  mutate(INDEX = seq(1,259200,1))

# SAU relations between INDEX and Country's EEZs
EEZIDs_List <- read_excel(paste(Data_Path,"Spatial_Data/V_Lam/Updated_EEZList_17June2016.xlsx",sep=""))
EEZ_CellID <- read_excel(paste(Data_Path,"Spatial_Data/V_Lam/EEZ_CellID.xlsx",sep=""))
colnames(EEZ_CellID) <- c("EEZID","INDEX")

Index_Code <- EEZIDs_List %>% 
  left_join(EEZ_CellID) %>% 
  rename(Territory = Name)

```

```{r subroutine_get_catch_data, eval = F, echo = T}
# Notes, needs GetSppDist
# Note, It averages the LAST 10 years oe ach gridcell and then adds them together
#  I think NA is the ABNJ

system.time(
  SAU_Catch_Data <- bind_rows(
    mclapply(
      Species_List,
      FUN = EstCatch,
      Coord = CoorG,  # Gabriels coordinate system for read species dist
      Index_Code = Index_Code
    )
  )
)

# Save dataframe for future
Save_Path <- paste(Data_Path,"Distribution_Data/sau_catch_country_taxon.csv",sep="")
write_csv(SAU_Catch_Data,
          Save_Path)

```

## Estimate Transboundary Species

This is the main step of the project. In this segment we estimate the number of transboundary species existing today in each of the worlds EEZs.

The data-sets SAU_D, OCC, and ESM are used as species presence while SAU_C is used as a confirmation of that. The `Model Index` is the first threshold and estimates the uncertainty of models. Goes from 0 - 1 as follows:

- 1 or 0; All models agree that in the presence of the species
- 0.75; Three out of four models agree on species presence
- 0.50; Two out of four models agree on species presence
- 0.25; Only one model flags a species presence
- 0; All models agree there is no species there

The `Area_Index` is the second threshold and estimates the proportion of each species distribution in each EEZ. It only take into consideration the distribution of the species within the two EEZs, rather than the complete distribution. Hence, is mirror (e.g. if EEZ#1 is 0.40, EEZ#2 is 0.6)

** NOTE THAT THE FUNCTION IS SETTED TO OUTPUT RESULTS FOR ALL MODELS **

### Function `EstTransIndex`

This is the main function of the project, the one that actually estimates if a species is transboundary or not

```{r funEstTransIndex, eval= T ,echo = T,warning = F,message = F}

# Function to determine whether or not a species is transboundary

# Varibales needed
# Spp: Species to be analized, data with presence/absence per gridcell
# Index_EEZ: Reference list of all INDEX that fall within EEZs
# EEZ_Size: Number of grid-cells that each EEZ has
# Neighbours: Reference list of neighbouring countries

# Needs getSppDist

#### For testing 
# Spp <- 600004  # Anchovie Equator, Chile and Peru
# Spp <- 600107 # Tuna Many RFMOs
# Spp <- 600004 # Argument 1 must have names (RFMO)
# Spp <- 600464 # Discrete species 
# Spp <- 600051 # One of Pierrs massive (wrong) discrete spp
# x <- c(600464, # Discreat spp in one nation
# 600059, # nowhere to be found spp
# 601352, # Discrete spp in two nations
# 600004) # transboundary spp
# Model <- "All"
# Coord <- CoorG
# 
# # For EEZ
# Neighbours <- Neighbours_Data
# Index_Code <- Index_Code

# __________________________________ #

# The function
EstTransIndex <- function(Spp,Model = "All",Neighbours,Coord,Index_Code,Save="Y",Jurisdiction_Type){

  # Get model data from spps
  SppDist <- GetSppDist(Spp,Model,Coord)

  # Result 1. Number of Countries that share the species

  #____________ ESTIMATING MODEL INDEX (TRESHOLD 1)_________ #
  Trans_Spp <- SppDist %>%
    filter(INDEX %in% Neighbours$INDEX,# Filter data only located within EEZs
           Model != "SAU_C") %>% # Only using observational and modelled data
    mutate(Value = ifelse(Value > 0, 1,0)) %>%  
    filter(Value > 0) %>%
    group_by(TaxonKey,
             INDEX
    ) %>%
    summarise(Model_Index = sum(Value,na.rm=T)/3
    ) %>%
    filter(Model_Index > 0.4) %>% # at least 2 sources agree
    # ____________ ESTIMATING FUNDAMENTAL NICHE (TRESHOLD 2)_________ #
    left_join(SppDist,
              by = c("TaxonKey","INDEX")) %>%
    filter(Model == "SAU_C", # Only keeping cells where SAU catch exists
           Value > 0) %>%
    mutate(Model_Index = Model_Index*100) %>%
    select(-Model, -Value, -Latitude,-Longitude) %>%
    left_join(Index_Code,
              by = "INDEX")

  # MODEL INDEX (TRESHOLDS 1 & 2) #
  Model_Index_D <- Trans_Spp %>% 
    group_by(Territory,
             TaxonKey,
             Model_Index
    ) %>% 
    summarise(n_cells_spp = n()) %>% 
    select(-n_cells_spp)

  #____________ ESTIMATING DISTRIBUTION INDEX (TRESHOLD 3)_________ #
  # The number of species' cells present within each country's EEZ

  # if(n_Territory > 1){

  #Step 1.  Get EEZ id and Neighbour
  Neighbours_List <- Neighbours %>% 
    group_by(Territory,Neighbour) %>% 
    summarise(n=n()) %>% 
    ungroup() %>% 
    select(-n)

  # Step 2. Determines the amount of grids present in each country
  Spp_Grid <- Trans_Spp %>% 
    group_by(TaxonKey,
             Model_Index, # Un-comment after producing models x datasets
             Territory) %>% 
    summarise(n_spp_eez = length(unique(INDEX))) %>% 
    left_join(Neighbours_List,
              by = "Territory") %>% 
    filter(Territory %in% Trans_Spp$Territory, #Filter out unwanted Neighbours (those who don't have grids within but get included because they are Neighbours)
           Neighbour %in% Trans_Spp$Territory)

  # IF SPECIES DISCRETE AND APEARS IN ONLY 1 REGION OR > 1 REGION BUT NOT NEIGHBOURS #

  # Step 2.1. Determines discrete spp
  if(nrow(Spp_Grid) == 0 & Save == "Y"){

    #  Get the Territory where species happen
    Discrete_Territory <- unique(Model_Index_D$Territory)
    n_Territory <- length(Discrete_Territory)

    # Reads dataset of Non transboundary species
    Non_Transboundary_Data <- fread(paste(Results_Path,"Trans_Results/Non_Transboundary_Data.csv",sep=""))

    # Set new data
    Current_Spp <- tibble(TaxonKey = as.integer(Spp),
                          Territory = Discrete_Territory)

    # Combine new spp with others
    New_Non <- Non_Transboundary_Data %>% 
      bind_rows(Current_Spp)

    # Set a name for file and saves data
    New_Non_Name <- paste(Results_Path,"Trans_Results/Non_Transboundary_Data.csv",sep="")
    write.csv(New_Non, New_Non_Name, row.names = F)

    EstTransIndex=Current_Spp

  }

  # Step 3. Sum total grids per Neighbours
  if(nrow(Spp_Grid) > 0 & Save == "Y"){

    # Split dataframes to merge latter
    Territory_T <- Spp_Grid %>% 
      ungroup() %>% 
      select(
        Model_Index, # Un-comment after producing models x datasets
        TaxonKey,
        Name=Territory,
        n_spp_eez
      )

    Neighbour_T <- Spp_Grid %>% 
      ungroup() %>% 
      select(
        Model_Index,
        TaxonKey,n_spp_eez,
        Name=Neighbour,
        Territory
      )

    # Merge dataframes to get totals per Neighbourds
    Area_Index_D <- full_join(Territory_T,
                              Neighbour_T, 
                              by = c("Model_Index","Name","TaxonKey")
    ) %>%
      rowwise() %>%
      mutate(Spp_Total = sum(n_spp_eez.x,n_spp_eez.y,na.rm=T)) %>% # Total gridcelles per Neighbours
      distinct() %>% # Removes false duplicates from `full_join()`
      rename(Territory = Name,
             Neighbour =Territory,
             n_spp_Country = n_spp_eez.x,
             n_spp_Neighbour = n_spp_eez.y) %>% 
      mutate(Area_Index = n_spp_Country/Spp_Total) %>%  #Estimates the proportion of grids per country
      select(1:3,6,8)

    #____________ FINAL INDEX TABLE_________ #
    Indexes <- Model_Index_D %>% 
      full_join(Area_Index_D,
                by = c("TaxonKey","Territory","Model_Index")
      ) %>% 
      filter(!is.na(Area_Index),
             !is.na(Model_Index) # Uncomment after producing Models x datasets
      ) %>%  # Remove cases where the country does not pass the area_index and viceversa
      select(TaxonKey,Territory,Neighbour,everything())

    ### Save spp dataframe

    if(nrow(Indexes) > 0 & Save == "Y"){

      File_Name <- paste(Spp,"_Transboundary.csv",sep = "")
      Save_Path <- paste(Results_Path,"Trans_Results/",File_Name,sep="")

      write_csv(Indexes,
                Save_Path)

    }

    if(nrow(Indexes) > 0 & Save == "N"){
      EstTransIndex=Indexes # A dataframe with   
    }

    ### Function return
    EstTransIndex=Indexes # A dataframe with

  } # Closes Treshold (Step 3)
} # closes function

# __________________________________
# Test function
# __________________________________

# Spp <-  # Tuna Many RFMOs
# Model <- "All"
# Coord = CoorG
# #
# 
# # For EEZ
# Neighbours <- Neighbours_EEZ
# Index_Code <- EEZ_Index_Code
# 
# 
# EstTransIndex(Spp = 600107,
#               Mode = "All",
#               Neighbours = ,Coord,Index_Code,Save="Y")

```

### Function `Mclapply_Hack`

This function will hack the `mclappy()` function to work in PC and use multiple cores for the analysis. It was created by someone else and the original version is here: 
https://www.r-bloggers.com/implementing-mclapply-on-windows-a-primer-on-embarrassingly-parallel-computation-on-multicore-systems-with-r/

```{r Mclapply_Hack_Fun, eval= T , echo = T, warning = F, message = F}

# The hack

Mclapply_Hack <- function(...){
  ## Create a cluster
  size.of.list <- length(list(...)[[1]])

  cl <- makeCluster(min(size.of.list, n_cores))

  ## Find out the names of the loaded packages 
  loaded.package.names <- c(
    ## Base packages
    sessionInfo()$basePkgs,
    ## Additional packages
    names(sessionInfo()$otherPkgs))
  tryCatch( {

    ## Copy over all of the objects within scope to
    ## all clusters. 
    this.env <- environment()
    while( identical( this.env, globalenv() ) == FALSE ) {
      clusterExport(cl,
                    ls(all.names=TRUE, env=this.env),
                    envir=this.env)
      this.env <- parent.env(environment())
    }
    clusterExport(cl,
                  ls(all.names=TRUE, env=globalenv()),
                  envir=globalenv())

    ## Load the libraries on all the clusters
    ## N.B. length(cl) returns the number of clusters
    parLapply( cl, 1:length(cl), function(xx){
      lapply(loaded.package.names, function(yy) {
        require(yy , character.only=TRUE)})
    })

    ## Run the lapply in parallel 
    return( parLapply( cl, ...) )
  }, finally = {        
    ## Stop the cluster
    stopCluster(cl)
  })

  ## Warn the user if they are using Windows
  if( Sys.info()[['sysname']] == 'Windows' ){
    message(paste(
      "\n", 
      "   *** Microsoft Windows detected ***\n",
      "   \n",
      "   For technical reasons, the MS Windows version of mclapply()\n",
      "   is implemented as a serial function instead of a parallel\n",
      "   function.",
      "   \n\n",
      "   As a quick hack, we replace this serial version of mclapply()\n",
      "   with a wrapper to parLapply() for this R session. Please see\n\n",
      "     http://www.stat.cmu.edu/~nmv/2014/07/14/implementing-mclapply-on-windows \n\n",
      "   for details.\n\n"))
  }

  ## If the OS is Windows, set mclapply to the
  ## the hackish version. Otherwise, leave the
  ## definition alone. 
  mclapply <- switch( Sys.info()[['sysname']],
                      Windows = {Mclapply_Hack}, 
                      Linux   = {mclapply},
                      Darwin  = {mclapply})

}
## end mclapply.hack.R

```

### **Principal Routine**

```{r Trans_Spp_Control_Pannel, eval = F, echo = F}

# Set Jurisdiction to EEZ or RFMO
Jurisdiction <- "EEZ"

# Determine forlder for saving results
Results_Path <- paste(Data_Path,"Results/",Jurisdiction,"/",sep = "")

# Removes everything before the number
Step_One <- gsub(".*_","",list.files(paste(Data_Path,"/Distribution_Data/SAU_Distribution/",sep="")))

# Species List
Species_List <- gsub("\\..*","",Step_One)

# Set Spceies list
Exploited_Species <- fread(paste(Data_Path,"Distribution_Data/Species/exploited_species_list.csv",sep="")) %>%
  filter(TaxonKey %in% Species_List)

# Step to run function for species not computed
Computed_Spp <- as.integer(gsub("\\_.*","",list.files(Results_Path)))

Exploited_Species <- Exploited_Species %>%
  filter(!TaxonKey %in% Computed_Spp)

# Gabriel's coordinate system
CoorG <- read.csv(paste(Data_Path,"Spatial_Data/coordinates_gab.csv",sep="")) %>%
  mutate(INDEX = seq(1,259200,1))

# For Function
Exploited_Species_List <- Exploited_Species$TaxonKey

# Load Jursidiction-dependant data

# Load neighbours data from previouse routine
Neighbours_Data <- read_excel(paste(Data_Path,"Spatial_Data/EEZ_Neighbour_List.xlsx", sep =""))

# Modifications have to be made if ran un Hall
# R<e9>union (Réunion) and C<f4>te d'Ivoire (Côte d'Ivoire)

# SAU relations between INDEX and Country's EEZs
EEZIDs_List <- read_excel(paste(Data_Path,"Spatial_Data/V_Lam/Updated_EEZList_17June2016.xlsx",sep=""))
EEZ_CellID <- read_excel(paste(Data_Path,"Spatial_Data/V_Lam/EEZ_CellID.xlsx",sep=""))
colnames(EEZ_CellID) <- c("EEZID","INDEX")

Index_Code <- EEZIDs_List %>% 
  left_join(EEZ_CellID) %>% 
  rename(Territory = Name)

# Creates dummy dataset for Non transboundary species
write.csv(tibble(), paste(Results_Path,"Non_Transboundary_Data.csv",sep=""), row.names = F)

# Number of cores to us 
n_cores <- detectCores()

# Just a warning for me not to forget
print(paste("You are running the model for",nrow(Exploited_Species),"Species at the",Jurisdiction,"level and will be saved at",Results_Path))
```

```{r Trans_Spp_Routine, eval = F, echo = F}

# Run model in parallel
system.time(
  Dis_Trans <- bind_rows(
    mclapply(
      Exploited_Species_List,
      FUN = EstTransIndex,
      Model = "All", # all data sources
      Neighbours = Neighbours_Data,
      Coord = CoorG,  # Gabriels coordinate system for read species dist
      Index_Code = Index_Code, 
      Save = "Y",  # For saving
      Jurisdiction_Type = Jurisdiction # For saving in proper path
    )
  )
)

head(Dis_Trans)
# Data for Hall with 8 cores on RFMOs
# user    system   elapsed 
# 11375.652  3726.515  8342.296

# Data on time for the Beast with 30 cores
# user   system  elapsed 
# 8117.72  1991.91 10296.00

# Data on time for the Beast with 32 cores
#  user  system elapsed 
# 7831.10 1327.27 9261.48

```

## Methods for straddling stocks

1.- Get transboundary species list

1.1- Remove highly migratory species based on literature

2- High seas shapefile

2.1- FAO Areas ID

3.- Identiy what transboundary species are actually straddling

4.- Be happy, happy like a hippo

### Shapefiles

```{r}

# World land maps
world_land <- ne_countries(scale = 'medium', returnclass = c("sf")) %>% 
  st_transform(crs = 4326)# 4326

# ggplot(world_land) +
#   geom_sf()

# FAO regions
fao_regions <- my_sf("FAO") %>% 
  filter(F_LEVEL == "MAJOR")
st_crs(fao_regions) = 4326

# head(fao_regions)

# fao_regions %>% 
#   ggplot() +
#   geom_sf(aes(fill = F_AREA))

# EEZs
world_eez <- my_sf("SAU",simple = 1000)

# Lon-lat DBEM

# First get Vicky's data to get high seas
EEZ_CellID <- my_path("Spa","DBEM", "EEZ_CellID.xlsx", read = TRUE)
colnames(EEZ_CellID) <- c("EEZID","index")

# Now our gridcell
abny <- my_path("Spa","DBEM","Lon_Lat_DBEM.txt",read = T, header = F) %>% 
  bind_cols(my_path("Spa","DBEM","WArea.txt",read = T, header = F)) %>% 
  filter(!index %in% EEZ_CellID$INDEX) %>% 
  mutate(cat = ifelse(v1>0, "abnj","land")) %>% 
  select(-v1)

# SAVE FOR COMPLETE DBEM
complete_lon_lat_dbem <- my_path("Spa","DBEM","Lon_Lat_DBEM.txt",read = T, header = F) %>% 
  left_join(abny) %>% 
  mutate(cat = ifelse(is.na(cat), "eez",cat))

# Save for fouture
# write_csv(complete_lon_lat_dbem, my_path("Spa", "DBEM","complete_lon_lat_dbem.csv"))

  ggplot(complete_lon_lat_dbem) +
  geom_tile(
    aes(
      x = lon,
      y = lat,
      fill = cat
    )
  )

# All maps test
# ggplot() +
#   geom_sf(data = fao_regions, aes()) +
#   geom_sf(data = world_land, aes()) +
#   geom_sf(data = world_eez, aes())

  # First get Vicky's data to get high seas
EEZ_SAU <- my_path("Spa","DBEM", "Updated_EEZList_17June2016.xlsx", read = TRUE)
colnames(EEZ_SAU) <- c("sau_code","eez")

# First get Vicky's data to get high seas
EEZ_CellID <- my_path("Spa","DBEM", "EEZ_CellID.xlsx", read = TRUE)
colnames(EEZ_CellID) <- c("sau_code","index")

complete_sau <- EEZ_SAU %>% 
  left_join(EEZ_CellID)

#   
# coords %>% 
#   left_join(EEZ_CellID) %>% 
#     rename(sau_eez_id = EEZID) %>% 
#     gather("code_na","code",f_code:sau_eez_id) %>%

  # write_csv(Coord, my_path("Spa", "DBEM","complete_lon_lat_dbem.csv"))
```

### Determining straddling stocks

```{r GetSppDistFun}

GetSppDist=function(Spp,Model,Coord){

  Distpath <- paste(Data_Path,"Data/Distribution/",sep="")
  INDEX = seq(1,259200,1)
  # INDEX <- Coordinates$INDEX

  # SAU Distributions
  if(Model == "SAU_D"){
    File_Name <- paste("SAU_Distribution/DIST_GOOD_SP_")
  }

  # SAU Catch
  if(Model == "SAU_C"){
    File_Name <- paste("SAU_data_per_species/CATCH_SP_")
  }

  # Occurence
  if(Model == "Occ"){
    File_Name <- paste("Occurence/OCCURENCE_JULIANO_")
  }

  # ENM Model
  if(Model == "ENM"){
    File_Name <- paste("ENM/ENM_JULIANO_")
  }

  # DBEM Data
  # if(Model == "DBEM"){
  #   Distpath <- "/Volumes/DATA/JULIANO_NEYMAR/PristineSeasData
  #   Final_Path <- "ENM"
  #   File_Name <- paste("ENM_JULIANO",Spp,".mat",sep="")
  # }

  if(Model == "All"){

    Models_List <- c(paste(Distpath,"SAU_Distribution/DIST_GOOD_SP_",Spp,".mat",sep =""),
                     paste(Distpath,"Occurence/OCCURENCE_JULIANO_",Spp,".mat",sep=""),
                     paste(Distpath,"ENM/ENM_JULIANO_",Spp,".mat",sep="")
    )

    # Jumps species not modeled. NOTE: We only need one as ENM, Occ and SAU Dis have all the same 939 spp
    if(file.exists(Models_List[1])){

      Load <- lapply(Models_List, FUN=R.matlab::readMat, na.strings=0)

      sppdist <- as.data.frame(bind_cols(Load)) %>% 
        mutate(
          INDEX = INDEX,
          TaxonKey = Spp
        )
      colnames(sppdist) <- c("SAU_D","Occ","ENM","INDEX","TaxonKey")

      #### Step for SAU catch data that has to be averaged
      File_Name <- paste("SAU_data_per_species/CATCH_SP_")
      SppPath <- paste(Distpath,File_Name,Spp,".mat",sep="")

      # For now we're using only the last 10 years average, basicaaly if the species has been fished in any of these years, its considered present
      SAU_C_data <- as.data.frame(R.matlab::readMat(SppPath)) %>% 
        select(CATCH.1:CATCH.65) %>% 
        mutate(INDEX = INDEX) %>% 
        gather("Year","Catch",CATCH.56:CATCH.65) %>% # Last 10 years of data
        group_by(INDEX) %>% 
        summarise(SAU_C = mean(Catch,na.rm=T))

      # Join both tables
      sppdist <- sppdist %>% 
        left_join(SAU_C_data,
                  by = "INDEX") %>% 
        select(TaxonKey,INDEX,everything()) %>% 
        left_join(CoorG,
                  by = "INDEX")

      # Fix coordinate system incompatibility between Gab and DBEM
      sppdist <- suppressWarnings(sppdist[order(sppdist$latitude, rev(sppdist$longitude),decreasing=TRUE), ] %>% 
                                    mutate(INDEX = seq(1,259200,1)) %>% 
                                    gather("Model","Value",3:6) %>%
                                    mutate(Value = ifelse(is.na(Value), 0, Value)) # Converting NA's to ceros
      )

      getSppDist = sppdist

    }

  }else{

    # Merge paths
    SppPath <- paste(Distpath,File_Name,Spp,".mat",sep="")

    if(file.exists(SppPath) == TRUE){

      #Install (if needed) R.matlab package
      if(!require(R.matlab)){
        install.packages("R.matlab")
      }

      # Read Files

      sppdist <- as.data.frame(R.matlab::readMat(SppPath)) %>% 
        mutate(INDEX = INDEX,
               Species = Spp) %>% 
        left_join(Coord)

      # Fix coordinate system incompatibility between Gab and DBEM
      sppdist <- sppdist[order(sppdist$latitude, rev(sppdist$longitude),decreasing=TRUE), ] %>% 
        mutate(INDEX = seq(1,259200,1))

      # Return 
      getSppDist=sppdist

    }else{
      print(paste("No info for this species",Spp, "in",Model))
    }

  }
}
```

```{r StraddIndexFun}

StraddIndex <- function(Spp,Model = "All",Neighbours,Coord,Index_Code){

  # Get model data from spps
  SppDist <- GetSppDist(Spp,Model,Coord)

    # Result 1. Number of Countries that share the species

  #____________ ESTIMATING MODEL INDEX (TRESHOLD 1)_________ #
  Trans_Spp <- SppDist %>%
    filter(#INDEX %in% Neighbours$INDEX,# Filter data only located within EEZs
      Model != "SAU_C") %>% # Only using observational and modelled data
    mutate(Value = ifelse(Value > 0, 1,0)) %>%  
    filter(Value > 0) %>%
    group_by(TaxonKey,
             INDEX
    ) %>%
    summarise(Model_Index = sum(Value,na.rm=T)/3
    ) %>%
    filter(Model_Index > 0.4) %>% # at least 2 sources agree
    # ____________ ESTIMATING FUNDAMENTAL NICHE (TRESHOLD 2)_________ #
    left_join(SppDist,
              by = c("TaxonKey","INDEX")) %>%
    filter(Model == "SAU_C", # Only keeping cells where SAU catch exists
           Value > 0) %>%
    mutate(Model_Index = Model_Index*100) %>%
    select(-Model, -Value, -latitude,-longitude) %>%
    rename(index = INDEX) %>% 
    # joint eez names
    left_join(Index_Code,
              by = "index") %>% 
    # joint fao area names
    left_join(fao_areas,
              by = "index") %>% 
    mutate(
      territory = ifelse(!is.na(name),name,name_en),
      category = ifelse(!is.na(name),"eez","fao"),
    ) %>% 
    select(1:3,territory,category) %>% 
    clean_names() %>% 
    filter(model_index == 100)

  # Distributional test
  # test <- Trans_Spp %>% 
    # left_join(coords)

  # ggplot(test) +
  #   geom_tile(
  #     aes(
  #       x = lon,
  #       y = lat,
  #       fill = category
  #     )
  #   )

  # Step 2.1. Determines not a straddling stock
  if("fao" %in% Trans_Spp$category == FALSE){

    # output <- Spp_Grid %>% 
    #   select(
    #     taxon_key,
    #     territory,
    #     neighbour
    #   ) %>% 
    #   mutate(straddling = "no")
    return(print("not a straddling stock"))
    stop()

  }else{

    # MODEL INDEX (TRESHOLDS 1 & 2) #
    Model_Index_D <- Trans_Spp %>% 
      group_by(territory,
               category,
               taxon_key,
               model_index
      ) %>% 
      summarise(n_cells_spp = n()) %>% 
      select(-n_cells_spp)

    #____________ ESTIMATING DISTRIBUTION INDEX (TRESHOLD 3)_________ #
    # The number of species' cells present within each country's EEZ

    # if(n_Territory > 1){

    #Step 1.  Get EEZ id and Neighbour
    Neighbours_List <- neighbours %>%
      group_by(territory,neighbour) %>%
      summarise(n=n()) %>%
      ungroup() %>%
      select(-n) %>% 
      left_join(fao_to_sau) %>% 
      filter(!is.na(territory))

    # Set FAO neighbours... 
    fao_neighbours <- Neighbours_List %>% 
      select(territory,fao_area_name) %>% 
      rename(territory = fao_area_name,
             neighbour = territory) %>% 
      group_by(territory, neighbour) %>% 
      summarise(n())

    fao_territory <- fao_neighbours %>% 
      rename(territory = neighbour,
             neighbour = territory)

    neighbours_list_fao <- fao_neighbours %>% 
      bind_rows(Neighbours_List,fao_territory) %>% 
      select(-fao_area_name)

    # Step 2. Determines the amount of grids present in each country
    Spp_Grid <- Trans_Spp %>% 
      group_by(taxon_key,
               model_index, # Un-comment after producing models x datasets
               territory,
               category) %>% 
      summarise(n_spp_eez = length(unique(index))) %>% 
      left_join(neighbours_list_fao,
                by = "territory") %>% 
      filter(territory %in% Trans_Spp$territory, #Filter out unwanted Neighbours (those who don't have grids within but get included because they are Neighbours)
             neighbour %in% Trans_Spp$territory)

    # Split dataframes to merge latter
    Territory_T <- Spp_Grid %>% 
      filter(category == "eez") %>% 
      ungroup() %>% 
      select(
        model_index, # Un-comment after producing models x datasets
        taxon_key,
        Name=territory,
        n_spp_eez
      )

    Neighbour_T <- Spp_Grid %>%
      filter(category == "eez") %>% 
      ungroup() %>% 
      select(
        model_index,
        taxon_key,
        n_spp_eez,
        territory,
        Name=neighbour,
        territory
      )

    fao_grid <- Spp_Grid %>% 
      filter(category == "fao") %>% 
      left_join(neighbours_list_fao) %>% 
      select(Name =neighbour,
             fao_name = territory,
             n_spp_eez.z = n_spp_eez)

    # Merge dataframes to get totals per Neighbourds
    Area_Index_D <- full_join(Territory_T,
                              Neighbour_T, 
                              by = c("model_index","Name","taxon_key")
    ) %>%
      left_join(fao_grid) %>% 
      filter(!is.na(fao_name)) %>% # remove non-straddling stocks
      rowwise() %>%
      mutate(spp_total = sum(n_spp_eez.x,n_spp_eez.y,n_spp_eez.z,na.rm=T)) %>% # Total gridcelles per Neighbours
      distinct() %>% # Removes false duplicates from `full_join()`
      rename(territory = Name,
             neighbour = territory,
             n_spp_country = n_spp_eez.x,
             n_spp_neighbour = n_spp_eez.y,
             n_spp_fao_area = n_spp_eez.z) %>%
      # Estimate if it is straddling
      mutate(stradling_index = (n_spp_fao_area/spp_total)*100)

    ### Save spp dataframe

    output <- Area_Index_D %>% 
      select(taxon_key,
             territory,
             neighbour,
             fao_name,
             stradling_index)

    File_Name <- paste(Spp,"_straddling.csv",sep = "")
    Save_Path <- my_path("R","EEZ/Stradling_Results/", File_Name)

    # Save_Path <- paste(Results_Path,"Trans_Results/",File_Name,sep="")

    write_csv(output,
              Save_Path)
  }
} # closes function

 #Test me  
# suppressMessages(
#   StraddIndex(600069, Model = "All", Coord = coords, Index_Code = Index_Code)
# )

```

### Straddling Control pannel

```{r}
# DBEM coordinate system
# complete_lat_lon <- my_path("Spa", "DBEM","complete_lon_lat_dbem.csv", read = T)

# Gabriel's coordinate system
CoorG <- my_path("G","FishForVisa/Data/Spatial/","coordinates_gab.csv",read =T) %>%
  mutate(INDEX = seq(1,259200,1))

# Get fao areas on DBEM grid
fao_areas <- my_path("Spa", "DBEM","complete_lon_lat_dbem.csv", read = T) %>% 
  filter(cat == "abnj")

# Fao combo
fao_to_sau <- read_excel("/Volumes/DATA/JULIANO_NEYMAR/Spatial/SAU/EEZ-FAO-Country-Combo.xlsx") %>% 
  select(territory = eez_name,eez_id,fao_area_name)

# Neighboring dataset
neighbours <- read_excel("/Volumes/DATA/JULIANO_NEYMAR/FishForVisa/Data/Spatial/EEZ_Neighbour_List.xlsx") %>% 
  clean_names() %>% 
  rename(eez_id = eezid) %>% 
  left_join(fao_to_sau,
            by = "eez_id") %>% 
  select(territory = territory.x,neighbour)

# Load FishForVisa Results
transboundary_spp <- read_csv("https://media.githubusercontent.com/media/jepa/FishForVisa/master/Data/Results/Clean_Results_Trans.csv") %>% 
  # Set same filters as in FishForVisa
  filter(
    area_index >= 0.25,
    area_index <= 1-0.25,
    model_index >= 100
  )

# DepPelingo

depth_profile <- my_path("G","FishForVisa/Data/Distribution/Species","SppTaxonName.csv", read = T)

# Double check we have 633 species
length(unique(transboundary_spp$taxon_key)) # 633

coords <- my_path("Spa","DBEM","Lon_Lat_DBEM.txt",read = T, header = F)

Data_Path <- "/Volumes/DATA/JULIANO_NEYMAR/FishForVisa/" #x

# SAU relations between INDEX and Country's EEZs
# First get Vicky's data to get high seas
EEZIDs_List <- my_path("Spa","DBEM", "Updated_EEZList_17June2016.xlsx", read = T) #x
EEZ_CellID <- my_path("Spa","DBEM", "EEZ_CellID.xlsx", read = TRUE) #x
colnames(EEZ_CellID) <- c("eezid","index") #x

Index_Code <- EEZIDs_List %>%  #x
  left_join(EEZ_CellID)

```

```{r run_me, eval = T}

suppressWarnings(
  suppressMessages(
    lapply(unique(transboundary_spp$taxon_key), StraddIndex, Model = "All", Coord = coords,
           Index_Code = Index_Code)
  )
)

```

# Results

## Functions needed

### Function `GetTransResults`

This function simply reads in the files and import them all as on large d.f

```{r Fun_GetTransResults, eval = F}

GetTransResults=function(Spp, type = "NA"){

  if(type == "strad"){
    # Set the path for each file
  Distpath <- paste(my_path("R","EEZ/Straddling_Results/"),Spp,"_straddling.csv",sep="")

  }else{
  # Set the path for each file
  Distpath <- paste(Results_Path,"Trans_Results/",Spp,sep="")

  }

  # Loads all files in a df
  Load_Data <- bind_rows(lapply(Distpath, FUN=fread))
  return(Load_Data)

}

```

### Function `CleanResults`

There are quite a few inconsistencies between the names of territories within the SAU shape file, the distribution data and the UN database. The following function cleans up the names, some of them manually some of the automatic. It also removes the "transboundary nature" of those species shared by two territories that belong to the same nation (e.g. Brazil mainland and Brazil (Fernando de Noronha) or US Gulf of Mexico and US East coast). Finally, there is an annoying thing coding issue between MAC and Windows related to characters so I have to also fix for that, manually.

```{r funCleanResults, eval = F, echo = F}

# Requires Clean_sau_eez_fes
# Data <- Results_Trans
CleanResults <- function(Data,names_data){

  clean_names_data <- names_data %>% 
    select(
      eez_id,
      fishing_entity_id,
      fishing_entity = sau_fishing_entity,
      eez_name = sf_eez_name
    )

  # Clean out countries that share with their own borders (e.g. Brazil mainland and Brazil (Fernando de Noronha))
  Auto_Remove_Territories <- Data %>% 
    filter(!is.na(territory),
           territory != "NA",
           !is.na(neighbour),
           neighbour != "NA") %>% 
    rename(eez_name=territory) %>% 
    left_join(clean_names_data,
              by = "eez_name") %>% 
    select(taxon_key,
           territory = eez_name,
           eez_name = neighbour, 
           model_index,
           area_index,
           ct_fishing_entity= fishing_entity) %>% 
    left_join(clean_names_data,
              by = "eez_name") %>% 
    select(1,2, 
           neighbour = eez_name,
           model_index:ct_fishing_entity,
           nt_fishing_entity= fishing_entity,
           everything()) %>% 
    mutate(
      ct_fishing_entity = ifelse(is.na(ct_fishing_entity),paste(territory),paste(ct_fishing_entity)),
      nt_fishing_entity = ifelse(is.na(nt_fishing_entity),paste(neighbour),paste(nt_fishing_entity)),
      Same_Nation = ifelse(ct_fishing_entity==nt_fishing_entity,"Yes","No")
    ) %>% 
    filter(neighbour != "NA")

  #### Manual removal of those left...

  # List of territories to eliminate (No other neighobors but mainland)
  Islands_to_Remove <- c(
    "Brazil (Fernando de Noronha)",
    "Brazil (St Paul and St. Peter Archipelago)",
    "Hawaii Main Islands (USA)",
    "Hawaii Northwest Islands (USA)",
    "Juan Fernandez Islands (Chile)",
    "Kermadec Isl. (New Zealand)"
  )

  # Complete list of territories to remove from dataset (Including islands)
  Territories_to_Remove <- Data %>% 
    filter(
      territory %in% Islands_to_Remove |
        neighbour %in% Islands_to_Remove |
        territory == "Balearic Island (Spain)" & neighbour == "Spain (Mediterranean and Gulf of Cadiz)" |
        territory == "British Virgin Isl. (UK)" & neighbour == "Anguilla" |
        neighbour == "British Virgin Isl. (UK)" & territory == "Anguilla" |
        territory == "Corsica (France)" & neighbour == "France (Mediterranean)" |
        neighbour == "Corsica (France)" & territory == "France (Mediterranean)" |
        territory == "Crete (Greece)" & neighbour == "Grece" |
        neighbour == "Crete (Greece)" & territory == "Grece" |
        territory == "Cyprus (North)" & neighbour == "Cyprus (South)" |
        neighbour == "Cyprus (North)" & territory == "Cyprus (South)" |
        territory == "Channel Isl. (UK)" & neighbour == "United Kingdom" |
        neighbour == "Channel Isl. (UK)" & territory == "United Kingdom" |
        territory == "Glorieuse Islands (France)" & neighbour == "Mayotte (France)" |
        neighbour == "Glorieuse Islands (France)" & territory == "Mayotte (France)" |
        territory == "Guam (USA)" & neighbour == "Northern Marianas (USA)" |
        neighbour == "Guam (USA)" & territory == "Northern Marianas (USA)" |
        territory == "Italy" & neighbour == "Sardinia (Italy)" |
        neighbour == "Italy" & territory == "Sardinia (Italy)" |
        territory == "Italy" & neighbour == "Sicily (Italy)" |
        neighbour == "Italy" & territory == "Sicily (Italy)" |
        territory == "Japan (Daito Islands)" & neighbour == "Japan (main islands)"|
        neighbour == "Japan (Daito Islands)" & territory == "Japan (main islands)"|
        territory == "Japan (Daito Islands)" & neighbour == "Japan (Ogasawara Islands)"|
        neighbour == "Japan (Daito Islands)" & territory == "Japan (Ogasawara Islands)"|
        territory == "Norway" & neighbour == "Svalbard Isl. (Norway)" |
        neighbour == "Norway" & territory == "Svalbard Isl. (Norway)" |
        territory == "Sicily (Italy)" & neighbour == "Sardinia (Italy)" |
        neighbour == "Sicily (Italy)" & territory == "Sardinia (Italy)" |
        territory == "South Africa (Atlantic Coast)" & neighbour == "South Africa (Indian Ocean Coast)" |
        neighbour == "South Africa (Atlantic Coast)" & territory == "South Africa (Indian Ocean Coast)" |
        territory == "USA (Alaska, Arctic)" & neighbour == "USA (Alaska, Subarctic)" |
        neighbour == "USA (Alaska, Arctic)" & territory == "USA (Alaska, Subarctic)"
    ) %>%
    group_by(territory,neighbour) %>% 
    summarise(n())

  # Cleaned results 
  Clean_Results <- Auto_Remove_Territories %>% 
    anti_join(Territories_to_Remove,
              by = c("territory","neighbour")
    ) %>% 
    filter(Same_Nation == "No") %>% 
    select(taxon_key,
          eez_name = territory,
          eez_neighbour = neighbour,
          model_index,
          area_index
          )

  return(Clean_Results)
}

```

### Function `Figure CreateFigureOne`

This function will create figure one for both catch or value

```{r fun_Figure_1, eval =F, echo =F}

CreateFigureOne <- function(Variable_Selected){

  if(Variable_Selected == "catch"){
    Map_Label <- "Catch (Thousand Tonnes)"
    Plot_Name <- "Catch_plot.png"
  }else{
    Map_Label <- "Revenue (Million USD)"
    Plot_Name <- "Rev_plot.png"
  }

  # Filter vatch/valye data for the variable selecter (Map)
  Transboundary_Data_Plot <- Value_Trans_Country %>% 
    filter(Variable %in% Variable_Selected)

  # Total catch/Revenue shapefile
  World_sf_si_land <- World_land_sf %>%
    # st_transform(crs = 4326) %>% # 4326
    st_transform(crs = "+proj=eck4") %>%
    st_simplify(preserveTopology = TRUE, dTolerance = 0.1) %>% 
    left_join(Value_Trans_Country, by ="fishing_entity") %>% 
    filter(Variable %in% Variable_Selected)

  # Get the NA's to be grey
  Grey_land <- World_land_sf %>%  
    filter(!fishing_entity %in% World_sf_si_land$fishing_entity)

  Bins <- Value_Trans_Country %>% 
    filter(Variable %in% Variable_Selected) %>% 
    arrange(Bins) %>% 
    pull(Bins) %>% 
    unique()

  # Labels_Trans <- c("(1-29]",paste(Bins_Trans[2:5]))
  Labels_Trans <- c("(1-29]","(29-57]","(57-85]","(85-113]","(113-141]")
  # Labels <- c("(1-10]",paste(Bins[2:5]))
  Labels <- c("(1-10]","(10-100]","(100-1000]","(1000-5000]","(5000-12000]")

  Map <- ggplot() +
    geom_sf(data = World_sf_si_land,
            aes(fill = reorder(Bins_Plot,Value)), 
            colour = "grey80",
            size = 0.1) +
    geom_sf(data = World_sf_si,
            aes(fill =  reorder(Bins_Plot,n_trans_spp)),
            colour = "grey20",
            size = 0.1) + # n_trans_spp or Bins
    geom_sf(data = Grey_land,aes(fill = Bins),fill = "grey80", colour = "grey50",size = 0.1) + # n_trans_spp or Bins
    scale_fill_manual("",values = c("#5BBCD6",#"#5BBCD6", # blue
                                    "#00A08A",#"#00A08A", #green
                                    "#F2AD00",#"#F2AD00", # Yellow
                                    "#F98400",#"#F98400", #Orange
                                    "#FF0000"#,"#FF0000" #Red
    ),
    na.translate = F,
    # labels = comma
    ) +
    ggtheme_map() +
    # Transboundary labels
    annotate("text", 
             # x =(c(-85.5,-30,25,81,136)+5),
             # y = -93.5,
             x =(c(-8500000,-3000000,2500000,7500000,12800000)),
             y = -9050000,
             label = Labels_Trans, 
             color="black", 
             size=3, 
             angle=0, 
             # fontface="bold",
             hjust=0) +
    # Revenue labels
    annotate("text",
             # x =(c(-85.5,-31,20,75,131)+5),
             # y = -110,
             x =(c(-8500000,-3200000,2000000,7000000,12500000)),
             y = -11000000,
             label = Labels,
             color="black",
             size=3,
             angle=0,
             # fontface="bold",
             hjust=0) +
    # Legend Titles
    annotate("text", 
             # x =-180,
             # y = -93.5,
             x =-17000000,
             y = -9050000,
             label = "Transboundary Spp. (n)", 
             color="black", 
             size=3, 
             angle=0, 
             fontface="bold",
             hjust=0) +
    annotate("text", 
             # x =-180,
             # y = -110,
             x =-17000000,
             y = -11000000,
             label = Map_Label,
             color="black", 
             size=3, 
             angle=0, 
             fontface="bold",
             hjust=0) +
    theme(
      legend.box.margin = margin(t = -58, l = 120),
      legend.text = element_blank(),
    ) +
    guides(fill = guide_legend(nrow = 1))

  ####___________________________###
  ### Now create the Two circluar bar plots
  ####___________________________###

  ### Labels

  # Label info Rev
label_data_rev <- Subregion_Trans_Value %>% 
  filter(Variable == "value") %>% 
  arrange(Value_sum) %>% # soo it looks like a nice snail :)
  mutate(
    id = seq(1,number_of_bar),
    angle = 90 -360 * (id-0.5)/number_of_bar,
    hjust = ifelse(angle < -90,1,0),
    angleb = ifelse(angle < -90, (angle +180), angle)
  ) %>% 
  mutate(
    sub_region = ifelse(str_detect(sub_region,"Latin"),"Ltn. Ame. &\nthe Car.",
                        ifelse(sub_region =="Northern America","Northern\nAmerica",
                               ifelse(sub_region =="Sub-Saharan Africa","Sub-Saharan\nAfrica",
                               paste(sub_region))
                        )
    )#,
    # sub_region_n = paste(sub_region," (n=",Trans_Entity_n_sum,")",sep = "")
  )

# Label info Catch
label_data_catch <- Subregion_Trans_Value %>% 
  filter(Variable == "catch") %>% 
  arrange(Value_sum) %>% # soo it looks like a nice snail :)
  mutate(
    id = seq(1,number_of_bar),
    angle = 90 -360 * (id-0.5)/number_of_bar,
    hjust = ifelse(angle < -90,1,0),
    angleb = ifelse(angle < -90, (angle +180), angle)
  ) %>% 
  mutate(
    sub_region = ifelse(str_detect(sub_region,"Latin"),"Ltn. Ame. &\nthe Car.",
                        ifelse(sub_region =="Northern America","Northern\nAmerica",
                               ifelse(sub_region =="Sub-Saharan Africa","Sub-Saharan\nAfrica",
                                      paste(sub_region))
                        )
    )
  )

  # Revenue Plot

  # Set plot limits

  max_Rev_lim <- Subregion_Trans_Value %>%
    filter(Variable == "value") %>%
    group_by() %>%
    summarise(max(Value_sum)) %>%
    pull()

  Subregion_Rev_Data <- Subregion_Trans_Value %>% 
    filter(Variable == "value")

  Rev_Plot <- ggplot(
    # Regional plots
    data = Subregion_Rev_Data,
    aes(
      x=reorder(id,Value_sum),
      y=Value_sum, 
      fill=Value_sum
    )
  ) +    
    ylim(-1500, max_Rev_lim) +
    geom_bar(stat="identity") +
    # Regional averages
    geom_point(data = Subregion_Rev_Data,
               aes(
                 x=reorder(id,Value_sum),
                 y=Value_m
               ),
               colour = "black",
               alpha = 0.5
    ) +
    # Regional sds
    geom_errorbar(data = Subregion_Rev_Data,
                  aes(
                    ymin=ifelse(Value_m-Value_sd <= 0,0,Value_m-Value_sd), # sets lower to 0
                    ymax=Value_m+Value_sd
                  ), 
                  width=.2,
                  position=position_dodge(.9),
                  alpha = 0.5) +
    # Looking good
    theme_minimal() +
    theme(
      legend.position = c(0.73,0.60),
      axis.text = element_blank(),
      axis.title = element_blank(),
      panel.grid = element_blank(),
      plot.margin = margin(0,-10,-6,-10,"cm")
    )  +
    # Base line
    # geom_segment(aes(x = 1, y = 1, xend = 17, yend = 17),  
    #              colour = "black", size=0.1, inherit.aes = FALSE) +
    coord_polar(start = 0) + 
    geom_text(data=label_data_rev, 
              aes(x=id,
                  y=Value_sum,
                  label=sub_region,
                  hjust=hjust), 
              color="black",
              fontface="bold",
              size=2, 
              angle= label_data_rev$angleb, 
              inherit.aes = FALSE) +
    scale_fill_gradientn("",
                         colours = pal,
                         limits = c(0,14100),
                         breaks = seq(0,14000,3500)
    ) +
    guides(fill = guide_colorbar(label.position = "left")
    );Rev_Plot

  ###____________###
  ### Catch plot
  ###____________###

  max_Catch_lim <- Subregion_Trans_Value %>% 
    filter(Variable == "catch") %>% 
    group_by() %>% 
    summarise(max(Value_sum)) %>% 
    pull()

  Subregion_Catch_Data <- Subregion_Trans_Value %>% 
    filter(Variable == "catch")

  Catch_Plot <- ggplot(data = Subregion_Catch_Data,
                       aes(x=reorder(id,Value_sum),
                           y=Value_sum, 
                           fill=Value_sum
                       )
  ) +    
    ylim(-1.5, max_Catch_lim) +
    geom_bar(stat="identity") +
    # Regional averages
    geom_point(data = Subregion_Catch_Data,
               aes(
                 x=reorder(id,Value_sum),
                 y=Value_m
               ),
               colour = "black",
               alpha = 0.5
    ) +
    # Regional sd's
    geom_errorbar(data = Subregion_Catch_Data,
                  aes(
                    ymin=ifelse(Value_m-Value_sd <= 0,0,Value_m-Value_sd),
                    ymax=Value_m+Value_sd
                  ), 
                  width=.2,
                  position=position_dodge(.9),
                  alpha = 0.5) +
    theme_minimal() +
    theme(
      legend.position = c(0.1,0.60),
      axis.text = element_blank(),
      axis.title = element_blank(),
      panel.grid = element_blank(),
      plot.margin = margin(0,-10,-6,-5,"cm")
    )  +
    # Base line
    # geom_segment(aes(x = 0, y = 1, xend = 17, yend = 17),
    #              colour = "black", size=0.1, inherit.aes = FALSE) +
    coord_polar(start = 0) + 
    geom_text(data=label_data_catch, 
              aes(x=id,
                  y=ifelse(Value_sum >= 12,12,Value_sum+0.1), # Fix names overlap
                  label=sub_region,
                  hjust=hjust), 
              color="black",
              fontface="bold",
              size=2, 
              angle= label_data_catch$angleb, 
              inherit.aes = FALSE) +
    scale_fill_gradientn("",
                         limits = c(0,12),
                         breaks = seq(0,12,3),
                         colours = pal
    );Catch_Plot

  ## Joining everything in one ploit

  if(Variable_Selected == "value"){

    plot_all <- ggdraw() +
      # map
      draw_plot(Map, x = 0, y = 0.45, width = 1, height = 0.58) +
      # Revenue circular
      draw_plot(Rev_Plot, x = 0.1, y = 0.08, width = 0.5, height = 0.4) +
      # Catch circular
      draw_plot(Catch_Plot, x = 0.45, y = 0.08, width = 0.5, height = 0.4) +
      # Labeks
      draw_plot_label(label = c("a", "b"), 
                      size = 20,
                      x = c(0, 0), 
                      y = c(1, 0.45)
      ) +
      draw_plot_label(label = c("Revenue (USD) | Catch (Tonnes)","(Million)"), 
                      size = 8,
                      x = c(0.325,0.485), 
                      y = c(0.31,0.29)
      )

    gc()
    save_plot(
      # paste(Figures_Path,"100_25_Bins_Value_Region.png",sep=""),
      Plot_Name,
      plot_all,
      base_height = 8,
      base_width = 8
    )

  }else{
    plot_all <- ggdraw() +
      # map
      draw_plot(Map, x = 0, y = 0, width = 1, height = 1) #+
      # Revenue circular
      # draw_plot(Rev_Plot, x = 0.1, y = 0.08, width = 0.5, height = 0.4) +
      # # Catch circular
      # draw_plot(Catch_Plot, x = 0.45, y = 0.08, width = 0.5, height = 0.4) +
      # # Labeks
      # draw_plot_label(label = c("A", "B"), 
      #                 size = 20,
      #                 x = c(0, 0), 
      #                 y = c(1, 0.45)
      # ) +
      # draw_plot_label(label = c("Revenue (USD) | Catch (Tons)","(Million)"), 
      #                 size = 8,
      #                 x = c(0.330,0.485), 
      #                 y = c(0.31,0.29)
      # )

    save_plot(
      # paste(Figures_Path,"100_25_Bins_Value_Region.png",sep=""),
      Plot_Name,
      plot_all,
      base_height = 5,
      base_width = 8
    )
  }

} # End of function
```

### Standarization of names

```{standarize_names, eval = F, echo = F}

### --------------------- ###
# Sea Around Us eez to fishing entity
Clean_sau_eez_fes <- my_path("G","Spatial/SAU/","EEZ_FishingEntityID_v2_9June2020.xlsx", read =T)
### --------------------- ###

### --------------------- ###
# UNited Nations Dataset
### --------------------- ###
UN_Regions <- read.csv(paste(my_path("D"),"Spatial/UN_Regions.csv",sep=""), header = TRUE) %>% 
  clean_names() %>% 
  # mutate(un_fishing_entity = country_or_area) %>% 
  group_by(#un_fishing_entity,
    fishing_entity = country_or_area,
    georegion = region_name,
    sub_region = sub_region_name,
    country_or_area) %>% 
  summarise(n=n()) %>% #merge china's 3 rows
  select(-n) %>% 
  filter(fishing_entity != "") %>%  # there are three rows with "developing" trend as in, the code
  ungroup() %>% 
  # Change names to match SAU data and for shorter versions
  mutate(
    fishing_entity = gsub("","",fishing_entity),
    fishing_entity = gsub("British Indian Ocean Territory","Brit. Indian Ocean Terr. (UK)",fishing_entity),
    fishing_entity = gsub("Bermuda","Bermuda (UK)",fishing_entity),
    fishing_entity = gsub("British Virgin Islands","British Virgin Isl. (UK)",fishing_entity),
    fishing_entity = gsub("Cayman Islands","Cayman Isl. (UK)",fishing_entity),
    fishing_entity = gsub("Christmas Island","Christmas Isl. (Australia)",fishing_entity),
    fishing_entity = gsub("Mayotte","Mayotte (France)",fishing_entity),
    fishing_entity = gsub("Guadeloupe","Guadeloupe (France)",fishing_entity),
    fishing_entity = gsub("Guam","Guam (USA)",fishing_entity),
    fishing_entity = gsub("Martinique","Martinique (France)",fishing_entity),
    fishing_entity = gsub("Montserrat","Montserrat (UK)",fishing_entity),
    fishing_entity = gsub("New Caledonia","New Caledonia (France)",fishing_entity),
    fishing_entity = gsub("Niue","Niue (New Zealand)",fishing_entity),
    fishing_entity = gsub("Norfolk Island","Norfolk Isl. (Australia)",fishing_entity),
    fishing_entity = gsub("Northern Mariana Islands","North Marianas (USA)",fishing_entity),
    fishing_entity = gsub("Pitcairn","Pitcairn (UK)",fishing_entity),
    fishing_entity = gsub("Puerto Rico","Puerto Rico (USA)",fishing_entity),
    fishing_entity = gsub("Saint Helena","Saint Helena (UK)",fishing_entity),
    fishing_entity = gsub("Anguilla","Anguilla (UK)",fishing_entity),
    fishing_entity = gsub("Svalbard and Jan Mayen Islands","Svalbard Isl. (Norway)",fishing_entity),
    fishing_entity = gsub("Tokelau","Tokelau (New Zealand)",fishing_entity),
    fishing_entity = gsub("Turks & Caicos Islands","Turks & Caicos Isl. (UK)",fishing_entity),
    fishing_entity = gsub("United States Virgin Islands","US Virgin Isl.",fishing_entity),
    fishing_entity = gsub("Wallis and Futuna Islands","Wallis & Futuna Isl. (France)",fishing_entity),
    fishing_entity = gsub("Saint Barthélemy","St Barthelemy (France)",fishing_entity),
    fishing_entity = gsub("Iran (Islamic Republic of)","Iran",fishing_entity),
    fishing_entity = gsub("Bonaire","Bonaire (Netherlands)",fishing_entity),
    fishing_entity = gsub("Aruba","Aruba (Netherlands)",fishing_entity),
    fishing_entity = gsub("Curaçao","Curacao",fishing_entity),
    fishing_entity = gsub("United Kingdom of Great Britain and Northern Ireland","United Kingdom",fishing_entity),
    fishing_entity = gsub("United States of America", "USA",fishing_entity),
    fishing_entity = gsub("Democratic People's Republic of Korea", "Korea (North)",fishing_entity),
    fishing_entity = gsub("Republic of Korea", "Korea (South)",fishing_entity),
    fishing_entity = gsub("United Republic of Tanzania", "Tanzania",fishing_entity),
    fishing_entity = gsub("Cabo Verde", "Cape Verde",fishing_entity),
    fishing_entity = gsub("Timor-Leste", "Timor Leste",fishing_entity),
    fishing_entity = gsub(" and ", " & ",fishing_entity),
    # fishing_entity = gsub("Taiwan", "Taiwan (China)",fishing_entity),
    fishing_entity = gsub("Côte d’Ivoire", "Côte d'Ivoire",fishing_entity),
    fishing_entity = gsub("Vietnam", "Viet Nam",fishing_entity),
    fishing_entity = gsub("Solomon Islands", "Solomon Isl.",fishing_entity),
    # fishing_entity = gsub("Marshall Islands", "Solomon Isl.",fishing_entity),
    fishing_entity = gsub("Congo","Congo, R. of",fishing_entity),
    # fishing_entity = gsub("Réunion","Réunion (France)",fishing_entity),
  ) %>%
  # Some of them do not work with gsub
  mutate(
    fishing_entity = ifelse(fishing_entity == "Venezuela (Bolivarian Republic of)","Venezuela",
                            ifelse(fishing_entity == "Democratic Republic of the Congo, R. of","Congo (ex-Zaire)",
                                   ifelse(fishing_entity == "Iran (Islamic Republic of)","Iran",
                                          ifelse(fishing_entity == "Saint Martin (French Part)","St Martin",
                                                 ifelse(fishing_entity == "Cocos (Keeling) Islands","Cocos (Keeling) Isl. (Australia)",
                                                        ifelse(fishing_entity == "Falkland Islands (Malvinas)","Falkland Isl. (UK)",
                                                               ifelse(fishing_entity =="Micronesia (Federated States of)","Micronesia",
                                                                      ifelse(fishing_entity == "Saint Pierre & Miquelon","Saint Pierre & Miquelon (France)",
                                                                             ifelse(fishing_entity == "Sint Maarten (Dutch part)","Sint Maarten",
                                                                                    paste(fishing_entity)
                                                                             )
                                                                      )
                                                               )
                                                        )
                                                 )
                                          )
                                   )
                            )
    )
  ) %>%
  filter(fishing_entity != "Solomon Isl." | sub_region != "Micronesia")

### --------------------- ###
# Natural Earth Shapefile
### --------------------- ###

World_land_sf <- rnaturalearth::ne_countries(scale = 'medium', returnclass = c("sf")) %>% 
  st_transform(crs = "+proj=eck4") %>%
  # st_transform(crs = 4326) %>% # 4326
  rename(fishing_entity = name_long) %>% 
  mutate(
    fishing_entity = gsub(" and ", " & ",fishing_entity),
    fishing_entity = gsub("Anguilla", "Anguilla (UK)",fishing_entity),
    fishing_entity = gsub("American Samoa (USA of America)", "American Samoa",fishing_entity),
    fishing_entity = gsub("Antigua & Barb.", "Antigua & Barbu",fishing_entity),
    fishing_entity = gsub("Aruba","Aruba (Netherlands)",fishing_entity),
    fishing_entity = gsub("Bermuda","Bermuda (UK)",fishing_entity),
    fishing_entity = gsub("Bonaire","Bonaire (Netherlands)",fishing_entity),
    fishing_entity = gsub("British Indian Ocean Territory","Brit. Indian Ocean Terr. (UK)",fishing_entity),
    fishing_entity = gsub("British Virgin Islands","British Virgin Isl. (UK)",fishing_entity),
    fishing_entity = gsub("Cayman Islands","Cayman Isl. (UK)",fishing_entity),
    fishing_entity = gsub("Christmas Island","Christmas Isl. (Australia)",fishing_entity),
    fishing_entity = gsub("Curaçao","Curacao",fishing_entity),
    fishing_entity = gsub("Democratic Republic of the Congo", "Congo (ex-Zaire)",fishing_entity),
    fishing_entity = gsub("Dem. Rep. Korea", "Korea (North)",fishing_entity),
    fishing_entity = gsub("Falkland Islands","Falkland Isl. (UK)",fishing_entity),
    fishing_entity = gsub("Faeroe Islands","Faeroe Isl. (Denmark)",fishing_entity),
    fishing_entity = gsub("Federated States of Micronesia", "Micronesia",fishing_entity),
    fishing_entity = gsub("Guam","Guam (USA)",fishing_entity),
    fishing_entity = gsub("Northern Mariana Islands","North Marianas (USA)",fishing_entity),
    fishing_entity = gsub("Marshall Islands", "Marshall Isl.",fishing_entity),
    fishing_entity = gsub("Montserrat", "Montserrat (UK)",fishing_entity),
    fishing_entity = gsub("New Caledonia", "New Caledonia (France)",fishing_entity),
    fishing_entity = gsub("Niue", "Niue (New Zealand)",fishing_entity),
    fishing_entity = gsub("Norfolk Island", "Norfolk Isl. (Australia)",fishing_entity),
    fishing_entity = gsub("Pitcairn Islands","Pitcairn (UK)",fishing_entity),
    fishing_entity = gsub("Puerto Rico", "Puerto Rico (USA)",fishing_entity),
    fishing_entity = gsub("Republic of Congo", "Congo, R. of",fishing_entity),
    fishing_entity = gsub("Republic of Korea","Korea (South)",fishing_entity),
    fishing_entity = gsub("Saint Helena","Saint Helena (UK)",fishing_entity),
    fishing_entity = gsub("Saint-Martin","St Martin",fishing_entity),
    fishing_entity = gsub("Saint-Barthélemy","St Barthelemy (France)",fishing_entity),
    fishing_entity = gsub("Saint Pierre & Miquelon","Saint Pierre & Miquelon (France)",fishing_entity),
    fishing_entity = gsub("Solomon Islands", "Solomon Isl.",fishing_entity),
    fishing_entity = gsub("Syria", "Syrian Arab Republic",fishing_entity),
    fishing_entity = gsub("Timor-Leste", "Timor Leste",fishing_entity),
    fishing_entity = gsub("Turks & Caicos Islands", "Turks & Caicos Isl. (UK)",fishing_entity),
    fishing_entity = gsub("The Gambia","Gambia",fishing_entity),
    fishing_entity = gsub("United States", "USA",fishing_entity),
    fishing_entity = gsub("Vietnam", "Viet Nam",fishing_entity),
    fishing_entity = gsub("Wallis & Futuna Islands","Wallis & Futuna Isl. (France)",fishing_entity),
    fishing_entity = ifelse(fishing_entity == "USA Virgin Islands","US Virgin Isl.",fishing_entity)
  ) %>% 
  as.data.frame() %>% 
  select(-geometry)

### --------------------- ###
# SAU Shapefile
### --------------------- ###
path_world <- paste(my_path("G"),"Spatial/SAU/SAU_Shapefile",sep="")

# The File
fnam_world <- "SAUEEZ_July2015.shp"

# Load it!
World_sf <- st_read(dsn = path_world,
                    layer =file_path_sans_ext(fnam_world)) %>% 
  rename(eez_name = Name,
         eez_id = EEZID) %>% 
  st_transform(crs = "+proj=eck4") %>% 
  mutate(
    eez_id = ifelse(eez_name == "Russia (Laptev to Chukchi Sea)",913,eez_id)
  ) %>% 
  clean_names()

World_df <- as.data.frame(World_sf) %>% 
  select(objectid, eez_id,eez_name)

World_sf %>% anti_join(Clean_sau_eez_fes, by = "eez_id")

# Join all tables with matching names

matching_names <- Clean_sau_eez_fes %>% 
  full_join(World_df,
            by = "eez_id") %>% 
  full_join(UN_Regions,
            by = "fishing_entity") %>% 
  full_join(World_land_sf,
            by = "fishing_entity") %>% 
  filter(!is.na(fishing_entity_id)) %>% # Remove duplicaates from UN_regions
  rename(sau_eez_name = eez_name.x,
         sau_fishing_entity = fishing_entity,
         sf_eez_name = eez_name.y,
         sf_object_id = objectid,
         un_country = country_or_area,
         un_geo_region = georegion,
         un_sub_region = sub_region
  ) %>% 
  select(
    eez_id,
    fishing_entity_id,
    sau_eez_name,
    sau_fishing_entity,
    sf_eez_name,
    sf_object_id,
    un_country,
    un_geo_region,
    un_sub_region,
    iso_a3
  )

# Save data for future
write_csv(matching_names,
          "matching_names.csv")

````

### High migratory species

Remove highly migratory species according to UNCLOS extracted from FAO (Maguire et all). The above tuna species listed as highly migratory species in Annex 1 of UNCLOS have extensive distribution on the high seas

Maguire, J-J., Sissenwine, M., Csirke, J., Grainger, R., & Garcia, S. (2006). The state of world highly migratory, straddling and other high seas fishery resources and associated species. FAO.

```{r highly_migratory_species}

high_mig_taxa <- c("Thunnus alalunga", "Thunnus thynnus","Thunnus orientalis","Thunnus obesus","Katsuwonus pelamis", "Thunnus albacares","Thunnus atlanticus","Euthynnus alletteratus","Euthynnus affinis","Thunnus maccoyii", "Auxis thazard", "Auxis rochei",# Tuna-likes
                   "Tetrapturus angustirostris", "Tetrapturus belone", "Tetrapturus pfluegeri", "Tetrapturus albidus", "Tetrapturus audax", "Tetrapturus georgei", "Makaira indica", "Makaira nigricans","Istiophorus platypterus", "Istiophorus albicans","Xiphias gladius")

length(high_mig_taxa) # 23

high_mig_spp <- my_path("D","Distribution/Species","SppTaxonName.csv", read = T) %>% 
  filter(taxon_name %in% high_mig_taxa)

nrow(high_mig_spp) # 21

# ---------------- 
# Missing 4 species
# missing_hm_spp <- as.data.frame(x = high_mig_taxa) %>% 
  # filter(!high_mig_taxa %in% high_mig_spp$taxon_name)

# Tetrapturus belone # only found in the med				
# Tetrapturus georgei # Misspselled, correct name Tetrapturus georgii but not identified in the SAU dataset

# ---------------- # missing data
```

## Load Results data

In here we load and cleaned the multiple external data sets required to produce the manuscript figures

```{r Load_Results_Data, eval = F}

# ----------------- #
# STRADDLING
# ----------------- #
fao_areas <- my_path("Spa", "DBEM","complete_lon_lat_dbem.csv", read = T) %>% 
  filter(cat == "abnj")

# Get species taxon data
taxon_list <- my_path("G","FishForVisa/Data/Distribution/Species","SppTaxonName.csv", read = T)

# Load straddling species
# Spp list
spp_complete <- list.files(my_path("R","EEZ/Straddling_Results/"))

stradling_spp <- str_replace(spp_complete,"\\_.*","")

stradd_data <- bind_rows(
  lapply(stradling_spp,GetTransResults, type = "strad")
  ) %>% 
  left_join(taxon_list) %>% 
  select(taxon_key,taxon_name,dem_pel, everything()) %>% 
  mutate(category = ifelse(taxon_key %in% high_mig_spp$taxon_key,"high_migratory","straddling"))

# Complete transboundary data

stradd_hs_spp <- stradd_data %>% 
  select(taxon_key,category,sf_eez_name = territory, eez_neighbour = neighbour) %>% 
  unique()

# ----------------- #
# Misc.
# ----------------- #

# Get species taxon data
taxon_list <- my_path("G","FishForVisa/Data/Distribution/Species","SppTaxonName.csv", read = T)

Results_Path <- my_path("R")

# Set Jurisdiction to EEZ or RFMO
Jurisdiction <- "EEZ"
# Jurisdiction <- "RFMO"

# Updates result's path for the jurisdiction
Results_Path <- paste(my_path("R"),Jurisdiction,"/",sep="")

# Dataset to match all SAU names
matching_names <- my_path("G","Spatial/SAU", "sau_matching_names.csv", read = T)

# Species list to get species names
Exploited_Species <- fread(paste(my_path("D"),"Distribution/Species/exploited_species_list.csv",sep="")) %>%
  clean_names()

# STRADDLING: needed
Species_Sampled <- fread(paste(my_path("D"),"Distribution/Sampled_Species.csv", sep =""))

Spp_Clasification <- fread(paste(my_path("D"),"Distribution/Species/SppTaxonName.csv",sep="")) %>% 
  clean_names()

# ----------------- #
# Data #
# ----------------- #

# Load Transbonudary results
Raw_Results_Trans <- my_path("R","EEZ","Raw_Results_Trans.xlsx", read = T)

# Clean_Results_Trans <- Raw_Results_Trans %>% 
#   CleanResults(.,matching_names) %>% 
#   rename(sf_eez_name = eez_name) %>% 
#   left_join(matching_names,
#             by = "sf_eez_name") %>% 
#   rename(eez_name = sf_eez_name,
#          fishing_entity = sau_fishing_entity)

Clean_Results_Trans <- my_path("R","EEZ","Clean_Results_Trans.csv", read = T) %>% 
  rename(sf_eez_name = eez_name) %>% 
  left_join(matching_names)

# For upload to github
# Clean_Results_Trans %>%
#   left_join(Exploited_Species) %>%
#   select(1,taxon_name,2:5,fishing_entity,un_sub_region) %>%
#   write_csv(.,"Clean_Results_Trans.csv")

# New data produced by Vicky to compare with Teh and Sumaila
# Clean_sau_eez_fes <- my_path("G","Spatial/SAU/","EEZ_FishingEntityID_v2_9June2020.xlsx", read =T)

# UN Regions dataset
UN_Regions <- read.csv(paste(my_path("D"),"Spatial/UN_Regions.csv",sep=""), header = TRUE) %>% 
  clean_names() %>% 
  group_by(
    un_country = country_or_area,
    georegion = region_name,
    sub_region = sub_region_name) %>% 
  summarise(n=n()) %>% 
  left_join(matching_names) %>% 
  ungroup() %>% 
  distinct(un_country,.keep_all = TRUE) %>% 
  select(fishing_entity = sau_fishing_entity,georegion,sub_region)

# Load Land shapefile
World_land_sf <- rnaturalearth::ne_countries(scale = 'medium', returnclass = c("sf")) %>% 
  st_transform(crs = "+proj=eck4") %>%
  # st_transform(crs = 4326) %>% # 4326
    # rename(sau_fishing_entity = name_long) %>%
  left_join(matching_names,
            by = "iso_a3") %>% 
  distinct(iso_a3,.keep_all = TRUE) %>% 
  rename(fishing_entity = sau_fishing_entity)

# Load SAU shapefile name and paths

# The path
path_world <- paste(my_path("G"),"Spatial/SAU/SAU_Shapefile",sep="")

# The File
fnam_world <- "SAUEEZ_July2015.shp"

# Load it!
World_sf <- st_read(dsn = path_world,
                    layer =file_path_sans_ext(fnam_world)) %>% 
  rename(eez_name = Name,
         eez_id = EEZID) %>% 
  st_transform(crs = "+proj=eck4") %>% 
  mutate(
    eez_id = ifelse(eez_name == "Russia (Laptev to Chukchi Sea)",913,eez_id)
  )

# Determine polygon areas

# EEZ Area
EEZ_Area <- World_sf %>% 
  as.data.frame() %>% 
  group_by(eez_id) %>% 
  summarise(Size_EEZ = sum(Area_km2))

# Entity Area (e.g. Mexico = Pacific EEZ + Atalntic EEZ)
Entity_Area <- EEZ_Area %>% 
  left_join(matching_names,
            by = "eez_id") %>% 
  group_by(fishing_entity = sau_fishing_entity,fishing_entity_id) %>% 
  summarise(Size_Entity = sum(Size_EEZ))

# Regional UN-sub region area
Regional_Area <- Entity_Area %>% 
  left_join(UN_Regions) %>% 
  group_by(sub_region) %>% 
  summarise(size_region = sum(Size_Entity,na.rm=T))%>% 
  arrange(size_region)

# ---------------------------- #
# Supplement table 2
# write_csv(Regional_Area,
#           "Table S1.csv")
# ---------------------------- #

# ---------------------------- #
# Transboundary Results ####
# ---------------------------- #

# Read Sau Value Data (Thank you Tim!) reconstructede data
# Gross Domestic Product Implicit Price Deflator according to the US Saint Louis Federal Reserve Bank (average of all 2019 quarters) to transform the original 2010 real USD revenue value to 2019 real USD (https://fred.stlouisfed.org/searchresults?st=GDP+Implicit+price+deflator).*

# CPI in 2019/ CPI in 2010 × 2010  value = 2019  value

# Used for STRADDLING too
SAU_Global_Data <-fread(paste(my_path("D"),"Distribution/sau_catch_value_country_taxon_JEPA.csv", sep ="")) %>%
  # Remove high seas and 207 and 208 that are unknown
  filter(!fishing_entity_id %in% c(213,223)) %>%  # 207,208
  # Allocate catch of islands to UK
  mutate(fishing_entity_id = ifelse(fishing_entity_id %in% c(219,220), 183,fishing_entity_id)) %>% 
  # Convert to 2019 real USD
  mutate(values = 1.16 * value) %>%
  # Average 10 years of data
  group_by(fishing_entity_id,taxon_key) %>% 
  summarise_at(c("catch","value"),
               mean,na.rm=T) %>% 
  gather("Variable","Value",catch:value)

# Read Scatch trend Data (Sub-routine Catch Trend)
Trend_Data <- fread(paste(my_path("R"),"EEZ/Species_Trend.csv", sep ="")) %>% 
  clean_names() %>% 
  rename(sf_eez_name=territory) %>% 
  left_join(matching_names) %>% 
  rename(eez_name = sf_eez_name)
```

## Determine Thresholds

Determine transboundary stocks and countries based on selected thresholds. We created two thresholds (levels of uncertainty), in here we can chose which level we wanna work with.

```{r Results_Tresholds, eval = F, echo = F}

#### Select Tresholds ###

# Area Index
Min_AI <- 0.25 # Min area index
Max_AI <- 1-Min_AI # Max Area Index
# Model Index
MI <- 100 # Model Index

#### Determine transboundary and species and nations

Tresholded_T_Data <- Clean_Results_Trans %>% 
  filter(
    area_index >= Min_AI,
    area_index <= Max_AI,
    model_index >= MI
  )

### Duble check for NA's
# map(Tresholded_T_Data, ~sum(is.na(.))) # No NA's

## STRADLING
shared_species <- Tresholded_T_Data %>% 
  left_join(stradd_hs_spp) %>% 
  mutate(category = replace_na(category,"transboundary"))

shared_classification <- shared_species %>% 
  select(taxon_key,category,sf_eez_name) %>% 
  unique()

```

## Overall results

### Straddling Species

```{r}

# Number of species that are straddling or highly migratory
stradd_data %>% 
  group_by(category) %>% 
  summarize(
    n_spp = length(unique(taxon_key)),
    prop = round((length(unique(taxon_key))/633)*100)
    )

# RESULT
# category  n_spp   prop of transboundary
# high_mig	 20	     3		
# stradd	   84	    13	
# total     104     16

EEZIDs_List <- my_path("Spa","DBEM", "Updated_EEZList_17June2016.xlsx", read = T) %>% 
  rename(eez_name = name)

# Econ value
Summary_Trans <- SAU_Sampled_Data %>% 
  semi_join(Tresholded_T_Data,
            by = c("taxon_key","fishing_entity_id")
  ) %>% 
  group_by(Variable) %>%
  summarise(Total_Trans = sum(Value, na.rm=T)) %>% 
  left_join(Total_SAU_Sampled) %>% 
  left_join(Total_SAU_Global) %>% 
  mutate(Per_sampled = Total_Trans/Total_Sampled*100,
         Per_SAU = Total_Trans/Total_Global*100)

```

### Transboundary Species

```{r Transboundary_Data_per_Territory, eval = F, echo = F}

# -------------------------- #
# Number of EEZs and countries
# -------------------------- #

# number of eezs
length(unique(World_sf$eez_name))
# Number of coastal territories
length(unique(matching_names$fishing_entity))

# -------------------------- #
# Number of Transboundary and Discrete Species
# -------------------------- #

n_trans <- Tresholded_T_Data %>% 
  pull(taxon_key) %>% 
  unique() %>% 
  length()

# Percentage of transboundary species from sampled
n_trans/nrow(Species_Sampled)*100

# -------------------------- #
# Percentage of capture analyzed
# -------------------------- #

# Filter Tim's data for only those species we estimate
Total_SAU_Sampled <- SAU_Global_Data %>%
  filter(taxon_key %in% Species_Sampled$TaxonKey) %>% 
  group_by(Variable) %>% 
  summarize(Total_Sampled = sum(Value,na.rm=T),
            Sampled_nspp = length(unique(taxon_key))
            )

# Get the total values for the world excluding not identified species
Total_SAU_id <- SAU_Global_Data %>% 
  filter(taxon_key %in% Exploited_Species$taxon_key) %>%
  group_by(Variable) %>% 
  summarize(Total_id = sum(Value,na.rm=T),
            id_nspp = length(unique(taxon_key))
  )

# Get the total values for the world including non id species
Total_SAU_Global <- SAU_Global_Data %>% 
  group_by(Variable) %>% 
  summarize(Total_Global = sum(Value,na.rm=T),
            Global_nspp = length(unique(taxon_key))
  )

# Percentage catch of species analyzed from the identified species
Total_Catch_Spp <- Total_SAU_Sampled$Total_Sampled[1]/Total_SAU_id$Total_id[1] *100 #96.49%

# Percentage species analyzed from the identified species
936/length(unique(SAU_Global_Data$taxon_key)) # 67.5

Total_spp_analyzed <- Total_SAU_Sampled$Sampled_nspp[1]/Total_SAU_id$id_nspp[1]*100 #63.23%

# -------------------------- #
# Total and percentage Catch and Revenue from Transboundary species
# -------------------------- #

# Filter SAU data for only those sampled species
SAU_Sampled_Data <- SAU_Global_Data %>% 
  filter(taxon_key %in% Exploited_Species$taxon_key)

# Numbers for paper

Summary_Trans <- SAU_Sampled_Data %>% 
  semi_join(Tresholded_T_Data,
            by = c("taxon_key","fishing_entity_id")
  ) %>% 
  group_by(Variable) %>%
  summarise(Total_Trans = sum(Value, na.rm=T)) %>% 
  left_join(Total_SAU_Sampled) %>% 
  left_join(Total_SAU_Global) %>% 
  mutate(Per_sampled = Total_Trans/Total_Sampled*100,
         Per_SAU = Total_Trans/Total_Global*100)

```

### Discussion with Teh

```{r comparison_with_Teh_2015, eval = T, echo = F}

# Read SAU non-reconstructed catch data from 2006 for comparrisonwith Sumaila
# Provided by Vicky Lam
SAU_FAO_Data <- fread(paste(my_path("D"),"Distribution/LVCatch_byYrEEZFETaxa_2005_09.csv", sep ="")) %>% 
  filter(#eezid != 0,
         year == 2006) %>%
  group_by(fishing_entity_id,taxon_key,eezid) %>% 
  summarise_at(c("sumcatch"),
               sum,na.rm=T)

## Get the total of global data
Total_SAU_FAO <- sum(SAU_FAO_Data$sumcatch,na.rm = T)

# Estimate numbers for paper
SAU_FAO_Data %>% 
  semi_join(Tresholded_T_Data,
            by = c("taxon_key","fishing_entity_id")
  ) %>% 
  group_by() %>% 
  summarise(Total_Trans = sum(sumcatch, na.rm=T)) %>% 
  mutate(Per_Sau_FAO = Total_Trans/Total_SAU_FAO*100)

```

### Non-shared Transboundary

Based on our assessment (see Methods), 582 transboundary species were also identified as "discrete" species, meaning that the same species could be shared by two neighboring EEZs (transboundary) but not with a third neighbor (discrete).

```{r Discrete_Data_per_Territory, eval = F, echo = F}

Species_Left_Out <- Species_Sampled %>% 
  filter(!TaxonKey %in% Clean_Results_Trans$taxon_key) %>% 
  select(taxon_key = TaxonKey)

###_________________###
# Discrete species
# Species that are transboundary in some cases but discrete in other cases
###_________________###

Discrete <- Clean_Results_Trans %>% 
  filter(
    area_index < Min_AI |
      area_index > Max_AI |
      model_index < MI
  ) %>% 
  # Remove Countries that share with other countries
  anti_join(Tresholded_T_Data,
            by = c("taxon_key","fishing_entity")
  )

### Duble check for NA's
# map(Discrete, ~sum(is.na(.))) # No NA's

# Both categories (trans and Discrete)
n_discrete <- Tresholded_T_Data %>% 
   # Filter species that are also discrete
  filter(taxon_key %in% Discrete$taxon_key) %>% 
  group_by(taxon_key) %>% 
  summarise(n()) %>% 
  pull(unique(taxon_key)) %>% 
  length()

# Summary results for paper
Summary_Discrete <- SAU_Sampled_Data %>% 
  semi_join(Discrete,
            by = c("taxon_key","fishing_entity_id")
  ) %>% 
  group_by(Variable) %>%
  summarise(Total_Trans = sum(Value, na.rm=T)) %>% 
  left_join(Total_SAU_Sampled) %>% 
  left_join(Total_SAU_Global) %>% 
  mutate(Per_sampled = Total_Trans/Total_Sampled*100,
         Per_SAU = Total_Trans/Total_Global*100)

```

### Discrete Species

Within these species, 305 were considered exclusively discrete, that is, they could be fished by neighboring nations but were not considered transboundary according to our methods of determination (See methods).

```{r Discrete_Exclusive_per_Territory, eval = F, echo = F}
###_________________###
# Discrete Exclusive
# Species that are only discrete and are not present in the transboundary category
###_________________###

  # Discrete species (not in Tresholded)
Discrete_Exclusive <- Discrete %>% 
  filter(!taxon_key %in% Tresholded_T_Data$taxon_key) %>% 
  bind_rows(Species_Left_Out) %>%
  group_by(taxon_key) %>%
  summarise(n()) %>% 
  left_join(Exploited_Species)

## Double check no double counting

# Discrete_Exclusive %>% 
#   semi_join(Tresholded_T_Data, by = "taxon_key") # 0 ok thes

# Number of discrete species
n_discrete_exclusive <- length(unique(Discrete_Exclusive$taxon_key))

# Double check n = total species (938)
n_discrete_exclusive + n_trans # 938

# Numbers for paper
Summary_Discrete_Exclusive <- Discrete_Exclusive %>% 
  left_join(SAU_Sampled_Data) %>% 
  group_by(Variable) %>%
  summarise(Total_Discrete = sum(Value, na.rm=T)) %>% 
  left_join(Total_SAU_Sampled) %>% 
  left_join(Total_SAU_Global) %>% 
  mutate(Per_sampled = Total_Discrete/Total_Sampled*100,
         Per_SAU = Total_Discrete/Total_Global*100)

```

## Results per Territory

### Fig. 1 - Data For Map

We group countries according to the United Nations sub-regions and found that transboundary species are particularly economically important for Northern America and Eastern Asia

```{r Data_per_Territory, eval = F, echo = F}

# Data per Territory (EEZ)
Transboundary_Spp <- Tresholded_T_Data %>% 
  group_by(eez_name, eez_id) %>% 
  summarise(
    n_trans_spp = length(unique(taxon_key)),
    n_neighbours = length(unique(eez_neighbour)),
    trans_Rrate = n_trans_spp/n_neighbours
  ) %>% 
  ungroup() %>% 
  mutate(
    Bins = cut(n_trans_spp,breaks = 5),
    Bins_Plot = cut(n_trans_spp,breaks = 5,
                    labels = c('super low', 'low','median', 'high','super high')
                    )
  )

# Get Catch and Revenue data for mapping

# Set manual breaks for bins
Map_Breaks <- c(-1,10,100,1000,5000,12000)

# Estimate value per fishing entity
Total_Value_Entity <- SAU_Global_Data %>% 
  group_by(fishing_entity_id,Variable) %>% 
  summarise(Total_Value = sum(Value,na.rm=T))

# Data for map

Value_Trans_Country <- Tresholded_T_Data %>% 
  group_by(fishing_entity_id,
           fishing_entity,
           taxon_key) %>% 
  summarise(n()) %>% # get each taxa per nation
  left_join(SAU_Sampled_Data,
            by = c("fishing_entity_id","taxon_key")) %>% 
  group_by(fishing_entity,fishing_entity_id,Variable) %>% # Estimate the amount of dollars of each nation 
  summarise(
    Trans_Entity_n = length(unique(taxon_key)),
    Trans_Total = sum(Value,na.rm=T), #million
  ) %>%
  ungroup() %>%
  filter(!is.na(Variable)) %>%
  mutate(
    Trans_Total = ifelse(Variable == "value",Trans_Total/1000000, #million /1000000
                         ifelse(Variable == "catch",Trans_Total/1000, #thousand Trans_Total/1000
                                Trans_Total)
    )
  ) %>%
  spread(Variable,Trans_Total) %>%
  left_join(Entity_Area, 
            by = c("fishing_entity","fishing_entity_id")
          # by = "fishing_entity"
  ) %>% 
  # Include the weighted method for supplementals
  mutate_at(
    vars(catch,value),
    funs(weight= ./Size_Entity) # back to USDs
  ) %>% 
  gather("Variable","Value",catch:value,catch_weight:value_weight) %>% 
  mutate(
    Bins = cut(Value, breaks = Map_Breaks, dig.lab = 5),
    Bins_Plot = cut(Value, breaks = Map_Breaks,
                    labels = c('super low', 'low','median','high','super high'),
                    dig.lab = 5 # Removes scientific notaion
    )
  ) %>% 
  left_join(Total_Value_Entity,
            by = c("fishing_entity_id", "Variable") 
  ) %>%
  # Convert catch to million tonnes
  ungroup() %>% 
  mutate(
    Value = ifelse(Variable == "catch",Value/1000, #million
                   Value),
    Trans_Per = (Value/Total_Value*100000000)#*100000000 #Return to percentage
  )

```

### Data for continental analysis (round bar plots)

```{r Data_per_Continent, eval = F, echo = F}

# Continental level for transboundary species
Continent_Trans_Value <- Value_Trans_Country %>% 
  left_join(UN_Regions,
            by = "fishing_entity") %>% 
  mutate(georegion = ifelse(is.na(georegion), "Other",paste(georegion))) %>% 
  group_by(georegion,Variable) %>% 
  summarise_if(is.numeric,sum,na.rm=T)

## Sub region leven
Subregion_Trans_raw_Value <- Value_Trans_Country %>% 
  left_join(UN_Regions,
            by = "fishing_entity") %>% 
  mutate(sub_region = ifelse(is.na(sub_region), "Other",paste(sub_region))) %>% 
  group_by(sub_region,Variable) %>% 
  summarise_if(is.numeric,c(sum = sum,
                            m = mean,
                            sd = sd),
               na.rm=T) %>%
  arrange(Variable) %>%
  ungroup() %>% 
  mutate(id = rep(seq(1,17),4)) %>% 
  mutate(Value_sum=ifelse(Variable == "catch" & Value_sum >= 12,12,Value_sum))

# Fix names for plot
Subregion_Trans_Value <- Subregion_Trans_raw_Value

Subregion_Trans_Value$sub_region <- Subregion_Trans_raw_Value$sub_region %>% 
  gsub("Northern","N.",.) %>% 
  gsub("Southern","S.",.) %>% 
  gsub("Western","W.",.) %>% 
  gsub("Eastern","E.",.)

# Transboundary Species shapefile
World_sf_si <- World_sf %>%
  st_simplify(preserveTopology = TRUE, dTolerance = 0.1) %>%
  left_join(Transboundary_Spp, by ="eez_id") %>% 
  st_transform(crs = "+proj=eck4")
  # st_transform(crs = 4326)

# Douuble check we are not missing EEZs
Transboundary_Spp %>% anti_join(World_sf, by = "eez_id")

### Prepare data for plot

number_of_bar <- 17 # Number of regions

# Create labels for legend
Bins_Trans <- Transboundary_Spp %>% 
  arrange(Bins) %>% 
  pull(Bins) %>% 
  unique()

```

### Map of Transboundary Species

Figure 1. Number of transboundary species and their contribution to global fisheries catches and revenue. A) The number of species and amount of revenue are represented by color coding of EEZs and land polygons, respectively. B) Contribution of transboundary species to regional revenue (left) and catch (right). Regions according to the United Nations sub-regions. Points = mean +- sd. Revenue in 2010 real USD

```{r Figure_1}
gc()
# Set color pal
pal <- wes_palette("Zissou1",100,type = "continuous")

# Variable to plot (value, catch, weighted_catch, weighted_value)
Variable_plot <- "catch"
CreateFigureOne(Variable_plot)

```

### China, US, Russia, and Peru

```{r Map_Particulars, eval =F, echo =F}

# Revenue details

Top_Rev_Countries <- c("China", "Peru", "Russian Federation", "Chile", "USA","Japan")

# Total revenue from Top countries
Total_Top_Rev <- Value_Trans_Country %>% 
  # filter(Variable == "value") %>% 
  filter(fishing_entity %in% Top_Rev_Countries) %>% 
  left_join(UN_Regions,
            by = "fishing_entity") %>% 
  left_join(Subregion_Trans_raw_Value) %>% 
  mutate(Per_Top = Value/Value_sum*100) %>% 
  select(1:3,Per_Top)

### Percentaga that top 5 represents

Top5 <- c("Peru", "Russian Federation", "Japan", "USA","China")

Total_Top5 <- Value_Trans_Country %>% 
  filter(Variable == "value") %>% 
  filter(fishing_entity %in% Top_Rev_Countries) %>% 
  ungroup() %>% 
  select(fishing_entity,Value) %>% 
  group_by() %>% 
  summarise(sum(Value)) %>% 
  pull()/(Total_SAU_Sampled$Total_Sampled[2]/1000000)*100

```

### Statistical differences within regions
We used a one-way ANOVA to determine significant differences within regions for transboundary fisheries catches and revenue with a Tukey's HSD post-hoc test.

```{r statistics_fig1, eval = F, echo = T}

### Using ANOVA to estimate significan differences within regions for revenue and catch of transboundary species

#-----------------#
# Sub-regional differences
#-----------------#

Subregion_Data_stats <- Value_Trans_Country %>%
  left_join(UN_Regions,
            by = "fishing_entity") %>%
  mutate(sub_region = ifelse(is.na(sub_region), "Other",paste(sub_region))) %>%
  select(fishing_entity,sub_region,Variable,Value) %>%
  filter()

### Reveneu ANOVA

Subregion_Rev_stats <- Subregion_Data_stats %>% 
  filter(Variable == "value")

Rev_aov <- aov(Value~sub_region, data = Subregion_Rev_stats)
Rev_aov

summary(Rev_aov)
#              Df    Sum Sq Mean Sq F value   Pr(>F)    
# sub_region   16  75738982 4733686   4.948 3.46e-08 ***
# Residuals   158 151144071  956608

#  Post-hoc analysis
TukeyHSD(Rev_aov) %>% 
  tidy() %>% 
  filter(adj.p.value < 0.05) %>% 
  View()

### Catch ANOVA

Subregion_Catch_stats <- Subregion_Data_stats %>% 
  filter(Variable == "catch")

Catch_aov <- aov(Value~sub_region, data = Subregion_Catch_stats)
Catch_aov

summary(Catch_aov)
#  Df Sum Sq Mean Sq F value  Pr(>F)   
# sub_region   16  27.51  1.7195   2.265 0.00539 **
# Residuals   158 119.97  0.7593

#  Post-hoc analysis
TukeyHSD(Catch_aov) %>% 
   tidy() %>% 
  filter(adj.p.value < 0.05) # Eastern Asia

# Weight ANOVA
Subregion_Weigh_stats <- Subregion_Data_stats %>% 
  filter(Variable == "value_weight")

Weight_aov <- aov(Value~sub_region, data = Subregion_Weigh_stats)
Weight_aov
summary(Weight_aov)

#  Df Sum Sq Mean Sq F value  Pr(>F)   
# sub_region   16   6722   420.1   2.411 0.00289 **
# Residuals   158  27532   174.3                   
# ---
# Signif. codes:  0 ‘***’ 0.001 ‘**’ 0.01 ‘*’ 0.05 ‘.’ 0.1 ‘ ’ 1

#### There are significant differences. But between what?

#  Post-hoc analysis
TukeyHSD(Weight_aov) %>% 
  tidy() %>% 
  filter(adj.p.value < 0.05) # Northern Europe
```

## Fig. 2 - Weighted bar plots

### Data for Weighted plot
```{r Weight_Revenue, eval =F, echo =F}

#--------------------------- #
### Create Suplemental Table S1
#--------------------------- #

Subregion_Trans_raw_Value %>% 
  select(sub_region,Size_Entity_sum) %>% 
  distinct(sub_region,.keep_all = TRUE) %>% 
  arrange(Size_Entity_sum)
  # write_csv(.,
            # "Table_S1.csv")

#--------------------------- #

Weighted_values <-c("value_weight","catch_weight")

Weighted_Data <- Subregion_Trans_Value %>% 
  mutate(
    Value_sum = ifelse(Variable == "value_weight", Value_sum*1000, Value_sum),
    Value_m = ifelse(Variable == "value_weight", Value_m*1000, Value_m),
    Value_sd = ifelse(Variable == "value_weight", Value_sd*1000, Value_sd),
    Value_sum = ifelse(Variable == "catch_weight", Value_sum*1000, Value_sum),
    Value_m = ifelse(Variable == "catch_weight", Value_m*1000, Value_m),
    Value_sd = ifelse(Variable == "catch_weight", Value_sd*1000, Value_sd)
    )

# Labels
label_data_rev <- Weighted_Data %>% 
  filter(Variable == Weighted_values[1]) %>% 
  arrange(Value_sum) %>% # soo it looks like a nice snail :)
  mutate(
    id = seq(1,number_of_bar),
    angle = 90 -360 * (id-0.5)/number_of_bar,
    hjust = ifelse(angle < -90,1,0),
    angleb = ifelse(angle < -90, (angle +180), angle)
  ) %>% 
  mutate(
    sub_region = ifelse(str_detect(sub_region,"Latin"),"Ltn. Ame. &\nthe Car.",
                        ifelse(sub_region =="Northern America","Northern\nAmerica",
                               ifelse(sub_region =="Sub-Saharan Africa","Sub-Saharan\nAfrica",
                               paste(sub_region)
                               )
    )
  )
  )
# Label info Catch
label_data_catch <- Weighted_Data %>% 
  filter(Variable == Weighted_values[2]) %>% 
  arrange(Value_sum) %>% # soo it looks like a nice snail :)
  mutate(
    id = seq(1,number_of_bar),
    angle = 90 -360 * (id-0.5)/number_of_bar,
    hjust = ifelse(angle < -90,1,0),
    angleb = ifelse(angle < -90, (angle +180), angle)
  ) %>% 
  mutate(
    sub_region = ifelse(str_detect(sub_region,"Latin"),"Ltn. Ame. &\nthe Car.",
                        ifelse(sub_region =="Northern America","Northern\nAmerica",
                               ifelse(sub_region =="Sub-Saharan Africa","Sub-Saharan\nAfrica",
                               paste(sub_region)
                               )
    )
  )
  )
# Revenue Plot
  # Set plot limits

  max_Rev_lim <- Weighted_Data %>%
    filter(Variable == Weighted_values[1]) %>%
    group_by() %>%
    summarise(max(Value_sum)) %>%
    pull()

  Subregion_Rev_Data <- Weighted_Data %>% 
    filter(Variable == Weighted_values[1]) %>% 
    arrange(Value_sum)

````

### Plot weighted bars

```{r, eval = T, echo = F}
  Rev_Plot <- ggplot(
    # Regional plots
    data = Subregion_Rev_Data,
    aes(
      x=reorder(id,Value_sum),
      y=Value_sum, 
      fill=Value_sum
    )
  ) +    
    ylim(-5, 30) +
    geom_bar(stat="identity") +
    # Regional averages
    geom_point(data = Subregion_Rev_Data,
               aes(
                 x=reorder(id,Value_sum),
                 y=Value_m
               ),
               colour = "black",
               alpha = 0.5,
               size = 0.5
    ) +
    # Regional sds
    geom_errorbar(data = Subregion_Rev_Data,
                  aes(
                    ymin=ifelse(Value_m-Value_sd <= 0,0,Value_m-Value_sd), # sets lower to 0
                    ymax=Value_m+Value_sd
                  ), 
                  width=.2,
                  position=position_dodge(.9),
                  alpha = 0.5) +
    # Looking good
    theme_minimal(base_size = 9) +
    theme(
      legend.key.width =unit(1,"line"),
      legend.key.height =unit(0.9,"line"),
      legend.position = c(0.86,0.55),
      axis.text = element_blank(),
      axis.title = element_blank(),
      panel.grid = element_blank(),
      plot.margin = margin(1,0,0,-2.5,"cm") # top,x,x,left
      # plot.margin = margin(0,-10,-6,-10,"cm")
    )  +
  coord_polar(start = 0) + 
  geom_text(data=label_data_rev, 
            aes(x=id,
                # y=ifelse(sub_region=="N. Europe",250,Value_sum+1), # Fix names overlap
                y = Value_sum,
                label=sub_region,
                hjust=hjust), 
            color="black",
            fontface="bold",
            size=2, 
            angle= label_data_rev$angleb, 
            inherit.aes = FALSE) +
    scale_fill_gradientn("",
                         colours = pal,
                         limits = c(0,30),
                         breaks = seq(0,30,6)
                         ) + # puts legend text on the left
    guides(fill = guide_colorbar(label.position = "left"));Rev_Plot

  ###____________###
  ### Catch plot
  ###____________###

  max_Catch_lim <- Weighted_Data %>% 
    filter(Variable == Weighted_values[2]) %>% 
    group_by() %>% 
    summarise(max(Value_sum)) %>% 
    pull()

  Subregion_Catch_Data <- Weighted_Data %>% 
    filter(Variable == Weighted_values[2])

  Catch_Plot <- ggplot(data = Subregion_Catch_Data,
                     aes(x=reorder(id,Value_sum),
                         y=Value_sum, 
                         fill=Value_sum
                     )
) +    
  ylim(-3.4,20) +
  geom_bar(stat="identity") +
  # Regional averages
  geom_point(data = Subregion_Catch_Data,
             aes(
               x=reorder(id,Value_sum),
               y=Value_m
             ),
             colour = "black",
             alpha = 0.5,
             size = 0.5
  ) +
  # Regional sd's
  geom_errorbar(data = Subregion_Catch_Data,
    aes(
      ymin=ifelse(Value_m-Value_sd <= 0,0,Value_m-Value_sd),
      ymax=Value_m+Value_sd
    ), 
    width=.2,
    position=position_dodge(.9),
    alpha = 0.5) +
  theme_minimal(base_size = 9) +
  theme(
    # legend.position = "bottom",
    legend.position = c(0.595,4.52), #0.55
    legend.key.width =unit(1,"line"),
    legend.key.height =unit(0.9,"line"),
    axis.text = element_blank(),
    axis.title = element_blank(),
    axis.text.x = element_text(size = 10),
    panel.grid = element_blank(),
    plot.margin = margin(1,0,0,0.8,"cm")
  )  +
  coord_polar(start = 0) + 
  geom_text(data=label_data_catch, 
            aes(x=id,
                # y=ifelse(sub_region=="N. Europe",250,Value_sum+1), # Fix names overlap
                y = Value_sum,
                label=sub_region,
                hjust=hjust), 
            color="black",
            fontface="bold",
            size=2, 
            angle= label_data_catch$angleb, 
            inherit.aes = FALSE) +
  scale_fill_gradientn("",
                       limits = c(0,20),
                       breaks = seq(0,20,4),
                       colours = pal
  ); Catch_Plot

  ## Joining everything in one ploit

  # Get legend for cowplot
legend_plot <- get_legend(Catch_Plot)
# Re-plot without legend
Catch_Plot <- Catch_Plot +
  theme(legend.position = "")
  p<- ggdraw() +
  # Revenue circular
  draw_plot(Rev_Plot, x = 0, y = 0, width = 0.8, height = 1) +
  # Catch circular
  draw_plot(Catch_Plot, x = 0.4, y = 0, width =0.8, height = 1) +
  # Legend
  draw_plot(legend_plot, x = 0.2, y = 0.05, width = 0.59, height = 0.1) +
  # Labeks
  draw_plot_label(label = c("a", "b"),
                  size = 15,
                  x = c(0.1, 0.55),
                  y = c(1, 1)
  ) +
  # draw_plot_label(
  #   label = c("Revenue (Thousand USD) | Catch (Tons)","Per Km2"),
  #                 size = 6,
  #                 x = c(0.27,0.49),
  #                 y = c(0.66,0.63)
  # )
    draw_plot_label(
    label = c("Revenue     |     Catch",
              "(Thousand)  |    (Tonnes)",
              "  (Per km  )"),
                  size = 6,
                  x = c(0.41,
                        0.395,
                        0.48),
                  y = c(0.675,
                        0.65,
                        0.623)
  )

  save_plot(
    "weight_plot.png",
    p,
    base_height = 4.2,
    base_width = 7
  )

```

## OHI-Territories

*We determined the exploitation category of each species according to the catch trend within each EEZ (Fig. 3).

### Data for Circular plot

```{r Trend_Plot_Data, eval = F, echo = F}

# -------------------------- #
# Transboundary data
# -------------------------- #
# Stock trend Data
Trend_Data_Plot <- Tresholded_T_Data %>% 
  left_join(Trend_Data#, 
            # by =c("eez_name","taxon_key","sf_object_id")
  ) %>%
  mutate(trend = ifelse(is.na(status),"No Trend",status)) %>% 
  group_by(fishing_entity,trend) %>% 
  summarise(value = length(unique(taxon_key))) %>% 
  left_join(UN_Regions, by ="fishing_entity") %>%
  mutate(Georegion = ifelse(is.na(georegion),"Other",paste(georegion))) %>% 
  mutate(value = ifelse(value > 100,100,value)) %>% 
  filter(!is.na(fishing_entity),
         !str_detect(fishing_entity, ".\\(.") # remove associated territories
  ) %>% 
  select(individual = fishing_entity,observation = trend,group = Georegion,value) %>% 
  arrange(group,value) %>% 
  mutate(value = ifelse(observation %in% c("Collapsed","Over Exploited"),value*-1,value),
         observation = ifelse(observation %in% c("Developing","Rebuilding"),"Category A",
                              ifelse(observation == "Max Exploited","Category B",
                                     ifelse(observation %in% c("Over Exploited","Collapsed","Category C"),"Category C", "No Category")
                              )
         )
  ) %>% 
  group_by(individual,observation,group) %>% 
  summarise(value = sum(value))

#### Convert trends to percentage for plot

Totals_trend <- Trend_Data_Plot %>% 
  group_by(individual) %>% 
  summarise(total_trend = sum(abs(value)))

Trend_Data_Plot <- Trend_Data_Plot %>% 
  left_join(Totals_trend) %>% 
  mutate(
    value = round(value/total_trend*100)
  )

length(unique(Trend_Data_Plot$individual))
length(unique(Trend_Data_Plot$observation))

Trend_Plot_Data <- tibble(
  fishing_entity = rep(unique(Trend_Data_Plot$individual),4),
  observation = rep(unique(Trend_Data_Plot$observation),157)
) %>% 
  left_join(UN_Regions) %>% 
  rename(individual = fishing_entity) %>% 
  full_join(Trend_Data_Plot,
            by = c("individual","observation")
  ) %>% 
  mutate(value = ifelse(is.na(value),0,
                        ifelse(value < -100,-100, # if less than 100, 100
                               ifelse(value > 100,100, # if more than 100, 100
                                      value)
                        )#, # Ser NAs to 0's
                        # group = ifelse(is.na(group),"Other",group)
  )
  ) %>% 
  select(individual,observation,group=georegion,value) %>% 
  filter(group != "other")

# Summary for manuscipt

T_Summary <- Trend_Data_Plot %>% 
  mutate(value = ifelse(value <0, value*-1,value)) %>% 
  group_by(observation) %>% 
  summarise(tot=sum(value),
            mean(value),
            sd(value)) %>% 
  arrange(desc(tot))

# -------------------------- #
# Discrete data
# -------------------------- #

Discrete_Trend_Data <- Discrete %>% 
  left_join(Trend_Data, 
            by =c("eez_name","taxon_key")
  ) %>%
  mutate(trend = ifelse(is.na(status),"No Trend",status)) %>% 
  group_by(fishing_entity,trend) %>% 
  summarise(value = length(unique(taxon_key))) %>% 
  left_join(UN_Regions, by ="fishing_entity") %>%
  mutate(Georegion = ifelse(is.na(georegion),"Other",paste(georegion))) %>% 
  # mutate(value = ifelse(value > 100,100,value)) %>% 
  filter(!is.na(fishing_entity),
         !str_detect(fishing_entity, ".\\(.") # remove associated territories
  ) %>% 
  select(individual = fishing_entity,observation = trend,group = Georegion,value) %>% 
  arrange(group,value) %>% 
  mutate(value = ifelse(observation %in% c("Collapsed","Over Exploited"),value*-1,value),
         observation = ifelse(observation %in% c("Developing","Rebuilding"),"Category A",
                              ifelse(observation == "Max Exploited","Category B",
                                     ifelse(observation %in% c("Over Exploited","Collapsed","Category C"),"Category C", "No Category")
                              )
         )
  ) %>% 
  group_by(individual,observation,group) %>% 
  summarise(value = sum(value))

# Summary for manuscipt

D_Summary <- Discrete_Trend_Data %>% 
  mutate(value = ifelse(value <0, value*-1,value)) %>% 
  group_by(observation) %>% 
  summarise(tot=sum(value),
            mean(value),
            sd(value)) %>% 
  arrange(desc(tot))

#### Supplemental Table S2

T_Summary %>%
  left_join(D_Summary, by = "observation") %>%
    write_csv("Table_S2.csv")

# -------------------------- #
# Soley Discrete data
# -------------------------- #

Exclusive_Discrete_Trend_Data <- Discrete_Exclusive %>% 
  left_join(Trend_Data, 
            by =c("eez_name","taxon_key")
  ) %>%
  mutate(trend = ifelse(is.na(trend),"No Trend",trend)) %>% 
  group_by(fishing_entity,trend) %>% 
  summarise(value = length(unique(taxon_key))) %>% 
  left_join(UN_Regions, by ="fishing_entity") %>%
  mutate(Georegion = ifelse(is.na(georegion),"Other",paste(georegion))) %>% 
  # mutate(value = ifelse(value > 100,100,value)) %>% 
  filter(!is.na(fishing_entity),
         !str_detect(fishing_entity, ".\\(.") # remove associated territories
  ) %>% 
  select(individual = fishing_entity,observation = trend,group = Georegion,value) %>% 
  arrange(group,value) %>% 
  mutate(value = ifelse(observation %in% c("Collapsed","Over Exploited"),value*-1,value),
         observation = ifelse(observation %in% c("Developing","Rebuilding"),"Category A",
                              ifelse(observation == "Max Exploited","Category B",
                                     ifelse(observation %in% c("Over Exploited","Collapsed","Category C"),"Category C", "No Category")
                              )
         )
  ) %>% 
  group_by(individual,observation,group) %>% 
  summarise(value = sum(value))

```

### Fig. 3- Circular plot

```{r Territories_Circluar_Plot, eval = F, echo = T}

# -------------------------- #
# Data manipulation for plot
# -------------------------- #

# Set a number of 'empty bar' to add at the end of each Georegion
Char <- as.factor(Trend_Plot_Data$group) # Set groups to factors

##### COPY PASTE

# Transform data in a tidy format (long format)
# data <- Trend_Data_Plot

# Set a number of 'empty bar' to add at the end of each group
empty_bar <- 2
nObsType <- nlevels(as.factor(Trend_Plot_Data$observation))
to_add <- data.frame(matrix(NA, empty_bar*nlevels(Char)*nObsType, ncol(Trend_Plot_Data)) )
colnames(to_add) <- colnames(Trend_Plot_Data)
to_add$group <- rep(levels(Char), each=empty_bar*nObsType)
Trend_Plot_Data <- rbind(Trend_Plot_Data, to_add) # Problem converts to list
# Trend_data <- bind_rows(data,to_add)
Trend_Plot_Data <- Trend_Plot_Data %>% arrange(group, individual)
Trend_Plot_Data$id <- rep( seq(1, nrow(Trend_Plot_Data)/nObsType) , each=nObsType)

# to_add$value <- NA
# to_add$observation <- NA
# to_add$individual <- to_add$group

# Trend_data <- Trend_Data_Plot %>%
#   bind_rows(to_add) %>%
#   ungroup() %>%
#   mutate(individual = forcats::fct_explicit_na(individual),
#          group = forcats::fct_explicit_na(group)) %>%
#   group_by(group,individual) %>%
#   mutate(id = as.integer(individual),
#          order = as.integer(group))

# Get the name and the y position of each label
label_data <- Trend_Plot_Data %>% 
  # filter(value > 0) %>% 
  group_by(id, individual) %>% 
  summarize(tot=sum(value,na.rm=T)) %>% 
  mutate(
    individual = ifelse(individual == "Saint Vincent & the Grenadines","Snt. Vincent & Grenadines",
                        ifelse(individual == "Brunei Darussalam","Burnei",
                               individual)
                        )
  )

number_of_bar <- nrow(label_data)
angle <- 90 - 360 * (label_data$id-0.5) /number_of_bar     # I substract 0.5 because the letter must have the angle of the center of the bars. Not extreme right(1) or extreme left (0)
label_data$hjust <- ifelse( angle < -90, 1, 0)
label_data$angle <- ifelse(angle < -90, angle+180, angle)

label_data <- label_data %>% 
  mutate(individual = ifelse(individual == "(Missing)",paste(" "),paste(individual)))

# prepare a data frame for base lines
base_data <-  Trend_Plot_Data %>% 
  group_by(group) %>% 
  summarize(start=min(id,na.rm=T), 
            end=max(id,na.rm=T) - empty_bar,
            median = median(value, na.rm = T)) %>% 
  rowwise() %>% 
  mutate(title=mean(c(start, end)))

# prepare a data frame for grid (scales)
grid_data <- base_data
grid_data$end <- grid_data$end[ c( nrow(grid_data), 1:nrow(grid_data)-1)] + 1
grid_data$start <- grid_data$start - 1
grid_data <- grid_data[-1,]

Trend_Plot_Data %>%
  ggplot() + 
  geom_bar(aes(x=id,
               y=value, 
               fill=observation),
           stat="identity", 
           alpha=0.5) +
  # scale_fill_manual("Catch Category",
  #                   values = c(rev(wes_palette(n = 5, name = "Rushmore1")),"grey")
  #                   ) +
  scale_fill_manual("Catch Category",
                    values = c(
                      "#046C9A", # Developing / Rebulding / Cat A / blue
                      "#0B775E", #Max exp. / Cat B / green
                      "#F2300F", # collapsed / Overexp. / Cat. C / Red
                      "#8D8680"
                    ),
                    na.value = "white"
  ) +
  # Add a val=100/75/50/25 lines. I do it at the beginning to make sur barplots are OVER it.
  geom_segment(data=grid_data, aes(x = end, y = 0, xend = start, yend = 0), colour = "black", alpha=1, size=0.3 , inherit.aes = FALSE ) +
  geom_segment(data=grid_data, aes(x = end, y = 50, xend = start, yend = 50), colour = "black", alpha=1, size=0.3, inherit.aes = FALSE ) +
  geom_segment(data=grid_data, aes(x = end, y = 100, xend = start, yend = 100), colour = "black", alpha=1, size=0.3, inherit.aes = FALSE ) +
  geom_segment(data=grid_data, aes(x = end, 
                                   y = -50, 
                                   xend = start, 
                                   yend = -50), colour = "black", alpha=1, size=0.3 , inherit.aes = FALSE ) +
  # Add text showing the value of each 100/75/50/25 lines
  ggplot2::annotate("text", 
                    # x = rep(max(data$id),3),
                    x =c(0.2,0,0,0.2,0.2),
                    y = c(-100,-50, 0, 50, 100),
                    label = c("100","50","0", "50", "100") , 
                    color="black", 
                    size=5, 
                    angle=0, 
                    # fontface="bold",
                    hjust=1) +
  # Add labels on top of each bar
  geom_text(data=label_data,
            aes(
              x = id,
              y = 105,
              label=individual,
              hjust=hjust),
            color="black",
            # fontface="bold",
            alpha=1,
            size=4,
            angle= label_data$angle,
            inherit.aes = FALSE ) +
  # Continents labels
  geom_text(data=base_data,
            aes(x = title,
                y = 85, 
                label=group), 
            hjust=c(1,1,1,0,0,0),
            colour = "black",
            # alpha=0, 
            size=5,
            fontface="bold",
            inherit.aes = FALSE) +
  ylim(-100,105) +
  theme_minimal() +
  theme(legend.position = c(0.93,0.14),
        legend.key.size = unit(1,"cm"),
        legend.title = element_text(size = 16),
        legend.text = element_text(size = 14),
        axis.text = element_blank(),
        axis.title = element_blank(),
        panel.grid = element_blank(),
        plot.margin = unit(rep(-1,4), "cm")) +
  coord_polar()

# Save plot

ggsave(
  plot = last_plot(),
  width = 17,
  height = 15,
  units = "in",
  # filename = "Plot.png"
  filename = "Territory_circular_trend.png"
)

```

### Statistical differences between Catch Trend

Using MANOVA to investigate differences in catch category between discrete and transboundary species

```{r statistics_fig2, eval = F, echo = T}

#-----------------# 
# Get Data
#-----------------#

# Transboundary Data
Transboundary_Data_Stats <- Trend_Plot_Data %>% 
filter(observation != "No Category") %>%
  mutate(value = abs(value))

# Discrete Data (trans.not shared)
Trans_Dis_Data_Stats <- Discrete_Trend_Data %>% 
  filter(observation != "No Category") %>%
  mutate(value = abs(value))

# Discrete Data (just desrete and not trans.not shared)
Discrete_Data_Stats <- Exclusive_Discrete_Trend_Data %>% 
  filter(observation != "No Category") %>%
  mutate(value = abs(value))

#### By Category (Manova)

# Use manova to test for a significant effect of Catch category on discrete and transboundary species simultaneously. Takes into account possible correlations between dependent variables.
# Decreases Tipe I errors

Manova_Differences <- Transboundary_Data_Stats %>% 
  rename(Trans_value = value) %>% 
  left_join(Discrete_Data_Stats, by = c("individual","observation","group")
  ) %>% 
  rename(Discrete_value= value) %>% 
  mutate(Discrete_value = ifelse(is.na(Discrete_value),0,Discrete_value)) %>% 
  left_join(Trans_Dis_Data_Stats, by = c("individual","observation","group")) %>% 
  rename(Trans_Discrete_value= value) %>% 
  mutate(Trans_Discrete_value = ifelse(is.na(Trans_Discrete_value),0,Trans_Discrete_value))

#-----------------# 
## Test assumptions
#-----------------#

Manova_Differences %>% 
  gather("cat","val",Trans_value,Discrete_value) %>% 
  ggplot() +
  geom_histogram(
    aes(
      x = val
    )
  ) +
  facet_wrap(~observation+cat)

### Correlation

plot(Manova_Differences$Trans_value~Manova_Differences$Discrete_value)
cor(Manova_Differences$Trans_value,Manova_Differences$Discrete_value)

#-----------------# 
## MANOVA test
#-----------------#

mano <- manova(cbind(Discrete_value,Trans_value,Trans_Discrete_value) ~ observation, data = Manova_Differences)

summary(mano)
# Df  Pillai approx F num Df den Df    Pr(>F)    
# observation   2 0.21621   27.091      4    894 < 2.2e-16 ***
# Residuals   447
# ---
# Signif. codes:  0 ‘***’ 0.001 ‘**’ 0.01 ‘*’ 0.05 ‘.’ 0.1 ‘ ’ 1
# 
#  Response Trans_value :
#              Df Sum Sq Mean Sq F value    Pr(>F)    
# observation   2  22778 11389.2  55.005 < 2.2e-16 ***
# Residuals   447  92555   207.1

# 
#             Df  Pillai approx F num Df den Df    Pr(>F)    
# observation   2 0.17329   21.771      4    918 < 2.2e-16 ***
# Residuals   459                                             
# ---
# Signif. codes:  0 ‘***’ 0.001 ‘**’ 0.01 ‘*’ 0.05 ‘.’ 0.1 ‘ ’ 1

summary.aov(mano)

#-----------------# 
## Run non-parametric to make sure
# Same result
#-----------------#

# kruskal.test(Trans_value~Discrete_value, data = Manova_Differences)
# Kruskal-Wallis chi-squared = 100.24, df = 45, p-value = 4.358e-06

# New only comparing shared trans to discrete exclusive
# Kruskal-Wallis chi-squared = 50.723, df = 13, p-value = 2.24e-06

#-----------------# 
## ANOVA test for each category
#-----------------#

# Transboundary Species

#### Test for normallity 
Transboundary_Data_Stats %>%
  ggplot() +
  geom_histogram(
    aes(
      x = value
    )
  ) +
  facet_wrap(~observation)

bartlett.test(value~observation, Transboundary_Data_Stats)

# Differences between categories of transboundary species

Stat_Cat_aov <- aov(value~observation, data = Transboundary_Data_Stats)
Stat_Cat_aov
summary(Stat_Cat_aov)
#              Df Sum Sq Mean Sq F value Pr(>F)    
# observation   2  22778   11389   55.01 <2e-16 ***
# Residuals   447  92555     207

     # Df Sum Sq Mean Sq F value Pr(>F)    
# observation   2  24776   12388   53.93 <2e-16 ***
# Residuals   459 105436     230

TukeyHSD(Stat_Cat_aov)
#                            diff       lwr       upr     p adj
# Category B-Category A  5.946667  2.039431  9.853902 0.0011125
# Category C-Category A 17.160000 13.252765 21.067235 0.0000000
# Category C-Category B 11.213333  7.306098 15.120569 0.0000000

#    diff       lwr      upr     p adj
# Category B-Category A  5.948052  1.886814 10.00929 0.0018096
# Category C-Category A 17.629870 13.568633 21.69111 0.0000000
# Category C-Category B 11.681818  7.620581 15.74306 0.0000000

#-----------------# 
# Double check the Non parametic test
# same result
#-----------------#

# Stat_Cat_kt <- kruskal.test(observation~value, data = Transboundary_Data_Stats)
# Stat_Cat_kt
# 
# # Post hock test
# kruskalmc(Transboundary_Data_Stats$value,Transboundary_Data_Stats$observation)

# -------------------------- #
# Discrete species
# -------------------------- #

#### Test for normallity 
Discrete_Data_Stats %>%
  ggplot() +
  geom_histogram(
    aes(
      x = value
    )
  ) +
  facet_wrap(~observation)

bartlett.test(value~observation, Discrete_Data_Stats)

# ANOVA of Discrete categories
Stat_discrete_cat_aov <- aov(value~observation, data = Discrete_Data_Stats)
Stat_discrete_cat_aov
summary(Stat_discrete_cat_aov)

#             Df Sum Sq Mean Sq F value   Pr(>F)    
# observation   2   6442    3221   21.86 1.03e-09 ***
# Residuals   380  55992     147

# f Sum Sq Mean Sq F value   Pr(>F)    
# observation   2   5322    2661   25.58 3.76e-11 ***
# Residuals   382  39740     104

#-----------------# 
# Double check the Non parametic test
# All different... But it does not show any distance, just the categorial difference
#-----------------#

Stat_discrete_cat_kt <- kruskal.test(observation~value, data = Discrete_Data_Stats)
Stat_discrete_cat_kt

kruskalmc(Discrete_Data_Stats$value,Discrete_Data_Stats$observation)

Trans_Dis_Data_Stats

```

## Results per Species

### Mackerel conflict

```{r mackrell, eval = F, echo =F}

Exploited_Species %>% 
  filter(taxon_name == "Scomber scombrus") # 600118

NW_Europe <- UN_Regions %>% 
  filter(sub_region  %in% c("Western Europe", "Northern Europe")) %>% 
  pull(fishing_entity) %>% 
  unique()

Conflict_nations <- c(NW_Europe,"Portugal","Spain")

Mackerel <- Clean_sau_eez_fes %>% 
  group_by(fishing_entity, fishing_entity_id) %>% 
  summarise(n()) %>% 
  left_join(SAU_Sampled_Data, by = "fishing_entity_id") %>% 
  filter(taxon_key == 600118,
         fishing_entity %in% Conflict_nations
  ) %>% 
  group_by(Variable) %>% 
  summarise(Mac = sum(Value))

### Mackerel percentage of region's total catch

Conflict_Values <- Clean_sau_eez_fes %>% 
  group_by(fishing_entity, fishing_entity_id) %>% 
  summarise(n()) %>% 
  left_join(SAU_Sampled_Data, by = "fishing_entity_id") %>% 
  filter(
         fishing_entity %in% Conflict_nations
  ) %>% 
  group_by(Variable) %>% 
  summarise(Conflict= sum(Value))

Mackerel %>% 
  left_join(Conflict_Values) %>% 
  mutate(Mac/Conflict*100)

```

Create figures that are based in species (e.g. number of countries each species has)

```{r Circular_Spp_Data, eval = F, echo =F}

# Number of countries shared by each Species
Species_Trans <- Tresholded_T_Data %>% 
  group_by(taxon_key) %>% 
  summarise(
    n_countries = length(unique(fishing_entity))
  ) %>% 
  left_join(Exploited_Species) %>%
  left_join(Spp_Clasification)

# hist(Species_Trans$n_countries)

# Use rfishbase to get species environment
Species_list <- Species_Trans %>% 
  mutate(scientific_name = paste(genus,species)) %>% 
  pull(scientific_name)

Species_environ_list <- rbind(species(Species_list, #species from rfishbase package
                                      fields = species_fields$habitat[4])
) %>% 
  mutate(scientific_name = Species_list)

Raw_Species_Data <- Tresholded_T_Data %>% 
  group_by(taxon_key) %>% 
  summarise(n_trans_spp = length(unique(fishing_entity))) %>% 
  left_join(Exploited_Species) %>%
  rename(scientific_name = taxon_name) %>% 
  left_join(Species_environ_list) %>%
  mutate(DemersPelag = ifelse(is.na(DemersPelag),"Other",
                              ifelse(DemersPelag %in% c("bathypelagic","benthopelagic"), "bathy-bentho-pelagic",paste(DemersPelag)
                              )
  ),
  n_trans_spp = ifelse(n_trans_spp > 100, 100, n_trans_spp),
  DemersPelag = str_to_title(DemersPelag)
  ) %>% 
  select(individual=scientific_name,
         value = n_trans_spp,
         group = DemersPelag) %>% 
  arrange(group,value)  %>% 
  mutate(
    individual = ifelse(individual == "Chloroscombrus chrysurus", "C. chrysurus",
                        ifelse(individual == "Scomberomorus commerson", "S. commerson",
                               individual)
    )
  ) %>% 
  filter(!group %in% c("Bathydemersal","Pelagic"))

#### Ecosystem based plots

Top_5_Ecosystem <- Raw_Species_Data %>% 
  rename(taxon_name = individual) %>% 
  group_by(group) %>% 
  top_n(5,value)

# Just catch data
Catch_Data <- SAU_Sampled_Data %>% 
  filter(Variable == "catch")

# Just country names
country_names <- matching_names %>% 
  select(fishing_entity=sau_fishing_entity,fishing_entity_id) %>% 
  distinct()

# If you want to show catch  
Country_Spp_Data <- Tresholded_T_Data %>%
  left_join(Exploited_Species) %>%
  left_join(Top_5_Ecosystem) %>% # includes the groups
  group_by(taxon_key,taxon_name,fishing_entity_id,fishing_entity,group) %>%
  summarise(n()) %>% 
  left_join(Catch_Data,
            by = c("fishing_entity_id","taxon_key")
  ) %>%
  filter(taxon_name %in% Top_5_Ecosystem$taxon_name) %>%
  group_by(fishing_entity_id,group) %>%
  summarise(top_catch = sum(Value,na.rm=T)) %>%
  group_by(group) %>%
  top_n(5) %>% 
  left_join(country_names, by = "fishing_entity_id") %>% 
  mutate(
    fishing_entity = ifelse(fishing_entity == "Korea (South)", "S. Korea",
                            ifelse(fishing_entity == "Russian Federation","Russia",
                                   ifelse(fishing_entity == "Papua New Guinea", "N.Guinea",
                                          ifelse(fishing_entity == "Solomon Isl.", "Solomon",fishing_entity)))
    )
  )

```

### OHI-Species

```{r species_circluar_plot, eval = F, echo = T}

Species_Data <- Raw_Species_Data %>%
  filter(value > 20)

# Set a number of 'empty bar' to add at the end of each group
Char <- as.factor(Species_Data$group)
empty_bar <- 2
to_add <- data.frame( matrix(NA, empty_bar*nlevels(Char), ncol(Species_Data)) )
colnames(to_add) <- colnames(Species_Data)
to_add$group <- rep(levels(Char), each=empty_bar)
Species_Data <- rbind(Species_Data, to_add)
Species_Data <- Species_Data %>% arrange(group,value)
Species_Data$id <- seq(1, nrow(Species_Data))

# Get the name and the y position of each label

# Create labels for all but top 5
label_data <- Species_Data 
number_of_bar <- nrow(label_data)
angle <- 90 - 360 * (label_data$id-0.5) /number_of_bar     # I substract 0.5 because the letter must have the angle of the center of the bars. Not extreme right(1) or extreme left (0)
label_data$hjust <- ifelse( angle < -90, 1, 0)
label_data$angle <- ifelse(angle < -90, angle+180, angle)

# Get the name and the y position of each label
# Top n of each group label
Top_spp <- Species_Data %>% 
  group_by(group) %>% 
  top_n(5,value)

# Create labels for all but top 5
label_data <- Species_Data
number_of_bar <- nrow(label_data)
angle <- 90 - 360 * (label_data$id-0.5) /number_of_bar     # I substract 0.5 because the letter must have the angle of the center of the bars. Not extreme right(1) or extreme left (0)
label_data$hjust <- ifelse( angle < -90, 1, 0)
label_data$angle <- ifelse(angle < -90, angle+180, angle)

# Remove top 5
label_data_rest <- label_data %>% 
  filter(!individual %in% Top_spp$individual)

# Create labels for top 5
Top_label_data <- label_data %>% 
  filter(individual %in% Top_spp$individual)

# prepare a data frame for base lines
base_data <- Species_Data %>% 
  group_by(group) %>% 
  summarize(start=min(id), 
            end=max(id) - empty_bar, 
            median = median(value, na.rm = T)
  ) %>% 
  rowwise() %>% 
  mutate(title=mean(c(start, end)))

# prepare a data frame for grid (scales)
grid_data <- base_data
grid_data$end <- grid_data$end[ c( nrow(grid_data), 1:nrow(grid_data)-1)] + 1
grid_data$start <- grid_data$start - 1
# grid_data <- grid_data[-1,]
grid_data[1,2] <- 5 # Manually include bathydemerssal
grid_data[1,3] <- 1

# Median data
median_data <- grid_data %>% 
  mutate_at(vars(start,end),
            .funs = as.numeric)

Seq_axis_spp <- seq(20,100,20)
Plot_Limit <- max(Species_Data$value,na.rm=T)
Ticks_Color <- "black"

### ------------------- ###
# Make the plot
### ------------------- ###
Spp_circular <- ggplot(Species_Data, 
                       aes(x=as.factor(id), 
                           y=value,
                           fill=group)
) +       # Note that id is a factor. If x is numeric, there is some space between the first bar
  geom_bar(aes(x=as.factor(id),
               y=value,
               fill=group),
           stat="identity",
           alpha=0.5) +
  # Add a val=100/75/50/25 lines. I do it at the beginning to make sur barplots are OVER it.
  geom_segment(data=grid_data, 
               aes(x = end, y = 100, xend = start, yend = 100), 
               colour = Ticks_Color, alpha=1, size=0.3, inherit.aes = FALSE) +
  geom_segment(data=grid_data,
               aes(x = end, y = 80, xend = start, yend = 80),  
               colour = Ticks_Color, alpha=1, size=0.3, inherit.aes = FALSE) +
  geom_segment(data=grid_data, 
               aes(x = end, y = 60, xend = start, yend = 60),  
               colour = Ticks_Color, alpha=1, size=0.3, inherit.aes = FALSE) +
  geom_segment(data=grid_data, 
               aes(x = end, y = 40, xend = start, yend = 40),  
               colour = Ticks_Color, alpha=1, size=0.3, inherit.aes = FALSE) +
  geom_segment(data=grid_data, 
               aes(x = end, y = 20, xend = start, yend = 20), 
               colour = Ticks_Color, alpha=1, size=0.3, inherit.aes = FALSE) +
  # Add median lines for each gropup
  geom_segment(data=median_data, 
               aes(x = end, y = median, xend = start+5, yend = median, colour = group),
               alpha=1, size=0.5 , inherit.aes = FALSE) +
  # Add text showing the value of each line
  annotate("text", 
           x = rep(max(Species_Data$id),5),
           y = Seq_axis_spp,
           label = c(Seq_axis_spp[-5],">100"),
           color=Ticks_Color, 
           size=3, 
           angle=0, 
           fontface="bold",
           hjust=1) +
  geom_bar(aes(x=as.factor(id),
               y=value,
               fill=group),
           stat="identity",
           alpha=0.5) +
  # Species names
  geom_text(data=label_data_rest,
            aes(x=id, y=101,
                label=individual,
                hjust=hjust), 
            color="black",
            fontface="italic",
            alpha=1,
            size=3,
            angle= label_data_rest$angle,
            inherit.aes = FALSE) +
  # Bold top 3 
  geom_text(data=Top_label_data,
            aes(x=id, y=101,
                label=individual,
                hjust=hjust#,
                # color = group
            ),
            color="black",
            fontface="bold.italic",
            alpha=1,
            size=3,
            angle= Top_label_data$angle,
            inherit.aes = FALSE,
            show.legend = FALSE) +
  # Add base line information
  geom_segment(data=base_data, 
               aes(x = start, 
                   y = -5, 
                   xend = end, 
                   yend = -5
               ), 
               colour = "black",
               alpha=0.8, 
               size=0.6 , 
               inherit.aes = FALSE) +
  ylim(-50,110) + # sets the circle the first value the central the last value how wide
  theme_minimal() +
  theme(
    legend.position = c(1,0.11),
    legend.text = element_text(size = 13),
    legend.title = element_text(size = 14, face = "bold"),
    axis.text = element_blank(),
    axis.title = element_blank(),
    panel.grid = element_blank(),
    plot.margin = unit(rep(-3.5,4.5),"cm")
  ) +
  coord_polar() + 
  scale_fill_manual("Ecosystem Preference",
                    values = c(wes_palette(n = 5, name = "Darjeeling1"),
                               wes_palette(n = 3, name = "Darjeeling2")
                    ),
                    na.value = "white"
  ) +
  scale_colour_manual("Ecosystem Preference",
                      values = c(wes_palette(n = 5, name = "Darjeeling1"),
                                 wes_palette(n = 3, name = "Darjeeling2")
                      ),
                      na.value = "white"
  )

#### Ecosystem  preference based plots

Ecosystem_plots <- Country_Spp_Data %>%
  ungroup() %>% 
  mutate(name = reorder_within(fishing_entity,top_catch,group)) %>% # to order the facet wrap
  ggplot() +
  geom_bar(
    aes(
      x = name,
      y = top_catch/1000, # thusand tons
      fill = group
    ), 
    stat = "identity"
  ) +
  labs(x = "",
       y = "Catch (Thousand tonnes)") +
  scale_fill_manual("",
                    values = c(wes_palette(n = 5, name = "Darjeeling1"),
                               wes_palette(n = 3, name = "Darjeeling2")
                    ),
                    na.value = "white"
  ) +
  facet_wrap(~group,
             nrow=2,
             scales = "free") +
  # ggtheme_plot() +
  my_ggtheme_p() +
  scale_x_reordered() +
  theme(legend.position = "",
        axis.text.x = element_text(size = 11,
                                   angle = 0,
                                   face = "plain"),
        strip.background = element_blank(),
        strip.text.x = element_blank()
  )

### Combined plot

spp_plot <- ggdraw() +
  # Circular plot
  draw_plot(Spp_circular, x = 0, y = 0.305, width = 1, height = 0.60) +
  # Environmental maps
  draw_plot(Ecosystem_plots, x = 0.02, y = -0.029, width = 0.95, height = 0.26) +
  # Labeks
  draw_plot_label(label = c("a", "b"), 
                  size = 30,
                  x = c(0, 0), 
                  y = c(1, 0.27)
  )

save_plot(
  paste("spp_circular_catch_country.png",sep=""),
  # "spp_plot.png",
  spp_plot,
  base_height = 14,
  base_width = 14
)

```

### Statistics on Species Ecosystem Preference

```{r Spp_Stats, eval = T, echo = F}

head(Raw_Species_Data)

# Differences between groups
Stat_spp_aov <- aov(value~group, data = Raw_Species_Data)
Stat_spp_aov
summary(Stat_spp_aov)
#              Df Sum Sq Mean Sq F value Pr(>F)    
# group         5  56721   11344   53.82 <2e-16 ***
# Residuals   597 125844     211

TukeyHSD(Stat_spp_aov) %>% 
  tidy() %>% 
  filter(adj.p.value < 0.05) # Pelagic-Oceanic

```

# Sensitivity Analysis

```{r Sensitivity_Area_Treshold, eval = T, echo = F}

# Model Index
MI <- 100 # Model Index

####_____End Select Tresholds_______ ###

Area_Index_Level <- c(seq(.1,0.5,0.1),0.25)
result <- c(40,35,28,18,4,31.5)

plot(result~Area_Index_Level)

p <- list()
i = NULL

for(i in 1:length(Area_Index_Level)){

  data <- Clean_Results_Trans %>% 
    filter(
      area_index >= Area_Index_Level[i],
      area_index <= 1-Area_Index_Level[i],
      model_index == MI
    ) %>% 
    group_by(eez_name) %>% 
    summarise(
      n_trans_spp = length(unique(taxon_key)),
      n_Neighbours = length(unique(eez_neighbour)),
      Trans_Rate = n_trans_spp/n_Neighbours
    )

  Median_data <- median(data$n_trans_spp)
  print(Median_data)

  p[[i]] <- ggplot(data) +
    geom_histogram(aes(n_trans_spp),
                   binwidth = 10, 
                   colour = "white",
                   fill = ifelse(i!= 6,"grey80","grey20")
    ) +
    labs(
      x = "Transboundary species (n)",
      y = "Frequency of EEZs",
      title= paste((Area_Index_Level[i]*100),"% area threshold"," (Median ",Median_data,")",sep="")
    ) +
    # ggtitle(paste("Using a",(Area_Index_Level[i]*100),"% area treshold. (Median",Median_data,")")) +
    ggtheme_plot()
}

ggsave(
  plot = do.call(gridExtra::grid.arrange,p),
  width = 8,
  height = 8,
  units = "in",
  filename = "Area_Thresholds.png",
  path = "/Volumes/DATA/JULIANO_NEYMAR/FishForVisa/Sensitivity_Analysis"
)

```